---
title: "Combine all pollutant data and calculate additional emissions metrics"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  if (!require("pacman")) install.packages("pacman")
  pacman::p_load(tidyverse, kableExtra)
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/',
                        warning = FALSE, message = FALSE, cache = FALSE)
```

# tidy data

## load files

pollutant data

```{r}
  samples <- readRDS("../../r_data/samples.RDS")
  oc_om <- readRDS("../../r_data/oc_om.RDS")
  ecoc <- readRDS("../../r_data/ecoc_merged.RDS")
  pm <- readRDS("../../r_data/pm_merged.RDS")
  carbonyls <- readRDS("../../r_data/carbonyls_merged.RDS")
  carbohydrates <- readRDS("../../r_data/carbohydrates_merged.RDS")
  ions <- readRDS("../../r_data/ions_merged.RDS")
  vocs <- readRDS("../../r_data/vocs_merged.RDS")
  pahs <- readRDS("../../r_data/pah_merged.RDS")
  ultrafines <- readRDS("../../r_data/ultrafines_merged.RDS")
  dilution <- readRDS("../../r_data/dilution_merged.RDS")
  gases <- readRDS("../../r_data/fivegas_merged_avg.RDS")
```

constants

```{r}
  hood_flow <- readRDS("../../r_data/hood_flow.RDS")
  mw_c <- readRDS("../../r_data/mw_c.RDS")
  pol_properties <- readRDS("../../r_data/pol_properties.RDS")
```

# combine pollutants

```{r}
  emissions <-
    ecoc %>%
    dplyr::bind_rows(pm, carbonyls, carbohydrates, ions, vocs, pahs, gases, ultrafines)
```

# calculate additional metrics

* correct for dilution

```{r}
  emissions <- 
    emissions %>%
    dplyr::rename("conc" = "val") %>%
    dplyr::left_join(pol_properties, by = "pol") %>%
    dplyr::mutate(mass_emitted = conc * hood_flow * dur,
                  carbon_fraction = num_c * mw_c / mw,
                  mass_carbon = mass_emitted * carbon_fraction) %>%
    dplyr::select(id, pol_category, pol, conc, mass_emitted, mass_carbon, qc, bg) %>%
    tidyr::gather("metric", "val", 4:6) %>%
    dplyr::inner_join(dilution, by = "id") %>%
    dplyr::mutate(val = ifelse(pol == "ultrafines", val * (1 / ratio), val),
                  units = ifelse(metric == "conc", "micrograms per meter cubed", "micrograms"),
                  units = ifelse(metric == "conc" & pol == "ultrafines", "number per meter cubed", units),
                  units = ifelse(metric == "mass_emitted" & pol == "ultrafines", "number per meter cubed", units)) %>%
    dplyr::select(-ratio)
```

```{r}
  om <- 
    emissions %>%
    dplyr::filter(pol == "oc") %>%
    dplyr::left_join(samples %>%
                       dplyr::select(id, stove), by = "id") %>%
    dplyr::left_join(oc_om , by = "stove") %>%
    dplyr::mutate(val = val * ratio,
                  pol_category = "organic matter") %>%
    dplyr::select(-ratio, -stove)
```

```{r}
  emissions <-
    emissions %>%
    dplyr::bind_rows(om)
```

# save

```{r}
  saveRDS(emissions, file = "../../r_data/emissions.RDS")
```

## plot: aerosol mass balance

add stove fuel info

```{r}
  emissions <-
    emissions %>%
    dplyr::left_join(samples %>%
                       dplyr::select(id, stove, fuel, stovecat, date, fuelcat), by = "id")
```

select pollutants

```{r}
  ec_oc_ions <-
    emissions %>%
    dplyr::filter(grepl("ions|ec|^oc$|pm", pol_category)) %>%
    dplyr::filter(metric == "conc") %>%
    dplyr::group_by(id, pol_category, stove, fuel, fuelcat) %>%
    dplyr::summarise(val = sum(val, na.rm = TRUE)) %>%
    tidyr::spread(pol_category, val) %>%
    na.omit %>%
    tidyr::gather(pol_category, val, 5:8) #%>%
    #dplyr::mutate(val = ifelse(pol_category == "oc", val * 1.3, val))
```

```{r}
  ratio <-
    emissions %>%
    dplyr::filter(grepl("ec|^oc$|pm", pol_category)) %>%
    dplyr::filter(metric == "conc") %>%
    dplyr::select(id, pol, val, stove, fuel, fuelcat) %>%
    tidyr::spread(pol, val) %>%
    dplyr::mutate(ratio = (ec + oc) / pm_mass)
```

```{r, fig.width=12, fig.height=10}
  ratio %>%
    dplyr::filter(id != "25C") %>%
    ggplot(aes(x = id, y = ratio)) +
      geom_point() + 
      theme_bw() +
      facet_wrap(stove~fuel, scales = "free", ncol = 5)
```

```{r figure_1, echo=FALSE, fig.width=12, fig.height=10}
  ec_oc_ions %>%
  dplyr::filter(pol_category == "pm_mass") %>%
  ggplot(aes(x = id, y = val, fill = pol_category)) +
    geom_histogram(stat = "identity") +
    geom_histogram(data = ec_oc_ions %>%
                     dplyr::filter(pol_category != "pm_mass"),
                     aes(x = id, y = val, fill = pol_category), stat = "identity", width = 0.6) +
    scale_fill_manual(values = c("gray30", "#E69F00", "#56B4E9", "#999999")) +
    theme_bw() +
    theme(text = element_text(size = 12),
          legend.position = "top") +
    guides(fill = guide_legend(title = NULL)) +
    facet_wrap(stove~fuel, scales = "free", ncol = 5)
```

## table: missing data

```{r, echo = FALSE} 
  dplyr::select(samples, id, stove, fuel) %>%
  dplyr::filter(!grepl("^G", id)) %>%
  dplyr::filter(!grepl("^27|^28|^29|^30", id)) %>%
  dplyr::left_join(emissions %>%
                     dplyr::select(id, pol_category) %>%
                     dplyr::distinct(),
                   by = "id") %>%
  dplyr::mutate(exists = "yes") %>%
  tidyr::spread(pol_category, exists) %>%
  knitr::kable("markdown", align = 'c')
```

## table: number of data points (by pollutant)

```{r, echo=FALSE}
  emissions %>%
    dplyr::filter(metric == "conc") %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = nrow(.)) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val < 10, "red", "blue"))) %>%
    dplyr::arrange(val) %>%
    knitr::kable("markdown", align = 'c')
```

## table: percent of data below background (by pollutant)

```{r, echo=FALSE}
  emissions %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., bg == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    dplyr::arrange(val) %>%
    knitr::kable("markdown", align = 'c')
```

## table: percent of data below background (by test)

```{r, echo=FALSE}
  emissions %>%
    dplyr::group_by(id) %>%
    dplyr::do(val = round(nrow(filter(., bg == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    dplyr::arrange(val) %>%
    knitr::kable("markdown", align = 'c')
```

## table: percent of data below background (for study)

```{r, echo=FALSE}
  emissions %>%
    dplyr::do(val = round(nrow(filter(., bg == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## total number of pollutants (including pm mass)

```{r, echo=FALSE}
    emissions %>%
    dplyr::select(pol) %>%
    dplyr::distinct() %>%
    nrow(.)
```
