---
title: "background analysis template"
author: "Nicholas Good"
date: "12/15/2016"
output: html_document
---

# Setup

1. Create working directory
2. Download lastest .Rda files folder from server (r_files) and place in working directory
3. Create RStudio project in working directory

# Libraries

The dataset has been optimized to work with `tidyverse`. It is strongly recommended you use `tidyverse` packages for data manipulation and plotting.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

# Source files

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

# Load data

* First we'll load the sample info data.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
```

* Next we'll load the sample times file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/test_times.Rda")
```

* Next we'll load the pollutant data we're interested in. In this case we're going to look at the CO2 data:

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/fivegas_merged.Rda")
  pol_data <- dplyr::filter(fivegas_merged, pol == "co2")
```

# Extract data

* by fivegas background log times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_pre <- dplyr::filter(test_times,
                             var == "bg_pre_start_fivegas" | var == "bg_pre_end_fivegas") %>%
               tidyr::spread(var, value) %>%
               dplyr::select(-date) %>%
               dplyr::rename(start = bg_pre_start_fivegas,
                             end = bg_pre_end_fivegas)

  times_post <- dplyr::filter(test_times,
                             var == "bg_post_start_fivegas" | var == "bg_post_end_fivegas") %>%
               tidyr::spread(var, value) %>%
               dplyr::select(-date) %>%
               dplyr::rename(start = bg_post_start_fivegas,
                             end = bg_post_end_fivegas)

```

* 5 minutes before and after each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_pre_5 <- dplyr::filter(test_times, var == "ignite") %>%
                tidyr::spread(var, value) %>%
                dplyr::select(-date) %>%
                dplyr::rename(end = ignite) %>%
                dplyr::mutate(start = end - 5*60)

  times_post_5 <- dplyr::filter(test_times, var == "end") %>%
                tidyr::spread(var, value) %>%
                dplyr::select(-date) %>%
                dplyr::rename(start = end) %>%
                dplyr::mutate(end = start + 5*60)
```

* join time data with precedence for the logged times over estimated times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_pre <- dplyr::anti_join(times_pre_5, times_pre, by = "id") %>%
               dplyr::bind_rows(times_pre)

  times_post <- dplyr::anti_join(times_post_5, times_pre, by = "id") %>%
               dplyr::bind_rows(times_post)
```

* sample times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_sample <- dplyr::filter(test_times,
                                var == "start_1" | var == "shutdown") %>%
                  tidyr::spread(var, value) %>%
                  dplyr::select(-date) %>%
                  dplyr::rename(start = start_1, end = shutdown)
```

* extract data for times windows

Append type column with "pre", "post" and "sample"

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  df_pre <- filter_times(times_pre, pol_data)
  df_pre <- dplyr::mutate(df_pre, type = "pre")

  df_post <- filter_times(times_post, pol_data)
  df_post <- dplyr::mutate(df_post, type = "post")

  df_sample <- filter_times(times_sample, pol_data)
  df_sample <- dplyr::mutate(df_sample, type = "sample")

  background_data <- dplyr::bind_rows(df_pre, df_post, df_sample)
```

## Analyze

Calculate mean pollutant value for each period

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
         background_means <- dplyr::group_by(background_data, id, type) %>%
         dplyr::summarise(mean = mean(val)) %>%
         dplyr::ungroup()
```

* Plot

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 12, fig.height=8}
  ggplot(background_means, aes(id, mean, colour = type)) +
  geom_point() +
  ggtitle("CO2") +
  theme_minimal()
```
