---
title: "background analysis template"
author: "Nicholas Good"
date: "12/15/2016"
output: html_document
---

# Setup

1. Create working directory
2. Download lastest .Rda files folder from server (r_files) and place in working directory
3. Create RStudio project in working directory

# Libraries

The dataset has been optimized to work with `tidyverse`. It is strongly recommended you use `tidyverse` packages for data manipulation and plotting.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

# Load data

* First we'll load the sample info data.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda") 
```

* Next we'll load the pollutnant data we're interested in. In this case we're going to look at the VOCs:

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/voc_merged.Rda")
```

# Extract background tests and benzene data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}  
  bz_bg <- filter(voc_merged, grepl("^G", id) == TRUE, pol == "voc_benzene")
```

## Plot background benzene

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}  
  ggplot(bz_bg, aes(date, val, color = qc)) +
    geom_point() +
    theme_minimal()
```

# Extract all benzene tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}  
  bz <- filter(voc_merged, grepl("^[0-9]|^G", id) == TRUE, pol == "voc_benzene", val != -8888)
```

## Plot normal and background distributions

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}  
  ggplot(bz, aes(x=val, y=..density.., fill = type)) +
    geom_histogram(alpha = 0.4) +
    theme_minimal() +
    scale_x_log10()
```

# Merge with fuel stove info

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}  
  bz <- left_join(bz, select(samples, id, fuel, stove), by = "id")
```


