---
title: "Process carbonyl data"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  if (!require("pacman")) install.packages("pacman")
  pacman::p_load(tidyverse, knitr, GGally, gridExtra)
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/', warning = FALSE, message = FALSE, cache = FALSE)
```

```{r, echo=FALSE}
  source("../../scripts/tidy.R")
  source("../../scripts/plots.R")
```

# load data

* carbonyls

```{r}
  carbonyls <- readRDS("../../r_data/ions.RDS")
```

* metadata

```{r}
  samples <- readRDS("../../r_data/samples.RDS")
  carb_flows <- readRDS("../../r_data/carb_flows.RDS")
  carb_times <- readRDS("../../r_data/carb_times.RDS")
  notes <- readRDS("../../r_data/notes.RDS")

  data_2 <- readRDS("../../r_data/data_2.RDS")
```

# organize

## extract ions data

* separate from carbonyls in same file
* remove G8B (breakthrough test)

```{r}
  carbonyls_merged <- dplyr::select(carbonyls, 27:42, type, id, id_ions) %>%
                      dplyr::filter(type == "test" | type == "bg", id != "G8B")
```

## flows and times

* average flows

```{r}
  flows <- dplyr::select(carb_flows, -date) %>%
           dplyr::group_by(id, type) %>%
           dplyr::summarise(flow = mean(value, na.rm = TRUE)) %>%
           dplyr::group_by(id) %>%
           dplyr::summarise(flow = mean(flow, na.rm = TRUE))
```

* match cassette timestamps to id

```{r}
  times <- dplyr::select(carb_times, -date) %>%
           tidyr::spread(type, value)
```

* merge ions with metadata

```{r}
  carbonyls_merged <- dplyr::left_join(carbonyls_merged,
                                  flows,
                                  by = "id") %>%
                 dplyr::left_join(times,
                                  by = "id")
```

# qc

load notes

```{r}
  notes <- dplyr::filter(notes, grepl("carb|carbonyl|carbonyls|all", inst) == TRUE)
```

```{r}
  knitr::kable(dplyr::filter(notes, grepl(".*carb.*", inst)) %>% 
               dplyr::select(-inst),
               "markdown", digits = 2)
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r}
  flags <- dplyr::select(notes, id, qc) %>%
           group_by(id) %>%
           arrange(qc) %>%
           summarise(qc = first(qc))
```

* merge

```{r}
  carbonyls_merged <- dplyr::left_join(carbonyls_merged, flags, by = "id") %>%
                      dplyr::mutate(id = as.factor(id)) %>%
                      dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

* additional bad tests

```{r}
  # carbonyls_merged$qc[carbonyls_merged$id == ""] <- "bad"
```

# reformat

* convert to longer format

```{r}
  carbonyls_merged <- tidyr::gather(carbonyls_merged, "pol", "mass_ug",
                                    -id, -id_ions, -flow,
                                    -start, -end, -type, -qc) %>%
                      dplyr::filter(!is.na(pol)) %>%
                      dplyr::mutate(pol = as.factor(pol)) %>%
                      dplyr::select(-id_ions)
```

# background analysis

* extract background data
* remove missing data
* calculate average concentration emitted ( and other stats)

```{r}
  bg <- dplyr::filter(carbonyls_merged, type == "bg", qc == "ok") %>%
        dplyr::mutate(mass_ug = ifelse(is.na(mass_ug), 0, mass_ug)) %>%
        na.omit() %>%
        dplyr::mutate(dur = end - start) %>%
        # ug * 1000 * 60 / (L/min * s) = ug / m^3
        dplyr::mutate(conc = (mass_ug * 1000 * 60) / (flow * dur)) %>%
        dplyr::group_by(pol) %>%
        dplyr::summarise(mean = mean(conc),
                         sd = sd(conc),
                         min = min(conc),
                         max = max(conc),
                         n = n())
```

* background concentration ($\mu g/m^3$)

```{r, echo=FALSE}
  knitr::kable(bg, "markdown", digits = 2)
```

* merge with test data

```{r}
  bg <- dplyr::select(bg, pol, mean) %>%
        dplyr::rename(conc_bg = mean)

  carbonyls_merged <- dplyr::left_join(carbonyls_merged, bg, by = "pol") %>%
                      dplyr::mutate(conc_bg = ifelse(is.na(conc_bg), 0, conc_bg))
```

## save merged file

```{r}
  saveRDS(carbonyls_merged, file = "../../r_data/carbonyls_merged.RDS")
```

## summary

Carbonyl data was collected during `r length(unique(carbonyls_merged$id))` experiments. There is no carbonyl data for tests: `r setdiff(as.character(samples$id), as.character(carbonyls_merged$id))`.

# plots

## mass by pollutant

```{r, echo=FALSE, fig.width=12, fig.height=8}
  ggplot(carbonyls_merged, aes(id, mass_ug, group = pol, color = pol)) +
         geom_point() +
         geom_line() +
         scale_y_log10() +
         theme_minimal() +
         ylab("ug") +
         theme(legend.position="top") +
         theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
         facet_wrap(~type, ncol = 1, scales = "free")
```

## qc

```{r, echo=FALSE, fig.width=12, fig.height=8}
  ggplot(carbonyls_merged, aes(id, mass_ug, group = pol, color = qc)) +
         geom_point() +
         geom_line() +
         scale_y_log10() +
         theme_minimal() +
         ylab("ug") +
         theme(legend.position="top") +
         theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
         facet_wrap(~type, ncol = 1, scales = "free")
```
