---
title: "Process ecoc data"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  if (!require("pacman")) install.packages("pacman")
  pacman::p_load(tidyverse, kableExtra)
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/', 
                        fig.width = 12, fig.height = 5,
                        warning = FALSE, message = FALSE, cache = FALSE)
```

# tidy data

## load files

load scripts

```{r}
  source("../../scripts/functions.R")
```

load data

```{r}
  ecoc <- readRDS("../../r_data/ecoc.RDS")
  filter_area <- readRDS("../../r_data/filter_area.RDS")
  times <- readRDS("../../r_data/filter_times.RDS")
  flows <- readRDS("../../r_data/filter_flows.RDS")
  notes <- readRDS("../../r_data/notes.RDS")
  samples <- readRDS("../../r_data/samples.RDS")
```

## fix incorrect ids

```{r}
  ecoc <- 
    ecoc %>%
    dplyr::mutate(id = as.character(id),
                  cassette = as.character(cassette),
                  cassette = ifelse(grepl("blank", id), NA, cassette)) %>%
    dplyr::filter(!grepl("system", id), !is.na(id), type != "pilot")

  ecoc$id[172] <- "11C"
  ecoc$id[173] <- "11C"
```

## flags

* extract flags

```{r}
  notes <- 
    notes %>%
    dplyr::filter(grepl("ecoc|all", inst) == TRUE)
```

* bad preceeds maybe preceeds good

```{r}
  flags <-
    notes %>%
    dplyr::select(id, qc) %>%
    dplyr::group_by(id) %>%
    dplyr::arrange(qc) %>%
    dplyr::summarise(qc = first(qc))
```

* add flag by id

```{r}
  ecoc <-
    ecoc %>%
    dplyr::left_join(flags, by = "id") %>%
    dplyr::mutate(id = as.factor(id)) %>%
    dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

## test type 

define test type variable

```{r}
  ecoc <- 
    ecoc %>%
    
    dplyr::mutate(test_type = ifelse(grepl("^[0-9]", id), "test", "bg"),
                  test_type = ifelse(grepl("^B", id), "blank", test_type)) %>%
    dplyr::filter(!grepl("MC", id))
```

# missing and bad data

## table: missing data

```{r, echo = FALSE}
  dplyr::select(samples, id, stove, fuel) %>%
    dplyr::anti_join(ecoc %>%
                     dplyr::select(id) %>%
                     dplyr::distinct(),
                   by = "id") %>%
    knitr::kable("markdown", align = 'c')
```

## table: tests with bad measurements

```{r, echo=FALSE}
  notes %>%
    dplyr::filter(grepl("ecoc|all", inst) == TRUE,
                  qc != "ok") %>%
    dplyr::select(-inst) %>%
    knitr::kable("markdown", align = 'c')
```

## remove bad tests

```{r}
  ecoc_merged <-
    ecoc %>% 
    dplyr::filter(qc != "bad")
```

## reformat

```{r}
  ecoc_merged <- 
    ecoc_merged %>%
      dplyr::select(id, oc_ug_sq_cm, ec_ug_sq_cm,
                    cassette, test_type, qc) %>%
      tidyr::gather("pol", "val", ec_ug_sq_cm, oc_ug_sq_cm) %>%
      dplyr::mutate(pol = ifelse(pol == "ec_ug_sq_cm", "ec", pol),
                    pol = ifelse(pol == "oc_ug_sq_cm", "oc", pol),
                    val = val * filter_area)
```

# blank analysis 

extract blanks

```{r}
  blanks <-
    ecoc_merged %>%
    dplyr::filter(test_type == "blank")
```

## table: lab blank outliers

outliers (z > 3) when blanks are grouped by pollutant and filter type

```{r, echo=FALSE}
  blanks %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(pol) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::select(id, pol, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

conclusions: 

## plot: time series of lab blanks

lab blanks (ug)

```{r ecoc_blank_timeseries, echo=FALSE, fig.height = 6, fig.width = 8}
 blanks %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::left_join(dplyr::select(samples, blank, date), by = c('id' = 'blank')) %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(color = qc), size = 2) +
      guides(fill = FALSE) +
      theme_bw() +
      theme(text = element_text(size = 20),
            legend.position = "top") +
      facet_wrap(~pol, scales = "free", ncol = 1) +
      ylab("micrograms")
```

conclusions: 

* blanks look good!
* no obvious trends in time.

## table: lab blanks with NA values

percent of lab blanks with NA values

```{r, echo=FALSE}
  blanks %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## table: blank summary

summarize blanks

```{r}
  blanks <-
    blanks %>%
    dplyr::group_by(pol) %>%
    dplyr::summarise(mean = mean(val, na.rm = TRUE),
                     median = median(val, na.rm = TRUE),
                     stdev = sd(val, na.rm = TRUE),
                     lod = stdev * 3 + mean,
                     loq = stdev * 10 + mean)
```

summary of lab blanks (ug)

```{r, echo=FALSE}
  blanks %>%
    knitr::kable(digits = 2, align = 'c')
```

## save blanks

```{r}
  saveRDS(blanks, file = "../../r_data/ecoc_lod.RDS")
```

remove lab blanks from merged file

```{r}
  ecoc_merged <-
    ecoc_merged %>%
    dplyr::filter(test_type != "blank") # remove blanks
```

add lod variable

```{r}
  ecoc_merged <-
    ecoc_merged %>%
    dplyr::left_join(dplyr::select(blanks, pol, lod) %>%
                     dplyr::rename("lod_val" = "lod"), by = "pol") %>%
    dplyr::mutate(lod = ifelse(val <= lod_val, "below", "above"),
                  val = ifelse(val <= lod_val, lod_val / sqrt(2), val)) %>%
    dplyr::select(-lod_val)
```

# concentration calculations

## extract data

extract flow rate data and average

* match cassette flows to id

```{r}
 flows <- 
  flows %>%
  dplyr::select(-date) %>%
  dplyr::filter(colour == "white" | colour == "orange") %>%
  dplyr::group_by(id, type, colour) %>%
  dplyr::summarise(flow = mean(value, na.rm = TRUE)) %>%
  dplyr::group_by(id, colour) %>%
  dplyr::summarise(flow = mean(flow, na.rm = TRUE)) %>%
  dplyr::mutate(colour = as_factor(colour)) %>%
  dplyr::mutate(cassette =  forcats::fct_recode(colour, "a" = "white", "e" = "orange")) %>%
  dplyr::select(-colour)
```

extract time data

```{r}
  times <-
    times %>%
    dplyr::select(-date) %>%
    dplyr::filter(color == "white" | color == "orange") %>%
    tidyr::spread(type, value) %>%
    dplyr::mutate(cassette = forcats::fct_recode(color, "a" = "white", "e" = "orange")) %>%
    dplyr::select(-color)
```

## merge and calculate

merge data with flows and times

```{r}
  ecoc_merged <-
    ecoc_merged %>%
    dplyr::left_join(flows, by = c("id", "cassette")) %>%
    dplyr::left_join(times, by = c("id", "cassette"))
```

calculate concentrations

```{r}
  ecoc_merged <-
    ecoc_merged %>%
    dplyr::mutate(dur = (end - start) / 60,
                  val = val * 1000 / (flow * dur))
```

# backgrounds

extract background and remove any data with no cassette

```{r}
  bg <-
    ecoc_merged %>%
    dplyr::filter(test_type == "bg", grepl("a|e", cassette))
```

## plot: lab background outliers

outliers (z > 3) when backgrounds are grouped by pollutant and filter type

```{r pah_bg_plot, echo=FALSE, fig.height = 8}
  bg %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(pol) %>%
    dplyr::mutate(z_score = (val- mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, pol, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    knitr::kable("markdown", align = 'c')
```

conclusions: 

* no backgrounds to investigate

## plot: time series of lab backgrounds

background concentrations (ug per meter cubed)

* shape indicates if sample was above or below instrument lod

```{r ecoc_bg_timeseries, echo=FALSE, fig.height = 6, fig.width = 12}
  bg %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::left_join(dplyr::select(samples, id, date), by = 'id') %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(shape = lod, color = qc), size = 4, stroke = 2) +
      scale_shape(solid = FALSE) +
      guides(fill = FALSE) +
      theme_bw() +
      theme(text = element_text(size = 20),
            legend.position = "top") +
      facet_wrap(pol~cassette, scales = "free", ncol = 2) +
      ylab("concentration") +
      xlab("")
```

conclusions: 

## table: lab backgrounds below lod for  analysis

percent of lab backgrounds below the lod of chemical analysis

```{r}
  bg %>%
    dplyr::group_by(pol, cassette) %>%
    dplyr::do(val = round(nrow(filter(., lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## table: lab backgrounds with NA values

percent of lab backgrounds with NA values

```{r}
  bg %>%
    dplyr::group_by(pol, cassette) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## table: background summary

summarize backgrounds

```{r}
  bg <-
    bg %>%
    dplyr::group_by(pol, cassette) %>%
    dplyr::summarise(mean = mean(val, na.rm = TRUE),
                     median = median(val, na.rm = TRUE),
                     stdev = sd(val, na.rm = TRUE),
                     min = min(val, na.rm = TRUE),
                     max = max(val, na.rm = TRUE),
                     count = n()) %>%
    dplyr::mutate(mean = ifelse(is.nan(mean), NA, mean),
                  median = ifelse(is.nan(median), NA, median),
                  stdev = ifelse(is.nan(stdev), NA, stdev),
                  min = ifelse(is.infinite(min), NA, min),
                  max = ifelse(is.infinite(max), NA, max),
                  count = ifelse(is.nan(count), NA, count))
```

summary of backgrounds (ug per meter cubed)

```{r}
  bg %>%
    knitr::kable(digits = 2, align = 'c')
```

## save backgrounds

```{r}
  saveRDS(bg, file = "../../r_data/ecoc_bg.RDS")
```


# sample analysis

remove bad tests

```{r}
  ecoc_merged <-
    ecoc_merged %>%
    dplyr::filter(!grepl("6D|8D", id))
```

## plot: time series of burns

burn concentrations (ug per meter cubed)

* samples are not background corrected
* shape indicates if sample was above or below instrument lod

```{r ecoc_samples_timeseries, echo=FALSE, fig.height = 6, fig.width = 14}
 ecoc_merged %>%
    dplyr::filter(test_type == "test") %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::left_join(dplyr::select(samples, id, date), by = 'id') %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(shape = lod, color = qc), size = 4, stroke = 2) +
      scale_shape(solid = FALSE) +
      guides(fill = FALSE) +
      theme_bw() +
      theme(text = element_text(size = 20),
            legend.position = "top") +
      labs(shape = "lod") +
      facet_wrap(pol~cassette, scales = "free", ncol = 2) +
      ylab("concentration") +
      xlab("")
```

conclusions:

* 6D and 8D have very high values, remove for now

## plot: comparing background and sample concentrations

test and background concentrations (ug per meter cubed)

* data is summed across all filter types
* plotted and colored by test type

```{r ecoc_test_bg_conc, echo=FALSE, fig.height = 8, fig.width = 10}
  ecoc_merged %>%
    dplyr::group_by(id, test_type) %>%
    dplyr::filter(!is.na(val)) %>%
    ggplot(aes_string(y = "val", x = "test_type")) +
      geom_boxplot(outlier.colour = NA) +
      geom_jitter(aes(color = qc, shape = lod), size = 3, stroke = 2) +
      theme_bw() +
      scale_shape(solid = FALSE) +
      guides(fill = FALSE) +
      scale_y_log10() +
      theme(text = element_text(size = 20),
            legend.position = "top") +
      facet_wrap(pol~cassette, scales = "free", ncol = 2) +
      ylab("concentration") +
      xlab("")
```

conclusions: 

* some overlap between burn and background concentrations
* need to investigate ways to account for background measurements (for now, subtract the median)

## subtract background

```{r}
  ecoc_merged <-
    ecoc_merged %>%
    dplyr::filter(test_type == "test")
```

```{r}
  ecoc_merged <-
    ecoc_merged %>%
    dplyr::left_join(bg, by = c("pol", "cassette"))
```

```{r}
  ecoc_merged <-
   ecoc_merged %>%
    dplyr::mutate(val = val - median) %>%
    dplyr::select(-min, -max, -count, -mean, -median, -stdev)
```

## table: measurements below the background median

percentage of test concentrations below background median

```{r, echo=FALSE}
  ecoc_merged %>%
    dplyr::left_join(dplyr::select(samples, id, stove, fuel), by = "id") %>%
    #dplyr::filter(stove == "lpg") %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., val < 0))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

## subtract artifact filters

```{r}
  ecoc_artifact <-
    ecoc_merged %>%
    dplyr::filter(cassette == "a") %>%
    dplyr::select(id, pol, val) %>%
    dplyr::rename("artifact" = "val")

  ecoc_merged <-
    ecoc_merged %>%
    dplyr::filter(cassette == "e") %>%
    dplyr::left_join(ecoc_artifact, by = c("id", "pol")) %>%
    dplyr::mutate(val = val - artifact) %>%
    dplyr::select(-artifact, -cassette)
```

## plot: PM concentrations by stove type

concentrations (ug per meter cubed)

* samples are background corrected

```{r ecoc_stovetype, echo=FALSE, fig.height = 10, fig.width = 16}
  ecoc_merged_p <-
    ecoc_merged %>%
    dplyr::filter(!is.na(val), !grepl("27|28|29|30", id)) %>%
    dplyr::left_join(dplyr::select(samples, id, stove, fuel, stovecat, fuelcat), by = "id") %>%
    dplyr::mutate(stove = 
                    stove %>% 
                    tolower %>%
                    gsub(" [:(:].*[:):]", "", .) %>%
                    forcats::fct_relevel("three stone fire",
                                         "clay ring",
                                         "ceramic charcoal",
                                         "rocket elbow",
                                         "built-in justa",
                                         "metal charcoal",
                                         "fan rocket elbow",
                                         "gasifier",
                                         "wick kerosene",
                                         "pressure kerosene",
                                         "lpg"),
                  fuel = 
                    fuel %>% 
                    tolower,
                  fuelcat =
                    fuelcat %>%
                    forcats::fct_recode("liquid" = "advanced"))

 ecoc_merged_p %>%
    ggplot(aes_string(x = "stove", y = "val")) +
      geom_boxplot(outlier.colour = NA) +
      geom_jitter(aes(color = qc, shape = lod), size = 4,  width = 0.2, stroke = 2) + 
      scale_shape(solid = FALSE) +
      guides(fill = FALSE) +
      theme_bw() +
      scale_y_log10() +
      theme(text = element_text(size = 20),
            axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            legend.position = "top") +
      facet_grid(pol~fuelcat, scales = "free", space = "free_x") +
      ylab("concentration") +
      xlab("")
```

conclusions: 

* some liquid fuel stoves have NA values for pollutants of interest (would rerunning these samples at higher concentrations allow us to resolve values for these stoves?)

## table: outliers by stove type

outliers (z > 3) when samples are grouped by pollutant and stove type

```{r, echo=FALSE}
  ecoc_merged_p %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(stove, pol) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, pol, stove, fuel, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    dplyr::rename("particle type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

conclusion:

* no outliers. Looks good! 

## plot: PM concentrations by fuel type

concentrations (ug per meter cubed):

* samples are background corrected

```{r ecoc_fueltype, echo=FALSE, fig.height = 10, fig.width = 16}
  ecoc_merged_p %>%
    ggplot(aes_string(x = "fuel", y = "val")) +
      geom_boxplot(outlier.colour = NA) +
      geom_jitter(aes(color = qc, shape = lod), size = 4,  width = 0.2, stroke = 2) + 
      scale_shape(solid = FALSE) +
      guides(fill = FALSE) +
      theme_bw() +
      scale_y_log10() +
      theme(text = element_text(size = 20),
            axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            legend.position = "top") +
      facet_grid(pol~fuelcat, scales = "free", space = "free_x") +
      ylab("concentration") +
      xlab("")
```

conclusions: 

## table: outliers by fuel type

outliers (z > 3) when samples are grouped by pollutant and fuel type

```{r, echo=FALSE}
  ecoc_merged_p %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(fuel, pol) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, stove, fuel, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    knitr::kable("markdown", align = 'c')
```

conclusions: 

* no repeat tests recommended

## table: burns below lod for chemical analysis

percent of burn samples below the lod

```{r}
  ecoc_merged %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## table: burns with NA values

percent of burn samples with NA values

```{r}
  ecoc_merged %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## save data

```{r}
  ecoc_merged <-
    ecoc_merged %>%
    dplyr::select(id, pol, val, lod, qc) %>%
    dplyr::mutate(units = "micrograms per meter cubed")

  saveRDS(ecoc_merged, file = "../../r_data/ecoc_merged.RDS")
```
