---
title: "background analysis template"
author: "Nicholas Good edited by Kelsey Bilsback"
date: "12/15/2016"
output: html_document
---

# Setup

1. Create working directory
2. Download lastest .Rda files folder from server (r_files) and place in working directory
3. Create RStudio project in working directory

# Libraries

The dataset has been optimized to work with `tidyverse`. It is strongly recommended you use `tidyverse` packages for data manipulation and plotting.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(wesanderson)
```

# Source files

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

# Load data

* First we'll load the sample info data.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
```

* Next we'll load the sample times file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/test_times.Rda")
```

* Next we'll load the pollutant data we're interested in. In this case we're going to look at the CO2 data:

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/smps_ultrafine_bkgd.Rda")
  pol_data <- dplyr::filter(smps_ultrafine_bkgd, pol == "ufp_conc")
```

# Extract data

* by fivegas background log times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_pre <- dplyr::filter(test_times,
                             var == "bg_pre_start_pax_smps" | var == "bg_pre_end_pax_smps") %>%
               tidyr::spread(var, value) %>%
               dplyr::select(-date) %>%
               dplyr::rename(start = bg_pre_start_pax_smps,
                             end = bg_pre_end_pax_smps)

  times_post <- dplyr::filter(test_times,
                             var == "bg_post_start_pax_smps" | var == "bg_post_end_pax_smps") %>%
               tidyr::spread(var, value) %>%
               dplyr::select(-date) %>%
               dplyr::rename(start = bg_post_start_pax_smps,
                             end = bg_post_end_pax_smps)

```

* 5 minutes before and after each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_pre_5 <- dplyr::filter(test_times, var == "ignite") %>%
                 tidyr::spread(var, value) %>%
                 dplyr::select(-date) %>%
                 dplyr::rename(end = ignite) %>%
                 dplyr::mutate(start = end - 5*60)

  times_post_5 <- dplyr::filter(test_times, var == "end") %>%
                  tidyr::spread(var, value) %>%
                  dplyr::select(-date) %>%
                  dplyr::rename(start = end) %>%
                  dplyr::mutate(end = start + 5*60)
```

* join time data with precedence for the logged times over estimated times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_pre <- dplyr::anti_join(times_pre_5, times_pre, by = "id") %>%
               dplyr::bind_rows(times_pre)

  times_post <- dplyr::anti_join(times_post_5, times_pre, by = "id") %>%
                dplyr::bind_rows(times_post)
```

* sample times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_sample <- dplyr::filter(test_times,
                                var == "start_1" | var == "shutdown") %>%
                  tidyr::spread(var, value) %>%
                  dplyr::select(-date) %>%
                  dplyr::rename(start = start_1, end = shutdown)
```

* extract data for times windows

Append type column with "pre", "post" and "sample"

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  df_pre <- filter_times(times_pre, pol_data)
  df_pre <- dplyr::mutate(df_pre, type = "pre")

  df_post <- filter_times(times_post, pol_data)
  df_post <- dplyr::mutate(df_post, type = "post")

  df_sample <- filter_times(times_sample, pol_data)
  df_sample <- dplyr::mutate(df_sample, type = "sample")

  background_data <- dplyr::bind_rows(df_pre, df_post, df_sample)
```

## Analyze

Calculate mean pollutant value for each period

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  background_means <- dplyr::group_by(background_data, id, type) %>%
                      dplyr::summarise(mean = mean(val), date = first(date),
                                       time = first(time))
```

* Compare test conc to background conc

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 12, fig.height=8}
  ggplot(background_means, aes(id, mean, colour = type)) +
         geom_point() +
         ggtitle("Ultrafine Particles (dp < 100 nm)") +
         theme_minimal() +
         xlab("test id") +
         ylab("concentration") + 
         scale_color_manual(values=wes_palette(n=3, name="GrandBudapest")) +
         theme(axis.text.x = element_text(angle=90, hjust=1))
```

* Sort by time of day

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 12, fig.height=8}
  background_time <- dplyr::filter(background_means, type != "sample")

  ggplot(background_time, aes(time, mean, colour = type)) +
         geom_point() +
         ggtitle("Ultrafine Particles (dp < 100 nm)") +
         theme_minimal() +
         xlab("test id") +
         ylab("concentration") +
         scale_color_manual(values=wes_palette(n=3, name="GrandBudapest")) +
         theme(axis.text.x = element_text(angle=45, hjust=1))
```

* ratio of pre to post background

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 12, fig.height=8}
  background_ratio <- dplyr::select(background_means, id, type, mean) %>%
                      tidyr::spread(type, mean) %>%
                      dplyr::mutate(bkgd_ratio = pre/post, 
                                    bkgd_avg = (pre + post)/2,
                                    bkgd_sample_ratio = sample/bkgd_avg)
                      
  ggplot(background_ratio, aes(id, bkgd_ratio)) +
         geom_point() +
         ggtitle("Ultrafine Particles (dp < 100 nm)") +
         theme_minimal() +
         xlab("test id") +
         ylab("pre to post background ratio") +
         scale_color_manual(values=wes_palette(n=3, name="GrandBudapest")) +
         theme(axis.text.x = element_text(angle=45, hjust=1))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 6, fig.height=8}
    ggplot(background_ratio, aes(x = 1, y = bkgd_sample_ratio)) +
         
         geom_dotplot(binaxis = "y", stackdir = "center") +
         ggtitle("Ultrafine Particles (dp < 100 nm)") +
         theme_minimal() +
         scale_y_log10() +
         xlab("") +
         ylab("average background to sample ratio") +
         scale_color_manual(values=wes_palette(n=3, name="GrandBudapest")) +
         theme(axis.text.x = element_blank())
```
