---
title: "background analysis carbon dioxide"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, include=FALSE}
  library(tidyverse)
  library(forcats)
```

```{r global_options, include=FALSE}
  knitr::opts_chunk$set(fig.path = 'figures/', warning = FALSE, message = FALSE,
                        fig.width=10, fig.height = 4, cache = FALSE)
```

```{r}
  source("../r_scripts/R_functions.R")
```

# Load data

* test times

```{r}
  test_times <- readRDS("../r_files/test_times_all.RDS")
```

* fivegas co2

```{r}
  fivegas_merged <- readRDS("../r_files/fivegas_merged.RDS")

  pol_data <- dplyr::filter(fivegas_merged, pol == "co2")
```

# Extract data

## sample times

* background times from logged timestamps
* trim first and last minute

```{r}
  times_pre <- dplyr::filter(test_times,
                             var == "bg_pre_start_fivegas" | var == "bg_pre_end_fivegas") %>%
               tidyr::spread(var, value) %>%
               dplyr::select(-date) %>%
               dplyr::rename(start = bg_pre_start_fivegas,
                             end = bg_pre_end_fivegas) %>%
               dplyr::mutate(start = start + 60,
                             end = end - 60)

  times_post <- dplyr::filter(test_times,
                             var == "bg_post_start_fivegas" | var == "bg_post_end_fivegas") %>%
               tidyr::spread(var, value) %>%
               dplyr::select(-date) %>%
               dplyr::rename(start = bg_post_start_fivegas,
                             end = bg_post_end_fivegas) %>%
               dplyr::mutate(start = start + 60,
                             end = end - 60)

```

* 5 minutes before and after each test

```{r}
  times_pre_5 <- dplyr::filter(test_times, var == "ignite") %>%
                tidyr::spread(var, value) %>%
                dplyr::select(-date) %>%
                dplyr::rename(end = ignite) %>%
                dplyr::mutate(start = end - 5*60)

  times_post_5 <- dplyr::filter(test_times, var == "end") %>%
                tidyr::spread(var, value) %>%
                dplyr::select(-date) %>%
                dplyr::rename(start = end) %>%
                dplyr::mutate(end = start + 5*60)
```

* join time data with precedence for the logged times over estimated times

```{r}
  times_pre <- dplyr::anti_join(times_pre_5, times_pre, by = "id") %>%
               dplyr::bind_rows(times_pre)

  times_post <- dplyr::anti_join(times_post_5, times_post, by = "id") %>%
                dplyr::bind_rows(times_post)
```

* sample times

```{r}
  times_sample <- dplyr::filter(test_times,
                                var == "set_1" | var == "end") %>%
                  tidyr::spread(var, value) %>%
                  dplyr::select(-date) %>%
                  dplyr::rename(start = set_1, end = end)
```

## filter data

* filter data inside times windows
* add `type` column with value: `pre`, `post` or `sample`

```{r}
  df_pre <- filter_times(times_pre, pol_data)
  df_pre <- dplyr::mutate(df_pre, type = "pre")

  df_post <- filter_times(times_post, pol_data)
  df_post <- dplyr::mutate(df_post, type = "post")

  df_sample <- filter_times(times_sample, pol_data)
  df_sample <- dplyr::mutate(df_sample, type = "sample")
```

* merge

```{r}
  background_data <- dplyr::bind_rows(df_pre, df_post, df_sample)
```

# Analyze

* mean value for each window

```{r}
  background_pre_post <- dplyr::group_by(background_data, id, type) %>%
                         dplyr::summarise(mean = mean(ppm, na.rm = TRUE)) %>%#, 
                                          #qc_val = mean(as.numeric(qc))) %>%
                         dplyr::ungroup() #%>%
	                       # dplyr::mutate(qc = ifelse(qc_val == 3, "ok", NA)) %>%
	                       # dplyr::mutate(qc = ifelse(qc_val == 2, "maybe", qc)) %>%
	                       # dplyr::mutate(qc = ifelse(qc_val == 1, "bad", qc)) %>%
	                       # dplyr::mutate(qc = as.factor(qc)) %>%
	                       # dplyr::ungroup() %>%
	                       # dplyr::select(-qc_val)
```

* average pre and post (good tests only)

```{r}
  background_means <- dplyr::filter(background_pre_post, type == "pre"| type == "post") %>%#, qc == "ok") %>%
                      dplyr::group_by(id) %>%
                      dplyr::summarise(ppm = mean(mean, na.rm = TRUE)) %>%
                      dplyr::ungroup() %>%
                      dplyr::mutate(mw = 44.01) %>%
                      dplyr::mutate(conc = convert_ppmv_ugpmc(ppm, mw, 25, 84))
```

Calculate mean background over all good tests

```{r}
  mean_bg <- dplyr::summarise(background_means,
                              ppm = mean(ppm, na.rm = TRUE),
                              conc = mean(conc, na.rm = TRUE),
                              mw = mean(mw, na.rm = TRUE))
```

* create list object with output data

```{r}
  background <- list(bg = background_means, mean = mean_bg)
```

# Save to file

```{r}
  saveRDS(background, file = "../r_files/background_co2.RDS")
```

# Plots

## outliers

```{r, fig.width=20, fig.height=12}
  p_data <- dplyr::mutate(background_means, 
                          value_norm = (conc - mean(conc)) / sd(conc),
                          outlier = ifelse(is_outlier(conc), as.character(id), NA))

  ggplot(p_data, aes(x = "background tests", y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

## background by type

```{r, fig.width = 12, fig.height=8}
  p_data <- background_pre_post

  ggplot(p_data, aes(id, mean, colour = type)) +
    geom_point() +
    ggtitle("CO2") +
    theme_minimal() +
    scale_colour_manual(values = 
                        c("pre" = "mediumaquamarine",
                          "sample" = "mediumorchid1",
                          "post" = "mistyrose4")) +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## background by qc flag

```{r, fig.width = 12, fig.height=8}
  p_data <- dplyr::filter(background_pre_post, type != "sample")

  ggplot(p_data, aes(id, mean)) + #, colour = qc)) +
    geom_point() +
    ggtitle("CO2") +
    theme_minimal() +
    scale_colour_manual(values = 
                        c("ok" = "mediumaquamarine",
                          "maybe" = "mediumorchid1",
                          "bad" = "mistyrose4")) +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## timeseries

```{r, fig.width=12, fig.height=40}
  p_data <- dplyr::mutate(background_data, type = as.factor(type)) %>%
            dplyr::filter(type != "sample")

  ggplot(p_data, aes(datetime, ppm, colour = type, group = type)) + 
    geom_line() +
    scale_colour_manual(values = 
                        c("pre" = "mediumaquamarine",
                          "post" = "mistyrose4")) +
    facet_wrap(~id, ncol = 3, scale = "free") +
    theme_minimal() +
    xlab("") +
    theme(legend.position="top")
```

