---
title: "carbonyls"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_load_data.R")
  source("../r_scripts/R_load_metadata.R")
  source("../r_scripts/R_tidy.R")
```

# Load data

* carbonyls

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/ions.Rda")
```

* metadata

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
  load("../r_files/carb_flows.Rda")
  load("../r_files/carb_times.Rda")
  load("../r_files/notes.Rda")

  load("../r_files/data_2.Rda")
```

# Organize

## extract ions data

* separate from carbonyls in same file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls_merged <- dplyr::select(ions, 24:39, type, id, id_ions) %>%
                      dplyr::filter(type == "test" | type == "bg")
```

## flows and times

* average flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flows <- dplyr::select(carb_flows, -date) %>%
           dplyr::group_by(id, type) %>%
           dplyr::summarise(flow = mean(value, na.rm = TRUE)) %>%
           dplyr::group_by(id) %>%
           dplyr::summarise(flow = mean(flow, na.rm = TRUE))
```

* match cassette timestamps to id

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times <- dplyr::select(carb_times, -date) %>%
           tidyr::spread(type, value)
```

* merge ions with metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls_merged <- dplyr::left_join(carbonyls_merged,
                                  flows,
                                  by = "id") %>%
                 dplyr::left_join(times,
                                  by = "id")
```

# QC

Load notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("carbonyls|all", notes$inst) == TRUE)
```

Set one flag per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           group_by(id) %>%
           arrange(qc) %>%
           summarise(qc = first(qc))
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls_merged <- dplyr::left_join(carbonyls_merged, flags, by = "id") %>%
                      dplyr::mutate(id = as.factor(id)) %>%
                      dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

Additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # carbonyls_merged$qc[carbonyls_merged$id == ""] <- "bad"
```

## Convert to long format

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  carbonyls_merged <- tidyr::gather(carbonyls_merged, "pol", "mass_ug",
                                    -id, -id_ions, -post_flow,
                                     -pre_flow, -time_start, -time_end, -type, -qc) %>%
                      dplyr::filter(!is.na(pol)) %>%
                      dplyr::mutate(pol = as.factor(pol))
```

## Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(carbonyls_merged, aes(id, mass_ug, group = pol, color = pol)) +
         geom_point() +
         geom_line() +
         scale_y_log10() +
         theme_minimal() +
         ylab("ug") +
         theme(legend.position="top")
```

## Plot QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(carbonyls_merged, aes(id, mass_ug, group = pol, color = qc)) +
         geom_point() +
         geom_line() +
         scale_y_log10() +
         theme_minimal() +
         ylab("ug") +
         theme(legend.position="top")
```

## Background analysis

* extract background data
* deal with LOD (not implemented)
* remove missing data
* calculate average concentration emitted ( and other stats)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg <- dplyr::filter(carbonyls_merged, type == "bg", qc == "ok") %>%
        dplyr::mutate(mass_ug = ifelse(is.na(mass_ug), 0, mass_ug)) %>%
        na.omit() %>%
        dplyr::select(-id_ions) %>%
        dplyr::mutate(flow = (post_flow + pre_flow) / 2) %>%
        dplyr::mutate(dur = time_end - time_start) %>%
        dplyr::mutate(conc = (mass_ug * 1000 * 60) / (flow * dur)) %>%  # ug * 1000 * 60 / (L/min * s) = ug / m^3
        dplyr::group_by(pol) %>%
        dplyr::summarise(mean = mean(conc),
                         sd = sd(conc),
                         min = min(conc),
                         max = max(conc),
                         n = n())
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  knitr::kable(bg, "markdown", digits = 2)
```

* merge with test data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # extract data for merge
  bg <- dplyr::select(bg, pol, mean) %>%
        dplyr::rename(conc_bg = mean)

 # merge
  carbonyls_merged <- dplyr::left_join(carbonyls_merged, bg, by = "pol") %>%
                      dplyr::mutate(conc_bg = ifelse(is.na(conc_bg), 0, conc_bg))
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(carbonyls_merged, file = "../r_files/carbonyls_merged.Rda")
```

## Summary

Carbonyl data was collected during `r length(unique(carbonyls_merged$id))` experiments. There is no carbonyl data for tests: `r setdiff(as.character(samples$id), as.character(carbonyls_merged$id))`.
