---
title: "carbonyls"
author: "Nicholas Good"
date: "1/19/2017"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_load_data.R")
  source("../r_scripts/R_load_metadata.R")
  source("../r_scripts/R_tidy.R")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")  # sample info

```

## Load carbonyl data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/ions.Rda")  # loads carbonyls file

  carbonyls <- dplyr::select(ions, 24:39, type, id, id_ions)  # select carbonlys only
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(carbonyls)
```

## Merge with filter meta data

Load metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/data_2.Rda")  # metadata
```

Extract carbonyl data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  meta <- dplyr::select(data_2, matches("_carb|_carb_|^id$|^date$"))

  flows <- dplyr::select(meta, matches("flow_|^id$")) %>%
           tidyr::gather("var", "value", -id) %>%
           dplyr::filter(grepl("_avg$", var)==FALSE) %>%
           dplyr::mutate(type = sub("flow.*", "", var),
                         rep = as.factor(gsub("[^0-9]", "", var))) %>%
           dplyr::select(-var)  %>%
           dplyr::filter(!is.na(id)) %>%
           dplyr::group_by(id, type) %>%
           dplyr::summarise(mean = mean(value, na.rm = TRUE))  %>%
           tidyr::spread(type, mean)

  times <- dplyr::select(meta, matches("^time_|^id$|^date$")) %>%
           tidyr::gather("var", "value", -id, -date) %>%
           dplyr::mutate(type = ifelse(grepl("_start_", var), "start", "NA"),
                         type = ifelse(grepl("_end_", var), "end", type)) %>%
           dplyr::arrange(id) %>%
           dplyr::filter(!is.na(id)) %>%
           dplyr::select(-var, -date) %>%
           tidyr::spread(type, value)
```

Merge carbonyls with metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls_merged <- dplyr::left_join(carbonyls, flows, by = "id") %>%
                      dplyr::left_join(times, by = "id") %>%
                      dplyr::filter(grepl("^[0-9]", id)==TRUE | grepl("^G[0-9]", id)==TRUE) %>%
                      dplyr::rename(pre_flow = pre, post_flow = post) %>%
                      dplyr::rename(time_start = start, time_end = end) %>%
                      dplyr::mutate(id = as.factor(id))
```

## QC

Load notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")  # notes

  notes <- dplyr::filter(notes, grepl("carbonyls|all", notes$inst) == TRUE) # update notes for carbs/ions?
```

Set one flag per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           group_by(id) %>%
           arrange(qc) %>%
           summarise(qc = first(qc))
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls_merged <- dplyr::left_join(carbonyls_merged, flags, by = "id") %>%
                      dplyr::mutate(id = as.factor(id)) %>%
                      dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

Additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # carbonyls_merged$qc[carbonyls_merged$id == ""] <- "bad"
```

## Convert to long format

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  carbonyls_merged <- tidyr::gather(carbonyls_merged, var, val,
                                    -id, -id_ions, -post_flow,
                                     -pre_flow, -time_start, -time_end, -type, -qc) %>%
                      dplyr::filter(!is.na(val)) %>%
                      dplyr::mutate(var = as.factor(var))
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(carbonyls_merged, file = "../r_files/carbonyls_merged.Rda")
```

## Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(carbonyls_merged, aes(id, val, group = var, color = var)) +
         geom_point() +
         geom_line() +
         scale_y_log10() +
         theme_minimal() +
         ylab("ug") +
         theme(legend.position="top")
```

## Plot QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(carbonyls_merged, aes(id, val, group = var, color = qc)) +
         geom_point() +
         geom_line() +
         scale_y_log10() +
         theme_minimal() +
         ylab("ug") +
         theme(legend.position="top")
```

## Summary

Carbonyl data was collected during `r length(unique(carbonyls_merged$id))` experiments. There is no carbonyl data for tests: `r setdiff(as.character(samples$id), as.character(carbonyls_merged$id))`.
