---
title: "ecoc"
author: "Nicholas Good"
date: "11/20/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

## Load ecoc data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/ecoc.Rda")    # ecoc dataset
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(ecoc)
```

## Merge with filter meta data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/data_1.Rda")    # metadata
```

Extract metadata
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  meta <- dplyr::select(data_1, matches("_orange_|_white_|^id$|^date$"))

  flows <- dplyr::select(meta, matches("flow_|^id$|^date$")) %>%
            tidyr::gather("var", "value", -id, -date) %>%
            dplyr::filter(grepl("_avg$", var)==FALSE) %>%
            dplyr::mutate(type = sub("flow.*", "", var),
                          cassette = as.factor(ifelse(grepl("_white_", var), "A", "E")),
                          rep = as.factor(gsub("[^0-9]", "", var))) %>%
            dplyr::select(-var)  %>%
            dplyr::filter(!is.na(id)) %>%
            dplyr::group_by(id, type, cassette) %>%
            dplyr::summarise(mean = mean(value, na.rm = TRUE))  %>%
            tidyr::spread(type, mean)
  
  times <- dplyr::select(meta, matches("^time_|^id$|^date$")) %>%
            tidyr::gather("var", "value", -id, -date) %>%
            dplyr::mutate(type = ifelse(grepl("_start_", var), "start", "NA"),
                          type = ifelse(grepl("_end_", var), "end", type)) %>%
            dplyr::filter(!is.na(id)) %>%  
            dplyr::select(-var, -date) %>%
            tidyr::spread(type, value)
```

Merge ecoc with metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc_merged <- dplyr::left_join(ecoc, flows, by = c("id", "cassette"))  %>%  
                  dplyr::left_join(times, by = c("id")) %>%  
                  dplyr::filter(grepl("^[0-9]", id)==TRUE | grepl("^G[0-9]", id)==TRUE) %>%
                  dplyr::rename(pre_flow = pre, post_flow = post)  %>%
                  dplyr::arrange(id)
  # convert back to factor
    ecoc_merged$id <- as.factor(ecoc_merged$id)
  
```

## QC

Load notes
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load(file="../r_files/notes.Rda")
  notes <- dplyr::filter(notes, grepl("ecoc|all", notes$inst)==TRUE)
```

Flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc <- "bad"
  notes$qc[1:9] <- "maybe"
  notes$qc[15:16] <- "maybe"
  notes$qc[17:22] <- "ok"
  notes$qc[23:27] <- "maybe"
  notes$qc[33:39] <- "maybe"
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc_merged <- dplyr::left_join(ecoc_merged, select(notes, id, qc), by = "id")
  ecoc_merged$id <- as.factor(ecoc_merged$id)
  ecoc_merged$qc <- ifelse(is.na(ecoc_merged$qc), "ok", as.character(ecoc_merged$qc))
  ecoc_merged$qc <- as.factor(ecoc_merged$qc)
```

Addtional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # co2_merged$qc[co2_merged$id == ""] <- "bad"
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(ecoc_merged, file="../r_files/ecoc_merged.Rda")
```

## Load sample info

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load(file="../r_files/samples.Rda") # load sample info
  samples <- dplyr::select(samples, id, stove, fuel)
  
# join ecoc with sample info data
  ecoc_p <- left_join(ecoc_merged, samples, by = "id")
  ecoc_p$id <- as.factor(ecoc_p$id)
```

## Plot by cassette (artifact)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12}
  ggplot(ecoc_p, aes(id, ec_ug_sq_cm, colour = cassette)) +
    geom_point() +
    scale_y_log10() +
    theme_minimal() +
    ylab("ec ug/sq.cm") +
    theme(legend.position = "top")

  ggplot(ecoc_p, aes(id, oc_ug_sq_cm, color = cassette)) +
    geom_point() +
    scale_y_log10() +
    theme_minimal() +
    ylab("oc ug/sq.cm") +
    theme(legend.position = "top")
```

## Plot by fuel

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12}
  ggplot(ecoc_p, aes(id, ec_ug_sq_cm, colour = fuel)) +
    geom_point() +
    scale_y_log10() +
    theme_minimal() +
    ylab("ec ug/sq.cm")+
    theme(legend.position = "top")

  ggplot(ecoc_p, aes(id, oc_ug_sq_cm, color = fuel)) +
    geom_point() +
    scale_y_log10() +
    theme_minimal() +
    ylab("oc ug/sq.cm") +
    theme(legend.position = "top")
```

## Plot QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12}
  ggplot(ecoc_p, aes(id, ec_ug_sq_cm, color = qc)) +
    geom_point() +
    scale_y_log10() +
    theme_minimal() +
    ylab("ec ug/sq.cm")+
    theme(legend.position = "top")
    
  ggplot(ecoc_p, aes(id, oc_ug_sq_cm, color = qc)) +
    geom_point() +
    scale_y_log10() +
    theme_minimal() +
    ylab("oc ug/sq.cm") +
    theme(legend.position = "top")
```

## Summary

$ECOC$ was measured for `r length(unique(ecoc_merged$id))` experiments between `r min(ecoc_merged$date, na.rm = TRUE)` and `r max(ecoc_merged$date, na.rm = TRUE)`. There is no $ecoc$ data for tests: `r setdiff(as.character(samples$id), as.character(ecoc_merged$id))`.
