---
title: "ecoc"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r global_options, include=FALSE}
  knitr::opts_chunk$set(fig.path = 'figures/',
                        warning = FALSE,
                        message = FALSE,
                        fig.width = 10, fig.height = 4,
                        cache = FALSE,
                        results='markup')
```

```{r}
  library(tidyverse)
  library(knitr)
```

```{r}
  source("../r_scripts/R_tidy.R")
  source("../r_scripts/R_functions.R")
```

# Load data

* ecoc

```{r,}
  ecoc <- readRDS("../r_files/ecoc.RDS")
```

* metadata

```{r}
  filter_times <- readRDS(file = "../r_files/filter_times.RDS")
  filter_flows <- readRDS(file = "../r_files/filter_flows.RDS")
  notes <- readRDS(file = "../r_files/notes.RDS")
  samples <- readRDS(file = "../r_files/samples.RDS")
```

# Filter data

* retain test and background data (blanks fall under test category)

```{r}
  ecoc_merged <- ecoc %>%
                 tidyr::separate(filter_id, c("id", "filter_type"), extra = "merge") %>%
                 dplyr::mutate(filter_type = ifelse(id == "blank", "blank", filter_type),
                               filter_type = ifelse(grepl(".*Blank.*", filter_type), "blank", filter_type),
                               filter_type = ifelse(grepl(".*of day.*", filter_type), "start_blank", filter_type),
                               filter_type = ifelse(grepl(".*Test.*", filter_type), "test", filter_type),
                               id = ifelse(grepl(".*blank.*", filter_type), "blank", id),
                               id = ifelse(grepl(".*start_blank.*", filter_type), "start_blank", id),
                               id = ifelse(grepl(".*test.*", filter_type), "test", id))
```

# Meta data

* match cassette flows to id

```{r}
  flows <- dplyr::select(filter_flows, -date) %>%
           dplyr::filter(colour == "white" | colour == "orange") %>%
           dplyr::group_by(id, type, colour) %>%
           dplyr::summarise(flow = mean(value, na.rm = TRUE)) %>%
           dplyr::group_by(id, colour) %>%
           dplyr::summarise(flow = mean(flow, na.rm = TRUE)) %>%
           dplyr::mutate(colour = factor(colour)) %>%
           dplyr::mutate(cassette = 
                         forcats::fct_recode(colour, "a" = "white",
                                                     "e" = "orange")) %>%
           dplyr::select(-colour)
```

* match cassette timestamps to id

```{r}
 times <- dplyr::select(filter_times, -date) %>%
           dplyr::filter(color == "white" | color == "orange") %>%
           tidyr::spread(type, value) %>%
           dplyr::mutate(cassette = 
                         forcats::fct_recode(color, "a" = "white",
                                                     "e" = "orange")) %>%
           dplyr::select(-color)
```

* merge ecoc with metadata

```{r}
  ecoc_merged <- dplyr::left_join(ecoc_merged,
                                  flows,
                                  by = c("id", "cassette")) %>%
                 dplyr::left_join(times,
                                  by = c("id", "cassette"))
```

# QC

## flags from notes

* extract ecoc notes

```{r}
  notes <- dplyr::filter(notes, grepl("ecoc|all", inst) == TRUE)
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

* merge flags with data

```{r}
  ecoc_merged <- dplyr::left_join(ecoc_merged, flags,
                                  by = "id") %>%
                 dplyr::mutate(id = as.factor(id)) %>%
                 dplyr::mutate(qc = ifelse(is.na(qc),
                                    "ok", as.character(qc))) %>%
                 dplyr::mutate(qc = as.factor(qc))
```

# Reformat

* select columns of interest (others?)
* longer format with columns: `pol = ec, oc or tc` containing the pollutant name and `ug_sq_cm` containing the measured value

```{r}
  ecoc_merged <- dplyr::select(ecoc_merged,
                               id, 
                               oc_ug_sq_cm, ec_ug_sq_cm, tc_ug_sq_cm,
                               datetime, date, time,
                               flow,
                               start, end,
                               qc,
                               type, cassette,
                               punch_area,
                               analyst) %>%
                  tidyr::gather("pol", "ug_sq_cm",
                                ec_ug_sq_cm,
                                oc_ug_sq_cm,
                                tc_ug_sq_cm) %>%
                  dplyr::mutate(pol = ifelse(pol == "ec_ug_sq_cm",
                                             "ec", pol),
                                pol = ifelse(pol == "oc_ug_sq_cm",
                                             "oc", pol),
                                pol = ifelse(pol == "tc_ug_sq_cm",
                                             "tc", pol))
```

# LOD / LOQ

```{r}
  blanks <- dplyr::filter(ecoc_merged, filter_type == "blank") %>%
            dplyr::group_by(pol) %>%
            dplyr::summarise(lod = sd(ug_sq_cm) * 3 + mean(ug_sq_cm),
                             loq = sd(ug_sq_cm) * 10 + mean(ug_sq_cm))
```

```{r}
  lod_ec <- dplyr::filter(blanks, pol == "ec")$lod[1]
  lod_oc <- dplyr::filter(blanks, pol == "oc")$lod[1]
  loq_ec <- dplyr::filter(blanks, pol == "ec")$loq[1]
  loq_oc <- dplyr::filter(blanks, pol == "oc")$loq[1]
```

* the EC limit of detection is `r round(lod_ec, 3)` ($\mu g/cm^2$) and limit of quantification is `r round(loq_ec, 3)` ($\mu g/cm^2$).

* the OC limit of detection is `r round(lod_oc, 3)` ($\mu g/cm^2$) and limit of quantification is `r round(loq_oc, 3)` ($\mu g/cm^2$).

* add ec lod and loq flags

```{r}
  ecoc_merged <- dplyr::mutate(ecoc_merged,
                                loq = NA,
                                lod = NA) %>%
                  dplyr::mutate(lod = ifelse(pol == "ec" &
                                            (type == "test" | type == "bg") &
                                            ug_sq_cm >= lod_ec,
                                            "above",
                                            lod)) %>%
                  dplyr::mutate(lod = ifelse(pol == "ec" &
                                            (type == "test" | type == "bg") &
                                            ug_sq_cm < lod_ec,
                                            "below",
                                            lod)) %>%
	                dplyr::mutate(loq = ifelse(pol == "ec" &
                                            (type == "test" | type == "bg") &
                                            ug_sq_cm >= lod_ec,
                                            "above",
                                            loq)) %>%
                  dplyr::mutate(loq = ifelse(pol == "ec" &
                                            (type == "test" | type == "bg") &
                                            ug_sq_cm < lod_ec,
                                            "below",
                                            loq))
```

* add oc lod and loq flags

```{r}

  ecoc_merged <- dplyr::mutate(ecoc_merged, lod = ifelse(pol == "oc" &
                                            (type == "test" | type == "bg") &
                                            ug_sq_cm >= lod_oc,
                                            "above",
                                            lod)) %>%
                  dplyr::mutate(lod = ifelse(pol == "oc" &
                                            (type == "test" | type == "bg") &
                                            ug_sq_cm < lod_oc,
                                            "below",
                                            lod)) %>%
	                dplyr::mutate(loq = ifelse(pol == "oc" &
                                            (type == "test" | type == "bg") &
                                            ug_sq_cm >= loq_oc,
                                            "above",
                                            loq)) %>%
                  dplyr::mutate(loq = ifelse(pol == "oc" &
                                            (type == "test" | type == "bg") &
                                            ug_sq_cm < loq_oc,
                                            "below",
                                            loq))
```

# Background analysis

```{r}
 area <- 11.79  # filter area (cm^2)
```

* extract "ok" background data
* remove missing data
* calculate average concentration emitted ( and other stats)
* average any repeat measurements (non in G8-G14)
* note G12 missing measurement data
* calculate averge background by pollutant / type
* the filter areas is `r area ` cm^2.

```{r}
  bg <- dplyr::filter(ecoc_merged, type == "bg", qc == "ok") %>%
        na.omit() %>%
        dplyr::mutate(dur = end - start) %>%
        dplyr::mutate(conc = ug_sq_cm * area * 1000 / (flow * dur)) %>%
        dplyr::group_by(pol, id, cassette) %>%
        dplyr::summarise(rep_mean = mean(conc)) %>%
        dplyr::ungroup() %>%
        dplyr::group_by(pol, cassette) %>%
        dplyr::summarise(mean = mean(rep_mean),
                         sd = sd(rep_mean),
                         min = min(rep_mean),
                         max = max(rep_mean),
                         n = n())
```

* account for artifact

```{r}
  bg <- dplyr::select(bg, pol, cassette, mean) %>%
        tidyr::spread(cassette, mean) %>%
        dplyr::mutate(conc_bg = e - a)
```

```{r}
  knitr::kable(bg, "markdown", digits = 2)
```

* merge with test data

```{r}
  ecoc_merged <- dplyr::left_join(ecoc_merged, 
                                  dplyr::select(bg, pol, conc_bg),
                                                by = "pol")
```

# Save merged file

```{r}
  save(ecoc_merged, file ="../r_files/ecoc_merged.Rda")
```

# Plots

```{r}
  ecoc_merged <- dplyr::left_join(ecoc_merged,
                                  dplyr::select(samples, id, stove, fuel),
                                                by = "id") %>%
                 dplyr::mutate(id= as.factor(id))
```

## values below lod


```{r, fig.width=12, fig.height=8}
  test_data <- dplyr::filter(ecoc_merged, pol != "tc" &
                                       qc == "ok" &
                                       type == "test" &
                                       grepl("^[0-9]", id) &
                                       cassette == "e")

  ggplot(test_data, aes(id, ug_sq_cm, colour = lod)) +
    geom_point() +
    theme_minimal() +
    theme(legend.position = "top") +
    scale_y_log10() +
    scale_colour_manual(values = c("ok" = "mediumorchid",
                                   "maybe" = "mediumaquamarine",
                                   "bad" = "mistyrose4")) +
    facet_wrap(~pol, ncol = 1, scales = "free") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

