---
title: "Emission Factor Summary"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(forcats)
```

# Load and tidy data

* Load data and meta data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/emission_factors.Rda")
  load("../r_files/samples.Rda")
```

* Load functions

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

* Output emissions factors column names 

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(emission_factors)
```

* Add stove/fuel information

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  emission_factors <- dplyr::left_join(emission_factors, 
                                       dplyr::select(samples, id, stove, fuel,
                                                              fuelcat, stovecat),
                                       by = "id")
```

* Use carbon balance method for missing emissions factors

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # Better variable name?
  emission_factors <- dplyr::mutate(emission_factors, 
                                    mass_ef_comb = 
                                      ifelse(is.na(mass_ef),
                                             mass_c_ef, mass_ef)) %>%
                      dplyr::mutate(energy_ef_comb =
                                      ifelse(is.na(energy_ef), energy_c_ef,
                                             energy_ef))
```

* Remove outliers

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  emission_factors <- dplyr::filter(emission_factors, mass_ef_comb > 0)
```

# Plot summary data

* Calculate summary statistics 
* Plot emissions factors by mass
* Error bars represent min and max values

## CO data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_box(dplyr::filter(emission_factors, grepl('^co$', pol)), "Carbon monoxide")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_bar(dplyr::filter(emission_factors, grepl('^co$', pol)), "Carbon monoxide")
```

## PM data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_box(dplyr::filter(emission_factors, grepl('grav$', pol)), 'Particulate matter')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_bar(dplyr::filter(emission_factors, grepl('grav$', pol)), 'Particulate matter')
```

## VOC data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_box(dplyr::filter(emission_factors, grepl('^voc_benzene$', pol)), 'Benzene')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_bar(dplyr::filter(emission_factors, grepl('^voc_benzene$', pol)), 'Benzene')
```

## Carbonyl data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_box(dplyr::filter(emission_factors, grepl('^formaldehyde$', pol)), 'Formaldehyde')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_bar(dplyr::filter(emission_factors, grepl('^formaldehyde$', pol)), 'Formaldehyde')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_box(dplyr::filter(emission_factors, grepl('^acetaldehyde$', pol)), 'Acetaldehyde')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_bar(dplyr::filter(emission_factors, grepl('^acetaldehyde$', pol)), 'Acetaldehyde')
```

## EC and OC data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_box(dplyr::filter(emission_factors, grepl('^ec$', pol)), 'Elemental carbon')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_bar(dplyr::filter(emission_factors, grepl('^ec$', pol)), 'Elemental carbon')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_box(dplyr::filter(emission_factors, grepl('^oc$', pol)), 'Organic carbon')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 8}
  plot_ef_bar(dplyr::filter(emission_factors, grepl('^oc$', pol)), 'Organic carbon')
```

# Exploring pollutant relationships

## CO vs formaldehyde

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
  plot_correlation(dplyr::filter(emission_factors, grepl('^co$', pol)),
                   dplyr::filter(emission_factors, grepl('^formaldehyde$', pol)), 
                   'Carbon monoxide', 'Formaldehyde')
```

## PM vs formaldehyde

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
  plot_correlation(dplyr::filter(emission_factors, grepl('^grav$', pol)),
                   dplyr::filter(emission_factors, grepl('^formaldehyde$', pol)), 
                   'Carbon monoxide', 'Formaldehyde')
```

## CO vs acetaldehyde

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
  plot_correlation(dplyr::filter(emission_factors, grepl('^co$', pol)),
                   dplyr::filter(emission_factors, grepl('^acetaldehyde$', pol)), 
                   'Carbon monoxide', 'Acetaldehyde')
```

## PM vs acetaldehyde

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
  plot_correlation(dplyr::filter(emission_factors, grepl('^grav$', pol)),
                   dplyr::filter(emission_factors, grepl('^acetaldehyde$', pol)), 
                   'Carbon monoxide', 'Acetaldehyde')
```

## CO vs benzene

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
  plot_correlation(dplyr::filter(emission_factors, grepl('^co$', pol)),
                   dplyr::filter(emission_factors, grepl('^voc_benzene$', pol)), 
                   'Carbon monoxide', 'Benzene')
```

## PM vs benzene

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
  plot_correlation(dplyr::filter(emission_factors, grepl('^grav$', pol)),
                   dplyr::filter(emission_factors, grepl('^voc_benzene$', pol)), 
                   'Particulate matter', 'Benzene')
```

## CO vs black carbon

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
  plot_correlation(dplyr::filter(emission_factors, grepl('^co$', pol)),
                   dplyr::filter(emission_factors, grepl('^ec$', pol)), 
                   'Carbon monoxide', 'Black carbon')
```

## PM vs benzene

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
  plot_correlation(dplyr::filter(emission_factors, grepl('^grav$', pol)),
                   dplyr::filter(emission_factors, grepl('^ec$', pol)), 
                   'Particulate matter', 'Black carbon')
```