---
title: "Emission factors"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r global_options, include=FALSE}
  knitr::opts_chunk$set(fig.path = 'figures/',
                        warning = FALSE,
                        message = FALSE,
                        fig.width = 10, fig.height = 4,
                        cache = FALSE,
                        results='markup')
```

```{r, echo=FALSE}
  library(tidyverse)
  library(knitr)
```

# Load

* sample info

```{r}
  samples <- readRDS("../r_files/samples.RDS")
```

* fuel burnt

```{r}
  fuel_burnt <- readRDS("../r_files/fuel_burnt.RDS")
```

* emissions

```{r}
  emissions_long <- readRDS("../r_files/emissions_long.RDS")
```

* hood flow

```{r}
  hood_flow <- readRDS("../r_files/hood_flow.RDS")
```

* polultant information

```{r}
  pol_properties <- readRDS("../r_files/pol_properties.RDS")
```

* fuel information

```{r}
  fuel_properties <- readRDS("../r_files/fuel_properties.RDS")
```

# Carbon calculations

* sum carbon over all pollutants

* calculate fuel burned using carbon balance method

```{r}
  carbon_emitted <- dplyr::filter(emissions_long, metric == "mass_carbon") %>%
                    dplyr::group_by(id) %>%
                    dplyr::summarise(mass_carbon = sum(value, na.rm = TRUE)) %>%
                    dplyr::left_join(dplyr::select(samples, id, fuel),
                                     by = "id") %>%
                    dplyr::left_join(fuel_properties, by = "fuel") %>%
                    dplyr::left_join(select(fuel_burnt, -units), by = "id") %>%
                    dplyr::filter(!is.na(fuel)) %>%
                    dplyr::mutate(mass_fuel_from_carbon_balance_kg =
                                  mass_carbon * (1e-9) * (1 / carbon_fraction),
                                  mass_fuel_scale_kg = mass_burnt * (1e-3)) %>%
                    dplyr::select(-mass_burnt)
```

* plot fuel from carbon balance versus measured fuel mass

```{r}
  ggplot(carbon_emitted, aes(mass_fuel_scale_kg, mass_fuel_from_carbon_balance_kg)) + 
  geom_point() +
  theme_minimal() +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_abline(intercept = 0, slope = 1, color = "grey90") +
  xlim(0, NA) + ylim(0, NA) +
  geom_text(aes(label = id), nudge_x = -0.01, nudge_y = 0.01)

```

```{r}
  df <- dplyr::select(carbon_emitted, mass_fuel_from_carbon_balance_kg, mass_fuel_scale_kg) %>%
      dplyr::rename(mass_carbon_balance = mass_fuel_from_carbon_balance_kg, mass_scale = mass_fuel_scale_kg)

  summary(lm(mass_carbon_balance ~ mass_scale, data = df))
```

# Pollutant mass per kg fuel
 
```{r}
  emission_factors <- dplyr::left_join(dplyr::filter(emissions_long, metric == "mass_emitted"), 
                                       dplyr::select(fuel_burnt, id, mass_burnt),
                                       by = "id" ) %>%
                      dplyr::mutate(mass_pol_mass_fuel = ((value * 1e-3) / mass_burnt)) # g pol / kg fuel
                            
```

## Plot emissions by id

```{r, fig.width=12}
  p_data <- dplyr::left_join(emission_factors,
                             dplyr::select(samples, id, type, mc_level),
                             by = "id") %>%
            dplyr::filter(type != "bg") %>%
            dplyr::mutate(mc_level = factor(p_data$mc_level,levels(p_data$mc_level)[c(3,1,2)]))

  ggplot(p_data, aes(id, mass_pol_mass_fuel, color = pol)) +
  geom_point(size = 4) +
  scale_y_log10() +
  theme_minimal() +
  theme(legend.position = "top",
        text = element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("g pol / kg fuel") + xlab("")
  
```

## Plot emissions by moisture content

```{r}
  ggplot(p_data, aes(mc_level, mass_pol_mass_fuel, color = pol)) +
  geom_boxplot(outlier.size=NA) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") +
  ylab("g pol / kg fuel") + xlab("moisture content") +
  facet_wrap(~pol, scales = "free")
  
 
  
```


# Save

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  saveRDS(mass_fuel, file = "../r_files/mass_fuel.RDS")
  saveRDS(emission_factors, file = "../r_files/emission_factors.RDS")
```

# Plots

## carbon emitted

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(mass_fuel, aes(x = id, y = mass_carbon)) +
    geom_point() +
    theme_minimal() +
    scale_y_log10() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
    ggtitle("mass carbon emitted")
```

## carbon from CO2

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(dplyr::filter(emission_factors, pol == "co2", metric == "mass_carbon"),
         aes(x = id, y = value)) +
    geom_point() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
    ggtitle("mass carbon emitted")
```

## fuel burnt vs carbon

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=10}  
  ggplot(mass_fuel, aes(x = mass_fuel_kg, y = mass_fuel_c, color = fuel)) +
    geom_point() +
    theme_minimal() +
    ylim(0, 2) +
    xlab("mass of fuel (measured) [kg]") +
    ylab("mass of fuel (carbon balance) [kg]")
```

## carbon emitted (bar)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE, fig.width=10, fig.height=40, eval=TRUE} 
  ggplot(dplyr::filter(emission_factors, metric == "mass_carbon"),
                       aes(x = id, y = value, fill = inst), na.rm = TRUE) +
    geom_bar(stat = "identity")+
    coord_flip() +
    theme_minimal() +
    ylab("test id") + 
    ylab("mass emitted") +
    theme(legend.position = "top")
```

## mass emitted (bar)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE, fig.width=10, fig.height=40, eval=TRUE} 
  ggplot(dplyr::filter(emission_factors, metric == "mass_emitted"),
        aes(x = id, y = value, fill = inst), na.rm = TRUE) +
    geom_bar(stat = "identity")+
    coord_flip() +
    theme_minimal() +
    ylab("test id") + 
    ylab("mass emitted") + 
    theme(legend.position = "top")
```
