---
title: "Emissions"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(knitr)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

# Load

* pollutant data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/ecoc_merged.Rda")
  load("../r_files/fivegas_merged.Rda")
  load("../r_files/background_co.Rda")
    bg_co <- background$bg
    bg_co_mean <- background$mean
  load("../r_files/background_co2.Rda")
    bg_co2 <- background$bg
    bg_co2_mean <- background$mean
  load("../r_files/background_ch4.Rda")
    bg_ch4 <- background$bg
    bg_ch4_mean <- background$mean
  load("../r_files/grav_merged.Rda")
  load("../r_files/smps_ultrafine.Rda")
    load("../r_files/smps_ultrafine_bg.Rda")
  load("../r_files/ions_merged.Rda")
  load("../r_files/carbonyls_merged.Rda")
  load("../r_files/voc_merged.Rda")
  load("../r_files/temp.Rda")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
  load("../r_files/test_times.Rda")
  load("../r_files/batch_wgts.Rda")
  load("../r_files/batch_pot.Rda")
  load("../r_files/wood_wgts.Rda")
  load("../r_files/wood_pot.Rda")
```

# Test timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times <- dplyr::filter(test_times, var == "start_1" | var == "shutdown") %>%
           tidyr::spread(var, value) %>%
           dplyr::select(-date) %>%
           dplyr::rename(start = start_1, end = shutdown)
```

# Constants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flow_hood <- 4  # m^3/min
```

# Fuel

## batch stoves

* fuel consumed

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  batch_mass <- dplyr::filter(batch_wgts, var == "fuel" |
                                          var == "refueled" |
                                          var == "preshutdown") %>%
                tidyr::spread(var, value) %>%
                dplyr::rename(wgt_fuel = fuel,
                              wgt_refuel = refueled,
                              wgt_shutdown = preshutdown) %>%
                dplyr::mutate(mass_fuel = wgt_fuel + wgt_refuel - wgt_shutdown)
```

## wood stoves

* fuel consumed

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  wood_mass <- dplyr::filter(wood_wgts, var == "fuel" |
                                        var == "ashpot_unusedfuel" |
                                        var == "ashpot_char_ash" |
                                        var == "ashpot_lid") %>%
                tidyr::spread(var, value) %>%
                dplyr::rename(wgt_fuel = fuel,
                              wgt_pot_unusedfuel = ashpot_unusedfuel,
                              wgt_pot_char_ash = ashpot_char_ash,
                              wgt_pot = ashpot_lid) %>%
                dplyr::mutate(mass_fuel = wgt_fuel -
                                          wgt_pot_unusedfuel -
                                          wgt_pot_char_ash +
                                          2 * wgt_pot)
```

* merge test types

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fuel_burnt <- dplyr::bind_rows(dplyr::select(batch_mass, id, mass_fuel),
                                 dplyr::select(wood_mass, id, mass_fuel)) %>%
                dplyr::mutate(id = as.factor(id)) %>%
                dplyr::filter(grepl("^[0-9]", id) == TRUE)
```

* plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  p_data <- dplyr::left_join(samples, fuel_burnt) %>%
            dplyr::mutate(id = as.factor(id)) %>%
            dplyr::filter(grepl("^[0-9]", id) == TRUE)

  ggplot(p_data, aes(x = id, y = mass_fuel, colour = stove)) +
        geom_point() +
        theme_minimal() + 
        theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
        theme(legend.position = "top")
        ylab("fuel burnt (g)")
```

## save fuel file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(fuel_burnt, file = "../r_files/fuel_burnt.Rda")
```

___

# Energy

* timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_times <- dplyr::filter(test_times, grepl("^start_[0-9]|^end_[0-9]", var)) %>%
             dplyr::select(-date) %>%
             tidyr::separate(var, c("var", "rep")) %>%
             tidyr::spread(var, value)
```

* which pot

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_pot <- dplyr::select(wood_pot, -date) %>%
           dplyr::bind_rows(dplyr::select(batch_pot, -date))
```

* pot mass

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_pot_mass <- dplyr::filter(wood_wgts,
                              grepl("^pot",
                              var)) %>%
                tidyr::separate(var, c("var", "pot")) %>%
                dplyr::select(-var, -date) %>%
                tidyr::spread(pot, value)

  e_pot_mass <- dplyr::bind_rows(e_pot_mass,
                                 dplyr::bind_rows(e_pot_mass,
                                 dplyr::filter(batch_wgts,
                                               grepl("^pot",
                                               var)) %>%
                                 tidyr::separate(var, c("var", "pot")) %>%
                                 dplyr::select(-var, -date) %>%
                                 tidyr::spread(pot, value)))

```

* pot + water mass

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_pot_water_mass <- dplyr::filter(wood_wgts,
                                    grepl("^on_[0-9]|^off_[0-9]", var)) %>%
                      tidyr::separate(var, c("var", "rep")) %>%
                      tidyr::spread(var, value) %>%
                      dplyr::select(-date) %>%
                      dplyr::rename(mass_on = on, mass_off = off)

  e_pot_water_mass <- dplyr::bind_rows(e_pot_water_mass,
                                       dplyr::filter(batch_wgts,
                                       grepl("^on_[0-9]|^off_[0-9]", var)) %>%
                                       tidyr::separate(var, c("var", "rep")) %>%
                                       tidyr::spread(var, value) %>%
                                       dplyr::select(-date) %>%
                                       dplyr::rename(mass_on = on, mass_off = off))
```

* extract temperature for periods of interest

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_temp <-  dplyr::select(temp, id, datetime, time, t_pot, thood)

  e_temp <- filter_temp(e_times, e_temp)
```

* temperature timeseries

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80}
  ggplot(e_temp, aes(datetime, t_pot, group = rep, colour = rep)) +
    geom_point() +
    facet_wrap(~id, scales = "free", ncol = 3) +
    theme_minimal()
```

* temperature change

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_temp <- dplyr::group_by(e_temp, id, rep) %>%
            dplyr::summarise(t_min = quantile(t_pot, 0.05, na.rm = TRUE),
                             t_max = quantile(t_pot, 0.95, na.rm = TRUE))
```

* merge data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_data <- dplyr::left_join(e_times, e_pot_water_mass, by = c("id", "rep")) %>%
            dplyr::left_join(e_temp, by = c("id", "rep")) %>%
            dplyr::left_join(e_pot_mass, by = "id") %>%
            dplyr::left_join(e_pot, by = c("id", "rep")) %>%
            dplyr::mutate(mass_pot = ifelse(pot == "a", a, NA),
                          mass_pot = ifelse(pot == "b", b, mass_pot),
                          id = as.factor(id)) %>%
            dplyr::select(-a, -b)
```

* energy applied

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_data <- dplyr::mutate(e_data,
                          mass_evap = mass_on - mass_off,
                          mass_h2o = mass_on - mass_pot,
                          dt = t_max - t_min,
                          j_heat = mass_h2o * 4.186 * dt, # g * J/(g.oC) * oC
                          j_vap = mass_evap * 2260,
                          joules = j_heat + j_vap)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  kable(dplyr::slice(e_data, 1:5),
        align = 'c', caption = "Energy Number")
```

* energy applied per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  energy <- dplyr::select(e_data, id, joules) %>%
            dplyr::group_by(id) %>%
            dplyr::summarise(joules = sum(joules, na.rm = TRUE))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  p_data <- dplyr::left_join(dplyr::filter(samples, type == "SF"),
                             energy,
                             by = "id")

  ggplot(p_data, aes(id, joules, colour = stove)) +
  geom_point() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = "top")
  ylab("energy applied to stove (J)")
  
```


```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
# template
# hyper-special kb code :-):-):-):-):-):-):-)
#  delta_m <- dplyr::filter(wood_wgts, grepl("^on|^off", var)) %>%
#             tidyr::separate(var, c("pot_status","rep")) %>%
#             dplyr::mutate(pot_id = paste(id, rep)) %>%
#             dplyr::select(id, pot_id, pot_status, value) %>%
#             tidyr::spread(pot_status, value) %>%
#             dplyr::mutate(delta_m = on - off)

# boring Nick code :-(:-(:-(:-(:-(:-(:-(:-(:-(
# mass H20 evaporated per test
#  delta_m <- dplyr::filter(wood_wgts, grepl("^on|^off", var)) %>%
#             tidyr::separate(var, c("pot_status","rep")) %>%
#             tidyr::spread(pot_status, value) %>%
#             dplyr::mutate(grams = on - off) %>%
#             dplyr::group_by(id) %>%
#             dplyr::summarise(grams = sum(grams, na.rm = TRUE))
```

# ecoc

* filter area

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  area <- 11.79  # cm^2
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::filter(ecoc_merged, type == "test", grepl("^[0-9]", id)) %>%
          dplyr::select(id, cassette,
                        start, end,
                        flow,
                        pol, ug_sq_cm,
                        conc_bg, type, qc)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::mutate(ecoc, dur = (end - start) / 60,
                              conc = ug_sq_cm * area * 1000 / (flow * dur))
```

* account for artifact

Use replicate average artifact (probably clean up spreadsheet at some point)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc_a <- dplyr::group_by(ecoc, pol, id, cassette) %>%
            dplyr::summarise(rep_mean = mean(conc, na.rm = TRUE)) %>%
            dplyr::ungroup() %>%
            tidyr::spread(cassette, rep_mean) %>%
            dplyr::mutate(a = ifelse(is.na(a), 0, a)) %>%
            dplyr::mutate(conc_cor = e - a)

  ecoc <- dplyr::left_join(
            dplyr::filter(ecoc, cassette == "e"),
            dplyr::select(ecoc_a, -a, -e), 
          by = c("id", "pol"))
```

* calculate mass emitted

Concentration is corrected for background by removing background test average concentration from each test.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::mutate(ecoc, 
                        mass_emitted = (conc_cor - conc_bg) * flow_hood * dur,
                        mass_carbon = mass_emitted)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ecoc, 2)
```

___

# fivegas

* select co, co2 and ch4.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::filter(fivegas_merged, pol == "co" | pol == "co2" | pol == "ch4")
```

* filter out data from outside emissions window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- filter_times(times, co_co2_ch4)
```

* calculate average for each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::group_by(co_co2_ch4, id, pol) %>%
                dplyr::summarise(ppm = mean(ppm, na.rm = TRUE),
                                 qc = first(qc)) %>%
                dplyr::ungroup()
```

* convert mixing ratio to mass (assuming constant T and P)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  mw <- data_frame(pol = c("co", "co2", "ch4"), mw = c(28.01, 44.01, 16.04))

  co_co2_ch4 <- dplyr::left_join(co_co2_ch4, mw, by = "pol") %>%
                dplyr::mutate(conc = convert_ppmv_ugpmc(ppm, mw, 25, 84))
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::left_join(co_co2_ch4, times, by = "id") %>%
                dplyr::mutate(dur = end - start) %>%
                dplyr::mutate(id = as.factor(id))
```

* background analysis
* replace NaN with between test mean value
* combine pollutants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg_co <- dplyr::mutate(bg_co,
                         ppm_bg = ifelse(is.na(ppm), bg_co_mean$ppm[1], ppm),
                         conc_bg = ifelse(is.na(conc), bg_co_mean$conc[1], conc),
                         pol = "co")

  bg_co2 <- dplyr::mutate(bg_co2,
                         ppm_bg = ifelse(is.na(ppm), bg_co2_mean$ppm[1], ppm),
                         conc_bg = ifelse(is.na(conc), bg_co2_mean$conc[1], conc),
                         pol = "co2")
 
  bg_ch4 <- dplyr::mutate(bg_ch4,
                         ppm_bg = ifelse(is.na(ppm), bg_ch4_mean$ppm[1], ppm),
                         conc_bg = ifelse(is.na(conc), bg_ch4_mean$conc[1], conc),
                         pol = "ch4")

  bg <- dplyr::bind_rows(bg_co, bg_co2, bg_ch4)
```

* merge background with fivegas data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::left_join(co_co2_ch4, 
                dplyr::select(bg, id, pol, conc_bg), by = c("id", "pol"))
```

* calculate micrograms emitted during tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::mutate(co_co2_ch4,
                              mass_emitted = (conc - conc_bg) * flow_hood * (dur / 60),
                              num_c = 1,
                              mass_carbon = mass_emitted * 12 * num_c / mw)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(co_co2_ch4, 2)
```

___

# grav

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::select(grav_merged,
                        id, pol,
                        start, end,
                        flow,
                        wgt_delta,
                        conc_bg, qc) %>%
          dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, dur = (end - start) / 60,
                              conc = wgt_delta * 1000 / (flow * dur), 
                              qc = as.factor(qc))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, 
                        mass_emitted = (conc - conc_bg) * flow_hood * dur,
                        mass_carbon = mass_emitted * 0)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(grav, 2)
```

___

# smps

* calculate test average number concentration
* calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ultrafine <- dplyr::filter(smps_ultrafine, n_bad == 0) %>%
               dplyr::group_by(id) %>%
               dplyr::summarise(conc = mean(number_conc, na.rm = TRUE),
                                n = n()) %>%
               dplyr::ungroup(id) %>%
               dplyr::left_join(times, by = "id") %>%
               dplyr::mutate(dur = end - start)
```

* calculate average background (use other method?)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg <- dplyr::filter(smps_ultrafine_bg, qc == "ok", n_bad == 0) %>%
        dplyr::group_by(id) %>%
        dplyr::summarise(conc_bg = mean(number_conc)) %>%
        dplyr::ungroup()

  conc_bg <- mean(bg$conc_bg)
```

* calculate number of particles emitted during tests (check this is all???)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # (N * 1,000,000 /cm^3) * m^3/min * (s / 60)
  ultrafine <- dplyr::mutate(ultrafine,
                             num_emitted = (conc - conc_bg) * 1e6 * flow_hood * (dur / 60))
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ultrafine, 2)
```

___

# ions

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions <- dplyr::select(ions_merged,
                        id, qc,
                        flow,
                        start, end,
                        pol, mass_ug,
                        conc_bg) %>%
	        dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions <- dplyr::mutate(ions, dur = (end - start) / 60,
                              conc = (mass_ug * 1000 / (flow * dur)))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions <- dplyr::mutate(ions, 
                        mass_emitted = (conc - conc_bg) * flow_hood * dur,
                        mass_carbon = mass_emitted * 0)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ions, 2)
```

# carbonyls

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls <- dplyr::select(carbonyls_merged,
                             id, qc,
                             flow,
                             start, end,
                             pol, mass_ug,
                             conc_bg) %>%
               dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls <- dplyr::mutate(carbonyls, dur = (end - start) / 60,
                             conc = (mass_ug * 1000 / (flow * dur)))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls <- dplyr::mutate(carbonyls, 
                             mass_emitted = (conc - conc_bg) * flow_hood * dur,
                             mass_carbon = mass_emitted * 0.95) # assume 95% carbon
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(carbonyls, 2)
```

# pah

* waiting on data

# voc

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flow <- 0.0135  # L/min
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::select(voc_merged,
                       id,
                       start, end,
                       pol, ppb,
                       ppb_bg, qc) %>%
         dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* moleculer weight data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  mw_voc <- tibble::data_frame(pol = unique(voc$pol)) %>%
            dplyr::mutate(mw = NA) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethane", 30.07, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethene", 28.05, mw)) %>%  
            dplyr::mutate(mw = ifelse(pol == "voc_propane", 44.1, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_propene", 42.08, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_butane", 58.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_butane", 58.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethyne", 26.04, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_t_2_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_cyclopentane", 70.1, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_pentane", 72.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_pentane", 72.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_t_2_pentene", 70.1329, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methyl_2_butene", 70.1329, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_pentene", 70.135, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_pentene", 70.135, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_cyclohexane", 84.16, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_hexane", 86.178, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_isoprene", 68.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_hexene", 84.1608, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_hexene", 84.162, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_4_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_methylcyclohexane", 98.186, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_heptane", 100.21, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_benzene", 78.11, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylpentane", 86.178, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_3_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "2_methylhexane", 100.2, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "3_methylpentane", 86.18, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "2_3_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methylhexane", 100.2, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylhexane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_2_4_trimethylpentane", 114.232, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_3_4_trimethylpentane", 114.232, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_toluene", 92.14, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methylheptane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylheptane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_octane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethylbenzene", 106.17, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_m_p_xylene", 106.168, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_styrene", 104.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_o_xylene", 106.16, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_nonane", 128.2, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_propylbenzene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_a_pinene", 136.24, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_propylbenzene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_4_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_3_5_trimethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_4_trimethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_decane", 142.29, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_3_trimethylbenzene", 30.07, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_3_diethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_4_diethylbenzene", 134.222, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_diethylbenzene", 134.222, mw))

  voc <- dplyr::left_join(voc, mw_voc, by = "pol")
```

* number of carbon atoms data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  num_carbon_voc <- tibble::data_frame(pol = unique(voc$pol)) %>%
                    dplyr::mutate(num_c = NA) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_ethane", 2, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_ethene", 2, num_c)) %>%  
                    dplyr::mutate(num_c = ifelse(pol == "voc_propane", 3, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_propene", 3, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_i_butane", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_butane", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_ethyne", 2, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_t_2_butene", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_butene", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_i_butene", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_c_2_butene", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_cyclopentane", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_i_pentane", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_pentane", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_t_2_pentene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_methyl_2_butene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_pentene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_c_2_pentene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_cyclohexane", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_hexane", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_isoprene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_hexene", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_c_2_hexene", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_4_dimethylpentane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_methylcyclohexane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_heptane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_benzene", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_3_methylpentane", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_3_dimethylpentane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "2_methylhexane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "3_methylpentane", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "2_3_dimethylpentane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_methylhexane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_3_methylhexane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_2_4_trimethylpentane", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_3_4_trimethylpentane", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_toluene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_methylheptane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_3_methylheptane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_octane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_ethylbenzene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_m_p_xylene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_styrene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_o_xylene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_nonane", 10, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_i_propylbenzene", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_a_pinene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_propylbenzene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_3_ethyltoluene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_4_ethyltoluene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_3_5_trimethylbenzene", 10, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_ethyltoluene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_2_4_trimethylbenzene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_decane", 10, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_2_3_trimethylbenzene", 2, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_3_diethylbenzene", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_4_diethylbenzene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_2_diethylbenzene", 9, num_c))

  voc <- dplyr::left_join(voc, num_carbon_voc, by = "pol")
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::mutate(voc, dur = (end - start) / 60,
                            ppm = ppb / 1000,
                            conc = convert_ppmv_ugpmc(ppm, mw, 25, 85),
                            conc_bg = convert_ppmv_ugpmc(ppb_bg / 1000, mw, 25, 85))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::mutate(voc, mass_emitted = (conc - conc_bg) * flow_hood * dur,
                            mass_carbon = mass_emitted * 12 * num_c / mw)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(voc, 2)
```

___

# Output

* combine pollutants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  emissions <- dplyr::select(co_co2_ch4, id, pol, conc,
                                         mass_emitted,
                                         mass_carbon, qc) %>%
               dplyr::bind_rows(dplyr::select(carbonyls, id, pol,
                                                         conc, conc_bg,
                                                         mass_emitted,
                                                         mass_carbon, qc)) %>%
               dplyr::bind_rows(dplyr::select(ecoc, id, pol,
                                                    conc, conc_bg,
                                                    mass_emitted,
                                                    mass_carbon, qc)) %>%
               dplyr::bind_rows(dplyr::select(grav, id, pol,
                                                    conc, conc_bg,
                                                    mass_emitted,
                                                    mass_carbon, qc)) %>%
               dplyr::bind_rows(dplyr::select(ions, id, pol,
                                                    conc, conc_bg,
                                                    mass_emitted,
                                                    mass_carbon, qc)) %>%
               dplyr::bind_rows(dplyr::select(voc, id, pol,
                                                   conc, conc_bg,
                                                   mass_emitted,
                                                   mass_carbon, qc))
```

* tests with negative values

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=10, fig.height=40} 
  kable(dplyr::filter(emissions, mass_emitted < 0),
        align = 'c', caption = "Tests with negative values")
```

* Save

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(emissions, file = "../r_files/emissions.Rda")
```
