---
title: "Pollutant emissions"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(knitr)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

# Load

* pollutant data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/ecoc_merged.Rda")
  load("../r_files/fivegas_merged.Rda")
  load("../r_files/background_co.Rda")
    bg_co <- background$bg
    bg_co_mean <- background$mean
  load("../r_files/background_co2.Rda")
    bg_co2 <- background$bg
    bg_co2_mean <- background$mean
  load("../r_files/background_ch4.Rda")
    bg_ch4 <- background$bg
    bg_ch4_mean <- background$mean
  load("../r_files/grav_merged.Rda")
  load("../r_files/smps_ultrafine.Rda")
  load("../r_files/smps_ultrafine_bg.Rda")
  load("../r_files/ions_merged.Rda")
  load("../r_files/carbonyls_merged.Rda")
  load("../r_files/voc_merged.Rda")
  pah_merged <- readRDS("../r_files/pah_merged.RDS")
```

* constants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/hood_flow.Rda")
  load("../r_files/pol_properties.Rda")
  load("../r_files/inst_constants.Rda")
  load("../r_files/calc_constants.Rda")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
  load("../r_files/test_times.Rda")
```

# timestamps

* integrated sample start & end times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times <- dplyr::filter(test_times, var == "start_1" | var == "shutdown") %>%
           tidyr::spread(var, value) %>%
           dplyr::select(-date) %>%
           dplyr::rename(start = start_1, end = shutdown)
```

# carbonyls

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls <- dplyr::select(carbonyls_merged,
                             id, qc,
                             flow,
                             start, end,
                             pol, mass_ug,
                             conc_bg) %>%
               dplyr::left_join(dplyr::select(pol_properties, 
                                              pol, 
                                              mw, 
                                              num_c), by = "pol") %>%
               dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls <- dplyr::mutate(carbonyls, dur = (end - start) / 60,
                             conc = (mass_ug * 1000 / (flow * dur)))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls <- dplyr::mutate(carbonyls, 
                             mass_emitted = (conc - conc_bg) * hood_flow * dur,
                             mass_carbon = mass_emitted * (num_c * mw_c / mw),
                             inst = "carbs")
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(carbonyls, 2)
```

# ecoc

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::filter(ecoc_merged, type == "test", qc != "bad", grepl("^[0-9]", id)) %>%
          dplyr::select(id, cassette,
                        start, end,
                        flow,
                        pol, ug_sq_cm,
                        conc_bg, type, qc)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::mutate(ecoc, dur = (end - start) / 60,
                              conc = (ug_sq_cm * filter_area * 1000) / (flow * dur))
```

* account for artifact

Use replicate average artifact (probably clean up spreadsheet at some point)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::group_by(ecoc, pol, id, cassette, qc, type) %>%
          dplyr::summarise_if(is.numeric, mean, na.rm = TRUE) %>%
          dplyr::ungroup()

  ecoc_a <- ecoc %>%
            dplyr::select(id, cassette, conc, pol) %>%
            tidyr::spread(cassette, conc) %>%
            dplyr::mutate(a = ifelse(is.na(a), 0, a)) %>%
            dplyr::mutate(conc_cor = e - a)

  ecoc <- dplyr::left_join(
            dplyr::filter(ecoc, cassette == "e"),
            dplyr::select(ecoc_a, -a, -e), 
          by = c("id", "pol"))
```

* calculate mass emitted

Concentration is corrected for background by removing background test average concentration from each test.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::mutate(ecoc, 
                        mass_emitted = (conc_cor - conc_bg) * hood_flow * dur,
                        mass_carbon = mass_emitted,
                        inst = "ecoc")
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ecoc, 2)
```

___

# fivegas

* select co, co2 and ch4.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::filter(fivegas_merged, pol == "co" | pol == "co2" | pol == "ch4")
```

* filter out data from outside emissions window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- filter_times(times, co_co2_ch4)
```

* calculate average for each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::group_by(co_co2_ch4, id, pol) %>%
                dplyr::summarise(ppm = mean(ppm, na.rm = TRUE),
                                 qc = first(qc)) %>%
                dplyr::ungroup()
```

* convert mixing ratio to mass (assuming constant T and P)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::left_join(co_co2_ch4,
                                 dplyr::select(pol_properties,
                                               pol,
                                               mw,
                                               num_c), by = "pol") %>%
                dplyr::mutate(conc = convert_ppmv_ugpmc(ppm, mw, 25, 84))
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::left_join(co_co2_ch4, times, by = "id") %>%
                dplyr::mutate(dur = end - start) %>%
                dplyr::mutate(id = as.factor(id))
```

* background analysis
* replace NaN with between test mean value
* combine pollutants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg_co <- dplyr::left_join(dplyr::select(samples, id), bg_co) %>%
           dplyr::mutate(ppm_bg = ifelse(is.na(ppm), bg_co_mean$ppm[1], ppm),
                         conc_bg = ifelse(is.na(conc), bg_co_mean$conc[1], conc),
                         pol = "co")

  bg_co2 <- dplyr::left_join(dplyr::select(samples, id), bg_co2) %>%
            dplyr::mutate(ppm_bg = ifelse(is.na(ppm), bg_co2_mean$ppm[1], ppm),
                          conc_bg = ifelse(is.na(conc), bg_co2_mean$conc[1], conc),
                          pol = "co2")

  bg_ch4 <- dplyr::left_join(dplyr::select(samples, id), bg_ch4) %>%
            dplyr::mutate(ppm_bg = ifelse(is.na(ppm), bg_ch4_mean$ppm[1], ppm),
                          conc_bg = ifelse(is.na(conc), bg_ch4_mean$conc[1], conc),
                          pol = "ch4")

  bg <- dplyr::bind_rows(bg_co, bg_co2, bg_ch4)
```

* merge background with fivegas data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::left_join(co_co2_ch4, 
                dplyr::select(bg, id, pol, conc_bg), by = c("id", "pol"))
```

* calculate micrograms emitted during tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::mutate(co_co2_ch4,
                              mass_emitted = (conc - conc_bg) * hood_flow * (dur / 60),
                              num_c = 1,
                              mass_carbon = mass_emitted * (mw_c * num_c / mw),
                              inst = "fivegas")
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(co_co2_ch4, 2)
```

___

# grav

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::select(grav_merged,
                        id, pol,
                        start, end,
                        flow,
                        wgt_delta,
                        conc_bg, qc) %>%
          dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, dur = (end - start) / 60,
                              conc = (wgt_delta * 1000) / (flow * dur), 
                              qc = as.factor(qc))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, 
                        mass_emitted = (conc - conc_bg) * hood_flow * dur,
                        mass_carbon = mass_emitted * 0,
                        inst = "grav")
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(grav, 2)
```

# ions

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions <- dplyr::select(ions_merged,
                        id, qc,
                        flow,
                        start, end,
                        pol, mass_ug,
                        conc_bg) %>%
	        dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions <- dplyr::mutate(ions, dur = (end - start) / 60,
                              conc = (mass_ug * 1000 / (flow * dur)))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions <- dplyr::mutate(ions, 
                        mass_emitted = (conc - conc_bg) * hood_flow * dur,
                        mass_carbon = mass_emitted * 0,
                        inst = "ions")
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ions, 2)
```

___

# smps

* calculate test average number concentration
* calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ultrafine <- dplyr::filter(smps_ultrafine, n_bad == 0) %>%
               dplyr::group_by(id) %>%
               dplyr::summarise(conc = mean(number_conc, na.rm = TRUE),
                                n = n(),
               	                qc = first(qc)) %>%
               dplyr::ungroup(id) %>%
               dplyr::left_join(times, by = "id") %>%
               dplyr::mutate(dur = end - start)
```

* calculate average background (use other method?)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg <- dplyr::filter(smps_ultrafine_bg, qc == "ok", n_bad == 0) %>%
        dplyr::group_by(id) %>%
        dplyr::summarise(conc_bg = mean(number_conc)) %>%
        dplyr::ungroup()

  conc_bg <- mean(bg$conc_bg)
```

* calculate number of particles emitted during tests (check this is all???)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # (N * 1,000,000 /cm^3) * m^3/min * (s / 60)
  ultrafine <- dplyr::mutate(ultrafine,
                             num_emitted = (conc - conc_bg) * 1e6 * hood_flow * (dur / 60),
                             inst = "smps")
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ultrafine, 2)
```

___

# pah

 * select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, eval=FALSE}
  pah <- pah_merged %>%
         dplyr::select(id,
                       lod,
                       flow,
                       start, end,
                       pol, mass_ug,
                       conc_bg) %>%
         dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  pah <- pah %>%
         dplyr::mutate(dur = (end - start) / 60,
                       conc = (mass_ug * 1000 / (flow * dur)))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  pah <- pah %>%
         dplyr::mutate(mass_emitted = (conc - conc_bg) * hood_flow * dur,
                       mass_carbon = mass_emitted * 0,
                       inst = "pah")
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(pah, 2)
```


# voc

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::select(voc_merged,
                       id,
                       start, end,
                       pol, ppb,
                       ppb_bg, qc) %>%
         dplyr::left_join(dplyr::select(pol_properties,
                                        pol,
                                        mw,
                                        num_c), by = "pol") %>% 
         dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::mutate(voc, dur = (end - start) / 60,
                            ppm = ppb / 1000,
                            conc = convert_ppmv_ugpmc(ppm, mw, 25, 85),
                            conc_bg = convert_ppmv_ugpmc(ppb_bg / 1000, mw, 25, 85))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::mutate(voc, mass_emitted = (conc - conc_bg) * hood_flow * dur,
                            mass_carbon = mass_emitted * (mw_c * num_c / mw),
                            inst = "voc")
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(voc, 2)
```

___

# Output

## combine pollutants

* mass based

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  emissions <- dplyr::select(co_co2_ch4, id, pol, 
                                         conc, conc_bg,
                                         mass_emitted,
                                         mass_carbon,
                                         inst, qc) %>%
               dplyr::bind_rows(dplyr::select(carbonyls, id, pol,
                                                         conc, conc_bg,
                                                         mass_emitted,
                                                         mass_carbon,
                                                         inst, qc)) %>%
               dplyr::bind_rows(dplyr::select(ecoc, id, pol,
                                                    conc, conc_bg,
                                                    mass_emitted,
                                                    mass_carbon,
                                                    inst, qc)) %>%
               dplyr::bind_rows(dplyr::select(grav, id, pol,
                                                    conc, conc_bg,
                                                    mass_emitted,
                                                    mass_carbon,
                                                    inst, qc)) %>%
               dplyr::bind_rows(dplyr::select(ions, id, pol,
                                                    conc, conc_bg,
                                                    mass_emitted,
                                                    mass_carbon,
                                                    inst, qc)) %>%
               dplyr::bind_rows(dplyr::select(voc, id, pol,
                                                   conc, conc_bg,
                                                   mass_emitted,
                                                   mass_carbon,
                                                   inst, qc)) %>%
               dplyr::bind_rows(dplyr::select(pah, id, pol,
                                                   conc, conc_bg,
                                                   mass_emitted,
                                                   mass_carbon,
                                                   inst, qc))
```

* convert to longer format

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  emissions_long <- tidyr::gather(emissions, "metric", "value", 3:6)
```

* number based

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  number <- dplyr::select(ultrafine, id, conc, num_emitted, inst, qc) %>%
            dplyr::mutate(pol = "number") %>%
            tidyr::gather("metric", "value", 2:3)
```

* combine mass and number instruments

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  emissions_long  <- dplyr::bind_rows(emissions_long, number)
```

## save

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  saveRDS(emissions_long, file = "../r_files/emissions_long.RDS")
```

## negative emissions

* tests with negative net emissions

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=10, fig.height=40} 
  kable(dplyr::filter(emissions, mass_emitted < 0),
        align = 'c', caption = "Tests with negative values")
```
