---
title: "emissions"
author: "Nicholas Good"
date: "1/6/2017"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

## Test time data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/test_times.Rda")

  times <- dplyr::filter(test_times, var == "start_1" | var == "shutdown") %>%
           tidyr::spread(var, value) %>%
           dplyr::select(-date) %>%
           dplyr::rename(start = start_1, end = shutdown)
```

## Sample info

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
```

## Constants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flow_hood <- 4  # m^3/min
```

## Fuel properties

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fuel_properties <- tibble::data_frame(fuel = c("Douglas Fir",
                                                  "Eucalyptus",
                                                  "Oak",
                                                  "Coconut Charcoal",
                                                  "Lump Hardwood (Med.)",
                                                  "Lump Hardwood (Sm.)",
                                                  "Eucalyptus Pellets",
                                                  "West Slope Pellets",
                                                  "Kerosene",
                                                  "Liquefied Petroleum Gas"),
                                         lhv = c(18658, 19427, 18658, 16628,
                                                 28611, 28611, 17501, 17501,
                                                 43100, 46100),
                                         carbon_fraction = c(0.497733333, 0.521166667,
                                                             0.5, 0.488, 0.840833333,
                                                             0.840833333, 0.4852,
                                                             0.4852, 0.82, 0.82))
```

## Mass of fuel burnt

* Load fuel mass data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/batch_wgts.Rda")
  load("../r_files/wood_wgts.Rda")
```

### Extract mass of fuel burnt during sampling periods

* batch stoves

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  batch_wgts <- dplyr::filter(batch_wgts, var == "fuel" |
                                          var == "refueled" |
                                          var == "preshutdown") %>%
                tidyr::spread(var, value) %>%
                dplyr::rename(wgt_fuel = fuel,
                              wgt_refuel = refueled,
                              wgt_shutdown = preshutdown) %>%
                dplyr::mutate(mass_fuel = wgt_fuel + wgt_refuel - wgt_shutdown)
```

* wood stoves

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  wood_wgts <- dplyr::filter(wood_wgts, var == "fuel" |
                                        var == "ashpot_unusedfuel" |
                                        var == "ashpot_char_ash" |
                                        var == "ashpot_lid") %>%
                tidyr::spread(var, value) %>%
                dplyr::rename(wgt_fuel = fuel,
                              wgt_pot_unusedfuel = ashpot_unusedfuel,
                              wgt_pot_char_ash = ashpot_char_ash,
                             wgt_pot = ashpot_lid) %>%
                dplyr::mutate(mass_fuel = wgt_fuel -
                                          wgt_pot_unusedfuel -
                                          wgt_pot_char_ash +
                                          2 * wgt_pot)
```

* merge test types

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fuel_burnt <- dplyr::bind_rows(dplyr::select(batch_wgts, id, mass_fuel),
                                 dplyr::select(wood_wgts, id, mass_fuel)) %>%
                dplyr::mutate(id = as.factor(id)) %>%
                dplyr::filter(grepl("^[0-9]", id) == TRUE)

```

* plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  p_data <- dplyr::left_join(samples, fuel_burnt) %>%
            dplyr::mutate(id = as.factor(id)) %>%
            dplyr::filter(grepl("^[0-9]", id) == TRUE)

  ggplot(p_data, aes(x = id, y = mass_fuel, colour = stove)) +
        geom_point() +
        theme_minimal() + 
        theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
        theme(legend.position = "top")
        ylab("fuel burnt (g)")

```

### Merge fuel burnt with fuel properties

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fuel_burnt_ <- dplyr::left_join(fuel_burnt, dplyr::select(samples, id, fuel), by = "id") %>%
	               dplyr::left_join(fuel_properties, by = "fuel")
```

### Save

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(fuel_burnt, file = "../r_files/fuel_burnt.Rda")
```

___

## co2

* Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/co2_merged.Rda")
```

* select sample data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co2 <- dplyr::filter(co2_merged, loc == "sample")
```

* filter out data from outside emissions window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co2 <- filter_times(times, co2)
```

* calculate average for each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co2 <- dplyr::group_by(co2, id) %>%
         dplyr::summarise(ppm = mean(ppm)) %>%
         dplyr::ungroup() %>%
         dplyr::mutate(mw = 44.01) %>%
         dplyr::mutate(num_c = 1)
```

* convert mixing ratio to mass (assuming constant T and P)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co2 <- dplyr::mutate(co2, conc = convert_ppmv_ugpmc(co2$ppm, mw, 25, 84))
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co2 <- dplyr::left_join(co2, times, by = "id") %>%
         dplyr::mutate(dur = end - start) %>%
         dplyr::mutate(id = as.factor(id))
```

* calculate micrograms of co2 emitted during test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co2 <- dplyr::mutate(co2, mass_emitted = conc * flow_hood * (dur / 60),  # ug/m^3 * m^3/min * (s / 60)
                            mass_carbon = mass_emitted * 12 * num_c / mw)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(co2, 2)
```

___

## ecoc

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/ecoc_merged.Rda")
```

* filter area

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  area <- 11.79  # cm^2
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::filter(ecoc_merged, type == "test") %>%
          dplyr::select(id, cassette,
                        start, end,
                        pre_flow, post_flow,
                        pol, ug_sq_cm,
                        conc_bg, type)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::mutate(ecoc, flow = (pre_flow + post_flow) / 2,
                              dur = (end - start) / 60,
                              conc = ug_sq_cm * area * 1000 / (flow * dur))
```

* account for artifact

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc_a <- dplyr::group_by(ecoc, pol, id, cassette) %>%
            dplyr::summarise(rep_mean = mean(conc)) %>%
            dplyr::ungroup() %>%
            tidyr::spread(cassette, rep_mean) %>%
            dplyr::mutate(A = ifelse(is.na(A), 0, A)) %>%
            dplyr::mutate(conc_cor = E - A)
```

* merge

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::filter(ecoc, cassette == "E") %>%
           dplyr::left_join(dplyr::select(ecoc_a, id, pol, conc_cor),
                            by = c("id", "pol"))
```

* calculate mass emitted

Concentration is corrected for background by removing background test average concentration from each test.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::mutate(ecoc, 
                        mass_emitted = (conc_cor - conc_bg) * flow_hood * dur,
                        mass_carbon = mass_emitted)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ecoc, 2)
```

___

## fivegas

### Load fivegas data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/fivegas_merged.Rda")

 # background co
  load("../r_files/background_co.Rda")

  bg_co <- background_means

 # background co2
  load("../r_files/background_co2.Rda")

  bg_co2 <- background_means
  
 # background ch4
  load("../r_files/background_ch4.Rda")

  bg_ch4 <- background_means
```

### Select fivegas pollutants

Select co , co2 and ch4.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::filter(fivegas_merged, pol == "co" | pol == "co2" | pol == "ch4") %>%
                dplyr::filter(qc != "bad")
```

### Process fivegas

* filter out data from outside emissions window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- filter_times(times, co_co2_ch4)
```

* calculate average for each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::group_by(co_co2_ch4, id, pol) %>%
                dplyr::summarise(ppm = mean(ppm)) %>%
                dplyr::ungroup()
```

* convert mixing ratio to mass (assuming constant T and P)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  mw <- data_frame(pol = c("co", "co2", "ch4"), mw = c(28.01, 44.01, 16.04))

  co_co2_ch4 <- dplyr::left_join(co_co2_ch4, mw, by = "pol") %>%
                dplyr::mutate(conc = convert_ppmv_ugpmc(ppm, mw, 25, 84))
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::left_join(co_co2_ch4, times, by = "id") %>%
                dplyr::mutate(dur = end - start) %>%
                dplyr::mutate(id = as.factor(id))
```

### Background

* merge background files
* calculate pre/post experiment average (ppm)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg <- dplyr::mutate(bg_co, pol = "co") %>%
        dplyr::bind_rows(dplyr::mutate(bg_co2, pol = "co2")) %>%
        dplyr::bind_rows(dplyr::mutate(bg_ch4, pol = "ch4")) %>%
        dplyr::filter(type == "pre"| type == "post") %>%
        dplyr::group_by(id, pol) %>%
        dplyr::summarise(bg_ppm = mean(mean, na.rm = TRUE)) %>%
        dplyr::ungroup() %>%
        dplyr::left_join(mw, by = "pol") %>%
        dplyr::mutate(bg_conc = convert_ppmv_ugpmc(bg_ppm, mw, 25, 84))
```

* merge background with fivegas data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::left_join(co_co2_ch4, 
                dplyr::select(bg, id, pol, bg_conc), by = c("id", "pol"))  # remove "NaNs"
```

* calculate micrograms emitted during tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_co2_ch4 <- dplyr::mutate(co_co2_ch4,
                              mass_emitted = (conc - bg_conc) * flow_hood * (dur / 60),
                              num_c = 1,
                              mass_carbon = mass_emitted * 12 * num_c / mw)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(co_co2_ch4, 2)
```

___

## grav

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/grav_merged.Rda")
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::select(grav_merged,
                        id, pol,
                        time_start, time_end,
                        flow_pre, flow_post,
                        wgt_delta,
                        conc_bg) %>%
          dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, flow = (flow_pre + flow_post) / 2,
                              dur = (time_end - time_start) / 60,
                              conc = wgt_delta * 1000 / (flow * dur))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, 
                        mass_emitted = (conc - conc_bg) * flow_hood * dur,
                        mass_carbon = mass_emitted * 0)  # all particle carbon in ec and oc
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(grav, 2)
```

___

## smps

* Load data

Note data is test average from desired time window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/smps_ultrafine.Rda")
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ultrafine <- dplyr::left_join(smps_ultrafine, times, by = "id") %>%
               dplyr::mutate(dur = end - start) %>%
               dplyr::mutate(id = as.factor(id))
```

* calculate number of particles emitted during tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # (N * 1,000,000 /cm^3) * m^3/min * (s / 60)
  ultrafine <- dplyr::mutate(ultrafine, emitted = mean_number_conc * 1e6 * flow_hood * (dur/60))
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ultrafine, 2)
```

___

## ions

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/ions_merged.Rda")
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions <- dplyr::select(ions_merged,
                        id, qc,
                        pre_flow, post_flow,
                        time_start, time_end,
                        pol, mass_ug,
                        conc_bg) %>%
          dplyr::rename(start = time_start,
                        end = time_end) %>%
	        dplyr::filter(grepl("^[A-Z]", id) == FALSE)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions <- dplyr::mutate(ions, dur = (end - start) / 60,
                              flow = (pre_flow + post_flow) / 2,
                              conc = (mass_ug * 1000 / (flow * dur)) - conc_bg)
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions <- dplyr::mutate(ions, 
                        mass_emitted = (conc - conc_bg) * flow_hood * dur,
                        mass_carbon = mass_emitted * 0)  # all particle carbon in ec and oc
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ions, 2)
```

## carbonyls

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/carbonyls_merged.Rda")
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls <- dplyr::select(carbonyls_merged,
                             id, qc,
                             pre_flow, post_flow,
                             time_start, time_end,
                             pol, mass_ug,
                             conc_bg) %>%
               dplyr::filter(grepl("^[A-Z]", id) == FALSE) %>%
               dplyr::rename(start = time_start,
                             end = time_end)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls <- dplyr::mutate(carbonyls, dur = (end - start) / 60,
                             flow = (pre_flow + post_flow) / 2,
                             conc = (mass_ug * 1000 / (flow * dur)) - conc_bg)
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  carbonyls <- dplyr::mutate(carbonyls, 
                             mass_emitted = (conc - conc_bg) * flow_hood * dur,
                             mass_carbon = mass_emitted * 0)  # all particle carbon in ec and oc
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(carbonyls, 2)
```

## pah

* waiting on data

## voc

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/voc_merged.Rda")

  flow <- 13.5 # L/min
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::select(voc_merged,
                       id,
                       time_start_voc, time_end_voc,
                       pol, ppb,
                       ppb_bg) %>%
         dplyr::filter(grepl("^[A-Z]", id) == FALSE) %>%
         dplyr::rename(start = time_start_voc,
                       end = time_end_voc)
```

* moleculer weight data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  mw_voc <- tibble::data_frame(pol = unique(voc$pol)) %>%
            dplyr::mutate(mw = NA) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethane", 30.07, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethene", 28.05, mw)) %>%  
            dplyr::mutate(mw = ifelse(pol == "voc_propane", 44.1, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_propene", 42.08, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_butane", 58.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_butane", 58.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethyne", 26.04, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_t_2_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_cyclopentane", 70.1, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_pentane", 72.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_pentane", 72.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_t_2_pentene", 70.1329, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methyl_2_butene", 70.1329, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_pentene", 70.135, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_pentene", 70.135, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_cyclohexane", 84.16, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_hexane", 86.178, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_isoprene", 68.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_hexene", 84.1608, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_hexene", 84.162, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_4_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_methylcyclohexane", 98.186, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_heptane", 100.21, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_benzene", 78.11, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylpentane", 86.178, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_3_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "2_methylhexane", 100.2, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "3_methylpentane", 86.18, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "2_3_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methylhexane", 100.2, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylhexane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_2_4_trimethylpentane", 114.232, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_3_4_trimethylpentane", 114.232, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_toluene", 92.14, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methylheptane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylheptane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_octane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethylbenzene", 106.17, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_m_p_xylene", 106.168, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_styrene", 104.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_o_xylene", 106.16, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_no_8888ne", NA, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_propylbenzene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_a_pinene", 136.24, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_propylbenzene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_4_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_3_5_trimethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_4_trimethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_decane", 142.29, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_3_trimethylbenzene", 30.07, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_3_diethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_4_diethylbenzene", 134.222, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_diethylbenzene", 134.222, mw))

  voc <- dplyr::left_join(voc, mw_voc, by = "pol")
```

* number of carbon atoms data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  num_carbon_voc <- tibble::data_frame(pol = unique(voc$pol)) %>%
                    dplyr::mutate(num_c = NA) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_ethane", 2, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_ethene", 2, num_c)) %>%  
                    dplyr::mutate(num_c = ifelse(pol == "voc_propane", 3, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_propene", 3, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_i_butane", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_butane", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_ethyne", 2, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_t_2_butene", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_butene", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_i_butene", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_c_2_butene", 4, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_cyclopentane", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_i_pentane", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_pentane", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_t_2_pentene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_methyl_2_butene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_pentene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_c_2_pentene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_cyclohexane", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_hexane", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_isoprene", 5, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_hexene", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_c_2_hexene", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_4_dimethylpentane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_methylcyclohexane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_heptane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_benzene", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_3_methylpentane", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_3_dimethylpentane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "2_methylhexane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "3_methylpentane", 6, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "2_3_dimethylpentane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_methylhexane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_3_methylhexane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_2_4_trimethylpentane", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_3_4_trimethylpentane", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_toluene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_methylheptane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_3_methylheptane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_octane", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_ethylbenzene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_m_p_xylene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_styrene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_o_xylene", 7, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_no_8888ne", NA, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_i_propylbenzene", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_a_pinene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_propylbenzene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_3_ethyltoluene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_4_ethyltoluene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_3_5_trimethylbenzene", 10, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_2_ethyltoluene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_2_4_trimethylbenzene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_n_decane", 10, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_2_3_trimethylbenzene", 2, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_3_diethylbenzene", 8, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_4_diethylbenzene", 9, num_c)) %>%
                    dplyr::mutate(num_c = ifelse(pol == "voc_1_2_diethylbenzene", 9, num_c))

  voc <- dplyr::left_join(voc, num_carbon_voc, by = "pol")
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::mutate(voc, dur = (end - start) / 60,
                            ppm = ppb / 1000,
                            conc = convert_ppmv_ugpmc(ppm, mw, 25, 85),
                            conc_bg = convert_ppmv_ugpmc(ppb_bg / 1000, mw, 25, 85))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::mutate(voc, mass_emitted = conc * flow_hood * dur,
                            mass_carbon = mass_emitted * 12 * num_c / mw)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(voc, 2)
```

___

## Combine pollutants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  emissions <- dplyr::select(co_co2_ch4, id, pol, mass_emitted, mass_carbon) %>%
               dplyr::bind_rows(dplyr::select(carbonyls, id, pol, mass_emitted, mass_carbon)) %>%
               dplyr::bind_rows(dplyr::select(ecoc, id, pol, mass_emitted, mass_carbon)) %>%
               dplyr::bind_rows(dplyr::select(grav, id, pol, mass_emitted, mass_carbon)) %>%
               dplyr::bind_rows(dplyr::select(ions, id, pol, mass_emitted, mass_carbon)) %>%
               dplyr::bind_rows(dplyr::select(voc, id, pol, mass_emitted, mass_carbon))
```

* Save

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(emissions, file = "../r_files/emissions.Rda")
```

## Mass carbon per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  mass_carbon <- dplyr::group_by(emissions, id) %>%
                dplyr::summarise(ug = sum(mass_carbon, na.rm = TRUE))
```

* Save

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(mass_carbon, file = "../r_files/mass_carbon.Rda")
```

* Plot carbon emitted

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(mass_carbon, aes(x = id, y = ug)) +
    geom_point() +
    theme_minimal() +
    scale_y_log10() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
    ggtitle("mass carbon emitted")
```
