---
title: "Emission factors"
subtitle: "Mass emitted per mass of fuel burned"
author: "Lizette van Zyl and Jessica Tryner"
date: "December 11, 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r global_options, include=FALSE}
  knitr::opts_chunk$set(fig.path = 'figures/',
                        warning = FALSE,
                        message = FALSE,
                        fig.width = 10, fig.height = 4,
                        cache = FALSE,
                        results='markup')
```

```{r, echo=TRUE, message=FALSE}
  library(tidyverse)
  library(knitr)
  library(broom)
```

# Load

Test metadata: 

```{r}
  samples <- readRDS("../r_files/samples.RDS")
```

Constants: 

```{r}
  hood_flow <- readRDS("../r_files/hood_flow.RDS")
  pol_properties  <- readRDS("../r_files/pol_properties.RDS")
  fuel_properties <- readRDS("../r_files/fuel_properties.RDS")
```

Emissions

```{r}
  emissions_long <- readRDS("../r_files/emissions_long.RDS")
```

# Caclulate mass of fuel burned

Calculate the dry mass of fuel burned from the mass of carbon measured as CO~2~, CO, and CH~4~ in the exhaust.  

```{r}

  df_fuel <- dplyr::filter(emissions_long, metric=="mass_carbon",
                                           pol %in% c("co2", "co", "ch4")) %>%
             dplyr::group_by(id) %>%
             dplyr::summarise(mass_c = (1e-6)*sum(value)) %>% # convert from ug to grams
             dplyr::left_join(dplyr::select(samples, id, fuel), by=c("id")) %>%
             dplyr::left_join(dplyr::select(fuel_properties, fuel, carbon_fraction), by=c("fuel")) %>%
             plyr::rename(c("carbon_fraction" = "y_c_fuel")) %>%
             dplyr::mutate(m_fuel_dry = mass_c/y_c_fuel) # grams
```

# Calculate emission factors

Calculate the mass of pollutant emitted per mass of fuel burned.  Note that `value` has units of micrograms (for mass emitted) or number (for number emitted).  Note that `m_fuel_dry` has units of grams.  

1. Select the `mass_emitted` and `num_emitted` metrics for actual tests.  
2. Add the mass of dry fuel burned to the dataframe.  
3. Calculate the emission factors as $\mu$g/g or number/g.  
4. Convert mass-based emission factors to g/MJ~d~.  
5. Convert number-based emission factors to number/kg.  
6. Add a column with units.  
 
```{r}

  df_ef <- dplyr::filter(emissions_long, metric %in% c("mass_emitted", # 1
                                                       "num_emitted"),
                                         !grepl("^BG|G", id)) %>%
           dplyr::left_join(dplyr::select(df_fuel, id, m_fuel_dry), by =c("id")) %>% # 2
           dplyr::mutate(ef = value / m_fuel_dry, # ug/g or number/g     3
                         ef = ifelse(metric == "mass_emitted",         # 4, 5
                                     ef * (1e-3), ef * (1e3)), # g/kg or number/kg
                         units = ifelse(metric == "mass_emitted",      # 6
                                       "g_pol_kg_fuel", "num_kg_fuel")) %>%
           dplyr::select(-metric, -value)
```

Add test metadata.

```{r}

  # Add metadata to data frame
  df_ef <- dplyr::left_join(df_ef, dplyr::select(samples, id, type, mc_level, mc_dry),
                            by=c("id")) %>% 
           dplyr::ungroup() %>%
           dplyr::mutate(id       = forcats::fct_relevel(id, "FL1", "FL2", "FL3", "FL4",
                                                             "FM1", "FM2", "FM3", "FM4",
                                                             "FH1", "FH2", "FH3", "FH4",
                                                             "LL1", "LL2", "LL3", "LL4",
                                                             "LM1", "LM2", "LM3", "LM4",
                                                             "LH1", "LH2", "LH3","LH4"),
                         fuel_mc_shape = paste(type, mc_level, sep="_")) %>%
    
          dplyr::filter(!is.na(type), # remove background tests
                        id != "FH2")
```

Save the data.

```{r}
  saveRDS(df_ef, file = "../r_files/ef_fuel.RDS")
```

# Calculate EC/OC ratio

```{r}
  df_ecoc <- emissions_long %>%
             dplyr::filter(pol %in% c("ec", "oc"), metric %in% "conc") %>%
             dplyr::select(-inst)
          
  df_ecoc <- unique(df_ecoc) %>%
             tidyr::spread(pol, value) %>%
             dplyr::mutate(ratio = ec / oc) %>%
             dplyr::left_join(dplyr::select(samples, id, type, mc_level, mc_dry), by=c("id")) %>%
             dplyr::mutate(fuel_mc_shape = paste(mc_level, type, sep = "_")) %>%
             dplyr::filter(type != "bg",
                           id   != "FH2") 
  
  df_ecoc_summary <- dplyr::group_by(df_ecoc, group_by=mc_level) %>%
                     dplyr::summarise(ratio_mean = mean(ratio))
  df_ecoc_summary  
```

# Plot results

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # Define the colors to use in the plots. 
  purple <- "#330099"
  gold   <- "#FF9900"
  teal1  <- "#33CC99"
  teal2  <- "#339999"
  teal3  <- "#0099CC"
  orange <- "#FF6633"
  pink1  <- "#CC0066"
  pink2  <- "#FF3366"
```

```{r, echo=FALSE, fig.width=6.5, fig.height=5, fig.align='center'}

  log_10_breaks <- c(0.0001,0.0002,0.0003,0.0004,0.0005,0.0006,0.0007,0.0008,0.0009,
                     0.001,0.002,0.003,0.004,0.005,0.006,0.007,0.008,0.009,
                     0.01,0.02,0.03,0.04,0.05,0.06,0.07,0.08,0.09,
                     0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,
                     1,2,3,4,5,6,7,8,9,
                     10,20,30,40,50,60,70,80,90,100)

  df_plot <- dplyr::filter(df_ef, pol %in% c("co", "ch4", "grav", "ec", "oc", "tc",
                                             "formaldehyde", "acetaldehyde",
                                             "benzene", "toluene", "ethylbenzene",
                                             "m+p-xylene", "o-xylene"))

  ggplot(df_plot, aes(id, ef, color=pol)) +
    geom_point(size = 3) +
    xlab("Test ID") + 
    ylab(expression(Emission ~ factor ~ "(" * g ~ kg^-1*")")) + 
    scale_y_log10(breaks=c(0.0001,0.001,0.01,0.1,1.0,10,100),
                  minor_breaks=log_10_breaks) +
    # format plot
    theme(panel.border     = element_rect(fill=NA),
          panel.background = element_rect(fill="white"),
          panel.grid.major = element_line(color="gray90"),
          panel.grid.minor = element_line(color="gray90"),
          text       = element_text(size=10, color="black"), 
          axis.text.x = element_text(size=10, color="black", angle=45, hjust=1),
          axis.text.y = element_text(size=10, color="black"),
          axis.title = element_text(size=10, color="black"),
          axis.ticks = element_line(color="black"),
          legend.position = "top")
   
```

```{r, fig.width=6.5, fig.height=4.1, fig.align='center'}

  df_ef_subset <- dplyr::filter(df_ef, pol %in% c("co2",  "co", "ch4",
                                                  "grav", "ec", "oc")) %>%
                   dplyr::mutate(pol = factor(pol, levels=c("co2", "co","ch4",
                                                            "grav","ec","oc")))

  # dummy data used to ensure that y-axis goes to zero on all panels except CO, PM, and CH4
  df_dummy <- data.frame(pol    = levels(df_ef_subset$pol),
                         mc_dry = c(5),
                         ef     = c(1820,10,4,0.9,0,0),
                         type   = c("Milled"))

  ggplot(df_ef_subset, aes(x=mc_dry, y=ef, shape=type, color=pol)) +
      facet_wrap(~pol, ncol=3, scales="free_y") + 
      # add data
      geom_point(size=2, alpha=0.7) +
      # format plot
      coord_cartesian(xlim=c(4,30)) + 
      scale_x_continuous(breaks=seq(5,25,10)) +
      geom_blank(data=df_dummy) + 
      xlab("Moisture content (%)") +
      ylab(expression(Emission ~ factor ~ "(" * g ~ kg^-1*")")) + 
      scale_color_manual(values = c("black", pink2, purple, gold, teal1, teal2)) +
      # label panels
      geom_text(aes(label="CO[2]",   x=30, y=1820), data=subset(df_ef_subset, pol=="co2"),
                color="black", hjust="right", vjust="bottom", size=3.5, parse=TRUE) +
      geom_text(aes(label="CO",      x=30, y=10),   data=subset(df_ef_subset, pol=="co"),
              color="black", hjust="right", vjust="bottom", size=3.5) +
      geom_text(aes(label="CH[4]",   x=30, y= 1.1), data=subset(df_ef_subset, pol=="ch4"),
                color="black", hjust="right", vjust="bottom", size=3.5, parse=TRUE) +
      geom_text(aes(label="PM[2.5]", x=30, y= 0.9), data=subset(df_ef_subset, pol=="grav"),
                color="black", hjust="right", vjust="bottom", size=3.5, parse=TRUE) +
      geom_text(aes(label="EC",      x= 4, y= 0),   data=subset(df_ef_subset, pol=="ec"),
                color="black", hjust="left", vjust="bottom", size=3.5) +
      geom_text(aes(label="OC",      x=30, y= 0),   data=subset(df_ef_subset, pol=="oc"),
                color="black", hjust="right", vjust="bottom", size=3.5) +
      # format plot
      theme(aspect.ratio=1,
            panel.border     = element_rect(fill=NA),
            panel.background = element_rect(fill="white"),
            strip.background = element_blank(),
            strip.text.x = element_blank(),
            text       = element_text(size=10, color="black"), 
            axis.text  = element_text(size=10, color="black"),
            axis.title = element_text(size=10, color="black"),
            axis.ticks = element_line(color="black"),
            legend.position = "none")
  
  ggsave(filename="../figures/emissions_fuel.pdf", plot=last_plot(), device="pdf", width=6.5, height=4.1)
```

```{r, fig.width=6.5, fig.height=8.25, fig.align='center'}
  
  df_ef_subset <- dplyr::filter(df_ef, pol %in% c("formaldehyde","acetaldehyde","acetone",
                                                  "acrolein","propionaldehyde","crotonaldehyde",
                                                  "butanone","methacrolein","butyraldehyde",
                                                  "benzaldehyde","isovaleraldehyde","valeraldehyde",
                                                  "o_tolualdehyde","m_p_tolualdehyde","hexaldehyde")) %>%
                   dplyr::mutate(pol  = ifelse(pol=="o_tolualdehyde","o-tolualdehyde",pol),
                                 pol  = ifelse(pol=="m_p_tolualdehyde","m+p-tolualdehyde",pol))

  ggplot(df_ef_subset, 
         aes(x=mc_dry, y=ef, shape=type)) +
         geom_point(size=3, alpha=0.7, color="black") +
         facet_wrap(~pol, scales="free_y", ncol=3) +
         # format plot
         coord_cartesian(xlim=c(4,30)) + 
         scale_x_continuous(breaks=seq(5,25,10)) +
         xlab("Moisture content (%)") +
         ylab(expression(Emission ~ factor ~ "(" * g ~ kg^-1*")")) + 
         labs(shape = "Fuel shape") + 
         theme(aspect.ratio=1,
               panel.border     = element_rect(fill=NA),
               panel.background = element_rect(fill="white"),
               strip.background = element_rect(fill="gray95", color="black"),
               strip.text.x = element_text(size=10, color="black"),
               panel.grid.major.x = element_line(color="gray95"),
               panel.grid.minor.x = element_line(color="gray95"),
               panel.grid.major.y = element_line(color="gray95"),
               panel.grid.minor.y = element_line(color="gray95"),
               axis.text    = element_text(size=10, color="black"),
               axis.title   = element_text(size=10, color="black"),
               axis.ticks   = element_line(color="black"),
               legend.text  = element_text(size=10, color='black'),
               legend.title = element_text(size=10, color='black'),
               legend.key   = element_blank())
  
  ggsave(filename="../figures/aldehydes_fuel.pdf", plot=last_plot(), device="pdf", width=6.5, height=8.25)
```

```{r, fig.width=6.5, fig.height=4.5, fig.align='center'}
  
  df_ef_subset <- dplyr::filter(df_ef, pol %in% c("benzene", "toluene", "ethylbenzene",
                                                  "m+p-xylene", "o-xylene")) %>%
                   dplyr::mutate(pol  = factor(pol, levels=c("benzene", "toluene", "ethylbenzene",
                                                             "m+p-xylene", "o-xylene")),
                                 type = forcats::fct_recode(type, "Split"   = "field",
                                                                   "Milled" = "lab"))

  ggplot(df_ef_subset, 
         aes(x=mc_dry, y=ef, shape=type)) +
         geom_point(size=3, alpha=0.7, color="black") +
         facet_wrap(~pol, scales="free_y", ncol=3) +
         # format plot
         coord_cartesian(xlim=c(4,30)) + 
         scale_x_continuous(breaks=seq(5,25,10)) +
         xlab("Moisture content (%)") +
         ylab(expression(Emission ~ factor ~ "(" * g ~ kg^-1*")")) + 
         labs(shape = "Fuel shape") + 
         theme(aspect.ratio=1,
               panel.border     = element_rect(fill=NA),
               panel.background = element_rect(fill="white"),
               strip.background = element_rect(fill="gray95", color="black"),
               strip.text.x = element_text(size=10, color="black"),
               panel.grid.major.x = element_line(color="gray95"),
               panel.grid.minor.x = element_line(color="gray95"),
               panel.grid.major.y = element_line(color="gray95"),
               panel.grid.minor.y = element_line(color="gray95"),
               axis.text    = element_text(size=10, color="black"),
               axis.title   = element_text(size=10, color="black"),
               axis.ticks   = element_line(color="black"),
               legend.text  = element_text(size=10, color='black'),
               legend.title = element_text(size=10, color='black'),
               legend.key   = element_blank(),
               legend.position = "none")
  
  ggsave(filename="../figures/voc_fuel.pdf", plot=last_plot(), device="pdf", width=6.5, height=4.5)
```


```{r, fig.width=3.15, fig.height=2.8, fig.align='center'}

  ggplot(dplyr::filter(df_ef, pol=="ultrafines"), 
         aes(x=mc_dry, y=ef, shape=type)) +
      geom_point(size=3, alpha=0.7) +
      coord_cartesian(xlim=c(4,30), ylim=c(0,2.5e16)) + 
      scale_x_continuous(breaks=seq(5,30,10)) + 
      #scale_y_continuous(breaks=seq(0.1,1.0,0.1)) + 
      xlab("Moisture content (%)") + 
      ylab(expression(Ultrafine ~ particles ~ "(" * "#" ~ kg^-1*")")) + 
      labs(shape = "Fuel shape") + 
      theme(aspect.ratio=1,
            panel.border     = element_rect(fill=NA),
            panel.background = element_rect(fill='white'),
            axis.text  = element_text(size=10, color='black'),
            axis.title = element_text(size=10, color='black'),
            axis.ticks = element_line(color='black'),
            legend.text  = element_text(size=10, color='black'),
            legend.title = element_text(size=10, color='black'),
            legend.key   = element_blank(),
            legend.box.background = element_rect(color="black", size=0.6),
            legend.position = c(0.79,0.215)) 

  ggsave(filename="../figures/ufp_fuel.pdf", plot=last_plot(), device="pdf", width=3.15, height=2.8)
```

```{r, fig.width=3.15, fig.height=3.15, fig.align='center'}

  ggplot(df_ecoc, aes(x=mc_dry, y=ratio, shape=type)) +
    geom_point(size=3, alpha=0.7) +
    coord_cartesian(xlim=c(4,30)) +
    scale_x_continuous(breaks=seq(5,30,10)) + 
    scale_y_continuous(breaks=seq(0,2,0.5)) + 
    xlab("Moisture content (%)") + 
    ylab("EC/OC") +
    labs(shape="Fuel shape") + 
    theme(aspect.ratio=1,
          panel.grid.major.x = element_line(color="gray95"),
          panel.grid.minor.x = element_line(color="gray95"),
          panel.grid.major.y = element_line(color="gray95"),
          panel.grid.minor.y = element_line(color="gray95"),
          panel.border     = element_rect(fill=NA),
          panel.background = element_rect(fill='white'),
          axis.text  = element_text(size=10, color='black'),
          axis.title = element_text(size=10, color='black'),
          axis.ticks = element_line(color='black'),
          legend.text  = element_text(size=10, color='black'),
          legend.title = element_text(size=10, color='black'),
          legend.key   = element_blank(),
          legend.box.background = element_rect(color="black", size=0.6),
          legend.position = c(0.8,0.8))
  
  ggsave(filename="../figures/ecoc_ratio.pdf", plot=last_plot(), device="pdf", width=8/2.54, height=8/2.54)
```

# Compare emission factors

### Between different fuel moisture levels and shapes

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="co2")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="co")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="ch4")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="grav")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="ultrafines")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="ec")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="oc")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="acetaldehyde")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="formaldehyde")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="benzene")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="toluene")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="ethylbenzene")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="m+p-xylene")))
```

```{r}
  TukeyHSD(aov(ef ~ fuel_mc_shape, dplyr::filter(df_ef, pol=="o-xylene")))
```

```{r}
  TukeyHSD(aov(ratio ~ fuel_mc_shape, df_ecoc))
```

### Between different fuel moisture levels

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="co2")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="co")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="ch4")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="grav")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="ultrafines")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="ec")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="oc")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="acetaldehyde")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="formaldehyde")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="benzene")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="toluene")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="ethylbenzene")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="m+p-xylene")))
```

```{r}
  TukeyHSD(aov(ef ~ mc_level, dplyr::filter(df_ef, pol=="o-xylene")))
```

```{r}
  TukeyHSD(aov(ratio ~ mc_level, df_ecoc))
```

# Summarize EFs by fuel moisture level

```{r}
  df_table <- dplyr::filter(df_ef, pol %in% c("co2","co","ch4","grav","ec","oc", 
                                              "acetaldehyde","formaldehyde",
                                              "benzene", "toluene", "ethylbenzene",
                                              "m+p-xylene", "o-xylene")) %>%
              dplyr::mutate(pol = factor(pol, levels=c("co2","co","ch4","grav","ec","oc", 
                                                       "acetaldehyde","formaldehyde",
                                                       "benzene", "toluene", "ethylbenzene",
                                                       "m+p-xylene", "o-xylene"))) %>%
              group_by(pol, mc_level) %>%
              summarise(mean   = mean(ef),
                        sd     = sd(ef))
```

```{r, echo=FALSE}
  kable(df_table, format="markdown", digits=4,
        col.names=c("Pollutant","Moisture level","Mean (g/kg)","SD (g/kg)"))
```

```{r}
  df_ecoc_summary <- group_by(df_ecoc, mc_level) %>%
                     summarise(mean   = mean(ratio),
                               sd     = sd(ratio))

  kable(df_ecoc_summary, format="markdown", digits=4,
        col.names=c("Moisture level","Mean","SD"))
```