
---
title: "Emission factors"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r global_options, include=FALSE}
  knitr::opts_chunk$set(fig.path = 'figures/',
                        warning = FALSE,
                        message = FALSE,
                        fig.width = 10, fig.height = 4,
                        cache = FALSE,
                        results='markup')
```

```{r, echo=FALSE}
  library(tidyverse)
  library(knitr)
  library(viridis)
  library(broom)
```

# Load

* sample info

```{r}
  samples <- readRDS("../r_files/samples.RDS")
```

* fuel burnt

```{r}
  fuel_burnt <- readRDS("../r_files/fuel_burnt.RDS")
```

* emissions

```{r}
  emissions_long <- readRDS("../r_files/emissions_long.RDS")
```

* hood flow

```{r}
  hood_flow <- readRDS("../r_files/hood_flow.RDS")
```

* polultant information

```{r}
  pol_properties <- readRDS("../r_files/pol_properties.RDS")
  fuel_properties <- readRDS("../r_files/fuel_properties.RDS")
```

* fuel information

```{r}
  fuel_properties <- readRDS("../r_files/fuel_properties.RDS")

fuel_prep_clean <- readRDS("../r_files/fuel_prep_clean.RDS")

mean_mc_df <- readRDS("../r_files/mean_mc_df.RDS")
 
mc_test_log <- readRDS("../r_files/mc_test_log.RDS")

test_times <- readRDS("../r_files/test_times.RDS")

grav <- readRDS("../r_files/grav_merged.RDS")

t_range <- readRDS("../r_files/t_range.RDS")
```

# Carbon calculations

* sum carbon over all pollutants

* calculate fuel burned using carbon balance method

```{r}
  carbon_emitted <- dplyr::filter(emissions_long, metric == "mass_carbon") %>% #grepl("co2", pol)) %>%
                    dplyr::group_by(id) %>%
                    dplyr::summarise(mass_carbon = sum(value, na.rm = TRUE)) %>%
                    dplyr::left_join(dplyr::select(samples, id, fuel),
                                     by = "id") %>%
                    dplyr::left_join(fuel_properties, by = "fuel") %>%
                    dplyr::left_join(select(fuel_burnt, -units), by = "id") %>%
                    dplyr::filter(!is.na(fuel)) %>%
                    dplyr::mutate(mass_fuel_from_carbon_balance_kg =
                                  mass_carbon * (1e-9) * (1/carbon_fraction),
                                  mass_fuel_scale_kg = mass_burned_correct * (1e-3)) %>%
                    dplyr::select(-mass_burned_correct)
                    # dplyr::filter(id != "FH2", id != "FL4", id != "LM4")
```

```{r}
char_ratio <- dplyr::mutate(fuel_burnt, ratio = mass_burned_correct/mass_dry_new_sum)
```

* plot fuel from carbon balance versus measured fuel mass

```{r}
  ggplot(carbon_emitted, aes(mass_fuel_scale_kg, mass_fuel_from_carbon_balance_kg)) + 
  geom_point() +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE, color = "black", formula = y ~ x) +
  geom_abline(intercept = 0, slope = 1, color = "grey90") +
  xlim(0, NA) + ylim(0, NA) +
  geom_text(aes(label = id), nudge_x = -0.01, nudge_y = 0.01)

```

# Carbon Balance
```{r}
fuel_prep_clean <- dplyr::rename(fuel_prep_clean, id = fuel_id)
df_carbon <- group_by(fuel_prep_clean, id) %>%
             summarise(mean_mc = mean(mc_actual))

mc_test_log <- select(mc_test_log, id, wgt_starter) %>%
               filter(!is.na(wgt_starter))

carbon_bal_2 <- dplyr::filter(emissions_long, metric == "mass_carbon", grepl("co2|co", pol)) %>%
  #grepl("co2", pol)) %>%
                    dplyr::group_by(id) %>%
                    dplyr::summarise(mass_carbon = 1e-6*sum(value, na.rm = TRUE)) %>%
                    dplyr::left_join(dplyr::select(samples, id, fuel),
                                     by = "id") %>%
                    dplyr::left_join(fuel_properties, by = "fuel") %>%
                    dplyr::left_join(select(fuel_burnt, -units), by = "id") %>%
                    dplyr::filter(!is.na(fuel)) %>%
                    dplyr::left_join(mc_test_log, by = "id")

carbon_bal_2 <- dplyr::left_join(carbon_bal_2, df_carbon, by = "id") %>%
                    dplyr::mutate(mass_carbon_burned = carbon_fraction *
                                    (mass_dry_new_sum - ((1- mean_mc/100)*unused_fuel-.8*.95*wgt_starter)) - (0.95 * char_ash)) # recheck the 0.8 multiple in the wgt_starter

ggplot(data = carbon_bal_2, aes(x = mass_carbon_burned, y = mass_carbon)) +
   geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "black", formula = y ~ x) +
  geom_abline(intercept = 0, slope = 1, color = "grey90") +
   theme_bw() +
  labs(x = "mass carbon burned", y = "mass carbon emitted")

  saveRDS(carbon_bal_2, file = "../r_files/carbon_bal_2.RDS")
```


```{r}
  df_carb <- dplyr::select(carbon_emitted,
                      mass_fuel_from_carbon_balance_kg,
                      mass_fuel_scale_kg) %>%
        dplyr::rename(mass_carbon_balance = mass_fuel_from_carbon_balance_kg,
                      mass_scale = mass_fuel_scale_kg)

  summary(lm(mass_carbon_balance ~ mass_scale, data = df_carb))
```

# Pollutant mass per kg fuel
 
```{r}
  emission_factors <- dplyr::left_join(dplyr::filter(emissions_long, metric == "mass_emitted"), 
                                       dplyr::select(fuel_burnt, id, mass_burned_correct),
                                       by = "id" ) %>%
                      dplyr::mutate(mass_pol_mass_fuel = ((value * 1e-3) / mass_burned_correct)) # g pol / kg fuel
                          
```

## Plot emissions by id

```{r, fig.width = 12, fig.height = 8}
  p_data <- dplyr::left_join(emission_factors,
                             dplyr::select(samples, id, type, mc_level),
                                           by = "id") %>%
            dplyr::filter(type != "bg") %>%
            dplyr::mutate(mc_level = forcats::fct_relevel(mc_level, "low",
                                                                    "med",
                                                                    "high")) %>%
            #dplyr::mutate(id = as.factor(id)) %>%
            dplyr::mutate(id = factor(id, levels = c("FL1", "FL2", "FL3", "FL4", "FM1", "FM2",
                                                     "FM3", "FM4", "FH1", "FH2", "FH3", "FH4",
                                                     "LL1", "LL2", "LL3", "LL4", "LM1", "LM2",
                                                     "LM3", "LM4", "LH1", "LH2", "LH3",
                                                     "LH4")))

   p_data_1 <- dplyr::filter(p_data, pol %in% c("co", "acetaldehyde", "benzene", "ch4",
                                        "ec", "ethylbenzene", "formaldehyde", "grav", "m+p-xylene",
                                        "o-xylene", "oc", "tc", "toluene"))

  ggplot(p_data_1, aes(id, mass_pol_mass_fuel, color = pol)) +
  geom_point(size = 4) +
  scale_y_log10() +
  theme_minimal() +
  theme(legend.position = "top",
        text = element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("g pol / kg fuel") + xlab("Test ID") +
  scale_color_viridis(discrete=TRUE)
  
```

## Remove tests

* remove FH2 from further plots

```{r}
 p_data <- p_data %>%
           dplyr::filter(id != "FH2")
```

## Gravimetric analysis

```{r}
 df <- dplyr::select(p_data, mass_pol_mass_fuel, type, mc_level, pol, id) %>%
       dplyr::mutate(type = as.factor(type),
                     mc_level = as.factor(mc_level)) %>%
       dplyr::filter(pol == "grav") %>%
       dplyr::mutate(test_fuel_type = paste(type, mc_level, sep = "_")) %>%
       dplyr::mutate(pol = "PM2.5")
 
 f <- formula(mass_pol_mass_fuel ~ mc_level + type + mc_level*type)
 
 mod_grav <- glm(f, data = df)
```

```{r}
  mod_grav$formula
```

```{r}
  summary(mod_grav)
```

```{r}
  grav_mean <- df %>%
               dplyr::group_by(test_fuel_type) %>%
               dplyr::summarise_if(is.numeric, mean, na.rm = TRUE)

  grav_mean_wide <- grav_mean %>%
                    tidyr::spread(test_fuel_type, mass_pol_mass_fuel)

  pairwise.t.test(df$mass_pol_mass_fuel, df$test_fuel_type, p.adj = "none")

  pairwise.t.test(df$mass_pol_mass_fuel, df$test_fuel_type, p.adj = "bonf")

  pairwise.t.test(df$mass_pol_mass_fuel, df$test_fuel_type, p.adj = "holm")

  pairwise.t.test(df$mass_pol_mass_fuel, df$mc_level, p.adj = "holm")

  TukeyHSD(aov(mass_pol_mass_fuel ~ test_fuel_type, df))

  TukeyHSD(aov(mass_pol_mass_fuel ~ mc_level, df))
```

```{r}
   plot(mod_grav)
```

```{r}
df_grav <- group_by(fuel_prep_clean,id) %>%
           summarise(mean_mc = mean(mc_actual))
           

df_grav <- left_join(df, df_grav)
df_grav_plot <- df_grav %>%
                dplyr::mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab")) %>%
                dplyr::rename(`MC Level` = mc_level, `Fuel Shape` = type)#dplyr::filter(df_grav, type == "lab")
  
ggplot(df, aes(mc_level, mass_pol_mass_fuel, color = "darkblue")) +
  geom_boxplot(outlier.size = NA) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") +
  ylab("g pol / kg fuel") + xlab("moisture content") +
  facet_wrap(~pol, scales = "free")

```

```{r}
ggplot(df_grav_plot, aes(x = mean_mc, y = mass_pol_mass_fuel, color = `MC Level`, shape = `Fuel Shape`)) +
  geom_point(size = 3.5) +
  theme_bw() +
  theme(text = element_text(size=15)) +
  #theme(legend.position = "none") +
  ylab("EF [g pol / kg fuel]") + xlab("Moisture Content (%)") +
  #title("Milled Fuel PM2.5") +
  facet_wrap(~pol, scales = "free") +
  scale_color_manual(values = c("darkorange", "darkorchid3", "blue")) +
  scale_x_continuous(breaks = c(5, 15, 25))
```

## ecoc

```{r}
  ecoc <- emissions_long %>%
          dplyr::filter(pol %in% c("ec", "oc"), metric %in% "conc") %>%
          dplyr::select(-inst)
          

  ecoc <- unique(ecoc) %>%
          tidyr::spread(pol, value) %>%
          dplyr::mutate(ratio = ec / oc) %>%
          dplyr::right_join(samples, by = "id") %>%
          dplyr::filter(type != "bg") %>%
          dplyr::mutate(mc_level = forcats::fct_relevel(mc_level, "low",
                                                                  "med",
                                                                  "high"),
                        test_fuel_type = paste(type, mc_level, sep = "_")) %>%
          dplyr::left_join(mean_mc_df, by = "id") %>%
          dplyr::mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab")) %>%
          dplyr::rename(`MC Level` = mc_level, `Fuel Shape` = type) %>%
          dplyr::filter(!grepl("FH2", id))

  ggplot(ecoc, aes(x = mean_mc, y = ratio, color = `MC Level`, shape = `Fuel Shape`)) +
  geom_point(size = 3.5) +
  theme_bw() +
  scale_color_manual(values = c("darkorange", "darkorchid3", "blue")) +
  theme(text = element_text(size = 15)) +
  ylab("EC/OC Ratio") + xlab("Moisture Content (%)") +
  scale_x_continuous(breaks = c(5, 15, 25))
  #ggtitle("EC ratio")

  #ecoc_summary <- ecoc %>%
                #  dplyr::group_by(mc_level) %>%
                 # dplyr::summarise_if(is.numeric, mean, na.rm = TRUE)

  #pairwise.t.test(ecoc$ratio, ecoc$mc_level, p.adj = "none")

 # pairwise.t.test(ecoc$ratio, ecoc$test_fuel_type, p.adj = "none")

```

```{r}
 df_ecoc <- dplyr::select(ecoc, ratio, `Fuel Shape`, `MC Level`, id) %>%
       dplyr::mutate(type = as.factor(`Fuel Shape`),
                     `MC Level` = as.factor(`MC Level`)) %>%
       dplyr::mutate(test_fuel_type = paste(type, `MC Level`, sep = "_"))

 median_ecoc <- group_by(df_ecoc, `MC Level`) %>%
                summarise(med_ecoc = median(ratio))
 
 f_ecoc <- formula(ratio ~ `MC Level` + type + `MC Level`*type)
 
 mod_ecoc <- glm(f_ecoc, data = df_ecoc)
 
 
 #df_ecoc <- rename(df_ecoc, mc_level = MC Level)
  ecoc_mean <- df_ecoc %>%
               dplyr::group_by(test_fuel_type) %>%
               dplyr::summarise_if(is.numeric, mean, na.rm = TRUE)

  ecoc_mean_wide <- ecoc_mean %>%
                    tidyr::spread(test_fuel_type, ratio)

  pairwise.t.test(df_ecoc$ratio, df_ecoc$test_fuel_type, p.adj = "none")

  pairwise.t.test(df_ecoc$ratio, df_ecoc$test_fuel_type, p.adj = "bonf")

  pairwise.t.test(df_ecoc$ratio, df_ecoc$test_fuel_type, p.adj = "holm")

  pairwise.t.test(df_ecoc$ratio, df_ecoc$`MC Level`, p.adj = "none")

  TukeyHSD(aov(ratio ~ test_fuel_type, df_ecoc))

  TukeyHSD(aov(ratio ~ df_ecoc$`MC Level`, df_ecoc))
```


## Plot emissions by moisture content

```{r}
mean_mc_df <- select(df_grav, id, mean_mc)

saveRDS(mean_mc_df, "../r_files/mean_mc_df.RDS")

p_data_ <- dplyr::filter(p_data, grepl("ec|oc",pol), !grepl("FH2", id))

p_data_ <- dplyr::left_join(p_data_, mean_mc_df) %>%
           dplyr::mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab")) %>%
           dplyr::rename(`MC Level` = mc_level, `Fuel Shape` = type) %>%
           dplyr::mutate(pol = as.factor(pol)) %>%
           dplyr::mutate(pol = forcats::fct_recode(pol, "Elemental Carbon" = "ec",
                                                      "Organic Carbon" = "oc"))

  ggplot(p_data_, aes(x = mean_mc, y = mass_pol_mass_fuel, color = `MC Level`, shape = `Fuel Shape`)) +
  geom_point(size = 3.5) +
  theme_bw() +
  scale_color_manual(values = c("darkorange", "darkorchid3", "blue")) +
  theme(text = element_text(size = 15)) + #legend.position = "none", 
  ylab("EF [g pol / kg fuel]") + xlab("Moisture Content (%)") +
  facet_wrap(~pol, scales = "free_y") +
  scale_x_continuous(breaks = c(5, 15, 25))
```

## Plot emissions by moisture content and fuel type

```{r}
library(forcats) 
 
p_data_gas <- dplyr::filter(p_data, pol %in% c("co", "co2", "ch4"), !grepl("FH2", id)) %>%
             dplyr::mutate(pol = factor(pol)) %>%
             dplyr::left_join(mean_mc_df, by = "id") %>%
             dplyr::mutate(pol = forcats::fct_recode(pol, "Carbon Monoxide" = "co",
                                                     "Carbon Dioxide" = "co2",
                                                     "Methane" = "ch4"))%>%
             dplyr::mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab")) %>%
             dplyr::rename(`Fuel Shape` = type, `MC Level` = mc_level)
           

  ggplot(p_data_gas, aes(x = mean_mc, y = mass_pol_mass_fuel, color = `MC Level`, shape = `Fuel Shape`)) +
  geom_point(size = 3.5) +
  theme_bw() +
  scale_color_manual(values = c("darkorange", "darkorchid3", "blue")) +
  theme(text = element_text(size = 15)) + 
  ylab("EF [g pol / kg fuel]") + xlab("Moisture Content (%)") +
  facet_wrap(~pol, scales = "free_y", ncol = 2) +
  scale_x_continuous(breaks = c(5, 15, 25))

```

## Plot tc vs grav

```{r, echo = FALSE, fig.width = 12, fig.height = 8}
  data_grav_tc <- dplyr::filter(emissions_long,
                               (pol == "grav" | pol == "tc") & metric == "conc") %>%
                  dplyr::select(-inst)%>%
                  tidyr::spread(pol, value)

  p <- ggplot(data_grav_tc, aes(grav, tc)) +
       geom_point(aes(color = id)) +
       theme_bw() +
       theme(legend.position = "top") +
       geom_abline(slope = 1, intercept = 0) +
       geom_text(aes(label = id), nudge_x = 200)

  p
```

```{r}
ultrafines <- filter(emissions_long, metric == "num_emitted") %>%
              filter(!grepl("^BG[1-9]", id)) %>%
              left_join(df_carbon, by = "id") %>%
              left_join(select(df_grav, id, mc_level, type), by = "id") %>%
              filter(!is.na(mc_level)) %>%
              mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab"))
              #rename(`Fuel Shape` = type, `MC Level` = mc_level)

ggplot(ultrafines, aes(x = mean_mc, y = value, color = mc_level, shape = type)) +
  geom_point(size = 3.5) +
  theme_bw() +
  scale_color_manual(values = c("darkorange", "darkorchid3", "blue")) +
  theme(text = element_text(size = 15)) + 
  ylab("EF [g pol / kg fuel]") + xlab("Moisture Content (%)") +
  scale_x_continuous(breaks = c(5, 15, 25))
```


# Save

```{r}
  saveRDS(carbon_emitted, file = "../r_files/carbon_emitted.RDS")
  saveRDS(emission_factors, file = "../r_files/emission_factors.RDS")
```
 
 * VOC EF's
 
```{r}
p_data_voc <- dplyr::filter(p_data, pol %in% c("benzene", "toluene", "ethylbenzene",
                                        "m+p-xylene", "o-xylene")) %>%
             dplyr::mutate(pol = factor(pol)) %>%
             dplyr::left_join(mean_mc_df, by = "id") %>%
             dplyr::mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab")) %>%
             dplyr::rename(`MC Level` = mc_level, `Fuel Shape` = type) %>%
             dplyr::mutate(pol = as.factor(pol)) %>%
             dplyr::mutate(pol = forcats::fct_relevel(pol, "benzene",
                                                      "toluene",
                                                      "ethylbenzene",
                                                      "m+p-xylene",
                                                      "o-xylene"))

  ggplot(p_data_voc, aes(x = mean_mc, y = mass_pol_mass_fuel, color = `MC Level`, shape = `Fuel Shape`)) +
  geom_point(size = 3) +
  theme_bw() +
  scale_color_manual(values = c("darkorange", "darkorchid3", "blue")) +
  theme(text = element_text(size = 15)) + #legend.position = "none", 
  ylab("EF [g pol / kg fuel]") + xlab("Moisture Content (%)") +
  facet_wrap(~pol, scales = "free_y") +
  scale_x_continuous(breaks = c(5, 15, 25))

```

```{r}
  p_data_carb <- dplyr::filter(p_data, grepl("formaldehyde|acetaldehyde",pol), !grepl("FH2", id)) %>%
                 dplyr::mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab")) %>%
                 dplyr::left_join(mean_mc_df, by = "id") %>%
                 dplyr::mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab")) %>%
                 dplyr::rename(`MC Level` = mc_level, `Fuel Shape` = type) %>%
                 dplyr::mutate(pol = as.factor(pol)) %>%
                 dplyr::mutate(pol = forcats::fct_recode(pol, "Acetaldehyde" = "acetaldehyde",
                                                      "Formaldehyde" = "formaldehyde"))

  ggplot(p_data_carb, aes(x = mean_mc, y = mass_pol_mass_fuel, color = `MC Level`, shape = `Fuel Shape`)) +
  geom_point(size = 4) +
  theme_bw() +
  scale_color_manual(values = c("darkorange", "darkorchid3", "blue")) +
  theme(text = element_text(size = 15)) + 
  ylab("EF [mg pol / kg fuel]") + xlab("Moisture Content (%)") +
  facet_wrap(~pol, scales = "free_y") +
  scale_x_continuous(breaks = c(5, 15, 25))

```

```{r eval= FALSE}
efs_spread <- dplyr::select(emission_factors, id, pol, mass_pol_mass_fuel) %>%
              dplyr::filter(!grepl("^BG|G", id)) %>%
              tidyr::spread(key = pol, value = mass_pol_mass_fuel)

library(ggraph)

test_data_dend <- efs_spread %>% 
  select(2:29) %>% 
  as.matrix() %>% 
  t() %>% 
  dist() %>% 
  hclust() %>% 
  as.dendrogram()

ggraph(test_data_dend, layout = "dendrogram") + 
  geom_edge_diagonal()
```

```{r}
emissions_long <- mutate(emissions_long, id = as.factor(id)) %>%
                  filter(!grepl("^BG|G", id))

anova_test <- dplyr::select(p_data, id, mc_level, type) %>%
              dplyr::distinct(id, mc_level, type)

anova_test <- dplyr::inner_join(emissions_long, anova_test) %>%
              dplyr::select(mc_level, type, pol, value) %>%
              dplyr::mutate(type = as.factor(type), pol = as.factor(pol))

anova_model <- anova_test %>%
                 filter(pol == "ch4") # , mc_level != "med")

                df_2way <- tidy(aov(value ~ mc_level + type + mc_level*type, data = anova_model))
                 
                df_1way <- tidy(aov(value ~ mc_level, data = anova_model)) 

#anova_test <- group_by(anova_test, pol) %>%
 #             summarise(two_way_model)
```

```{r}
# anova tests for pollutants efs               
df_anova <- dplyr::select(p_data, mass_pol_mass_fuel, type, mc_level, pol, id) %>%
       dplyr::mutate(type = as.factor(type),
                     mc_level = as.factor(mc_level)) %>%
       dplyr::filter(pol == "formaldehyde") %>%
       dplyr::mutate(test_fuel_type = paste(type, mc_level, sep = "_"))
                
#TukeyHSD(aov(mass_pol_mass_fuel ~ test_fuel_type, df_anova))
#TukeyHSD(aov(mass_pol_mass_fuel ~ mc_level, df_anova))
pairwise.t.test(df_anova$mass_pol_mass_fuel, df_anova$test_fuel_type, p.adj = "none")
pairwise.t.test(df_anova$mass_pol_mass_fuel, df_anova$test_fuel_type, p.adj = "holm")
pairwise.t.test(df_anova$mass_pol_mass_fuel, df_anova$mc_level, p.adj = "holm")
```

```{r}
mean_emissions <- select(p_data, mc_level, pol, mass_pol_mass_fuel) %>%
                  group_by(mc_level, pol) %>%
                  summarize(sd_pol = sd(mass_pol_mass_fuel))
```

```{r fig.height= 11, fig.width=6}
#plot of all carbonyls

carb_plot <- filter(p_data, !grepl("ec|oc|co2|co|ch4|benzene|toluene|ethylbenzene|o-xylene|m+p-xylene|tc|grav|2_5_dimethylbenzaldehyde", pol))
             #rename(dimethylbenzaldehyde = `)

ggplot(carb_plot, aes(x = mc_level, y = mass_pol_mass_fuel, color = mc_level, shape = type)) +
  geom_point(size = 2) +
  theme_bw() +
  scale_color_manual(values = c("darkorange", "darkorchid3", "blue")) +
  theme(text = element_text(size = 12)) + 
  ylab("EF [mg pol / kg fuel]") + xlab("Moisture Content (%)") +
  facet_wrap(~pol, scales = "free_y", ncol = 3) +
  scale_x_discrete(labels = c("low", "med", "high"))
```

```{r}
# for  ISO tier comparisons

dur <- select(grav, id, start, end) %>%
       filter(!is.na(start)) %>%
       mutate(dur = (end - start)/60)

iso_pm <- filter(p_data, grepl("grav", pol)) %>%
          left_join(dur, by = "id") %>%
          select(id, type, mc_level, value, dur) %>%
          left_join(mean_mc_df, by = "id") %>%
          mutate(mg_per_min = (value/10^3)/dur) %>%
          mutate(type = as.factor(type)) %>%
          mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab")) #%>%
          #rename(MC Level = mc_level, Fuel Shape = type)

iso_pm_mean <- group_by(iso_pm, mc_level) %>%
              summarise(med_pm = mean(mg_per_min))

  ggplot(iso_pm, aes(x = mean_mc, y = mg_per_min, color = mc_level, shape = type)) +
  geom_point(size = 4) +
  theme_bw() +
  scale_color_manual(values = c("darkorchid3", "darkorchid3", "blue")) +
  theme(text = element_text(size = 15)) + 
  ylab("PM2.5 [mg/min]") + xlab("Moisture Content (%)") +
  scale_x_continuous(breaks = c(5, 15, 25))
```

```{r}
iso_co <- filter(p_data, grepl("co", pol)) %>%
          filter(!grepl("co2", pol)) %>%
          left_join(dur, by = "id") %>%
          select(id, type, mc_level, value, dur) %>%
          left_join(mean_mc_df, by = "id") %>%
          mutate(g_per_min = (value/10^6)/dur)%>%
          mutate(type = as.factor(type)) %>%
          mutate(type = forcats::fct_recode(type, "Split" = "field",
                                                      "Milled" = "lab"))

iso_co_mean <- group_by(iso_co, mc_level) %>%
              summarise(mean_co = mean(g_per_min))

ggplot(iso_co, aes(x = mean_mc, y = g_per_min, color = mc_level, shape = type)) +
  geom_point(size = 4) +
  theme_bw() +
  scale_color_manual(values = c("darkorange", "darkorchid3", "blue")) +
  theme(text = element_text(size = 15)) + 
  ylab("CO [g/min]") + xlab("Moisture Content (%)") +
  scale_x_continuous(breaks = c(5, 15, 25))
```

```{r}

```

