---
title: "fivegas"
author: "Nicholas Good"
date: "11/20/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
  source("../r_scripts/R_functions.R")
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")    
```

## Load fivegas data

Mixing ratio files

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/fivegas_conc.Rda")    
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_conc, 2)
```

Voltage files

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/fivegas_volts.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_volts, 2)
```

Voltage calibration files

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/fivegas_calibration.Rda")    # fivegas voltage calibration data files
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_calibration, 2)
```

## Convert to longer format

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_conc <- as_tibble(dplyr::select(fivegas_conc, -datetime_secs, -time_str) %>%
                    tidyr::gather(pol, val, ch4:co))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_conc, 2)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_volts <- as_tibble(tidyr::gather(fivegas_volts, pol, val, co2:ch4))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_volts, 2)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_calibration <- as_tibble(tidyr::gather(fivegas_calibration, pol, val, co2:ch4))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_calibration, 2)
```

## Load calibration metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/cal_1.Rda")        # meta data
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(cal_1, 2)
```

## Split calibration data

Calibration concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal_conc <- tidy.date(cal_1, "^conc_", "conc_")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_cal_conc, 2)
```

Calibration time periods

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal_times <- tidy.date(cal_1, "^time_", "time_")
  fivegas_cal_times <- split.fivegascaltimes(fivegas_cal_times)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_cal_times, 2)
```

## Apply calibration

Calculate calibration voltages

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal <- tidyr::spread(fivegas_cal_times, type, value) %>%
                        dplyr::filter(!is.na(start) & pol != "zero")  %>%
                        dplyr::rowwise() %>%
                        dplyr::mutate(v_mean = time_window_stats(fivegas_calibration, date, start, end, pol, stat_var = "mean"))  %>%
                        dplyr::mutate(v_sd = time_window_stats(fivegas_calibration, date, start, end, pol, stat_var = "sd"))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_cal, 2)
```

Calculate zero voltages

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # extract zero times
    fivegas_cal_zero <- tidyr::spread(fivegas_cal_times, type, value) %>%
                        dplyr::filter(!is.na(start) & pol == "zero") %>%
                        dplyr::select(-pol) %>%
                        dplyr::rename(start_zero = start, end_zero = end)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_cal_zero, 2)
```

## Append zero times to cal data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal <- dplyr::left_join(fivegas_cal, fivegas_cal_zero, by = "date")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_cal, 2)
```
 
## Append zero voltages to cal data
    
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal <- dplyr::mutate(fivegas_cal, v_zero_mean = time_window_stats(fivegas_calibration, date, start_zero, end_zero, pol, stat_var = "mean"))  %>%
                  dplyr::mutate(v_zero_sd = time_window_stats(fivegas_calibration, date, start_zero, end_zero, pol, stat_var = "sd"))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_cal, 2)
```

Merge calibration voltages with standard mixing ratios

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal_conc <- dplyr::rename(fivegas_cal_conc, pol = var, standard = value)
  fivegas_cal <- merge(fivegas_cal, fivegas_cal_conc, by=c("date", "pol"))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_cal, 2)
```

Calculate calibration factors

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal$cal_beta <- fivegas_cal$standard/(fivegas_cal$v_mean-fivegas_cal$v_zero_mean)
  fivegas_cal$cal_int <- -(fivegas_cal$standard*fivegas_cal$v_zero_mean)/(fivegas_cal$v_mean-fivegas_cal$v_zero_mean)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_cal, 2)
```

Add calibration date to voltage data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_volts <- add_caldate(fivegas_volts, fivegas_cal)
  fivegas_volts$pol <- as.factor(fivegas_volts$pol)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_volts, 2)
```

Merge voltage data and calibration data, then apply calibration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_volts <- dplyr::left_join(fivegas_volts, 
                      select(fivegas_cal, date, pol, cal_beta, cal_int),
                      by = c("cal_date" = "date", "pol" = "pol")) %>%
                      
                      dplyr::rename(volts = val) %>%
  
                      mutate(val = volts*cal_beta + cal_int)

  fivegas_volts$pol <- as.factor(fivegas_volts$pol)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_volts, 2)
```

Drop variables

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_volts <- dplyr::select(fivegas_volts, datetime, date, time, id, pol, val)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(fivegas_volts, 2)
```

## Merge voltage and mixing ratio data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_merged <- dplyr::arrange(rbind(fivegas_volts, fivegas_conc), datetime)
```

## Save merged data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(fivegas_merged, file = "../r_files/fivegas_merged.Rda")
```

## Plot data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=40}
  df <- dplyr::filter(fivegas_merged, !is.na(val) & pol != "o2" & pol != "nox")  # filter out NAs, o2 and nox
  ggplot(df, aes(datetime, val, colour = pol)) + 
        geom_line() +
        scale_y_log10(limits = c(1, 10^5)) +
        facet_wrap(~id, ncol = 3, scale = "free") +
        theme_minimal()
```

## QC

Load notes
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")    # notes
  notes <- dplyr::filter(notes, grepl("fivegas|all", notes$inst) == TRUE)
```

Flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc[33:35] <- "ok"
```

Set one flag per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
            group_by(id) %>%
            arrange(qc) %>%
            summarise(qc = first(qc))
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_merged <- dplyr::left_join(fivegas_merged, flags, by = "id")
  fivegas_merged$id <- as.factor(fivegas_merged$id)
  fivegas_merged$qc <- ifelse(is.na(fivegas_merged$qc), "ok", as.character(fivegas_merged$qc))
  fivegas_merged$qc <- as.factor(fivegas_merged$qc)
```

Addtional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_merged$qc[fivegas_merged$id == "23B"] <- "bad" # CO:CH4
  fivegas_merged$qc[fivegas_merged$id == "17C"] <- "bad" # CO2
  fivegas_merged$qc[fivegas_merged$id == "4B"] <- "bad" # CO2
  fivegas_merged$qc[fivegas_merged$id == "18C"] <- "bad" # CH4
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(fivegas_merged, file = "../r_files/fivegas_merged.Rda")
```

## Plot QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80}
  df <- dplyr::filter(fivegas_merged, !is.na(val) & pol == "co")  # filter
  ggplot(df,  aes(datetime, val, colour = qc, group = pol)) +
        geom_line() +
        scale_y_log10(limits = c(1, 10^5)) +
        facet_wrap(~id, ncol = 3, scale = "free") +
        theme_minimal() +
        xlab("") + ylab("co ppm")
  
  df <- dplyr::filter(fivegas_merged, !is.na(val) & pol == "co2")  # filter
  ggplot(df,  aes(datetime, val, colour = qc, group = pol)) +
        geom_line() +
        scale_y_log10(limits = c(1, 10^5)) +
        facet_wrap(~id, ncol = 3, scale = "free") +
        theme_minimal() +
        xlab("") + ylab("co2 ppm")
  
  df <- dplyr::filter(fivegas_merged, !is.na(val) & pol == "ch4")  # filter
  ggplot(df,  aes(datetime, val, colour = qc, group = pol)) +
        geom_line() +
        scale_y_log10(limits = c(1, 10^5)) +
        facet_wrap(~id, ncol = 3, scale = "free") +
        theme_minimal() +
        xlab("") + ylab("ch4 ppm")
```

## Summary

The $Fivegas$ measured during `r length(unique(fivegas_merged$id))` experiments between `r min(fivegas_merged$date, na.rm = TRUE)` and `r max(fivegas_merged$date, na.rm = TRUE)`. There is no $Fivegas$ data for tests: `r setdiff(as.character(samples$id), as.character(fivegas_merged$id))`.

$Fivegas$ data is expected to be missing for: 

All "G" tests: five gas not run during backgrounds.

1B, 2A: "no data" noted on QC log

4A: only calibration file for this test day was found (2/24/2016); five gas file is lost.
