---
title: "Fivegas Data Processing"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
  source("../r_scripts/R_functions.R")
```

# Load data

* data logged in ppm units.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/fivegas_conc.Rda")    
```

* data logged in voltage units

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/fivegas_volts.Rda")
```

* voltage calibration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/fivegas_calibration.Rda")
```

* calibration metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/cal_1.Rda")
```

* sample information

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
```

* notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")
```

# Reformat data

## concentration

Convert concentration data to longer format

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_conc <- tibble::as_tibble(dplyr::select(fivegas_conc,
                                                  -datetime_secs,
                                                  -time_str) %>%
                  tidyr::gather("pol", "ppm", ch4:co))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', cache=FALSE}
  head(fivegas_conc, 2)
```

## voltage

Convert voltage data to longer format

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_volts <- as_tibble(tidyr::gather(fivegas_volts, pol, volts, co2:ch4))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  head(fivegas_volts, 2)
```

## voltage calibration

Convert voltage calibration data to longer format

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  fivegas_calibration <- as_tibble(tidyr::gather(fivegas_calibration,
                                                 pol,
                                                 volts,
                                                 co2:ch4))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  head(fivegas_calibration, 2)
```

# Wrangle data

* calibration standard mixing ratios

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal_conc <- tidy_date(cal_1, "^conc_", "conc_")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  head(fivegas_cal_conc, 2)
```

* calibration timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal_times <- tidy_date(cal_1, "^time_", "time_")

  fivegas_cal_times <- split_fivegas_cal_times(fivegas_cal_times)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  head(fivegas_cal_times, 2)
```

# Calibration

## calculate and organize

* response to standard

Calculate mean voltage over last 30 seconds of each calibration window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal <- tidyr::spread(fivegas_cal_times, type, value) %>%
                 dplyr::filter(!is.na(start) & pol != "zero")  %>%
                 dplyr::rowwise() %>%
                 dplyr::mutate(v_mean = 
                               time_window_stats(fivegas_calibration,
                                                 date,
                                                 start, end,
                                                 pol_var = pol,
                                                 val_var = "volts",
                                                 stat_var = "mean")) %>%
                 dplyr::mutate(v_sd =
                               time_window_stats(fivegas_calibration,
                                                 date,
                                                 start, end,
                                                 pol_var = pol,
                                                 val_var = "volts",
                                                 stat_var = "sd"))
```

* zero time windows

Extract zero time periods

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal_zero <- tidyr::spread(fivegas_cal_times, type, value) %>%
                      dplyr::filter(!is.na(start) & pol == "zero") %>%
                      dplyr::select(-pol) %>%
                      dplyr::rename(start_zero = start, end_zero = end)
```

* merge standard and zero 

Join zero time periods with the calibration data 

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal <- dplyr::left_join(fivegas_cal, fivegas_cal_zero, by = "date")
```

* response to zero

Calculate mean voltage during last 30 seconds of each zero calibration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal <- dplyr::mutate(fivegas_cal,
                        v_zero_mean =
                        time_window_stats(fivegas_calibration,
                                          date,
                                          start_zero, end_zero,
                                          pol_var = pol,
                                          val_var = "volts",
                                          stat_var = "mean")) %>%
                  dplyr::mutate(v_zero_sd =
                         time_window_stats(fivegas_calibration,
                                           date,
                                           start_zero, end_zero,
                                           pol_var = pol,
                                           val_var = "volts",
                                           stat_var = "sd"))
```

* remove NaNs

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal <- na.omit(fivegas_cal)
```

* combine voltage and standard info

Merge measured voltage with standard mixing ratio

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal_conc <- dplyr::rename(fivegas_cal_conc,
                                    pol = var,
                                    standard = value)

  fivegas_cal <- dplyr::left_join(fivegas_cal, fivegas_cal_conc,
                                  by = c("date", "pol"))
```

* calibration coefficients

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_cal <- dplyr::mutate(fivegas_cal,
                        cal_beta = (standard / (v_mean - v_zero_mean))) %>%
                 dplyr::mutate(cal_int = - (v_zero_mean * cal_beta))
```

* add calibration date

Add calibration date to voltage data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_volts <- add_caldate(fivegas_volts, fivegas_cal)

  fivegas_volts <- dplyr::mutate(fivegas_volts, pol = as.factor(pol))
```

## plot span

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=6}
 p_df <- tidyr::gather(fivegas_cal,
                       "type", "volts",
                       v_mean, v_zero_mean) %>%
         dplyr::mutate(sd = ifelse(type == "v_mean", v_sd, v_zero_sd)) %>%
         dplyr::select(-v_sd, -v_zero_sd)

 # standard response voltage error bars
  v_limits <- aes(ymax = volts + sd, ymin = volts - sd)  # error bars

 # plot
  ggplot(p_df, aes(x = date, y = volts, group = type, colour = type)) +
    geom_point() +
    geom_errorbar(v_limits, width = 0.2) +
    facet_wrap(~pol, 1, scales = "free") +
    theme_minimal() +
    xlab("") +
    ylab("span") +
    theme(legend.position="top")

```

## apply calibration

Merge voltage data and calibration data, then apply calibration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_volts <- dplyr::left_join(fivegas_volts, 
                   dplyr::select(fivegas_cal, date, pol, cal_beta, cal_int),
                                 by = c("cal_date" = "date", "pol" = "pol")) %>%
                   dplyr::mutate(ppm = volts * cal_beta + cal_int) %>%
                   dplyr::mutate(pol = as.factor(pol))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  head(fivegas_volts, 2)
```

# Combine file types

Drop variables prior to merge with mixing ration data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_volts <- dplyr::select(fivegas_volts,
                                 datetime, date, time, id, pol, ppm)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  head(fivegas_volts, 2)
```

Merge voltage and mixing ratio data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_merged <- dplyr::arrange(dplyr::bind_rows(fivegas_volts,
                                                    fivegas_conc),
                                   datetime)
```

Remove ozone and nox rows (channels not used during study)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_merged <- dplyr::filter(fivegas_merged, pol != "o2" & pol != "nox")
```


# QC

Extract notes relevant to  fivegas analyzer

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("fivegas|all", notes$inst) == TRUE)
```

Flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc[33:35] <- "ok"
```

Make one flag per test precidented from bad to ok.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_merged <- dplyr::left_join(fivegas_merged, flags, by = "id") %>%
                    dplyr::mutate(id = as.factor(id),
                                  qc = as.factor(ifelse(is.na(qc),
                                                 "ok", as.character(qc))))

```

Additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_merged$qc[fivegas_merged$id == "23B"] <- "bad"  # CO:CH4
  fivegas_merged$qc[fivegas_merged$id == "17C"] <- "bad"  # CO2
  fivegas_merged$qc[fivegas_merged$id == "4B"] <- "bad"   # CO2
  fivegas_merged$qc[fivegas_merged$id == "18C"] <- "bad"  # CH4
```

# Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(fivegas_merged, file = "../r_files/fivegas_merged.Rda")
```

# Summary

The Fivegas analyzer measured during `r length(unique(fivegas_merged$id))` experiments between `r min(fivegas_merged$date, na.rm = TRUE)` and `r max(fivegas_merged$date, na.rm = TRUE)`. There is no $Fivegas$ data for tests: `r setdiff(as.character(samples$id), as.character(fivegas_merged$id))`.

$Fivegas$ data is expected to be missing for: 

* all "G" tests: five gas not run during backgrounds.

* 1B, 2A: "no data" noted on QC log

* 4A: only calibration file for this test day was found (2/24/2016); five gas file is lost.

* calibration files for 2016/03/09 and 2016/04/01 missing (check)

# Plots

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=40, eval=TRUE}
  df <- dplyr::mutate(fivegas_merged, id = as.character(id)) %>%
        dplyr::arrange(id) %>%
        dplyr::mutate(id = as.factor(id))

  ggplot(df, aes(datetime, ppm, colour = qc, group = pol)) + 
    geom_line() +
    scale_y_log10(limits = c(1, 10^5)) +
    scale_colour_manual(values = 
                        c("ok" = "mediumaquamarine",
                          "maybe" = "mediumorchid1",
                          "bad" = "darkorange1")) +
    facet_wrap(~id, ncol = 3, scale = "free") +
    theme_minimal() +
    xlab("") +
    theme(legend.position = "top")
```
