---
title: "gravimetric"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(forcats)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

# Load data

* gravimetric

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- readRDS("../r_files/grav.RDS")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  test_info <- readRDS("../r_files/test_info.RDS")
  times_fuel <- readRDS("../r_files/times_fuel.RDS")
  times_bg <- readRDS("../r_files/times_bg.RDS")
  times_test <- readRDS("../r_files/times_test.RDS")
  flow <- readRDS("../r_files/flows.RDS")
  notes <- readRDS("../r_files/qc.RDS")
```

# Organize

* make longer

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::filter(grav, !is.na(id)) %>%
          tidyr::gather("var", "val", matches("^wgt.*")) %>%
          dplyr::filter(grepl(".*avg.*|.*dif.*", var) == FALSE) %>%
          tidyr::separate("var", c("type", "when", "rep")) %>%
          dplyr::mutate(type = ifelse(when == "cal", "cal", "sample"),
                        when = ifelse(when == "cal", rep, when),
                        rep = ifelse(rep == "pre" | rep == "post",
                              "1", rep))
```

* average

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::group_by(grav, id, type, 
                           when, date_pre, date_post,
                           toc_pre, rh_pre,
                           toc_post, rh_post,
                           id_filter,
                           id_grav,
                           id_blank) %>%
           dplyr::summarise(mean = mean(val, na.rm = TRUE)) %>%
           dplyr::ungroup() %>%
           tidyr::spread(when, mean) %>%
           dplyr::mutate(delta = post - pre) %>%
           dplyr::rename(id_test = id)
```

* merge with id and flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::left_join(grav, test_info)

  flow <- dplyr::filter(flow, inst == "white") %>%
          dplyr::select(id, mean, delta) %>%
          dplyr::rename(flow_delta = delta,
                         flow_mean = mean)
  
  grav <- left_join(grav, flow, by = "id")
```

* add pollutant variable `pol`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, pol = "grav")
```

## Calibration weight

Spread data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  cal <- dplyr::filter(grav, type == "cal") %>%
         dplyr::select(id, delta) %>%
         dplyr::rename(cal_delta = delta)

  grav <- dplyr::filter(grav, type == "sample") %>%
          dplyr::left_join(cal)
```

## Test times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 grav <- dplyr::left_join(grav,
                           dplyr::select(times_test, id, start, end))
```

## Test types

Test are either a sample: `sample`, a background experiment: `bg`, or a filter blank: `bk`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 grav <- dplyr::mutate(grav, conc = ((delta - cal_delta) * 60 * 1000) / (flow_mean / (end - start))) # ug/l converted to ug/m^3. time converted from seconds to minutes since flow rates in lpm. 
```

## Test type

Is B[0-9] are blank filters.BG[0-9] are background tests.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, 
                         type = ifelse(grepl("^BG.*", id_test),
                         "bg", type),
                         type = ifelse(grepl("^B[0-9].*", id_test), "bk", type))

```

# QC

* extract notes for gravimetric data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("grav|all", instrument))

  grav <- dplyr::left_join(grav, notes, by = "id")
```

# Blanks (LOD/LOQ)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bk <- dplyr::filter(grav, type == "bk") %>%
        dplyr::summarise(lod = 3 * sd(delta, na.rm = TRUE),
                         loq = 10 * sd(delta, na.rm = TRUE),
                         mean_bk = mean(delta, na.rm = TRUE))
```

The average blank has `r bk$mean_bk[1]`$\mu g$ of mass.

The LOD is `r bk$lod[1]`$\mu g$ and the LOQ is `r bk$loq[1]`$\mu g$.

## Conc calculation

* adjusting concentration to account for calibration weight delta and blank delta

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # (ug - ug) * 60 * 1000 / ((L / min) * (s - s)) = ug / m^3
  grav <- dplyr::mutate(grav, 
                        conc = (delta - cal_delta - bk$mean_bk[1]) * 60 * 1000 /
                               (flow_mean * (end - start)))
```

# Background

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg <- dplyr::filter(grav, type == "bg") %>%
        dplyr::summarise(mean = mean(conc, na.rm = TRUE),
                         bg_sd = sd(conc, na.rm = TRUE))

  grav <- dplyr::mutate(grav, bg_conc = bg$mean[1])
```

* calculate background-adjusted concentration
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav$conc_bgadj <- grav$conc - grav$bg_conc
```

* calculate mass per startup replicate 
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  dplyr::mutate(grav, 
                mass_rep = ((delta - cal_delta - bk$mean_bk[1]) / fuel_quant))

```
# Summary

Gravimetric data was collected during `r length(unique(grav$id))` experiments between `r min(grav$date, na.rm = TRUE)` and `r max(grav$date, na.rm = TRUE)`. There is no Gravimetric data for tests: `r setdiff(as.character(test_info$id), as.character(grav$id))`.(ALL DATA ACCOUNTED FOR!)

# Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  saveRDS(grav, file = "../r_files/grav_merged.RDS")
```

# Plots

## plot environ

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(grav) +
  geom_point(aes(id, toc_pre, colour = "t_pre")) +
  geom_point(aes(id, toc_post, colour = "t_post")) +
  geom_point(aes(id, rh_pre, colour = "rh_pre")) +
  geom_point(aes(id, rh_post, colour = "rh_post")) +
  scale_colour_manual("", 
                      breaks = c("t_pre", "t_post",
                                 "rh_pre", "rh_post"),
                      values = c("dodgerblue1", "dodgerblue4",
                                 "tomato1", "tomato4")) +
  theme_minimal() +
  theme(text = element_text(size=20)) +
  ylab("t (oC) / RH (%)")
```

## plot mass

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(grav) +
  geom_point(aes(id, pre, colour = "pre")) +
  geom_point(aes(id, post, colour = "post")) +
  scale_colour_manual("", 
                      breaks = c("pre", "post"),
                      values = c("dodgerblue1", "tomato4")) +
  theme_minimal() +
  theme(text = element_text(size = 20)) +
  ylab("filter mass (ug)")
```

## plot delta mass by fuel type, qc status noted

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(grav) +
    geom_point(aes(id, delta, colour = qc)) +
    theme_minimal() +
    theme(text = element_text(size = 20)) +
    ylab("delta mass (ug)") +
    facet_wrap(~fuel_type, scales = "free")
```

## plot concentration (ug/m3)

Not background adjusted -  mean background shown in black. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(grav) +
    geom_point(aes(id, conc, colour = type)) +
    geom_point(aes(id, bg_conc), colour = "black") +
    theme_minimal() +
    theme(text = element_text(size = 20)) +
    ylab("conc. ug/m^3")
```

## Background-corrected concentration (ug/m3) by fuel type, QC status noted 

(note: assume "NA" for QC means the grav is good)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=20, fig.height=15}
  ggplot(filter(grav, fuel_type!="background"), aes(id, conc_bgadj, colour = qc)) +
  geom_point(size = 4) +
  theme_minimal() +
  theme(text = element_text(size = 28)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("ug/m^3") +
  #scale_y_log10() +
  facet_wrap(~fuel_type, ncol = 5, scales = "free")
```
## plot flows

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(grav) +
    geom_point(aes(id, flow_mean, colour = "mean flow")) +
    geom_point(aes(id, flow_delta, colour = "flow delta")) +
    scale_colour_manual("", 
                        breaks = c("mean flow", "flow delta"),
                        values = c("dodgerblue1", "tomato4")) +
    theme_minimal() +
    theme(text = element_text(size = 20)) +
    ylab("flow rate (L/min)")
```


# Data Quality Conclusions (KF on 5/26/17)

Test 26 (L2) has lower average flow. Pre-test flow was normal but post-test flow was low (`r flow$flow_delta[26]`). Potentially indicative of a clog in the line. The background-corrected concentration is much higher than the other "L" tests. Consider changing QC of this test to "bad" and dropping from dataset? 

Test 8 (W1) is QC of maybe, with note that a different type of wood shims were used compared to the other tests, and the shims kept going out and having to be relit (at least 6 relits each). THe delta_mass looks aligned with other wood shim tests, but the background-adjusted concentration is much higher. Consider changing QC of this test to "bad" and dropping from dataset?



