---
title: "grav"
author: "Nicholas Good"
date: "11/20/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")    # sample info
```

## Load gravimetric data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/grav.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(grav)
```

## Merge with filter metadata

Load metadata
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/data_1.Rda")  # metadata
```

Extract gravimetric filter cartridge data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  meta <- dplyr::select(data_1, matches("_white_|^id$|^date$"))

  flows <- dplyr::select(meta, matches("flow_|^id$|^date$")) %>%
            tidyr::gather("var", "value", -id, -date) %>%
            dplyr::filter(grepl("_avg$", var)==FALSE) %>%
            dplyr::mutate(type = as.factor(sub("flow.*", "", var)),
                          rep = as.factor(gsub("[^0-9]", "", var))) %>%
            dplyr::select(-var)  %>%
            dplyr::filter(!is.na(id)) %>%
            dplyr::group_by(id, type, date) %>%
            dplyr::summarise(mean = mean(value))  %>%
            tidyr::spread(type, mean)
  
  times <- dplyr::select(meta, matches("^time_|^id$")) %>%
            tidyr::gather("var", "value", -id) %>%
            dplyr::mutate(type = ifelse(grepl("_start_", var), "start", "NA"),
                          type = ifelse(grepl("_end_", var), "end", type)) %>%
            dplyr::filter(!is.na(id)) %>%  
            dplyr::select(-var) %>%  
            tidyr::spread(type, value)
```

## Filter gravimatric data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- filter(grav, grepl("^[0-9]|^G[0-9]", grav$id))
```

Merge with filter metadata and drop average values

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::left_join(grav, flows, by = "id")  %>%  
                  dplyr::left_join(times, by = "id") %>%  
                  dplyr::filter(grepl("^[0-9]", id)==TRUE | grepl("^G[0-9]", id)==TRUE) %>%
                  dplyr::rename(flow_pre = pre, flow_post = post) %>%
                  dplyr::rename(time_start = start, time_end = end) %>%
                  dplyr::select(-wgt_pre_avg,
                                -wgt_post_avg,
                                -lod, 
                                -wgt_dif) %>%
                  dplyr::arrange(id)

  grav_merged$id <- as.factor(grav_merged$id)
```

## Melt and clean

Split pre and post measurements

Pre-weight average

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_pre <- dplyr::select(grav_merged, wgt_pre_1, wgt_pre_2, wgt_pre_3, id) %>%
              tidyr::gather("var", "val", wgt_pre_1, wgt_pre_2, wgt_pre_3) %>%
              dplyr::filter(!is.na(val)) %>%
              dplyr::group_by(id) %>%
              dplyr::summarise(wgt_pre = mean(val))
```

Post-weight average

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_post <- dplyr::select(grav_merged, wgt_post_1, wgt_post_2, wgt_post_3, id) %>%
              tidyr::gather("var", "val", wgt_post_1, wgt_post_2, wgt_post_3) %>%
              dplyr::filter(!is.na(val)) %>%
              dplyr::group_by(id) %>%
              dplyr::summarise(wgt_post = mean(val))
```

Weight difference

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_dif <- dplyr::full_join(grav_pre, grav_post, by = "id") %>%
                dplyr::mutate(wgt_delta = wgt_post - wgt_pre)
```

Merge differences

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::left_join(grav_merged, grav_dif, by = "id") %>%
                  dplyr::ungroup()
```

## Add variables

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::mutate(grav_merged, 
                               wgt_cal_delta = wgt_cal_post - wgt_cal_pre) # calibration weight change

  grav_merged <- dplyr::mutate(grav_merged,
                               wgt_blank_delta = wgt_blank_avg_post - wgt_blank_avg_pre) # blank change
```

## QC

Load notes.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")    # notes
  notes <- dplyr::filter(notes, grepl("grav|all", notes$inst) == TRUE)
```

Flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc[33] <- "bad"
  notes$qc[34] <- "bad"
```

Set one flag per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
            group_by(id) %>%
            arrange(qc) %>%
            summarise(qc = first(qc))
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::left_join(grav_merged, flags, by = "id")
  grav_merged$id <- as.factor(grav_merged$id)
  grav_merged$qc <- ifelse(is.na(grav_merged$qc), "ok", as.character(grav_merged$qc))
  grav_merged$qc <- as.factor(grav_merged$qc)
```

Addtional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged$qc[grav_merged$id == "28A"] <- "bad"
  grav_merged$qc[grav_merged$id == "28B"] <- "bad"
  grav_merged$qc[grav_merged$id == "29A"] <- "bad"
```

Missing data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  missing <- dplyr::select(grav_merged, -notes) 
  missing <- missing[!complete.cases(missing),]
```

A total of `r nrow(missing)` tests are missing data. The test id are: `r missing$id`.

## Plots

Weights

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(dplyr::filter(grav_merged, grepl("^[0-9][A-Z]", id)==TRUE),
         aes(id, wgt_delta, color= qc)) +
  geom_point() +
  theme_minimal()
  
  ggplot(dplyr::filter(grav_merged, grepl("^[1][0-9][A-Z]", id)==TRUE),
         aes(id, wgt_delta, color= qc)) +
  geom_point() +
  theme_minimal()
  
  ggplot(dplyr::filter(grav_merged, grepl("^[2-3][0-9][A-Z]", id)==TRUE),
         aes(id, wgt_delta, color= qc)) +
  geom_point() +
  theme_minimal()
  
  ggplot(dplyr::filter(grav_merged, grepl("^[G]", id)==TRUE),
         aes(id, wgt_delta, color= qc)) +
  geom_point() +
  theme_minimal()
```

Calibration weight

Tests `r dplyr::filter(grav_merged, wgt_cal_delta > 10 | wgt_cal_delta < -10)$id` have large differences (> 10 ug) in the calibration weight values.

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(grav_merged, aes(id, wgt_cal_delta)) +
  geom_point() +
  theme_minimal() +
  ylab("Cal weight delta (mg)") +
  theme(text = element_text(size=8), axis.text.x = element_text(angle=-45, hjust=1))
```

Blank weights

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(grav_merged, aes(id, wgt_blank_delta)) +
  geom_point() +
  theme_minimal() +
  ylab("Blanks weight delta (mg)") +
  theme(text = element_text(size=8), axis.text.x = element_text(angle=-45, hjust=1)) 
```

## LOD

The limit of detection is calculated to be `r round(sd(grav_merged$wgt_blank_delta, na.rm = TRUE)*3, 2)` ug.

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(grav_merged, file = "../r_files/grav_merged.Rda")
```

## Summary

$Gravimetric$ data was collected during `r length(unique(grav_merged$id))` experiments between `r min(grav_merged$date, na.rm = TRUE)` and `r max(grav_merged$date, na.rm = TRUE)`. There is no $Gravimetric$ data for tests: `r setdiff(as.character(samples$id), as.character(grav_merged$id))`.

$PAX$ data is expected to be missing for: 17A only. Filter ruptured so could not be post-weighed. 
