---
title: "gravimetric"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(forcats)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```


# Load data

* gravimetric

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/grav.Rda")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
  load("../r_files/filter_times.Rda")
  load("../r_files/filter_flows.Rda")
  load("../r_files/notes.Rda")
```

# Organize

## filter tests

* select stove and background tests only. Remove pre-calculated LOD.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::filter(grav, grepl("^[0-9]|^G[0-9]", id)) %>%
                 dplyr::select(-lod) %>%
                 dplyr::mutate(type = ifelse(grepl("^[0-9]", id) == TRUE, "test", "bg"))
```

## set levels

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  levels_type <- c("test", "bg")
```

# Metadata

* match cassette flows to id

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 flows <- dplyr::select(filter_flows, -date) %>%
           dplyr::filter(colour == "white") %>%
           dplyr::group_by(id, type, colour) %>%
           dplyr::summarise(flow = mean(value, na.rm = TRUE)) %>%
           dplyr::group_by(id, colour) %>%
           dplyr::summarise(flow = mean(flow, na.rm = TRUE)) %>%
           dplyr::select(-colour)
```

* match cassette timestamps to id

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 times <- dplyr::select(filter_times, -date) %>%
           dplyr::filter(color == "white") %>%
           tidyr::spread(type, value) %>%
           dplyr::select(-color)
```

* merge gravimetric with metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::left_join(grav_merged,
                                  flows,
                                  by = "id") %>%
                 dplyr::left_join(times,
                                  by = "id")
```

# Reformat

* pre-average

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_pre <- dplyr::select(grav_merged, wgt_pre_1, wgt_pre_2, wgt_pre_3, id) %>%
              tidyr::gather("var", "val", wgt_pre_1, wgt_pre_2, wgt_pre_3) %>%
              dplyr::filter(!is.na(val)) %>%
              dplyr::group_by(id) %>%
              dplyr::summarise(wgt_pre = mean(val))
```

* post-average

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_post <- dplyr::select(grav_merged, wgt_post_1, wgt_post_2, wgt_post_3, id) %>%
               tidyr::gather("var", "val", wgt_post_1, wgt_post_2, wgt_post_3) %>%
               dplyr::filter(!is.na(val)) %>%
               dplyr::group_by(id) %>%
               dplyr::summarise(wgt_post = mean(val))
```

* merge averages

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::left_join(grav_merged, grav_pre, by = "id") %>%
                 dplyr::left_join(grav_post, by = "id")
```

* mass added

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::mutate(grav_merged, 
                            wgt_delta = wgt_post - wgt_pre)
```

* change in calibration weight

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::mutate(grav_merged,
                            wgt_cal_delta = wgt_cal_post - wgt_cal_pre)
```

* change in blank weight

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::mutate(grav_merged,
                               wgt_blank_delta = wgt_blank_avg_post - wgt_blank_avg_pre)
```

# QC

* extract notes for gravimetric data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("grav|all", notes$inst) == TRUE)
```

* mark `bad` tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc[33] <- "bad"
  notes$qc[34] <- "bad"
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

* merge

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::left_join(grav_merged, flags, by = "id") %>%
                 dplyr::mutate(id = as.factor(id)) %>%
                 dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

Additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged$qc[grav_merged$id == "28A"] <- "bad"
  grav_merged$qc[grav_merged$id == "28B"] <- "bad"
  grav_merged$qc[grav_merged$id == "29A"] <- "bad"
```

# LOD/LOQ

* remove backgrounds and bad data then calculate the limit of detection and quantification

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  signal <- dplyr::filter(grav_merged, type == "test", qc != "bad") %>%
            dplyr::summarise(lod = sd(wgt_blank_delta) * 3 - mean(wgt_blank_delta),
                             loq = sd(wgt_blank_delta) * 10 - mean(wgt_blank_delta))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  lod_var <- signal$lod[1]
  loq_var <- signal$loq[1]
```

The LOD is `r lod_var`, the LOQ is `r loq_var`.

* flag the measurements

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::mutate(grav_merged,
                               lod = ifelse(wgt_delta >= lod_var, "ok", "low"),
                               loq = ifelse(wgt_delta >= loq_var, "ok", "low"))
```

# Background analysis

* extract background data
* remove missing data
* calculate average concentration emitted ( and other stats)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg <- dplyr::filter(grav_merged, type == "bg", qc == "ok") %>%
        dplyr::select(id, flow,
                         start, end,
                         wgt_delta) %>%
        na.omit() %>%
        dplyr::mutate(dur = (end - start) / 60,
                      conc = wgt_delta * 1000 / (flow * dur)) %>%
        dplyr::summarise(bg_mean = mean(conc),
                         sd = sd(conc),
                         min = min(conc),
                         max = max(conc),
                         n = n()) %>%
        dplyr::mutate(type = factor("test", levels_type))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  knitr::kable(bg, "markdown", digits = 2)
```

* add background to main dataframe

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav_merged <- dplyr::left_join(grav_merged,
                                  dplyr::select(bg, type, bg_mean),
                                                by = "type")
```

# Summary

Gravimetric data was collected during `r length(unique(grav_merged$id))` experiments between `r min(grav_merged$date, na.rm = TRUE)` and `r max(grav_merged$date, na.rm = TRUE)`. There is no $Gravimetric$ data for tests: `r setdiff(as.character(samples$id), as.character(grav_merged$id))`.

Gravimetric data is expected to be missing for: 17A only. Filter ruptured so could not be post-weighed. 

# Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(grav_merged, file = "../r_files/grav_merged.Rda")
```

# Plots

## weights

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(dplyr::filter(grav_merged, grepl("^[0-9][A-Z]", id)==TRUE),
         aes(id, wgt_delta, color= qc)) +
  geom_point() +
  theme_minimal()
  
  ggplot(dplyr::filter(grav_merged, grepl("^[1][0-9][A-Z]", id)==TRUE),
         aes(id, wgt_delta, color= qc)) +
  geom_point() +
  theme_minimal()
  
  ggplot(dplyr::filter(grav_merged, grepl("^[2-3][0-9][A-Z]", id)==TRUE),
         aes(id, wgt_delta, color= qc)) +
  geom_point() +
  theme_minimal()
  
  ggplot(dplyr::filter(grav_merged, grepl("^[G]", id)==TRUE),
         aes(id, wgt_delta, color= qc)) +
  geom_point() +
  theme_minimal()
```

## calibration

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(grav_merged, aes(id, wgt_cal_delta)) +
  geom_point() +
  theme_minimal() +
  ylab("Cal weight delta (mg)") +
  theme(text = element_text(size=8), axis.text.x = element_text(angle=-45, hjust=1))
```

## blanks

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(grav_merged, aes(id, wgt_blank_delta)) +
  geom_point() +
  theme_minimal() +
  ylab("Blanks weight delta (mg)") +
  theme(text = element_text(size=8), axis.text.x = element_text(angle=-45, hjust=1)) 
```
