---
title: "gravimetric"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(forcats)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

# Load data

* gravimetric

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- readRDS("../r_files/grav.RDS")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  test_info <- readRDS("../r_files/test_info.RDS")
  times_fuel <- readRDS("../r_files/times_fuel.RDS")
  times_bg <- readRDS("../r_files/times_bg.RDS")
  times_test <- readRDS("../r_files/times_test.RDS")
  flow <- readRDS("../r_files/flows.RDS")
  notes <- readRDS("../r_files/notes.RDS")
```

# Organize

* make longer

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::filter(grav, !is.na(id)) %>%
          tidyr::gather("var", "val", matches("^wgt.*")) %>%
          dplyr::filter(grepl(".*avg.*|.*dif.*", var) == FALSE) %>%
          tidyr::separate("var", c("type", "when", "rep")) %>%
          dplyr::mutate(type = ifelse(when == "cal", "cal", "sample"),
                        when = ifelse(when == "cal", rep, when),
                        rep = ifelse(rep == "pre" | rep == "post",
                                     "1", rep))
```

* average

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::group_by(grav, id, type, 
                           when, date_pre, date_post,
                           toc_pre, rh_pre,
                           toc_post, rh_post,
                           id_filter,
                           id_grav,
                           id_blank) %>%
           dplyr::summarise(mean = mean(val, na.rm = TRUE)) %>%
           dplyr::ungroup() %>%
           tidyr::spread(when, mean) %>%
           dplyr::mutate(delta = post - pre) %>%
           dplyr::rename(id_test = id)
```

* merge with id and flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::left_join(grav, test_info)

  flow <- dplyr::filter(flow, inst == "white") %>%
          dplyr::select(id, mean, delta) %>%
          dplyr::rename(flow_delta = delta,
                         flow_mean = mean)
  
  grav <- left_join(grav, flow, by = "id")
```

* add pollutant variable `pol`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, pol = "grav")
```

## Calibration weight

Spread data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  cal <- dplyr::filter(grav, type == "cal") %>%
         dplyr::select(id, delta) %>%
         dplyr::rename(cal_delta = delta)

  grav <- dplyr::filter(grav, type == "sample") %>%
          dplyr::left_join(cal)
```

## Test times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 grav <- dplyr::left_join(grav,
                           dplyr::select(times_test, id, start, end))
```

## Test types

Test are either a sample: `sample`, a background experiment: `bg`, or a filter blank: `bk`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav,
                        type = ifelse(grepl("^BG.*", id_test),
                        "bg", type),
                        type = ifelse(grepl("^B[0-9].*", id_test),
                                      "bk",
                                      type))
```

# QC

* extract notes for gravimetric data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("grav|all", notes$notes) == TRUE)
```

The LOD is `r bk$lod[1]`, the LOQ is `r bk$loq[1]`.

## Conc calculation

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # (ug - ug) * 60 * 1000 / ((L / min) * (s - s)) = ug / m^3
  grav <- dplyr::mutate(grav, 
                        conc = (delta - cal_delta) * 60 * 1000 /
                               (flow_mean * (end - start)))
```

# Blanks (LOD/LOQ)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bk <- dplyr::filter(grav, type == "bk") %>%
        dplyr::summarise(lod = 3 * sd(delta, na.rm = TRUE),
                         loq = 10 * sd(delta, na.rm = TRUE),
                         mean_bk = mean(delta, na.rm = TRUE))
```

# Background

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg <- dplyr::filter(grav, type == "bg") %>%
        dplyr::summarise(bg = mean(conc, na.rm = TRUE))
```
# Summary

Gravimetric data was collected during `r length(unique(grav$id))` experiments between `r min(grav$date, na.rm = TRUE)` and `r max(grav$date, na.rm = TRUE)`. There is no Gravimetric data for tests: `r setdiff(as.character(test_info$id), as.character(grav$id))`.

# Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  saveRDS(grav, file = "../r_files/grav_merged.RDS")
```

# Plots

## weights

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(dplyr::filter(grav, type == "sample"),
         aes(id, conc)) +
  geom_point() +
  theme_minimal()
```

