---
title: "filter extraction"
author: "Nicholas Good"
date: "11/20/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(tibble)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("r_scripts/R_load_data.R")
  source("r_scripts/R_load_metadata.R")
  source("r_scripts/R_tidy.R")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("r_files/samples.Rda")    # sample info
```

## Load ions and organics

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("r_files/ions.Rda")    # ions and organics 
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(ions)
```

## Merge with filter meta data

Load metadata
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("r_files/data_1.Rda")    # metadata
```

Extract ions filter cartridge data
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  meta <- dplyr::select(data_1, matches("_green_|^id$|^date$"))

  flows <- dplyr::select(meta, matches("flow_|^id$")) %>%
            tidyr::gather("var", "value", -id) %>%
            dplyr::filter(grepl("_avg$", var)==FALSE) %>%
            dplyr::mutate(type = sub("flow.*", "", var),
                          rep = as.factor(gsub("[^0-9]", "", var))) %>%
            dplyr::select(-var)  %>%
            dplyr::filter(!is.na(id)) %>%
            dplyr::group_by(id, type) %>%
            dplyr::summarise(mean = mean(value, na.rm = TRUE))  %>%
            tidyr::spread(type, mean)
  
  times <- dplyr::select(meta, matches("^time_|^id$|^date$")) %>%
            tidyr::gather("var", "value", -id, -date) %>%
            dplyr::mutate(type = ifelse(grepl("_start_", var), "start", "NA"),
                          type = ifelse(grepl("_end_", var), "end", type)) %>%
            dplyr::arrange(id) %>%
            dplyr::filter(!is.na(id)) %>%
            dplyr::select(-var, -date) %>%
            tidyr::spread(type, value)
```

Merge ion with filter metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions_merged <- dplyr::left_join(ions, flows, by = "id")  %>%  
                  dplyr::left_join(times, by = "id") %>%  
                  dplyr::filter(grepl("^[0-9]", id)==TRUE | grepl("^G[0-9]", id)==TRUE) %>%
                  dplyr::rename(pre_flow = pre, post_flow = post) %>%
                  dplyr::rename(time_start = start, time_end = end)

  ions_merged$id <- as.factor(ions_merged$id)
```

## QC

Load notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("r_files/notes.Rda")    # notes
  notes <- dplyr::filter(notes, grepl("^ions$|^all$", notes$inst) == TRUE)
```

Flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc <- "bad"
  notes$qc[1:9] <- "maybe"
  notes$qc[15:16] <- "maybe"
  notes$qc[23:27] <- "maybe"
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions_merged <- dplyr::left_join(ions_merged, select(notes, id, qc), by = "id")
  ions_merged$id <- as.factor(ions_merged$id)
  ions_merged$qc <- ifelse(is.na(ions_merged$qc), "ok", as.character(ions_merged$qc))
  ions_merged$qc <- as.factor(ions_merged$qc)
```

Addtional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # ions_merged$qc[ions_merged$id == ""] <- "bad"
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(ions_merged, file="r_files/ions_merged.Rda")
```

## Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ions_p <- tidyr::gather(ions_merged, var, val, 
                   -id, -id_ions, -post_flow,
                   -pre_flow, -time_start, -time_end, -type, -qc) %>%
            dplyr::filter(!is.na(val))

  ions_p$var <- as.factor(ions_p$var)

  ggplot(ions_p, aes(id, val, group = var, color = var)) +
    geom_point() +
    geom_line() +
    scale_y_log10() +
    theme_minimal() +
    ylab("ug") +
    theme(legend.position="top")
```

## Plot QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ions_p <- tidyr::gather(ions_merged, var, val, 
                   -id, -id_ions, -post_flow,
                   -pre_flow, -time_start, -time_end, -type, -qc) %>%
            dplyr::filter(!is.na(val))

  ions_p$var <- as.factor(ions_p$var)

  ggplot(ions_p, aes(id, val, group = var, color = qc)) +
    geom_point() +
    geom_line() +
    scale_y_log10() +
    theme_minimal() +
    ylab("ug") +
    theme(legend.position="top")
```

## Summary

$Ion$ data was collected during `r length(unique(ions_merged$id))` experiments. There is no $ion$ data for tests: `r setdiff(levels(samples$id), levels(ions_merged$id))`.


