---
title: "Ions"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(knitr)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_load_data.R")
  source("../r_scripts/R_load_metadata.R")
  source("../r_scripts/R_tidy.R")
```

# Load data

* load ions

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/ions.Rda")
```

* metadata

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
  load("../r_files/filter_times.Rda")
  load("../r_files/filter_flows.Rda")
  load("../r_files/notes.Rda")
```

# Organize

## extract ions data

* separate from carbonyls in same file
* remove G8B (breakthrough test)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions_merged <- dplyr::select(ions, 1:23, type, id) %>%
                 dplyr::filter(type == "test" | type == "bg", id != "G8B")
```

## flows and times

* match cassette flows to id

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flows <- dplyr::select(filter_flows, -date) %>%
           dplyr::filter(colour == "green") %>%
           dplyr::group_by(id, type, colour) %>%
           dplyr::summarise(flow = mean(value, na.rm = TRUE)) %>%
           dplyr::group_by(id, colour) %>%
           dplyr::summarise(flow = mean(flow, na.rm = TRUE)) %>%
           dplyr::select(-colour)
```

* match cassette timestamps to id

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times <- dplyr::select(filter_times, -date) %>%
           dplyr::filter(color == "green") %>%
           tidyr::spread(type, value) %>%
           dplyr::select(-color)
```

* merge ions with metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions_merged <- dplyr::left_join(ions_merged,
                                  flows,
                                  by = "id") %>%
                 dplyr::left_join(times,
                                  by = "id")
```

# QC

* extract notes for ion data 

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("ions|all", inst) == TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  knitr::kable(dplyr::filter(notes, grepl(".*ions.*", inst)) %>% 
               dplyr::select(-inst, -date),
               "markdown", digits = 2)
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           group_by(id) %>%
           arrange(qc) %>%
           summarise(qc = first(qc))
```

* merge

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ions_merged <- dplyr::left_join(ions_merged, flags, by = "id") %>%
                 dplyr::mutate(id = as.factor(id)) %>%
                 dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

* additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # ions_merged$qc[ions_merged$id == ""] <- "bad"
```

# Reformat

* convert to longer format

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ions_merged <- tidyr::gather(ions_merged, "pol", "mass_ug",
                               -id, -id_ions,
                               -flow, -start, -end,
                               -type, -qc) %>%
                 dplyr::filter(!is.na(mass_ug)) %>%
                 dplyr::mutate(pol = as.factor(pol))
```

# Background analysis

* extract background data
* remove missing data
* calculate average concentration emitted ( and other stats)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg <- dplyr::filter(ions_merged, type == "bg", qc == "ok") %>%
        dplyr::mutate(mass_ug = ifelse(is.na(mass_ug), 0, mass_ug)) %>%
        na.omit() %>%
        dplyr::select(-id_ions) %>%
        dplyr::mutate(dur = end - start) %>%
        dplyr::mutate(conc = (mass_ug * 1000 * 60) / (flow * dur)) %>%  # ug * 1000 * 60 / (L/min * s) = ug / m^3
        dplyr::group_by(pol) %>%
        dplyr::summarise(mean = mean(conc),
                         sd = sd(conc),
                         min = min(conc),
                         max = max(conc),
                         n = n())
```

* background concentration ($\mu g/m^3$)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  knitr::kable(bg, "markdown", digits = 2)
```

* merge with test data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg <- dplyr::select(bg, pol, mean) %>%
        dplyr::rename(conc_bg = mean)

  ions_merged <- dplyr::left_join(ions_merged, bg, by = "pol") %>%
                 dplyr::mutate(conc_bg = ifelse(is.na(conc_bg), 0, conc_bg))
```

# Summary

Ion data was collected during `r length(unique(ions_merged$id))` experiments. There is no ion data for tests: `r setdiff(as.character(samples$id), as.character(ions_merged$id))`.

# Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(ions_merged, file = "../r_files/ions_merged.Rda")
```

# Plots

## mass

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=12}
  ggplot(ions_merged, aes(id, mass_ug, group = pol, color = pol)) +
         geom_point() +
         geom_line() +
         scale_y_log10() +
         theme_minimal() +
         ylab("ug") +
         theme(legend.position="top") +
         theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
         facet_wrap(~type, ncol = 1, scales = "free")
```

## qc

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(ions_merged, aes(id, mass_ug, group = pol, color = qc)) +
         geom_point() +
         geom_line() +
         scale_y_log10() +
         theme_minimal() +
         ylab("ug") +
         theme(legend.position="top") +
         theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
         facet_wrap(~type, ncol = 1, scales = "free")
```
