---
title: "ISEE 2017 Presentation"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  library(tidyverse)
  library(knitr)
```

```{r global_options, include=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/', fig.height = 6, fig.width = 8,
                        warning = FALSE, message = FALSE, cache = FALSE)
```

# tidy data

## load data

```{r}
  emissions <- readRDS("../r_files/emissions_factors.RDS")
  load("../r_files/samples.Rda")
```

## load functions

```{r load_functions}
  source("../r_scripts/R_plots.R")
  source("../r_scripts/R_functions.R")
```

## select metric for plotting

```{r}
  emissions_metric <- "energy_fuel_ef_comb"
  y_label <- "mg emitted per MJ of fuel"
```

```{r}
  # remove extra gasifier tests
  emissions %<>%
    dplyr::filter(metric == emissions_metric, !grepl("27|28|29|30", id)) %>%
    tidyr::spread(metric, value)
```

## clean up sample info

* id to character type

```{r}
  samples <-
    samples %>%
    dplyr::mutate(id = as.character(id))
```

* format fuel information

```{r}
  samples <-
    samples %>%
    dplyr::mutate(fuel =
                    fuel %>% 
                    factor %>%
                    tolower %>% 
                    forcats::fct_recode("lpg" = "liquefied petroleum gas") %>%
                    forcats::fct_relevel("oak",
                                         "eucalyptus",
                                         "douglas fir",
                                         "lump hardwood (med.)",
                                         "lump hardwood (sm.)",
                                         "coconut charcoal",
                                         "west slope pellets",
                                         "eucalyptus pellets",
                                         "lpg",
                                         "kerosene"),
                  fuelcat = 
                    fuelcat %>% 
                    forcats::fct_recode("liquid" = "advanced"))
```

* format stove information

```{r}
  samples <-
    samples %>%
    dplyr::mutate(stove = 
                    stove %>% 
                    factor %>% 
                    tolower %>%
                    gsub(" [:(:].*[:):]", "",.) %>%
                    forcats::fct_relevel("three stone fire",
                                         "clay ring",
                                         "ceramic charcoal",
                                         "rocket elbow",
                                         "built-in justa",
                                         "metal charcoal",
                                         "fan rocket elbow",
                                         "gasifier",
                                         "wick kerosene",
                                         "pressurized kerosene",
                                         "pressurized lpg"),
                  stovecat = 
                    stovecat %>%
                    forcats::fct_recode("insulated" = "improved",
                                        "insulated" = "gasifiers",
                                        "liquid" = "advanced"))
```

## extract specific emissions data

* ions 

```{r}
  # sum ion data
  ions <-
    emissions %>% 
    dplyr::filter(
      grepl("ammonium|chloride|potassium|sodium|nitrite|calcium|magnesium|
            |chloride|nitrate|sulfate", pol)) %>%
    dplyr::group_by(id, qc) %>%
    dplyr::summarise_if(is.numeric, sum) %>%
    dplyr::mutate(pol = "ions")
```

* all aerosols 

```{r}
  # create aerosol variable
  aerosols <-
    emissions %>% 
    dplyr::filter(grepl("ecoc", inst)) %>%
    dplyr::filter(pol != "tc") %>%
    dplyr::select(-inst) %>%
    dplyr::bind_rows(ions)
```

* btex

```{r}
  # sum btex data
  btex <-
    emissions %>% 
    dplyr::filter(grepl("^voc_benzene$|^voc_toluene$|
                        |^voc_ethylbenzene$|xylene", pol)) %>%
    dplyr::group_by(id) %>%
    dplyr::summarise_if(is.numeric, sum) %>%
    dplyr::mutate(pol = "btex")
```

* vocs

```{r}
  vocs <-
    emissions %>% 
    dplyr::filter(inst == "voc") %>%
    dplyr::group_by(id) %>%
    dplyr::summarise_if(is.numeric, sum) %>%
    dplyr::mutate(pol = "vocs")
```

* carbonyls 

```{r}
 carbs <- 
   emissions %>%
   dplyr::filter(inst == "carbs") %>%
   dplyr::group_by(id) %>%
   dplyr::summarise_if(is.numeric, sum) %>%
   dplyr::mutate(pol = "carbonyls")
  
```

* all gases

```{r}
  gases <-
    emissions %>%
    dplyr::filter(inst == "fivegas", grepl("^co$|^ch4$", pol)) %>%
    dplyr::select(-inst, -qc) %>%
    dplyr::bind_rows(carbs, vocs)
```

* carcinogenic compounds

```{r}
  carcinogens <-
    emissions %>%
    dplyr::filter(grepl("^voc_benzene$|formaldehyde|acetaldehyde", pol)) %>%
    dplyr::mutate(inst = "carcinogens")
```

# carcinogen bar chart

## tidy carcinogen categories

```{r carcinogens_tidy}
  carcinogens <- 
    carcinogens %>% 
    dplyr::mutate(pol = forcats::fct_recode(pol, "benzene" = "voc_benzene"))
```

## group carcinogenic pollutants by stove type plot

```{r carcinogens_stoves}
  isee_summarize(carcinogens) %>% 
    isee_bar_charts(y_lab = y_label)
```

# aerosol composition bar charts

## tidy aerosol categories

```{r}
  aerosols <-
    aerosols %>%
    dplyr::mutate(pol =
                    pol %>%
                    forcats::fct_recode("elemental carbon" = "ec",
                                        "organic carbon" = "oc",
                                        "ions" = "ions") %>%
                    forcats::fct_relevel("organic carbon",
                                         "elemental carbon",
                                         "ions"))
```

## summarize by stove type

```{r aerosols_stove}
  isee_summarize(aerosols) %>% 
    isee_bar_charts(y_lab = y_label)
```

## summarize by fuel type

```{r aerosols_wood}
  isee_summarize(aerosols, group_var = "fuel", cat_filter = "wood") %>% 
    isee_bar_charts(y_lab = y_label, facet_var = "fuel")
```

```{r aerosols_charcoal}
  isee_summarize(aerosols, group_var = "fuel", cat_filter = "charcoal") %>% 
    isee_bar_charts(y_lab = y_label, facet_var = "fuel")
```

```{r aerosols_advanced}
  isee_summarize(aerosols, group_var = "fuel", cat_filter = "pellets|liquid") %>% 
    isee_bar_charts(y_lab = y_label, facet_var = "fuel")
```

# gas composition bar charts

## tidy gas categories

```{r gases_relevel}
  gases <- 
    gases %>% 
    dplyr::mutate(pol = 
                    pol %>% 
                    forcats::fct_relevel("carbonyls",
                                         "vocs",
                                         "ch4",
                                         "co"))
```

## summarize by stove type

```{r gases_stoves}
  isee_summarize(gases) %>% 
    isee_bar_charts(y_lab = y_label)
```

## summarize by fuel type

## summarize by fuel type

```{r gases_wood}
  isee_summarize(gases, group_var = "fuel", cat_filter = "wood") %>% 
    isee_bar_charts(y_lab = y_label, facet_var = "fuel")
```

```{r gases_charcoal}
  isee_summarize(gases, group_var = "fuel", cat_filter = "charcoal") %>% 
    isee_bar_charts(y_lab = y_label, facet_var = "fuel")
```

```{r gases_advanced}
  isee_summarize(gases, group_var = "fuel", cat_filter = "pellets|liquid") %>% 
    isee_bar_charts(y_lab = y_label, facet_var = "fuel")
```