---
title: "ISEE 2017 Presentation"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  library(tidyverse)
  library(knitr)
```

```{r global_options, include=FALSE}
  knitr::opts_chunk$set(fig.path='figures/', fig.height = 4, fig.width = 10, warning=FALSE, message=FALSE, cache=FALSE)
  cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#CC79A7")
```

# tidy data

## load data

```{r}
  emissions <- readRDS("../r_files/emissions_factors.RDS")
  load("../r_files/samples.Rda")
```

## load functions

```{r load_functions}
  source("../r_scripts/R_plots.R")
```

## select metric

```{r}
 emissions_metric <- "energy_ef"
```

```{r}
  emissions %<>%
    dplyr::filter(metric == emissions_metric, !grepl("27|28|29|30", id)) %>%
    dplyr::distinct() %>%
    tidyr::spread(metric, value)
```

## clean up sample info

```{r}
  samples$fuel <- factor(tolower(samples$fuel))

  samples$fuel <- forcats::fct_recode(samples$fuel, lpg = "liquefied petroleum gas")

  samples$fuel <- forcats::fct_relevel(samples$fuel, "oak",
                                                 "eucalyptus",
                                                 "douglas fir",
                                                 "lump hardwood (med.)",
                                                 "lump hardwood (sm.)",
                                                 "coconut charcoal",
                                                 "west slope pellets",
                                                 "eucalyptus pellets",
                                                 "lpg",
                                                 "kerosene")
```

```{r}
  samples$stove <- factor(tolower(samples$stove))
  
  samples <- dplyr::mutate(samples, stove = gsub(" [:(:].*[:):]", "", stove))
  
  samples$stove <- forcats::fct_relevel(samples$stove,
                                           "three stone fire",
                                           "clay ring",
                                           "ceramic charcoal",
                                           "rocket elbow",
                                           "built-in justa",
                                           "metal charcoal",
                                           "fan rocket elbow",
                                           "gasifier",
                                           "wick kerosene",
                                           "pressurized kerosene",
                                           "pressurized lpg")

  samples$stovecat <- forcats::fct_recode(samples$stovecat, insulated = "improved",
                                                            insulated = "gasifiers",
                                                            liquid = "advanced")
  
  samples$fuelcat <- forcats::fct_recode(samples$fuelcat, liquid = "advanced")
```

## extract specific emissions data

* pm and co

```{r}
  pm_co <-
    emissions %>% 
    dplyr::filter(grepl("^grav$|^co$", pol)) %>%
    dplyr::select(-inst, -qc)

  pm_co$pol <- forcats::fct_recode(pm_co$pol, "pm2.5" = "grav")

  pm_co$pol <- forcats::fct_relevel(pm_co$pol, "pm2.5", "co")
  
   pm_co <-
     pm_co %>% tidyr::spread_("pol", emissions_metric)
```

* ions 

```{r}
  # sum ion data
  ions <-
    emissions %>% 
    dplyr::filter(grepl("ammonium|chloride|potassium|sodium|nitrite|calcium|magnesium|chloride|nitrate|sulfate", pol)) %>%
    dplyr::group_by(id, qc) %>%
    dplyr::summarise_if(is.numeric, sum) %>%
    dplyr::mutate(pol = "ions")
```

* all aerosols 

```{r}
  # create aerosol variable
  aerosols <-
    emissions %>% 
    dplyr::filter(grepl("ecoc", inst)) %>%
    dplyr::filter(pol != "tc") %>%
    dplyr::select(-inst) %>%
    dplyr::bind_rows(ions)
```

* btex

```{r}
  # sum ion data
  btex <-
    emissions %>% 
    dplyr::filter(grepl("benzene|toluene|ethylbenzene|xylene", pol)) %>%
    dplyr::group_by(id) %>%
    dplyr::summarise_if(is.numeric, sum) %>%
    dplyr::mutate(pol = "btex")
```

* vocs

```{r}
  vocs <-
    emissions %>% 
    dplyr::filter(inst == "voc") %>%
    dplyr::filter(!grepl("benzene|toluene|ethylbenzene|xylene", pol)) %>%
    dplyr::group_by(id) %>%
    dplyr::summarise_if(is.numeric, sum) %>%
    dplyr::mutate(pol = "vocs (no btex)") %>%
    dplyr::bind_rows(btex)
```

* carbonyls 

```{r}
 carbs <- 
   emissions %>%
   dplyr::filter(inst == "carbs") %>%
   dplyr::group_by(id) %>%
   dplyr::summarise_if(is.numeric, sum) %>%
   dplyr::mutate(pol = "carbonyls")
  
```

* all gases

```{r}
  gases <-
    emissions %>%
    dplyr::filter(inst == "fivegas", grepl("^co$|^ch4$", pol)) %>%
    dplyr::select(-inst, -qc) %>%
    dplyr::bind_rows(carbs, vocs)
```

* carcinogenic compounds

```{r}

```

# aerosol composition bar charts

## tidy aerosol categories

```{r}
  aerosols$pol <- forcats::fct_recode(aerosols$pol, "elemental carbon" = "ec",
                                      "organic carbon" = "oc",
                                      "ions" = "ions")
  aerosols$pol <- forcats::fct_relevel(aerosols$pol, "organic carbon", "elemental carbon", "ions")
  
  aerosols$pol <- forcats::fct_relevel(aerosols$pol,
                                       "ions",
                                       "elemental carbon",
                                       "organic carbon")
```

## summarize by stove type

```{r}
  # plot ec + oc vs. grav
  aerosols_p <-
    aerosols %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>% # only keep test cases with all the the data
    #dplyr::mutate(unknown = grav - ec - oc) %>%
    #dplyr::select(-grav) %>%
    dplyr::left_join(
    dplyr::select(samples, id, stove), by = "id")
    #dplyr::mutate_if(is.numeric, funs(. / 1000))
```

```{r}
  aerosols_p <- 
    aerosols_p %>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, funs(mean, min, max), na.rm = TRUE) %>%
    tidyr::gather("pol", "value", 2:4) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, stovecat)), by = "stove")
```

```{r aerosols_stoves}
  ggplot(aerosols_p, aes(x = stove, y = value, fill = pol)) +
    geom_bar(stat = "identity") +
    #geom_errorbar(aes(ymin = min, ymax = max)) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 15, vjust = 0.6, hjust = 0.5),
          text = element_text(size=15),
          legend.title = element_blank()) +
    scale_fill_manual(values = cbPalette) +
    facet_grid(~stovecat, scales="free", space="free_x") +
    ylab("mg emitted per MJ fuel") +
    xlab("")
```

## summarize by fuel type

```{r}
  # plot ec + oc vs. grav
  aerosols_p <-
    aerosols %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>% # only keep test cases with all the the data
    #dplyr::mutate(unknown = grav - ec - oc) %>%
    #dplyr::select(-grav) %>%
    dplyr::left_join(
    dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(fuelcat == "wood")
    #dplyr::mutate_if(is.numeric, funs(. / 1000))
```

```{r}
  aerosols_p %<>%
    dplyr::group_by(stove, fuel) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 3:5)
```

```{r aerosols_wood}
  ggplot(aerosols_p, aes(x = stove, y = value, fill = pol)) +
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 15, vjust = 0.5, hjust = 0.5),
          text = element_text(size=15),
          legend.title = element_blank()) +
    scale_fill_manual(values=cbPalette) +
    facet_grid(~fuel, scales = "free", space = "free_x") +
    ylab("mg emitted per MJ fuel") +
    xlab("")
```

```{r}
  # plot ec + oc vs. grav
  aerosols_p <-
    aerosols %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>% # only keep test cases with all the the data
    #dplyr::select(-grav) %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(fuelcat == "charcoal")
```

```{r}
  aerosols_p %<>%
    dplyr::group_by(stove, fuel) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 3:5)
```

```{r aerosols_charcoal}
  ggplot(aerosols_p, aes(x = stove, y = value, fill = pol)) +
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 15, vjust = 0.5),
          text = element_text(size=15),
          legend.title = element_blank()) +
      scale_fill_manual(values=cbPalette) +
    facet_grid(~fuel, scales = "free", space = "free_x") +
    ylab("mg emitted per MJ fuel") +
    xlab("")
```

```{r}
  # plot ec + oc vs. grav
  aerosols_p <-
    aerosols %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>% # only keep test cases with all the the data
    #dplyr::select(-grav) %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(grepl("pellets|liquid", fuelcat))
```

```{r}
  aerosols_p %<>%
    dplyr::group_by(stove, fuel) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 3:5)
```

```{r aerosols_advanced}
  ggplot(aerosols_p, aes(x = stove, y = value, fill = pol)) +
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 15, vjust = 0.5),
          text = element_text(size=15),
          legend.title = element_blank()) +
      scale_fill_manual(values=cbPalette) +
    facet_grid(~fuel, scales = "free", space = "free_x") +
    ylab("mg emitted per MJ fuel") +
    xlab("")
```

# voc composition bar charts

## tidy gas categories

```{r}
  gases_p$pol <- forcats::fct_relevel(gases_p$pol,
                                      "carbonyls",
                                      "btex",
                                      "vocs (no btex)",
                                      "ch4",
                                      "co")
```

## summarize by stove type

```{r}
  gases_p <-
    gases %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>%
    dplyr::left_join(
    dplyr::select(samples, id, stove), by = "id")
```

```{r}
  gases_p %<>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 2:6) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, stovecat)), by = "stove") %>%
    dplyr::mutate(pol = as.factor(pol))
```

```{r gases_stoves}
  ggplot(gases_p, aes(x = stove, y = value, fill = pol)) +
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 15, vjust = 0.5),
          text = element_text(size=15),
          legend.title = element_blank()) +
    facet_grid(~stovecat, scales = "free", space = "free_x") +
      scale_fill_manual(values=cbPalette) +
    ylab("mg emitted per MJ fuel") +
    xlab("")
```

## summarize by fuel type

```{r}
  gases_p <-
    gases %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(fuelcat == "wood")
```

```{r}
  gases_p %<>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 2:6) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, fuel)), by = "stove")
```

```{r gases_wood}
  ggplot(gases_p, aes(x = stove, y = value, fill = pol)) +
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 15, vjust = 0.5),
          text = element_text(size=15),
          legend.title = element_blank()) +
    facet_grid(~fuel, scales = "free", space = "free_x") +
      scale_fill_manual(values=cbPalette) +
    ylab("mg emitted per MJ") +
    xlab("")
```

```{r}
  gases_p <-
    gases %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(fuelcat == "charcoal")
```

```{r}
  gases_p %<>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 2:6) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, fuel)), by = "stove")
```

```{r gases_charcoal}
  ggplot(gases_p, aes(x = stove, y = value, fill = pol)) +
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 15, vjust = 0.5),
          text = element_text(size=15),
          legend.title = element_blank()) +
    facet_grid(~fuel, scales = "free", space = "free_x") +
      scale_fill_manual(values=cbPalette) +
    ylab("mg emitted per MJ fuel") +
    xlab("")
```

```{r}
  gases_p <-
    gases %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(grepl("pellets|liquid", fuelcat))
```

```{r}
  gases_p %<>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 2:6) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, fuel)), by = "stove")
```

```{r gases_advanced}
  ggplot(gases_p, aes(x = stove, y = value, fill = pol)) +
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 15, vjust = 0.5),
          text = element_text(size=15),
          legend.title = element_blank()) +
    facet_grid(~fuel, scales = "free", space = "free_x") +
      scale_fill_manual(values=cbPalette) +
    ylab("mg emitted per MJ fuel") +
    xlab("")
```