---
title: "ISEE 2017 Presentation"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  library(tidyverse)
  library(knitr)
```

```{r global_options, include=FALSE}
  knitr::opts_chunk$set(fig.path='figures/', fig.height = 6, fig.width = 8,
                        warning=FALSE, message=FALSE, cache=FALSE)
```

# tidy data

## load data

```{r}
  emissions <- readRDS("../r_files/emissions_factors.RDS")
  load("../r_files/samples.Rda")
```

## load functions

```{r load_functions}
  source("../r_scripts/R_plots.R")
```

## select metric for plotting

```{r}
 emissions_metric <- "energy_fuel_ef_comb"
 y_label <- "mg emitted per MJ of fuel"
```

```{r}
  emissions %<>%
    dplyr::filter(metric == emissions_metric, !grepl("27|28|29|30", id)) %>%
    dplyr::distinct() %>%
    tidyr::spread(metric, value)
```

## clean up sample info

* format fuel information

```{r}
  samples <-
    samples %>%
    dplyr::mutate(fuel = factor(tolower(fuel)),
                  fuel = forcats::fct_recode(fuel,
                                             "lpg" = "liquefied petroleum gas"),
                  fuel = forcats::fct_relevel(fuel,
                                              "oak",
                                              "eucalyptus",
                                              "douglas fir",
                                              "lump hardwood (med.)",
                                              "lump hardwood (sm.)",
                                              "coconut charcoal",
                                              "west slope pellets",
                                              "eucalyptus pellets",
                                              "lpg",
                                              "kerosene"))
```

* format stove information

```{r}
  samples$stove <- factor(tolower(samples$stove))
  
  samples <- dplyr::mutate(samples, stove = gsub(" [:(:].*[:):]", "", stove))
  
  samples$stove <- forcats::fct_relevel(samples$stove,
                                           "three stone fire",
                                           "clay ring",
                                           "ceramic charcoal",
                                           "rocket elbow",
                                           "built-in justa",
                                           "metal charcoal",
                                           "fan rocket elbow",
                                           "gasifier",
                                           "wick kerosene",
                                           "pressurized kerosene",
                                           "pressurized lpg")

  samples$stovecat <- forcats::fct_recode(samples$stovecat, insulated = "improved",
                                                            insulated = "gasifiers",
                                                            liquid = "advanced")
  
  samples$fuelcat <- forcats::fct_recode(samples$fuelcat, liquid = "advanced")
```

## extract specific emissions data

* ions 

```{r}
  # sum ion data
  ions <-
    emissions %>% 
    dplyr::filter(
      grepl("ammonium|chloride|potassium|sodium|nitrite|calcium|magnesium|
            |chloride|nitrate|sulfate", pol)) %>%
    dplyr::group_by(id, qc) %>%
    dplyr::summarise_if(is.numeric, sum) %>%
    dplyr::mutate(pol = "ions")
```

* all aerosols 

```{r}
  # create aerosol variable
  aerosols <-
    emissions %>% 
    dplyr::filter(grepl("ecoc", inst)) %>%
    dplyr::filter(pol != "tc") %>%
    dplyr::select(-inst) %>%
    dplyr::bind_rows(ions)
```

* btex

```{r}
  # sum btex data
  btex <-
    emissions %>% 
    dplyr::filter(grepl("^voc_benzene$|^voc_toluene$|
                        |^voc_ethylbenzene$|xylene", pol)) %>%
    dplyr::group_by(id) %>%
    dplyr::summarise_if(is.numeric, sum) %>%
    dplyr::mutate(pol = "btex")
```

* vocs

```{r}
  vocs <-
    emissions %>% 
    dplyr::filter(inst == "voc") %>%
    dplyr::group_by(id) %>%
    dplyr::summarise_if(is.numeric, sum) %>%
    dplyr::mutate(pol = "vocs")
```

* carbonyls 

```{r}
 carbs <- 
   emissions %>%
   dplyr::filter(inst == "carbs") %>%
   dplyr::group_by(id) %>%
   dplyr::summarise_if(is.numeric, sum) %>%
   dplyr::mutate(pol = "carbonyls")
  
```

* all gases

```{r}
  gases <-
    emissions %>%
    dplyr::filter(inst == "fivegas", grepl("^co$|^ch4$", pol)) %>%
    dplyr::select(-inst, -qc) %>%
    dplyr::bind_rows(carbs, vocs)
```

* carcinogenic compounds

```{r}
  carcinogens <-
    emissions %>%
    dplyr::filter(grepl("^voc_benzene$|formaldehyde|acetaldehyde", pol)) %>%
    dplyr::mutate(inst = "carcinogens")
```

# carcinogen bar chart

## tidy carcinogen categories

```{r}
  carcinogens$pol <- forcats::fct_recode(carcinogens$pol, "benzene" = "voc_benzene")
```

## group carcinogenic pollutants by stove type and calculate mean emissions for each stove

```{r}
  carcinogens_p <-
    carcinogens %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>% 
    dplyr::left_join(
    dplyr::select(samples, id, stove), by = "id")
```

```{r}
  carcinogens_p <- 
    carcinogens_p %>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean, na.rm = TRUE) %>%
    tidyr::gather("pol", "value", 2:4) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, stovecat)), by = "stove")
```

## plot carcinogenic compounds

```{r carcinogens_stoves}
  # pass string variables
  isee_bar_charts(carcinogens_p, y_lab = y_label)
```

# aerosol composition bar charts

## tidy aerosol categories

```{r}
  aerosols$pol <- forcats::fct_recode(aerosols$pol, "elemental carbon" = "ec",
                                      "organic carbon" = "oc",
                                      "ions" = "ions")
  aerosols$pol <- forcats::fct_relevel(aerosols$pol, "organic carbon", "elemental carbon", "ions")
  
  aerosols$pol <- forcats::fct_relevel(aerosols$pol,
                                       "ions",
                                       "elemental carbon",
                                       "organic carbon")
```

## summarize by stove type

```{r}
  # plot ec + oc vs. grav
  aerosols_p <-
    aerosols %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>% # only keep test cases with all the the data
    #dplyr::mutate(unknown = grav - ec - oc) %>%
    #dplyr::select(-grav) %>%
    dplyr::left_join(
    dplyr::select(samples, id, stove), by = "id")
    #dplyr::mutate_if(is.numeric, funs(. / 1000))
```

```{r}
  aerosols_p <- 
    aerosols_p %>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean, na.rm = TRUE) %>%
    tidyr::gather("pol", "value", 2:4) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, stovecat)), by = "stove")
```

```{r aerosols_stoves}
  isee_bar_charts(aerosols_p, y_lab = y_label)
```

## summarize by fuel type

```{r}
  # plot ec + oc vs. grav
  aerosols_p <-
    aerosols %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>% # only keep test cases with all the the data
    #dplyr::mutate(unknown = grav - ec - oc) %>%
    #dplyr::select(-grav) %>%
    dplyr::left_join(
    dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(fuelcat == "wood")
    #dplyr::mutate_if(is.numeric, funs(. / 1000))
```

```{r}
  aerosols_p %<>%
    dplyr::group_by(stove, fuel) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 3:5)
```

```{r aerosols_wood}
  isee_bar_charts(aerosols_p, y_lab = y_label, facet_var = "fuel")
```

```{r}
  # plot ec + oc vs. grav
  aerosols_p <-
    aerosols %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>% # only keep test cases with all the the data
    #dplyr::select(-grav) %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(fuelcat == "charcoal")
```

```{r}
  aerosols_p %<>%
    dplyr::group_by(stove, fuel) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 3:5)
```

```{r aerosols_charcoal}
  isee_bar_charts(aerosols_p, y_lab = y_label, facet_var = "fuel")
```

```{r}
  # plot ec + oc vs. grav
  aerosols_p <-
    aerosols %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>% # only keep test cases with all the the data
    #dplyr::select(-grav) %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(grepl("pellets|liquid", fuelcat))
```

```{r}
  aerosols_p %<>%
    dplyr::group_by(stove, fuel) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 3:5)
```

```{r aerosols_advanced}
  isee_bar_charts(aerosols_p, y_lab = y_label, facet_var = "fuel")
```

# gas composition bar charts

## tidy gas categories

```{r}
  gases$pol <- forcats::fct_relevel(gases$pol,
                                      "carbonyls",
                                      "btex",
                                      "vocs (no btex)",
                                      "ch4",
                                      "co")
```

## summarize by stove type

```{r}
  gases_p <-
    gases %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>%
    dplyr::left_join(
    dplyr::select(samples, id, stove), by = "id")
```

```{r}
  gases_p %<>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 2:6) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, stovecat)), by = "stove") %>%
    dplyr::mutate(pol = as.factor(pol))
```

```{r gases_stoves}
  isee_bar_charts(gases_p, y_lab = y_label)
```

## summarize by fuel type

```{r}
  gases_p <-
    gases %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(fuelcat == "wood")
```

```{r}
  gases_p %<>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 2:6) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, fuel)), by = "stove")
```

```{r gases_wood}
  isee_bar_charts(gases_p, y_lab = y_label, facet_var = "fuel")
```

```{r}
  gases_p <-
    gases %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(fuelcat == "charcoal")
```

```{r}
  gases_p %<>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 2:6) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, fuel)), by = "stove")
```

```{r gases_charcoal}
  isee_bar_charts(gases_p, y_lab = y_label, facet_var = "fuel")
```

```{r}
  gases_p <-
    gases %>%
    tidyr::spread_("pol", emissions_metric) %>%
    na.omit %>%
    dplyr::left_join(
      dplyr::select(samples, id, stove, fuel, fuelcat), by = "id") %>%
    dplyr::filter(grepl("pellets|liquid", fuelcat))
```

```{r}
  gases_p %<>%
    dplyr::group_by(stove) %>%
    dplyr::summarise_if(is.numeric, mean) %>%
    tidyr::gather("pol", "value", 2:6) %>%
    dplyr::left_join(dplyr::distinct(dplyr::select(samples, stove, fuel)), by = "stove")
```

```{r gases_advanced}
  isee_bar_charts(gases_p, y_lab = y_label, facet_var = "fuel")
```