---
title: "Load and save data"
author: "Nicholas Good"
date: "1/31/2017"
output: html_document
---

## Libraries

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

## Source files

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_load_data.R")
  source("../r_scripts/R_load_metadata.R")
  source("../r_scripts/R_tidy.R")
  source("../r_scripts/R_functions.R")
```

# Pollutant data

## CO2

Loads and saves the CO2 data used for dilution calculations. The data was logged one of two two ways, as separate voltage files or a combined mixing ratio file. Three output files are created containing the room (lab) data (in volts), the diluted sample (sample) data (in volts) and the combined lab and sample data recorded as a mixing ratio (ppm)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  co2_lab <- load_multifile(fldr = "../data/co2", pattern = "DILUTION1.csv$", inst = "co2")

  save(co2_lab, file = "../r_files/co2_lab.Rda")
  
  co2_sample <- load_multifile(fldr = "../data/co2", pattern = "DILUTION2.csv$", inst = "co2")
  
  save(co2_sample, file = "../r_files/co2_sample.Rda")
    
  co2_lab_sample <- load_multifile(fldr = "../data/co2", pattern = "DILUTION.csv$", inst = "co2")

  save(co2_lab_sample, file = "../r_files/co2_lab_sample.Rda")
```

## ECOC

Loads ecoc data file and resaves as .Rda.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  ecoc <- load_singlefiles("ecoc")  # ecoc filter analysis

  save(ecoc, file = "../r_files/ecoc.Rda")
```  

## Fivegas

The fivegas analyzer recorded co, co2 and ch4. Three file types are loaded; raw voltages, pre-calibrated mixing ratios (ppm) and voltage calibrations.The data is loaded and resaved as .Rda files.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  fivegas_conc <- load_fivegas(type = "conc")

  save(fivegas_conc, file = "../r_files/fivegas_conc.Rda")

  fivegas_volts <- load_fivegas(type = "volts")

  save(fivegas_volts, file = "../r_files/fivegas_volts.Rda")

  fivegas_calibration <- load_fivegas(pattern = "_CALIBRATION_",type = "volts")

  save(fivegas_calibration, file = "../r_files/fivegas_calibration.Rda")
```

## Gravimetric

Loads filter weight data and resaves as .Rda.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- load_singlefiles("grav")

  save(grav, file = "../r_files/grav.Rda")
```

## Ions and carbonyls

Loads and resaves data as .Rda.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  ions <- load_singlefiles("ions")

  save(ions, file = "../r_files/ions.Rda")
```

## PAHS

Loads and resaves data as .Rda.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}   
  pah <- load_singlefiles("pah")  # pah data

  save(pah, file = "../r_files/pah.Rda")
```

## PAX

Loads and resaves data as .Rda.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  pax <- load_multifile(fldr = "../data/pax", pattern = "ALLDAY_PAX.csv$", inst = "pax")

  save(pax, file = "../r_files/pax.Rda")
```

## SCALE

Loads and resaves data as .Rda.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  scale <- load_multifile(fldr = "../data/scale", pattern = "_SCALE.xlsx$", inst = "scale")

  save(scale, file = "../r_files/scale.Rda")
```

## SMPS

Loads and resaves data as an .Rda.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  smps <- load_multifile(fldr = "../data/smps", pattern = "_SMPS.csv$", inst = "smps")

  save(smps, file = "../r_files/smps.Rda")
```

## TEMP

Loads and resaves data as an .Rda.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  temp <- load_multifile(fldr = "../data/temp", pattern = "_TEMP.csv|_TEMP.CSV", inst = "temp")

  save(temp, file = "../r_files/temp.Rda")
```

## TRANS

Loads and resaves data as an .Rda.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  trans <- load_singlefiles("trans")

  save(trans, file = "../r_files/trans.Rda")
```
 
## VOC

Loads the microgram sheet  and resaves data as an .Rda.
 
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  voc <- load_singlefiles("voc")

  save(voc, file = "../r_files/voc.Rda")
```

# Metadata

## Sample logs

Reference for test id, stove and fuel types.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  samples <- load_singlefiles("sample")

  samples$fuelcat <- fct_collapse(samples$fuel,
  wood = c("Oak", "Eucalyptus", "Douglas Fir"),
  charcoal = c("Coconut Charcoal", "Lump Hardwood (Med.)", 
               "Lump Hardwood (Sm.)"),
  pellets = c("Eucalyptus Pellets", "West Slope Pellets"),
  advanced = c("Kerosene", "Liquefied Petroleum Gas"),
  none = c("Background")
  )
  samples$fuelcat <- lvls_revalue(samples$fuelcat, c("wood", "charcoal", 
                                  "pellets", "advanced", "none"))

  samples$stovecat <- fct_collapse(samples$stove,
  traditional = c("Clay Ring (Artisan)", 
                  "Three Stone Fire (Artisan)", 
                  "Ceramic Charcoal (Artisan)"),
  improved = c("Fan Rocket Elbow (Biolite, HomeStove)", 
               "Metal Charcoal (Burn, Jikokoa)", 
               "Rocket Elbow (Envirofit, G3300)",
               "Built-in Justa (Envirofit, HM5000)"),
  gasifiers = c("Gasifier (Envirofit, Prototype)", 
                "Gasifier (Philips, HD4012)",
                "Gasifer (Mimi Moto, Mimi Moto)"),
  advanced = c("Pressurized LPG (WokSmith, WS-C1-25K)", 
               "Wick Kerosene (Cook 'n Lite, 82)", 
               "Pressurized Kerosene (Butterfly, 2412)"),
  none = c("Background")
  )
  samples$stovecat <- lvls_revalue(samples$stovecat, c("traditional", "improved",
                                   "gasifiers", "advanced", "none"))

  save(samples, file = "../r_files/samples.Rda")
```

## Batch sample meta data

Weights, times, notes...
 
```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}  
  batch <- load_singlefiles("batch")

  save(batch, file = "../r_files/batch.Rda")
```

## Wood sample meta data

Weights, times, notes...

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  wood <- load_singlefiles("wood")

  save(wood, file = "../r_files/wood.Rda")
```

## Calibration log 1

Fivegas calibration mixing ratios and timestamps.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  cal_1 <- load_singlefiles("cal_1")

  save(cal_1, file = "../r_files/cal_1.Rda")
```

## Calibration log 2

* CO2 calibration information
* Pax and smps flow measurements

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}  
  cal_2 <- load_singlefiles("cal_2")

  save(cal_2, file = "../r_files/cal_2.Rda")
```

## Data log 1

* Filter cartridge flows and times
* Voc times
* Fivegas backgriund times
* Notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  data_1 <- load_singlefiles("data_1")

  save(data_1, file = "../r_files/data_1.Rda")
```
 
## Data log 2

* Carbonyls and isokinetic sampler times and flows
* Smps and pax times (including background)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE} 
  data_2 <- load_singlefiles("data_2")

  save(data_2, file = "../r_files/data_2.Rda")
```
