---
title: "mass_emissions"
author: "Nicholas Good"
date: "1/6/2017"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

## Test time data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/test_times.Rda")
  times <- dplyr::filter(test_times, var == "start_1" | var == "shutdown") %>%
            tidyr::spread(var, value) %>%
            dplyr::select(-date) %>%
            dplyr::rename(start = start_1, end = shutdown)
```

## Constants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flow_hood <- 4 # m^3/min
```

___

## Co2

* Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/co2_merged.Rda")
```

* select sample data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- dplyr::filter(co2_merged, loc == "sample")
```

* filter out data from outside emissions window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- filter_times(times, co2)
```

* calculate average for each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- dplyr::group_by(co2, id) %>%
          dplyr::summarise(ppm = mean(ppm))
```

* convert mixing ratio to mass (assuming constant T and P)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2$conc <- convert_ppmv_ugpmc(co2$ppm, 28.01, 25, 84)
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- dplyr::left_join(co2, times, by = "id") %>%
          dplyr::mutate(dur = end - start) %>%
          dplyr::mutate(id = as.factor(id))
```

* calculate micrograms of co2 emitted during test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co2 <- mutate(co2, mass_emitted_co2 = conc * flow_hood * (dur/60)) # ug/m^3 * m^3/min * (s / 60)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(co2, 2)
```

___

## ecoc

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/ecoc_merged.Rda")
```

* filter area

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  area <- 11.8 # cm^2
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  ecoc <- dplyr::select(ecoc_merged,
                        id, cassette,
                        start, end,
                        pre_flow, post_flow,
                        oc_ug_sq_cm, ec_ug_sq_cm, tc_ug_sq_cm)
```


* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  ecoc <- dplyr::mutate(ecoc, flow = (pre_flow + post_flow)/2,
                              dur = (end - start)/60,
                              conc_oc = oc_ug_sq_cm * 11.8 * 1000 / (flow * dur),
                              conc_ec = ec_ug_sq_cm * 11.8 * 1000 / (flow * dur),
                              conc_tc = tc_ug_sq_cm * 11.8 * 1000 / (flow * dur))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::mutate(ecoc, 
                        mass_emitted_oc = conc_oc * flow_hood * dur,
                        mass_emitted_ec = conc_ec * flow_hood * dur,
                        mass_emitted_tc = conc_tc * flow_hood * dur)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ecoc, 2)
```

___

