---
title: "mass_emissions"
author: "Nicholas Good"
date: "1/6/2017"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

## Test time data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/test_times.Rda")
  times <- dplyr::filter(test_times, var == "start_1" | var == "shutdown") %>%
            tidyr::spread(var, value) %>%
            dplyr::select(-date) %>%
            dplyr::rename(start = start_1, end = shutdown)
```

## Constants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flow_hood <- 4 # m^3/min
```

___

## Co2

* Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/co2_merged.Rda")
```

* select sample data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- dplyr::filter(co2_merged, loc == "sample")
```

* filter out data from outside emissions window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- filter_times(times, co2)
```

* calculate average for each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- dplyr::group_by(co2, id) %>%
          dplyr::summarise(ppm = mean(ppm))
```

* convert mixing ratio to mass (assuming constant T and P)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2$conc <- convert_ppmv_ugpmc(co2$ppm, 44.01, 25, 84)
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- dplyr::left_join(co2, times, by = "id") %>%
          dplyr::mutate(dur = end - start) %>%
          dplyr::mutate(id = as.factor(id))
```

* calculate micrograms of co2 emitted during test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co2 <- mutate(co2, mass_emitted_co2 = conc * flow_hood * (dur/60)) # ug/m^3 * m^3/min * (s / 60)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(co2, 2)
```

___

## ecoc

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/ecoc_merged.Rda")
```

* filter area

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  area <- 11.8 # cm^2
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  ecoc <- dplyr::select(ecoc_merged,
                        id, cassette,
                        start, end,
                        pre_flow, post_flow,
                        oc_ug_sq_cm, ec_ug_sq_cm, tc_ug_sq_cm)
```


* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  ecoc <- dplyr::mutate(ecoc, flow = (pre_flow + post_flow)/2,
                              dur = (end - start)/60,
                              conc_oc = oc_ug_sq_cm * 11.8 * 1000 / (flow * dur),
                              conc_ec = ec_ug_sq_cm * 11.8 * 1000 / (flow * dur),
                              conc_tc = tc_ug_sq_cm * 11.8 * 1000 / (flow * dur))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::mutate(ecoc, 
                        mass_emitted_oc = conc_oc * flow_hood * dur,
                        mass_emitted_ec = conc_ec * flow_hood * dur,
                        mass_emitted_tc = conc_tc * flow_hood * dur)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ecoc, 2)
```

___

## fivegas


* Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/fivegas_merged.Rda")
```

* select co and ch4

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co_ch4 <- dplyr::filter(fivegas_merged, pol == "co" | pol == "ch4")
```

* filter out data from outside emissions window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co_ch4 <- filter_times(times, co_ch4)
```

* calculate average for each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co_ch4 <- dplyr::group_by(co_ch4, id, pol) %>%
            dplyr::summarise(ppm = mean(val)) %>%
            dplyr::ungroup()
```

* convert mixing ratio to mass (assuming constant T and P)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  mw <- data_frame(pol = c("co", "ch4"), mw = c(28.01, 16.04))
  co_ch4 <- dplyr::left_join(co_ch4, mw, by = "pol")
  co_ch4$conc <- convert_ppmv_ugpmc(co_ch4$ppm, co_ch4$mw, 25, 84)
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co_ch4 <- dplyr::left_join(co_ch4, times, by = "id") %>%
          dplyr::mutate(dur = end - start) %>%
          dplyr::mutate(id = as.factor(id))
```

* calculate micrograms emitted during tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_ch4 <- dplyr::mutate(co_ch4, 
                          mass_emitted = conc * flow_hood * (dur/60)) %>%
            tidyr::spread(pol, mass_emitted) %>%
            dplyr::rename(mass_emitted_co = co, mass_emitted_ch4 = ch4)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(co_ch4, 2)
```

___

## grav

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/grav_merged.Rda")
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  grav <- dplyr::select(grav_merged,
                        id,
                        time_start, time_end,
                        flow_pre, flow_post,
                        wgt_delta)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  grav <- dplyr::mutate(grav, flow = (flow_pre + flow_post)/2,
                              dur = (time_end - time_start)/60,
                              conc_pm = wgt_delta * 1000 / (flow * dur))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, 
                        mass_emitted_pm = conc_pm * flow_hood * dur)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(grav, 2)
```

___

## Smps

* Load data

Note data is test average from desired time window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/smps_ultrafine.Rda")
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  ultrafine <- dplyr::left_join(smps_ultrafine, times, by = "id") %>%
                dplyr::mutate(dur = end - start) %>%
                dplyr::mutate(id = as.factor(id))
```

* calculate number of particles emitted during tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ultrafine <- mutate(ultrafine, emitted = mean_number_conc * 1e6 * flow_hood * (dur/60)) # (N * 1,000,000 /cm^3) * m^3/min * (s / 60)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ultrafine, 2)
```

___

## voc

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/voc_merged.Rda")

  flow <- 13.5 # L/min
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  voc <- dplyr::select(voc_merged,
                       id,
                       time_start_voc, time_end_voc,
                       pol, val) %>%
         dplyr::rename(start = time_start_voc,
                       end = time_end_voc)
```

* moleculer weight data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  mw_voc <- data_frame(pol = unique(voc$pol)) %>%
            dplyr::mutate(mw = NA) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethane", 30.07, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethene", 28.05, mw)) %>%  
            dplyr::mutate(mw = ifelse(pol == "voc_propane", 44.1, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_propene", 42.08, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_butane", 58.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_butane", 58.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethyne", 26.04, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_t_2_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_cyclopentane", 70.1, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_pentane", 72.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_pentane", 72.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_t_2_pentene", 70.1329, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methyl_2_butene", 70.1329, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_pentene", 70.135, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_pentene", 70.135, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_cyclohexane", 84.16, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_hexane", 86.178, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_isoprene", 68.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_hexene", 84.1608, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_hexene", 84.162, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_4_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_methylcyclohexane", 98.186, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_heptane", 100.21, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_benzene", 78.11, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylpentane", 86.178, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_3_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "2_methylhexane", 100.2, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "3_methylpentane", 86.18, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "2_3_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methylhexane", 100.2, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylhexane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_2_4_trimethylpentane", 114.232, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_3_4_trimethylpentane", 114.232, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_toluene", 92.14, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methylheptane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylheptane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_octane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethylbenzene", 106.17, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_m_p_xylene", 106.168, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_styrene", 104.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_o_xylene", 106.16, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_no_8888ne", NA, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_propylbenzene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_a_pinene", 136.24, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_propylbenzene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_4_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_3_5_trimethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_4_trimethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_decane", 142.29, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_3_trimethylbenzene", 30.07, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_3_diethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_4_diethylbenzene", 134.222, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_diethylbenzene", 134.222, mw))

  voc <- dplyr::left_join(voc, mw_voc, by = "pol")
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  voc <- dplyr::mutate(voc, dur = (end - start) / 60,
                            ppm = val / 1000,
                            conc = convert_ppmv_ugpmc(ppm, mw, 25, 85))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::mutate(voc, 
                        mass_emitted = conc * flow_hood * dur)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(voc, 2)
```

___