---
title: "mass_emissions"
author: "Nicholas Good"
date: "1/6/2017"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

## Test time data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/test_times.Rda")

  times <- dplyr::filter(test_times, var == "start_1" | var == "shutdown") %>%
           tidyr::spread(var, value) %>%
           dplyr::select(-date) %>%
           dplyr::rename(start = start_1, end = shutdown)
```

## Sample info

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
```

## Constants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flow_hood <- 4  # m^3/min
```

## Mass of fuel burnt

* Load fuel mass data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/batch_wgts.Rda")
  load("../r_files/wood_wgts.Rda")
```

### Extract mass of fuel burnt during sampling periods

* batch stoves

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  batch_wgts <- dplyr::filter(batch_wgts, var == "fuel" |
                                          var == "refueled" |
                                          var == "preshutdown") %>%
                tidyr::spread(var, value) %>%
                dplyr::rename(wgt_fuel = fuel,
                              wgt_refuel = refueled,
                              wgt_shutdown = preshutdown) %>%
                dplyr::mutate(mass_fuel = wgt_fuel + wgt_refuel - wgt_shutdown)
```

* wood stoves

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  wood_wgts <- dplyr::filter(wood_wgts, var == "fuel" |
                                        var == "ashpot_unusedfuel" |
                                        var == "ashpot_char_ash" |
                                        var == "ashpot_lid") %>%
                tidyr::spread(var, value) %>%
                dplyr::rename(wgt_fuel = fuel,
                              wgt_pot_unusedfuel = ashpot_unusedfuel,
                              wgt_pot_char_ash = ashpot_char_ash,
                             wgt_pot = ashpot_lid) %>%
                dplyr::mutate(mass_fuel = wgt_fuel -
                                          wgt_pot_unusedfuel -
                                          wgt_pot_char_ash +
                                          2 * wgt_pot)
```

* merge test types

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  fuel_burnt <- dplyr::bind_rows(dplyr::select(batch_wgts, id, mass_fuel),
                                 dplyr::select(wood_wgts, id, mass_fuel)) %>%
                dplyr::mutate(id = as.factor(id)) %>%
                dplyr::filter(grepl("^[0-9]", id) == TRUE)

```

* plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=TRUE, fig.width=12, fig.height=8}
  p_data <- dplyr::left_join(samples, fuel_burnt) %>%
            dplyr::mutate(id = as.factor(id)) %>%
            dplyr::filter(grepl("^[0-9]", id) == TRUE)

  ggplot(p_data, aes(x = id, y = mass_fuel, colour = stove)) +
        geom_point() +
        theme_minimal() + 
        theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
        theme(legend.position = "top")
        ylab("fuel burnt (g)")

```

* save

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
 save(fuel_burnt, file="../r_files/fuel_burnt.Rda")
```

___

## co2

* Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/co2_merged.Rda")
```

* select sample data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- dplyr::filter(co2_merged, loc == "sample")
```

* filter out data from outside emissions window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- filter_times(times, co2)
```

* calculate average for each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- dplyr::group_by(co2, id) %>%
         dplyr::summarise(ppm = mean(ppm))
```

* convert mixing ratio to mass (assuming constant T and P)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2$conc <- convert_ppmv_ugpmc(co2$ppm, 44.01, 25, 84)
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co2 <- dplyr::left_join(co2, times, by = "id") %>%
         dplyr::mutate(dur = end - start) %>%
         dplyr::mutate(id = as.factor(id))
```

* calculate micrograms of co2 emitted during test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co2 <- dplyr::mutate(co2, mass_emitted_co2 = conc * flow_hood * (dur/60))  # ug/m^3 * m^3/min * (s / 60)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(co2, 2)
```

___

## ecoc

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/ecoc_merged.Rda")
```

* filter area

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  area <- 11.8  # cm^2
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  ecoc <- dplyr::select(ecoc_merged,
                        id, cassette,
                        start, end,
                        pre_flow, post_flow,
                        pol, val)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  ecoc <- dplyr::mutate(ecoc, flow = (pre_flow + post_flow)/2,
                              dur = (end - start)/60,
                              conc = val * area * 1000 / (flow * dur))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  ecoc <- dplyr::mutate(ecoc, 
                        mass_emitted = conc * flow_hood * dur)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ecoc, 2)
```

___

## fivegas

* Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/fivegas_merged.Rda")
```

* select co and ch4

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co_ch4 <- dplyr::filter(fivegas_merged, pol == "co" | pol == "ch4")
```

* filter out data from outside emissions window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co_ch4 <- filter_times(times, co_ch4)
```

* calculate average for each test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co_ch4 <- dplyr::group_by(co_ch4, id, pol) %>%
            dplyr::summarise(ppm = mean(val)) %>%
            dplyr::ungroup()
```

* convert mixing ratio to mass (assuming constant T and P)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  mw <- data_frame(pol = c("co", "ch4"), mw = c(28.01, 16.04))

  co_ch4 <- dplyr::left_join(co_ch4, mw, by = "pol") %>%
            dplyr::mutate(conc = convert_ppmv_ugpmc(ppm, mw, 25, 84))
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  co_ch4 <- dplyr::left_join(co_ch4, times, by = "id") %>%
            dplyr::mutate(dur = end - start) %>%
            dplyr::mutate(id = as.factor(id))
```

* calculate micrograms emitted during tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  co_ch4 <- dplyr::mutate(co_ch4, 
                          mass_emitted = conc * flow_hood * (dur/60)) %>%
            tidyr::spread(pol, mass_emitted) %>%
            dplyr::rename(mass_emitted_co = co, mass_emitted_ch4 = ch4)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(co_ch4, 2)
```

___

## grav

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/grav_merged.Rda")
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  grav <- dplyr::select(grav_merged,
                        id,
                        time_start, time_end,
                        flow_pre, flow_post,
                        wgt_delta)
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  grav <- dplyr::mutate(grav, flow = (flow_pre + flow_post)/2,
                              dur = (time_end - time_start)/60,
                              conc_pm = wgt_delta * 1000 / (flow * dur))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  grav <- dplyr::mutate(grav, 
                        mass_emitted_pm = conc_pm * flow_hood * dur)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(grav, 2)
```

___

## Smps

* Load data

Note data is test average from desired time window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/smps_ultrafine.Rda")
```

* merge with sample time data and calculate test duration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  ultrafine <- dplyr::left_join(smps_ultrafine, times, by = "id") %>%
               dplyr::mutate(dur = end - start) %>%
               dplyr::mutate(id = as.factor(id))
```

* calculate number of particles emitted during tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # (N * 1,000,000 /cm^3) * m^3/min * (s / 60)
  ultrafine <- mutate(ultrafine, emitted = mean_number_conc * 1e6 * flow_hood * (dur/60))
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(ultrafine, 2)
```

___

## voc

* load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  load("../r_files/voc_merged.Rda")

  flow <- 13.5 # L/min
```

* select data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  voc <- dplyr::select(voc_merged,
                       id,
                       time_start_voc, time_end_voc,
                       pol, val) %>%
         dplyr::rename(start = time_start_voc,
                       end = time_end_voc)
```

* moleculer weight data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  mw_voc <- data_frame(pol = unique(voc$pol)) %>%
            dplyr::mutate(mw = NA) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethane", 30.07, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethene", 28.05, mw)) %>%  
            dplyr::mutate(mw = ifelse(pol == "voc_propane", 44.1, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_propene", 42.08, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_butane", 58.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_butane", 58.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethyne", 26.04, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_t_2_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_butene", 56.108, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_cyclopentane", 70.1, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_pentane", 72.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_pentane", 72.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_t_2_pentene", 70.1329, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methyl_2_butene", 70.1329, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_pentene", 70.135, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_pentene", 70.135, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_cyclohexane", 84.16, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_hexane", 86.178, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_isoprene", 68.12, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_hexene", 84.1608, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_c_2_hexene", 84.162, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_4_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_methylcyclohexane", 98.186, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_heptane", 100.21, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_benzene", 78.11, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylpentane", 86.178, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_3_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "2_methylhexane", 100.2, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "3_methylpentane", 86.18, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "2_3_dimethylpentane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methylhexane", 100.2, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylhexane", 100.205, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_2_4_trimethylpentane", 114.232, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_3_4_trimethylpentane", 114.232, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_toluene", 92.14, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_methylheptane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_methylheptane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_octane", 114.23, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_ethylbenzene", 106.17, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_m_p_xylene", 106.168, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_styrene", 104.15, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_o_xylene", 106.16, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_no_8888ne", NA, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_i_propylbenzene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_a_pinene", 136.24, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_propylbenzene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_3_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_4_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_3_5_trimethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_2_ethyltoluene", 120.195, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_4_trimethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_n_decane", 142.29, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_3_trimethylbenzene", 30.07, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_3_diethylbenzene", 120.19, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_4_diethylbenzene", 134.222, mw)) %>%
            dplyr::mutate(mw = ifelse(pol == "voc_1_2_diethylbenzene", 134.222, mw))

  voc <- dplyr::left_join(voc, mw_voc, by = "pol")
```

* calculate concentrations

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  voc <- dplyr::mutate(voc, dur = (end - start) / 60,
                            ppm = val / 1000,
                            conc = convert_ppmv_ugpmc(ppm, mw, 25, 85))
```

* calculate mass emitted

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::mutate(voc, 
                        mass_emitted = conc * flow_hood * dur)
```

* output

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(voc, 2)
```

___