---
title: "Overall Manuscript Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(forcats)
  library(reshape2)
  library(MASS)
  library(GGally)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # Load data and meta data
  load("../r_files/emission_factors.Rda")
  load("../r_files/samples.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # Load functions
  source("../r_scripts/R_functions.R")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # Add stove/fuel information
  emission_factors <- dplyr::left_join(emission_factors, 
                                       dplyr::select(samples, id, stove, fuel,
                                                     fuelcat, stovecat),
                                       by = "id")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # Use carbon balance method for missing emissions factors
  emission_factors <- dplyr::mutate(emission_factors, 
                                    mass_ef_comb = 
                                      ifelse(is.na(mass_ef),
                                             mass_c_ef, mass_ef)) %>%
                      dplyr::mutate(energy_ef_comb =
                                      ifelse(is.na(energy_ef), energy_c_ef,
                                             energy_ef))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # Remove outliers
  emission_factors <- dplyr::filter(emission_factors, mass_ef_comb > 0) %>%
                      dplyr::filter(energy_ef_comb > 0)
```

# Bar plots in polar coordinates

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 20,  fig.height = 30}
  plot_ef_polar_all(emission_factors)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 20,  fig.height = 30}
  plot_ef_polar(dplyr::filter(emission_factors, grepl('wood', fuelcat)))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 20,  fig.height = 30}
  plot_ef_polar(dplyr::filter(emission_factors, grepl('charcoal', fuelcat)))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 20,  fig.height = 30}
  plot_ef_polar(dplyr::filter(emission_factors, grepl('pellets', fuelcat)))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 20,  fig.height = 30}
  plot_ef_polar(dplyr::filter(emission_factors, grepl('advanced', fuelcat)))
```

# Correlation heat maps for pollutant categories
* Spearman correlation plots (Spearman's rho)
* Correlations are for emissions factors on a energy delivered from the fuel basis

## All stove types

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::distinct(emission_factors) %>%
          dplyr::filter(pol != "co2") %>% dplyr::filter(pol != "ch4") %>%
          dplyr::filter(inst != "ions") %>%
          dplyr::mutate(inst = ifelse(inst == "fivegas", as.character(pol), as.character(inst))) %>%
          dplyr::mutate(inst = ifelse(inst == "ecoc", as.character(pol), as.character(inst))) %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::select(id, inst, energy_ef_comb) %>%
          dplyr::group_by_(.dots = c("id", "inst")) %>% 
          dplyr::summarise(energy_ef_comb = sum(energy_ef_comb, na.rm = TRUE)) %>%
          tidyr::spread(inst, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

## Wood fuels

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::filter(emission_factors, fuelcat == "wood") %>%
          dplyr::distinct() %>%
          dplyr::filter(pol != "co2") %>% dplyr::filter(pol != "ch4") %>%
          dplyr::filter(inst != "ions") %>%
          dplyr::mutate(inst = ifelse(inst == "fivegas", as.character(pol), as.character(inst))) %>%
          dplyr::mutate(inst = ifelse(inst == "ecoc", as.character(pol), as.character(inst))) %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::select(id, inst, energy_ef_comb) %>%
          dplyr::group_by_(.dots = c("id", "inst")) %>% 
          dplyr::summarise(energy_ef_comb = sum(energy_ef_comb, na.rm = TRUE)) %>%
          tidyr::spread(inst, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

## Charcoal fuels

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::filter(emission_factors, fuelcat == "charcoal") %>%
          dplyr::distinct() %>%
          dplyr::filter(pol != "co2") %>% dplyr::filter(pol != "ch4") %>%
          dplyr::filter(inst != "ions") %>%
          dplyr::mutate(inst = ifelse(inst == "fivegas", as.character(pol), as.character(inst))) %>%
          dplyr::mutate(inst = ifelse(inst == "ecoc", as.character(pol), as.character(inst))) %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::select(id, inst, energy_ef_comb) %>%
          dplyr::group_by_(.dots = c("id", "inst")) %>% 
          dplyr::summarise(energy_ef_comb = sum(energy_ef_comb, na.rm = TRUE)) %>%
          tidyr::spread(inst, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

## Pellet fuels

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::filter(emission_factors, fuelcat == "pellets") %>%
          dplyr::distinct() %>%
          dplyr::filter(pol != "co2") %>% dplyr::filter(pol != "ch4") %>%
          dplyr::filter(inst != "ions") %>%
          dplyr::mutate(inst = ifelse(inst == "fivegas", as.character(pol), as.character(inst))) %>%
          dplyr::mutate(inst = ifelse(inst == "ecoc", as.character(pol), as.character(inst))) %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::select(id, inst, energy_ef_comb) %>%
          dplyr::group_by_(.dots = c("id", "inst")) %>% 
          dplyr::summarise(energy_ef_comb = sum(energy_ef_comb, na.rm = TRUE)) %>%
          tidyr::spread(inst, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

## Advanced fuels

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::filter(emission_factors, fuelcat == "advanced") %>%
          dplyr::distinct() %>%
          dplyr::filter(pol != "co2") %>% dplyr::filter(pol != "ch4") %>%
          dplyr::filter(inst != "ions") %>%
          dplyr::mutate(inst = ifelse(inst == "fivegas", as.character(pol), as.character(inst))) %>%
          dplyr::mutate(inst = ifelse(inst == "ecoc", as.character(pol), as.character(inst))) %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::select(id, inst, energy_ef_comb) %>%
          dplyr::group_by_(.dots = c("id", "inst")) %>% 
          dplyr::summarise(energy_ef_comb = sum(energy_ef_comb, na.rm = TRUE)) %>%
          tidyr::spread(inst, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

# Correlation heat maps for carcinogenic compounds

## All stove types

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::distinct(emission_factors) %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::filter(grepl("^co$|^grav$|^voc_benzene$|^formaldehyde$|^acetaldehyde$", pol)) %>%
          dplyr::select(id, pol, energy_ef_comb) %>%
          tidyr::spread(pol, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

## Wood fuels

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::filter(emission_factors, fuelcat == "wood") %>%
          dplyr::distinct() %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::filter(grepl("^co$|^grav$|^voc_benzene$|^formaldehyde$|^acetaldehyde$", pol)) %>%
          dplyr::select(id, pol, energy_ef_comb) %>%
          tidyr::spread(pol, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

## Charcoal fuels

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::filter(emission_factors, fuelcat == "charcoal") %>%
          dplyr::distinct() %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::filter(grepl("^co$|^grav$|^voc_benzene$|^formaldehyde$|^acetaldehyde$", pol)) %>%
          dplyr::select(id, pol, energy_ef_comb) %>%
          tidyr::spread(pol, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

## Pellet fuels

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::filter(emission_factors, fuelcat == "pellets") %>%
          dplyr::distinct() %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::filter(grepl("^co$|^grav$|^voc_benzene$|^formaldehyde$|^acetaldehyde$", pol)) %>%
          dplyr::select(id, pol, energy_ef_comb) %>%
          tidyr::spread(pol, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

## Advanced fuels

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 14,  fig.height = 10}
    # convert to long format
  ef_w <- dplyr::filter(emission_factors, fuelcat == "advanced") %>%
          dplyr::distinct() %>%
          dplyr::group_by_(.dots = c("id", "pol","inst")) %>% 
          dplyr::summarise(energy_ef_comb = mean(energy_ef_comb, na.rm = TRUE)) %>%
          dplyr::filter(grepl("^co$|^grav$|^voc_benzene$|^formaldehyde$|^acetaldehyde$", pol)) %>%
          dplyr::select(id, pol, energy_ef_comb) %>%
          tidyr::spread(pol, energy_ef_comb)

  plot_cormap(ef_w, "spearman")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  #emission_factors_w <- dplyr::select(emission_factors, id, pol, mass_ef_comb)

 # emission_factors_w$row <- 1:nrow(emission_factors_w)

 # emission_factors_w <- tidyr::spread(emission_factors_w, pol, mass_ef_comb) %>%
                  #      dplyr::group_by(id) %>%
                 #       dplyr::summarise_each(funs(mean(., na.rm = TRUE))) %>%
                   #     dplyr::left_join(dplyr::select(emission_factors, id, fuelcat))
  
 # emission_factors_w <- unique(emission_factors_w)
  #emission_factors_w <- emission_factors_w[-2]
  #emission_factors_w <- emission_factors_w[-1]
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width = 25,  fig.height = 25}
  
  #ggpairs(
        #emission_factors_w, 1:10,
        #mapping = aes(color = fuelcat, alpha = 0.4),
        #types = list(continuous = "smooth_lm"),
        #lower = list(continuous = add_smooth)) +
    #theme_bw() +
    #theme(legend.position = "top",
          #axis.text.y = element_text(size = 16),
          #axis.text.x = element_text(angle = 45, vjust = 1, size = 16, hjust = 1),
          #strip.text.x = element_text(size = 16),
          #strip.text.y = element_text(size = 16))
    
```