---
title: "PAHs"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  library(tidyverse)
  library(knitr)
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/', fig.height = 5, fig.width = 12,
                        warning = FALSE, message = FALSE, cache = FALSE)
```

# Load files

Source files

```{r load_source}
  source("../r_scripts/R_load_data.R")
  source("../r_scripts/R_load_metadata.R")
  source("../r_scripts/R_tidy.R")
```

PAH data

```{r load_data}
  pah_raw <- readRDS("../r_files/pah.RDS")    # pah dataset
  pah_lod <- readRDS("../r_files/pah_lod.RDS")    # pah lod dataset
```

Metadata

```{r load_metadata}
  load("../r_files/data_1.Rda")    # metadata
  load("../r_files/notes.Rda")    # metadata
```

# Tidy data

## Merge raw and LOD data

Extract chemical analysis LOD data

```{r}
  lod <-
    pah_lod %>%
    dplyr::mutate(lod = ifelse(val == "BDL", "below", "above")) %>%
    dplyr::select(id, lod, filter_type, pol)
```

Calculate new value if measurement below LOD

```{r}
  pah <- 
    pah_raw %>% 
    dplyr::left_join(lod, by = c("id", "filter_type", "pol")) %>%
    dplyr::mutate(val = as.numeric(val),
                  val_lod_adj = ifelse(lod == "below", val / sqrt(2), val))
```

## Merge data with qc info

Extract notes for pah

```{r}
  notes <- 
    notes %>%
    dplyr::filter(grepl("pah|all", inst) == TRUE)
```

Apply flags: bad preceeds maybe preceeds good

```{r}
  flags <-
    notes %>%
    dplyr::select(id, qc) %>%
    dplyr::group_by(id) %>%
    dplyr::arrange(qc) %>%
    dplyr::summarise(qc = first(qc))
```

Merge data with flags

```{r}
  pah_merged <-
    pah %>%
    dplyr::left_join(flags, by = "id") %>%
    dplyr::mutate(id = as.factor(id)) %>%
    dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

Filter out bad tests

```{r}
  pah_merged <-
    pah_merged %>% 
    dplyr::filter(qc != "bad", !is.na(val_lod_adj))
```

# Blank analysis 

Extract blank pufs and blank filters

```{r}
  blanks <-
    pah_merged %>%
    dplyr::filter(grepl("^B", id))
```

Plot blank mass in ug:

* color indicates if sample was above or below instrument lod
* mean shown with an x
* median shown with a square

```{r pah_blank_plot, echo=FALSE, fig.height = 15}
  blanks <-
    blanks %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    as.factor %>%
                    forcats::fct_recode("puf" = "b",
                                        "puf" = "f",
                                        "filter" = "ba"))

  ggplot(blanks, aes_string(x = "filter_type", y = "val_lod_adj", colour = "lod")) +
    geom_point(size = 3, shape = 1, stroke = 1) +
    geom_point(stat = "summary", fun.y = "mean", colour = "black",
               shape = 4, size = 4, stroke = 2) +
    geom_point(stat = "summary", fun.y = "median", colour = "black",
               shape = 0, size = 4, stroke = 2) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
          text = element_text(size = 18),
          strip.text.x = element_text(size = 12)) +
    facet_wrap(~pol, ncol = 4, scales = "free") +
    ylab("mass (ug)") +
    xlab("")
```

Summarize blank data

```{r}
  blanks %<>%
    dplyr::group_by(id, pol, filter_type) %>%
    dplyr::summarise(mean = mean(val_lod_adj),
                     std = sd(val_lod_adj),
                     median = median(val_lod_adj))
```

# Calculate concentrations

Extract pah filter cartridge data

```{r extract_cartridge_data}
  meta <-
    data_1 %>%
    dplyr::select(matches("_red|^id$|^date$"))
```

Extra flow rate data

```{r extract_flows}
  flows <-
    meta %>%
    dplyr::select(matches("flow_|^id$|^date$")) %>%
    tidyr::gather("var", "value", -id, -date) %>%
    dplyr::filter(grepl("_avg$", var)==FALSE) %>%
    dplyr::mutate(type = sub("flow.*", "", var),
                  rep = as.factor(gsub("[^0-9]", "", var))) %>%
    dplyr::select(-var)  %>%
    dplyr::filter(!is.na(id)) %>%
    dplyr::group_by(rep, id, type) %>%
    dplyr::mutate(mean = mean(value, na.rm = TRUE))  %>%
    dplyr::ungroup() %>%
    dplyr::filter(rep == 1) %>%
    dplyr::select(-value, -rep, -date)  %>%
    tidyr::spread(type, mean)
```

Extra time data

```{r extract_times}
  times <-
    meta %>% 
    dplyr::select(matches("^time_|^id$|^date$")) %>%
    tidyr::gather("var", "value", -id, -date) %>%
    dplyr::mutate(type = ifelse(grepl("_start_", var), "start", "NA"),
                  type = ifelse(grepl("_end_", var), "end", type)) %>%
    dplyr::arrange(id) %>%
    dplyr::filter(!is.na(id)) %>%  
    dplyr::select(-var, -date) %>%  
    tidyr::spread(type, value)
```

Merge pah with filter metadata

```{r}
  pah_merged %<>%
    dplyr::filter(!grepl("^B", id)) %>%
    dplyr::left_join(flows, by = "id") %>%
    dplyr::left_join(times, by = "id") %>%
    dplyr::rename(pre_flow = pre, post_flow = post) %>%
    dplyr::mutate(flow = (pre_flow + post_flow) / 2)
```



# Save merged file

```{r}
  # saveRDS(pah, file = "../r_files/pah_merged.RDS")
```
