---
title: "PAHs"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE, warning = FALSE}
  library(tidyverse)
  library(knitr)
  library(kableExtra)
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/', fig.height = 5, fig.width = 12,
                        echo = FALSE,
                        warning = FALSE, message = FALSE, cache = FALSE)
```

```{r load_source}
  # load source files
  source("../r_scripts/R_load_data.R")
  source("../r_scripts/R_load_metadata.R")
  source("../r_scripts/R_tidy.R")
```

```{r load_data}
  # load pah data
  pah_raw <- readRDS("../r_files/pah.RDS")    # pah dataset
  pah_lod <- readRDS("../r_files/pah_lod.RDS")    # pah lod dataset
```

```{r load_metadata}
  # load metadata
  load("../r_files/data_1.Rda")
  load("../r_files/notes.Rda")
  load("../r_files/samples.Rda")
```

```{r}
  # Extract chemical analysis lod data
  lod <-
    pah_lod %>%
    dplyr::mutate(lod = ifelse(val == "BDL", "below", "above")) %>%
    dplyr::select(id, lod, filter_type, pol)
```

```{r}
  # Calculate new value if measurement below instrument lod
  # Replace values below the lod and loq with lod/sqrt(2)
  pah <- 
    pah_raw %>% 
    dplyr::mutate(val = as.numeric(val)) %>%
    dplyr::group_by(id, pol, filter_type) %>%
    dplyr::summarise(val = last(val)) %>%
    dplyr::ungroup() %>%
    dplyr::left_join(lod, by = c("id", "filter_type", "pol")) %>%
    dplyr::mutate(val_lod_adj = ifelse(lod == "below", val / sqrt(2), val))
```

```{r}
  # Extract notes for pah
  notes <- 
    notes %>%
    dplyr::filter(grepl("pah|all", inst) == TRUE)
```

```{r}
  # Apply flags: bad preceeds maybe preceeds good
  flags <-
    notes %>%
    dplyr::select(id, qc) %>%
    dplyr::group_by(id) %>%
    dplyr::arrange(qc) %>%
    dplyr::summarise(qc = first(qc))
```

```{r}
  # Merge data with flags
  pah <-
    pah %>%
    dplyr::left_join(flags, by = "id") %>%
    dplyr::mutate(id = as.factor(id)) %>%
    dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

```{r}
  # Filter out bad tests
  pah <-
    pah %>% 
    dplyr::filter(qc != "bad")
```

```{r}
  # Define new variable called test type
  pah <- 
    pah %>%
    dplyr::mutate(test_type = ifelse(!grepl("^B|^G", id), "test", "bg"),
                  test_type = ifelse(grepl("^B", id), "blank", test_type))
```

# Blank analysis 

```{r}
  # Extract blank pufs and blank filters
  blanks <-
    pah %>%
    dplyr::filter(test_type == "blank")
```

## Plot: Lab blanks

Lab blanks (ug):

* plotted by pollutant and filter type
* color indicates if sample was above or below instrument lod
* mean shown with an x
* median shown with a square

```{r pah_blank_plot, echo=FALSE, fig.height = 15}
 blanks %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "ba")) %>%
    ggplot(aes_string(x = "filter_type", y = "val_lod_adj", colour = "lod")) +
      geom_point(size = 3, shape = 1, stroke = 1) +
      geom_point(stat = "summary", fun.y = "mean", colour = "black",
                 shape = 4, size = 4, stroke = 2) +
      geom_point(stat = "summary", fun.y = "median", colour = "black",
                 shape = 0, size = 4, stroke = 2) +
      theme_bw() +
      theme(axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            text = element_text(size = 20),
            strip.text.x = element_text(size = 12),
            legend.position = "top") +
      labs(colour = "chemical analysis lod") +
      facet_wrap(~pol, ncol = 4, scales = "free") +
      ylab("") +
      xlab("")
```

## Table: Lab blanks below lod for chemical analysis

Lab blanks below lod for chemical analysis (%):

* values given by pollutant and filter type

```{r}
  blanks %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "ba")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of blanks below chemical analysis lod: `r round(nrow(filter(blanks, lod == "below")) / nrow(blanks), 2) * 100`%.

## Table: Lab blanks with NA values

Lab blanks with NA values (%):

* values given by pollutant and filter type

```{r}
  blanks %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "ba")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val_lod_adj) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of blanks with NA measurements: `r round(sum(is.na(blanks$val)) / nrow(blanks), 2) * 100`%.
* Total percent of 'bad' blank measurements: `r round(nrow(filter(blanks, lod == "below", is.na(val))) / nrow(blanks), 2) * 100`%.

```{r}
  blanks <-
    blanks %>%
    dplyr::group_by(pol, filter_type) %>%
    dplyr::summarise(bk_mean = mean(val_lod_adj, na.rm = TRUE),
                     bk_median = median(val_lod_adj, na.rm = TRUE),
                     bk_sd = sd(val_lod_adj, na.rm = TRUE)) %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("p" = "ba"))
```

```{r, include=FALSE}
  # summary table of blank data
  blanks %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    knitr::kable(digits = 2, align = 'c',
                 col.names = c("type of pah", "filter type", "mean", "median", "standard deviation"))
```

```{r}
  # Lab blank correction 
  # replace values below the lod and loq with lod/sqrt(2)
  # replace values below loq and above lod with loq/sqrt(2)
  pah_merged <-
    pah %>%
    dplyr::filter(test_type != "blank") %>%  # remove blanks
    dplyr::left_join(blanks, by = c("pol", "filter_type")) %>%
    dplyr::mutate(val_bk_adj = val_lod_adj - bk_mean,
                  bk_lod = ifelse(val_bk_adj <= 3 * bk_sd, "below", "above"),
                  bk_loq = ifelse(val_bk_adj <= 10 * bk_sd, "below", "above"),
                  val_bk_adj = ifelse(bk_lod == "below" & bk_lod == "below",
                                      (3 * bk_sd) / sqrt(2), val_bk_adj),
                  val_bk_adj = ifelse(bk_lod == "above" & bk_loq == "below",
                                      (10 * bk_sd) / sqrt(2), val_bk_adj))
```

```{r extract_cartridge_data}
  # Extract pah filter cartridge data
  meta <-
    data_1 %>%
    dplyr::select(matches("_red|^id$|^date$"))
```

```{r extract_flows}
  # Extra flow rate data
  flows <-
    meta %>%
    dplyr::select(matches("flow_|^id$|^date$")) %>%
    tidyr::gather("var", "value", -id, -date) %>%
    dplyr::filter(grepl("_avg$", var)==FALSE) %>%
    dplyr::mutate(type = sub("flow.*", "", var),
                  rep = as.factor(gsub("[^0-9]", "", var))) %>%
    dplyr::select(-var)  %>%
    dplyr::filter(!is.na(id)) %>%
    dplyr::group_by(rep, id, type) %>%
    dplyr::mutate(mean = mean(value, na.rm = TRUE))  %>%
    dplyr::ungroup() %>%
    dplyr::filter(rep == 1) %>%
    dplyr::select(-value, -rep, -date)  %>%
    tidyr::spread(type, mean)
```

```{r extract_times}
  # Extra time data
  times <-
    meta %>% 
    dplyr::select(matches("^time_|^id$|^date$")) %>%
    tidyr::gather("var", "value", -id, -date) %>%
    dplyr::mutate(type = ifelse(grepl("_start_", var), "start", "NA"),
                  type = ifelse(grepl("_end_", var), "end", type)) %>%
    dplyr::arrange(id) %>%
    dplyr::filter(!is.na(id)) %>%  
    dplyr::select(-var, -date) %>%  
    tidyr::spread(type, value)
```

```{r}
  # Merge pah with filter metadata
  pah_merged <-
    pah_merged %>%
    dplyr::filter(test_type != "blank") %>%
    dplyr::left_join(flows, by = "id") %>%
    dplyr::left_join(times, by = "id") %>%
    dplyr::rename(pre_flow = pre, post_flow = post) %>%
    dplyr::mutate(flow = (pre_flow + post_flow) / 2)
```

```{r}
  # Calculate concentrations
  pah_merged <-
    pah_merged %>%
    dplyr::mutate(dur = (end - start) / 60,
                  conc = val_bk_adj * 1000 / (flow * dur))
```

# Background analysis

```{r}
  # Extract background pufs (front and back) and background filters
  bg <-
    pah_merged %>%
    dplyr::filter(test_type == "bg")
```

## Plot: Lab backgrounds

Background concentrations (ug per meter cubed):

* values given by pollutant and filter type
* color indicates if sample was above or below lod calculated from lab blanks
* mean shown with an x
* median shown with a square

```{r pah_bg_plot, echo=FALSE, fig.height = 15}
  bg %>%
    dplyr::filter(!is.na(conc)) %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    as.factor %>%
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    ggplot(aes_string(x = "filter_type", y = "conc", colour = "bk_lod")) +
      geom_point(size = 3, shape = 1, stroke = 1) +
      geom_point(stat = "summary", fun.y = "mean", colour = "black",
                 shape = 4, size = 4, stroke = 2) +
      geom_point(stat = "summary", fun.y = "median", colour = "black",
                 shape = 0, size = 4, stroke = 2) +
      theme_bw() +
      theme(axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            text = element_text(size = 20),
            strip.text.x = element_text(size = 12),
            legend.position = "top") +
      labs(colour = "cookstove lab blank lod") +
      facet_wrap(~pol, ncol = 4, scales = "free") +
      ylab("") +
      xlab("")
```

## Table: Lab backgrounds below lod for chemical analysis

Lab backgrounds below the lod of chemical analysis (%):

* values given by pollutant and filter type

```{r}
  bg %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of backgrounds below lod for chemical analysis: `r round(nrow(filter(bg, lod == "below")) / nrow(bg), 2) * 100`%.

## Table: Lab backgrounds with NA values

Lab backgrounds with NA values (%):

* values given by pollutant and filter type

```{r}
  bg %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val_lod_adj) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of backgrounds with NA values: `r round(nrow(filter(bg, is.na(val_lod_adj) == TRUE))/nrow(bg), 2) * 100)`%.

## Table: Lab backgrounds below lab blank lod

Backgrounds below lab blank lod (%):

* values given by pollutant and filter type

```{r}
  bg %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., bk_lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of backgrounds below blank lod: `r round(nrow(filter(bg, bk_lod == "below")) / nrow(bg), 2) * 100`%.

## Table: Lab backgrounds below lab blank loq

Backgrounds below lab blank loq (%):

* values given by pollutant and filter type

```{r}
  bg %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "ba")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., bk_loq == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of backgrounds below blank loq: `r round(nrow(filter(bg, bk_loq == "below")) / nrow(bg), 2) * 100`%.
* Total percent of 'bad' background values: `r (1 - round(nrow(filter(bg, lod == "above", bk_lod == "above", bk_loq == "above", is.na(val_lod_adj) == FALSE)) / nrow(bg), 2)) * 100`%.

```{r}
  bg <-
    bg %>%
    dplyr::group_by(pol, filter_type) %>%
    dplyr::summarise(bg_mean = mean(conc, na.rm = TRUE),
                     bg_median = median(conc, na.rm = TRUE),
                     bg_sd = sd(conc, na.rm = TRUE))
```

```{r, include=FALSE}
  # Summary statistics of lab backgrounds
  bg %>%
    dplyr::mutate(filter_type = 
                  filter_type %>% 
                  forcats::fct_recode("back puf" = "b",
                                      "front puf" = "f",
                                      "filter" = "p")) %>%
    knitr::kable(digits = 2, align = 'c',
                 col.names = c("type of pah", "filter type", "mean", "median", "standard deviation"))
```

# Sample analysis

```{r}
  # Extract sample pufs (front and back) and sample filters
  pah_samples <-
    pah_merged %>%
    dplyr::filter(test_type == "test")
```

```{r}
  # Subtract background concentrations
  pah_samples <-
    pah_samples %>%
    dplyr::left_join(bg, by = c("pol", "filter_type")) %>%
    dplyr::mutate(conc = conc - bg_mean)
```

## Plot: Lab burns

Burn concentrations (ug per meter cubed):

* values given by pollutant and filter type
* negative values indicate the sample was below the background
* mean shown with an x
* median shown with a square

```{r, echo=FALSE, fig.height = 15}
  pah_samples %>%
    dplyr::filter(!is.na(conc)) %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    as.factor %>%
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    ggplot(aes_string(x = "filter_type", y = "conc", colour = "bk_lod")) +
      geom_point(size = 3, shape = 1, stroke = 1) +
      geom_point(stat = "summary", fun.y = "mean", colour = "black",
                 shape = 4, size = 4, stroke = 2) +
      geom_point(stat = "summary", fun.y = "median", colour = "black",
                 shape = 0, size = 4, stroke = 2) +
      theme_bw() +
      theme(axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            text = element_text(size = 20),
            strip.text.x = element_text(size = 12),
            legend.position = "top") +
      labs(colour = "lab blank lod") +
      facet_wrap(~pol, ncol = 4, scales = "free") +
      xlab("") +
      ylab("")
```

## Plot: Summed gas and particle phase concentrations

Total particle and gas phase pah concentrations (ug per meter cubed):

* grouped by stove type and colored by fuel category
* negative values indicate the sample was below the background
* mean shown with an x
* median shown with a square

```{r, echo=FALSE, fig.height = 60, fig.width = 16}
  pah_samples %>%
    dplyr::filter(!is.na(conc), !grepl("27|28|29|30", id)) %>%
    dplyr::group_by(id, pol, filter_type) %>%
    dplyr::summarise(conc = sum(conc)) %>%
    dplyr::left_join(dplyr::select(samples, id,
                                   stove, fuel,
                                   stovecat, fuelcat), by = "id") %>%
    dplyr::mutate(stove = 
                    stove %>% 
                    factor %>% 
                    tolower %>%
                    gsub(" [:(:].*[:):]", "",.) %>%
                    forcats::fct_recode("lpg" = "pressurized lpg") %>%
                    forcats::fct_recode("pressure kerosene" =
                                          "pressurized kerosene") %>% 
                    gsub(" ", " \n ",.) %>% 
                    forcats::fct_relevel("three \n stone \n fire",
                                         "clay \n ring",
                                         "ceramic \n charcoal",
                                         "rocket \n elbow",
                                         "built-in \n justa",
                                         "metal \n charcoal",
                                         "fan \n rocket \n elbow",
                                         "gasifier",
                                         "wick \n kerosene",
                                         "pressure \n kerosene",
                                         "lpg"),
                  fuelcat =
                    fuelcat %>% 
                    factor %>%
                    tolower,
                  stovecat = 
                    stovecat %>%
                    forcats::fct_recode("insulated [improved] stoves" = "improved",
                                        "insulated [improved] stoves" = "gasifiers",
                                        "liquid-fuel stoves" = "advanced",
                                        "traditional stoves" = "traditional")) %>%
    ggplot(aes_string(x = "stove", y = "conc", colour = "fuelcat")) +
      geom_point(size = 3, shape = 1, stroke = 2) +
      geom_point(stat = "summary", fun.y = "mean", colour = "black",
                 shape = 4, size = 4, stroke = 2) +
      geom_point(stat = "summary", fun.y = "median", colour = "black",
                 shape = 0, size = 4, stroke = 2) +
      theme_bw() +
      theme(text = element_text(size = 22),
            legend.position = "top") +
      labs(colour = "fuel category") +
      facet_wrap(pol~stovecat, ncol = 3, scales = "free") +
      ylab("") +
      xlab("")
```

## Plot: Ratio of particle phase pahs to total pahs

Ratio of particle phase pahs to total pahs (particle + gas):

* grouped by stove type and colored by fuel category
* mean shown with an x
* median shown with a square

```{r, echo=FALSE, fig.height = 60, fig.width = 16}
  pah_samples %>%
    dplyr::filter(!is.na(conc), !grepl("27|28|29|30", id)) %>%
    dplyr::mutate(filter_type = 
                    filter_type %>%
                    forcats::fct_recode("gas" = "b",
                                        "gas" = "f",
                                        "particle" = "p")) %>%
    dplyr::group_by(id, pol, filter_type) %>%
    dplyr::summarise(conc = sum(conc)) %>%
    tidyr::spread(filter_type, conc) %>%
    dplyr::mutate(sigma = particle / (particle + gas)) %>%
    dplyr::left_join(dplyr::select(samples, id,
                                   stove, fuel,
                                   stovecat, fuelcat), by = "id") %>%
    dplyr::mutate(stove = 
                    stove %>% 
                    factor %>% 
                    tolower %>%
                    gsub(" [:(:].*[:):]", "",.) %>%
                    forcats::fct_recode("lpg" = "pressurized lpg") %>%
                    forcats::fct_recode("pressure kerosene" =
                                          "pressurized kerosene") %>% 
                    gsub(" ", " \n ",.) %>% 
                    forcats::fct_relevel("three \n stone \n fire",
                                         "clay \n ring",
                                         "ceramic \n charcoal",
                                         "rocket \n elbow",
                                         "built-in \n justa",
                                         "metal \n charcoal",
                                         "fan \n rocket \n elbow",
                                         "gasifier",
                                         "wick \n kerosene",
                                         "pressure \n kerosene",
                                         "lpg"),
                  fuelcat =
                    fuelcat %>% 
                    factor %>%
                    tolower,
                  stovecat = 
                    stovecat %>%
                    forcats::fct_recode("insulated [improved] stoves" = "improved",
                                        "insulated [improved] stoves" = "gasifiers",
                                        "liquid-fuel stoves" = "advanced",
                                        "traditional stoves" = "traditional")) %>%
    ggplot(aes_string(x = "stove", y = "sigma", colour = "fuelcat")) +
      geom_point(size = 3, shape = 1, stroke = 2) +
      geom_point(stat = "summary", fun.y = "mean", colour = "black",
                 shape = 4, size = 4, stroke = 2) +
      geom_point(stat = "summary", fun.y = "median", colour = "black",
                 shape = 0, size = 4, stroke = 2) +
      theme_bw() +
      theme(text = element_text(size = 22),
            legend.position = "top") +
      labs(colour = "fuel category") +
      facet_wrap(pol~stovecat, ncol = 3, scales = "free") +
      ylab("") +
      xlab("")
```

## Table: Burn samples below lod for chemical analysis

Burn samples below the lod of chemical analysis (%):

* values given by pollutant and filter type

```{r}
  pah_samples %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of samples below lod for chemical analysis: `r round(nrow(filter(pah_samples, lod == "below")) / nrow(pah_merged), 2) * 100`%.

## Table: Burn samples with NA values

Burn samples with NA values (%):

* values given by pollutant and filter type

```{r}
  pah_samples %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val_lod_adj) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of samples with NA values: `r round(sum(is.na(pah_samples$conc)) / nrow(pah_merged), 2) * 100`%.

## Table: Burn samples below background

Burn samples below the background (%):

* values given by pollutant and filter type

```{r}
  pah_samples %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., conc <= 0))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of samples that are negative after background correction: `r round(sum(pah_samples$conc <= 0, na.rm = TRUE) / nrow(pah_merged), 3) * 100`%.

## Table: Burn samples below lab blank lod

Burn samples below the lab blank lod (%):

* values given by pollutant and filter type

```{r}
  pah_samples %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                        "front puf" = "f",
                                        "filter" = "p")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., bk_lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of samples below blank lod: `r round(nrow(filter(pah_samples, bk_lod == "below")) / nrow(pah_merged), 2) * 100`%.

## Table: Burn samples below lab blank loq

Burn samples below lab blank loq (%):

* values given by pollutant and filter type

```{r}
  pah_samples %>%
    dplyr::mutate(filter_type = 
                    filter_type %>% 
                    forcats::fct_recode("back puf" = "b",
                                       "front puf" = "f",
                                        "filter" = "p")) %>%
    dplyr::group_by(filter_type, pol) %>%
    dplyr::do(val = round(nrow(filter(., bk_loq == "below"))/nrow(.), 2) * 100) %>%
    plyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    tidyr::spread(filter_type, val) %>%
    dplyr::rename("pah type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

* Total percent of samples below blank loq: `r round(nrow(filter(pah_samples, bk_loq == "below")) / nrow(pah_merged), 2) * 100`%.
* Total percent of 'bad' sample values: `r (1 - round(nrow(filter(pah_samples, lod == "above", bk_lod == "above", bk_loq == "above", is.na(conc) == FALSE, conc > 0)) / nrow(pah_merged), 2)) * 100`%.

# Additional analysis

## Ratio of front to back filter

Ratio of front to back filter:

* any values less than one could indicate break through
* not sure how to deal with cases of repeat measurements (could be causing some of these problems)

```{r, echo=FALSE, fig.height = 8}
  # pah_merged %>%
  #   dplyr::mutate(complete.cases(.)) %>%
  #   dplyr::filter(filter_type != "p") %>%
  #   #dplyr::group_by(id, filter_type, pol, test_type) %>%
  #   #dplyr::summarise(conc = mean(conc, na.rm = TRUE)) %>%
  #   #dplyr::ungroup() %>%
  #   dplyr::select(id, filter_type, pol, conc, test_type) %>%
  #   tidyr::spread(filter_type, conc) %>%
  #   dplyr::mutate(ratio = f / b) %>%
  #   ggplot(aes_string(x = "pol", y = "ratio", colour = "test_type")) +
  #     geom_point(size = 3, shape = 1, stroke = 2) +
  #     scale_y_log10() +
  #     theme_bw() +
  #     theme(text = element_text(size = 20),
  #           legend.position = "top",
  #           axis.text.x = element_text(size = 12, angle = 45, hjust = 1, vjust = 1)) +
  #     labs(colour = "test type") +
  #     ylab("") +
  #     xlab("")
```

## Histogram of all data by pollutant and filter type

Blank, test, and background data (ug):

* x-axis on a log base 10 scale

```{r, echo=FALSE, fig.height = 60}
  pah %>%
    dplyr::filter(!is.na(val_lod_adj)) %>%
    dplyr::mutate(filter_type = 
                   filter_type %>% 
                   as.factor %>%
                   forcats::fct_recode("back puf" = "b",
                                       "front puf" = "f",
                                       "filter" = "p",
                                       "filter" = "ba")) %>%
    ggplot(aes_string("val_lod_adj", fill = "test_type")) +
      geom_histogram(bins = 25, stroke = 0,
                     position = "dodge") + 
      theme_bw() +
      scale_x_log10() +
      ylab("") +
      xlab("") + 
      theme(text = element_text(size = 20),
            legend.position = "top") +
      labs(fill = "test type") +
      facet_wrap(pol~filter_type, ncol = 3, scales = "free")
```

```{r}
  # Save merged files
  # saveRDS(pah_merged, file = "../r_files/pah_merged.RDS")
  # saveRDS(bg, file = "../r_files/pah_bg.RDS")
```
