---
title: "pax"
author: "Nicholas Good"
date: "11/21/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")    # sample info
```

## Load pax data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/pax.Rda")    # pax dataset
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(pax)
```

## Load flow data and test times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/cal_2.Rda")         # measured flow
  load("../r_files/test_times.Rda")    # test times and ids
```

Extract measured flows.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  pax_flows <- tidy.date(cal_2, "^preflow_pax|^postflow_pax", "")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(pax_flows,2)
```

Split variable to get measurement type, location and replicate information.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  pax_flows <- split.paxflows(pax_flows)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(pax_flows,2)
```

## Average measured flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # select inlet flow (others not required)
    pax_flows <- dplyr::filter(pax_flows, loc == "inlet")
    
  # group by date and type, and then order
    pax_flows <- dplyr::arrange(group_by(pax_flows, date, type), date)
  
  # calculate average and sdev of measured flows
    pax_flows <- dplyr::mutate(pax_flows, mean = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE))
    
  # drop duplicates
    pax_flows <- dplyr::select(filter(pax_flows, rep == 1), -rep, -loc, -value)
    
  # spread
    pax_flows <- gather(pax_flows, temp, val, mean, sd) %>% 
              unite(temp1, type, temp, sep = "_") %>% 
              spread(temp1, val)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
    head(pax_flows,2)
```

## Append measured flows to pax data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
    pax <- dplyr::left_join(pax, pax_flows, "date")
```

## Map experiment id to pax data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  pax_id <-  map.id(pax, test_times)
  
  pax_noid <- dplyr::setdiff(pax, select(pax_id, -id))

  pax_noid$id <- NA

  pax_merged <- dplyr::arrange(rbind(pax_id, pax_noid), datetime) # merged file is too long and contains duplicate rows and datetime for unknown reasons
```

## QC

Load notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")    # notes
  notes <- dplyr::filter(notes, grepl("^pax$|^all$", notes$inst) == TRUE)
```

Flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc <- "bad"
  notes$qc[1:9] <- "maybe"
  notes$qc[15:16] <- "maybe"
  notes$qc[17:22] <- "ok"
  notes$qc[23:27] <- "maybe"
```

Merge pax with qc flags

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  pax_merged <- dplyr::left_join(pax_merged, select(notes, id, qc), by = "id")
  pax_merged$id <- as.factor(pax_merged$id)
  pax_merged$qc <- ifelse(is.na(pax_merged$qc), "ok", as.character(pax_merged$qc))
  pax_merged$qc <- as.factor(pax_merged$qc)
```

Addtional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  pax_merged$qc[pax_merged$id == ""] <- "bad"
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(pax_merged, file="../r_files/pax_merged.Rda")
```

## Summary

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=100}
  p_data <- filter(pax_merged, !is.na(id) & mode == 0)
  ggplot(p_data) +
    geom_line(mapping = aes(datetime, babs_1_mm), colour = "black") +
    geom_line(mapping = aes(datetime, bscat_1_mm), colour = "royalblue2") +
    facet_wrap(~id, ncol = 2, scales = "free") + 
    theme_minimal() +
    labs(x = "", y = expression(Mm^{-1}))
```


## Plot QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=100}
  p_data <- filter(pax_merged, !is.na(id) & mode == 0)
  ggplot(p_data) +
    geom_line(mapping = aes(datetime, babs_1_mm, colour = qc)) +
    geom_line(mapping = aes(datetime, bscat_1_mm, colour = qc)) +
    facet_wrap(~id, ncol = 2, scales = "free") + 
    theme_minimal() +
    labs(x = "", y = expression(Mm^{-1}))
```

## Data summary

The $PAX$ measured during `r length(unique(pax_merged$id))` experiments between `r min(pax_merged$date, na.rm = TRUE)` and `r max(pax_merged$date, na.rm = TRUE)`. There is no $PAX$ data for tests: `r setdiff(levels(samples$id), levels(pax_merged$id))`.


