---
title: "Data file summary"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
  library(tidyverse)
  library(GGally)
  library(gridExtra)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
  source("../r_scripts/R_functions.R")
  source("../r_scripts/R_plots.R")
  source("../r_scripts/R_tidy.R")
```  
 
# Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  load("../r_files/pax.Rda")         # pax data
  load("../r_files/co2_lab.Rda")     # lab co2 levels
  load("../r_files/co2_sample.Rda")  # diluted sample line co2 levels
  load("../r_files/scale.Rda")       # scale data
  load("../r_files/temp.Rda")        # temperature data
  load("../r_files/smps.Rda")        # smps data

  load("../r_files/ecoc.Rda")        # ecoc analysis
  load("../r_files/ions.Rda")        # ion and carbonyls
  load("../r_files/trans.Rda")       # transmissometer data
  load("../r_files/voc.Rda")         # voc canister data

  load("../r_files/samples.Rda")     # sample info
  load("../r_files/batch.Rda")       # batch sampling log
  load("../r_files/wood.Rda")        # wood sampling log
  load("../r_files/cal_1.Rda")       # meta data
  load("../r_files/data_1.Rda")      # meta data
  load("../r_files/cal_2.Rda")       # meta data
  load("../r_files/data_2.Rda")      # meta data
```

# Stove-fuel combinations

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  test_list <- dplyr::filter(samples, type == "SF") %>%
               dplyr::select(id, stove, fuel, type) %>%
               dplyr::group_by(stove, fuel) %>%
               dplyr::summarise(id = paste(id, collapse = ","))

  ggplot(test_list, aes(y=stove, x=fuel)) + 
    geom_tile(colour = "white", width= 0.9, height = 0.9, aes(fill = id)) +
    scale_fill_discrete(na.value = 'grey95') +
    theme_minimal() +
    theme(legend.position="none") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95)) +
    theme(axis.text = element_text(size = 14)) +
    geom_text(aes(label = id, size = 8)) +
    xlab("") + ylab("")
```

# Clean and save metadata

## batch stove weights

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  batch_wgts <- dplyr::filter(tidy_id_date(batch, "^wgt_", "wgt_"), !is.na(value))

  save(batch_wgts, file="../r_files/batch_wgts.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(batch_wgts, var) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  ggplot(p_data, aes(x = factor(var), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 4) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

## batch stove timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  batch_times <- dplyr::filter(tidy_id_date(batch, "^time_", "time_"), !is.na(value))

  save(batch_times, file="../r_files/batch_times.Rda")

```
  
 * unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(batch_times, value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

Outliers shown on graph (18A, 16B, 17D, 23D, 17A, 25B, 25C, 16C) were all  checked by KF in February 2017; no transcription errors exist nor unusual (out of expected range) values.

## batch lab conditions

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  batch_lab <- dplyr::filter(tidy_id_date(batch, "^lab_", "lab_"), !is.na(value))

  save(batch_lab, file="../r_files/batch_lab.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=12, fig.height=6}
  p_data <- dplyr::group_by(batch_lab, var) %>%
            dplyr::mutate(value_norm =
                         (value - mean(value, na.rm = TRUE)) / sd(value, na.rm = TRUE),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  p_box <- ggplot(p_data, aes(x = factor(var), y = value_norm)) +
           geom_boxplot() +
           geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
           theme_minimal() +
           ylab("z score normalized value") +
           xlab("") +
           theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
           theme(axis.text.y = element_text(size = 30),
           axis.title=element_text(size=40))

  p_hist <- ggplot(p_data, aes(x = value)) +
            geom_histogram(binwidth = 15) +
            theme_minimal() +
            facet_wrap(~var, scales = "free") +
            xlab("")

  grid.arrange(p_hist, p_box, ncol = 2)
```

Outliers shown on graph (22A, 16C, 23A [p] / 29A, 30A [rh] / 25A, 25B [temperature]) were all  checked by KF in February 2017; no transcription errors exist nor unusual (out of expected range) values.

## batch pot number

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  batch_pot <- dplyr::filter(tidy_id_date(batch, "^pot_", "pot_"),
                             !is.na(as.numeric(value))) %>%
               dplyr::mutate(value = as.factor(value)) %>%
               dplyr::rename(pot = value, rep = var)

  batch_pot$pot <- forcats::fct_recode(batch_pot$pot,
                                   a = "1", b = "2", c = "3", d = "4")

  save(batch_pot, file="../r_files/batch_pot.Rda")
```

## carbonyl flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  carb_flows <- dplyr::filter(tidy_id_date(data_2, "_carb_.*[^avg]$", "_carb_"), !is.na(value))
  
  carb_flows <- split_flows(carb_flows)
  
  save(carb_flows, file="../r_files/carb_flows.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(carb_flows, type) %>%
            dplyr::mutate(value_norm =
                           (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value),
                                    as.character(id), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

Outlier 16B checked by KF in February 2017. There is no transcription error, but this measure is QC-flagged as bad due to an issue with the carbonyl line being clogged/having known sampling issues that impacted flow rates. 

## carbonyl timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  carb_times <- dplyr::filter(tidy_id_date(data_2, "^time_.*carb$", "^time_"), !is.na(value))
  
  carb_times <- split_times(carb_times)
  
  save(carb_times, file="../r_files/carb_times.Rda")
```

* unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(batch_times, value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

## isokinetic sampler flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  iso_flows <- dplyr::filter(tidy_id_date(data_2, "_iso_.*[^avg]$", "_iso_"), !is.na(value))
  
  iso_flows <- split_flows(iso_flows)
  
  save(iso_flows, file="../r_files/iso_flows.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(iso_flows, type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

Outlier 1C checked by KF in February 2017. There is no transcription error and the value is within an expected normal range. 

## wood stove weights

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  wood_wgts <- dplyr::filter(tidy_id_date(wood, "^wgt_", "wgt_"), !is.na(value))
  
  save(wood_wgts, file="../r_files/wood_wgts.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(wood_wgts, var) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  ggplot(p_data, aes(x = factor(var), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 4) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

Outliers 10A, 6D, 5A, 2A, 1A checked by KF in February 2017. There are no transcription errors. 

Outliers 6A and 13A were transcription errors that were fixed in wood sampling log on 2/18/17 so should not appear in subsequent re-run versions of this file.

## wood stove times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  wood_times <- dplyr::filter(tidy_id_date(wood, "^time_", "time_"), !is.na(value))

  save(wood_times, file="../r_files/wood_times.Rda")
```

 * unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(wood_times, value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```


## wood lab conditions

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  wood_lab <- dplyr::filter(tidy_id_date(wood, "^lab_", "lab_"), !is.na(value))
  
  save(wood_lab, file="../r_files/wood_lab.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=12, fig.height=6}
  p_data <- dplyr::group_by(wood_lab, var) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  p_box <- ggplot(p_data, aes(x = factor(var), y = value_norm)) +
           geom_boxplot() +
           geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
           theme_minimal() +
           ylab("z score normalized value") +
           xlab("") +
           theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
           theme(axis.text.y = element_text(size = 30),
           axis.title=element_text(size=40))

  p_hist <- ggplot(p_data, aes(x = value)) +
            geom_histogram(binwidth = 15) +
            theme_minimal() +
            facet_wrap(~var, scales = "free") +
            xlab("")

  grid.arrange(p_hist, p_box, ncol = 2)
```

All outliers checked by KF in February 2017. There are no transcription errors. 

## wood pot number

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  wood_pot <- dplyr::filter(tidy_id_date(wood, "^pot_", "pot_"),
                             !is.na(as.numeric(value))) %>%
               dplyr::mutate(value = as.factor(value)) %>%
               dplyr::rename(pot = value, rep = var)

  wood_pot$pot <- forcats::fct_recode(wood_pot$pot,
                                   a = "1", b = "2", c = "3", d = "4")
  
  save(wood_pot, file="../r_files/wood_pot.Rda")
```


## fivegas calibration standards

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}    
  fivegas_cal_conc <- dplyr::filter(tidy_date(cal_1, "^conc_", "conc_"), !is.na(value))
  
  save(fivegas_cal_conc, file="../r_files/fivegas_cal_conc.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(fivegas_cal_conc, var) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(var), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

Outlier date of 3-11-16: No transcription error, but KF thinks there was a mistake in the original data recording. It should have been written as 49.9 but instead was written as 49.5.

## fivegas calibration times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}    
  fivegas_cal_times <- dplyr::filter(tidy_date(cal_1, "^time_", "time_"),
                                     !is.na(value))

  fivegas_cal_times <- split_fivegas_cal_times(fivegas_cal_times)

  save(fivegas_cal_times, file="../r_files/fivegas_cal_times.Rda")
```

 * unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(fivegas_cal_times,
                            value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

## filter cartridge flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  filter_flows <- dplyr::filter(tidy_id_date(data_1,
                                  "^preflow_.*[^avg]$|postflow.*[^avg]$", ""),
                                  !is.na(value))

  filter_flows <- split_filter_flows(filter_flows)

  save(filter_flows, file="../r_files/filter_flows.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=16}
  p_data <- dplyr::group_by(filter_flows, type, colour) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40)) +
    facet_wrap(~colour, scales = "free")
```

## filter cartridge timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  filter_times <- dplyr::filter(tidy_id_date(data_1, "^time_.*cart", "time_"), !is.na(value))

  filter_times <- split_filter_times(filter_times)

  save(filter_times, file="../r_files/filter_times.Rda")
```

 * unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(filter_times,
                            value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

Outlier 23A was a transcription error (AM vs PM) that was fixed on 2/18/17. G3 is correct (~5-7pm test).

## co2 (dilution) calibration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  co2_lab_cal <- dplyr::filter(tidy_date(cal_2, "^sensor_1", "sensor_1_"), !is.na(value))

  co2_lab_cal <- split_co2_cal(co2_lab_cal)

  save(co2_lab_cal, file="../r_files/co2_lab_cal.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(co2_lab_cal, type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40)) +
    ggtitle("lab")
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  co2_sample_cal <- dplyr::filter(tidy_date(cal_2, "^sensor_2", "sensor_2_"), !is.na(value))
  
  co2_sample_cal <- split_co2_cal(co2_sample_cal)
  
  save(co2_sample_cal, file="../r_files/co2_sample_cal.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(co2_sample_cal, type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40)) +
    ggtitle("sample line")
```

## pax flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  pax_flows <- dplyr::filter(tidy_date(cal_2, "^preflow_pax|^postflow_pax", ""),
                             !is.na(value))
  
  pax_flows <- split_pax_flows(pax_flows)
  
  save(pax_flows, file="../r_files/pax_flows.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::filter(pax_flows, loc == "inlet") %>%
            dplyr::group_by(type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

## smps flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  smps_flows <- dplyr::filter(tidy_date(cal_2, "^preflow_smps|^postflow_smps", ""),
                              !is.na(value))

  smps_flows <- split_flows(smps_flows)
  
  save(smps_flows, file="../r_files/smps_flows.Rda")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(smps_flows, type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

## smps / pax times


```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  smps_pax_bg_times <- dplyr::filter(tidy_id_date(data_2,
                                                  "^time_.*smps.*", "^time_"),
                                                  !is.na(value))

  smps_pax_bg_times <- split_times_smps_pax(smps_pax_bg_times)

  save(smps_pax_bg_times, file="../r_files/smps_pax_bg_times.Rda")
```

* unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(smps_pax_bg_times,
                            value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

# Summary

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  count_samples <- dplyr::tally(group_by(samples, type))

  count_stoves <- dplyr::tally(group_by(samples, stove)) 
  
  count_fuels <- dplyr::tally(group_by(samples, fuel))
```

A total of `r (as.numeric(count_samples[1,2]) + as.numeric(count_samples[2,2]))` tests were performed between `r min(samples$date)` and `r max(samples$date)`. A total of `r nrow(count_stoves)` stoves types were tested. A total of `r nrow(count_fuels)` fuels were tested (on the appropriate stove).
