---
title: "Data file summary"
author: "Nicholas Good"
date: "11/10/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
# libraries
  library(tidyverse)
  library(GGally)

# source files
  source("../r_scripts/R_functions.R")
  source("../r_scripts/R_plots.R")
  source("../r_scripts/R_tidy.R")
```  
  
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=TRUE}
# load time-resolved data data
  load("../r_files/pax.Rda")          # pax data
  load("../r_files/co2_lab.Rda")      # lab co2 levels
  load("../r_files/co2_sample.Rda")   # sample co2 levels
  load("../r_files/scale.Rda")        # scale data
  load("../r_files/temp.Rda")         # temperature data
  load("../r_files/smps.Rda")         # smps data
  
# load time-integrated data
  load("../r_files/ecoc.Rda")         # ecoc filter analysis
  load("../r_files/ions.Rda")         # ion filter data
  load("../r_files/trans.Rda")        # transmissometer data
  load("../r_files/voc.Rda")          # voc canister data
  load("../r_files/samples.Rda")      # sample info
  load("../r_files/batch.Rda")        # batch sampling log
  load("../r_files/wood.Rda")         # wood sampling log
  load("../r_files/cal_1.Rda")        # meta data
  load("../r_files/data_1.Rda")       # meta data
  load("../r_files/cal_2.Rda")        # meta data
  load("../r_files/data_2.Rda")       # meta data
```

## Clean and save metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=TRUE}
# tidy by prefix and retain id and date

# batch test weights
  batch_wgts <- dplyr::filter(tidy.id.date(batch, "^wgt_", "wgt_"), !is.na(value))
    save(batch_wgts, file="../r_files/batch_wgts.Rda")

# batch test times   
  batch_times <- dplyr::filter(tidy.id.date(batch, "^time_", "time_"), !is.na(value))
    save(batch_times, file="../r_files/batch_times.Rda")

# lab pressure, temperature and relative humidity
  batch_lab <- dplyr::filter(tidy.id.date(batch, "^lab_", "lab_"), !is.na(value))
    save(batch_lab, file="../r_files/batch_lab.Rda")
    
# flow for ion and carbonyl filters
  carb_flows <- dplyr::filter(tidy.id.date(data_2, "_carb_.*[^avg]$", "_carb_"), !is.na(value))
    carb_flows <- split.flows(carb_flows)
    save(carb_flows, file="../r_files/carb_flows.Rda")
  
# times for ion and carbonyl filters
  carb_times <- dplyr::filter(tidy.id.date(data_2, "^time_.*carb$", "^time_"), !is.na(value))
    carb_times <- split.times(carb_times)
    save(carb_times, file="../r_files/carb_times.Rda")
 
# flow for isokinetic sampler
  iso_flows <- dplyr::filter(tidy.id.date(data_2, "_iso_.*[^avg]$", "_iso_"), !is.na(value))
    iso_flows <- split.flows(iso_flows)
    save(iso_flows, file="../r_files/iso_flows.Rda")
    
# wood test weights  
  wood_wgts <- dplyr::filter(tidy.id.date(wood, "^wgt_", "wgt_"), !is.na(value))
    save(wood_wgts, file="../r_files/wood_wgts.Rda")
  
# wood test times
  wood_times <- dplyr::filter(tidy.id.date(wood, "^time_", "time_"), !is.na(value))
    save(wood_times, file="../r_files/wood_times.Rda")

# lab pressure, temperature and relative humidity
  wood_lab <- dplyr::filter(tidy.id.date(wood, "^lab_", "lab_"), !is.na(value))
    save(wood_lab, file="../r_files/wood_lab.Rda")
    
# Tidy by prefix and retain date
    
# fivegas calibration
    
  fivegas_cal_conc <- dplyr::filter(tidy.date(cal_1, "^conc_", "conc_"), !is.na(value))
    save(fivegas_cal_conc, file="../r_files/fivegas_cal_conc.Rda")
    
  fivegas_cal_times <- dplyr::filter(tidy.date(cal_1, "^time_", "time_"), !is.na(value))
    fivegas_cal_times <- split.fivegas.cal.times(fivegas_cal_times)
    save(fivegas_cal_times, file="../r_files/fivegas_cal_times.Rda")
    
# filters
  filter_flows <- dplyr::filter(tidy.date(data_1, "^preflow_.*[^avg]$|postflow.*[^avg]$", ""), !is.na(value))
    filter_flows <- split.filter.flows(filter_flows)
    save(filter_flows, file="../r_files/filter_flows.Rda")
    
  filter_times <- dplyr::filter(tidy.date(data_1, "^time_.*cart", "time_"), !is.na(value))
    filter_times <- split.filter.times(filter_times)
    save(filter_times, file="../r_files/filter_times.Rda")
    
# co2
  co2_lab_cal <- dplyr::filter(tidy.date(cal_2, "^sensor_1", "sensor_1_"), !is.na(value))
    co2_lab_cal <- split.co2.cal(co2_lab_cal)
    save(co2_lab_cal, file="../r_files/co2_lab_cal.Rda")
    
  co2_sample_cal <- dplyr::filter(tidy.date(cal_2, "^sensor_2", "sensor_2_"), !is.na(value))
    co2_sample_cal <- split.co2.cal(co2_sample_cal)
    save(co2_sample_cal, file="../r_files/co2_sample_cal.Rda")
    
# pax
  pax_flows <- dplyr::filter(tidy.date(cal_2, "^preflow_pax|^postflow_pax", ""), !is.na(value))
    pax_flows <- split.pax.flows(pax_flows)
    save(pax_flows, file="../r_files/pax_flows.Rda")
    
# smps
  smps_flows <- dplyr::filter(tidy.date(cal_2, "^preflow_smps|^postflow_smps", ""), !is.na(value))
    smps_flows <- split.flows(smps_flows)
    save(smps_flows, file="../r_files/smps_flows.Rda")
```

## Stove-fuel combinations

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  count_samples <- dplyr::tally(group_by(samples, type))

  count_stoves <- dplyr::tally(group_by(samples, stove)) 
  
  count_fuels <- dplyr::tally(group_by(samples, fuel))
```

A total of `r (as.numeric(count_samples[1,2]) + as.numeric(count_samples[2,2]))` tests were performed between `r min(samples$date)` and `r max(samples$date)`. A total of `r nrow(count_stoves)` stoves types were tested using `r nrow(count_fuels)` fuels. The graphic below, shows the stove-fuel combinations tested. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=TRUE, fig.width=20, fig.height=14}
    # remove columns
      stove_fuel_list <- subset(complete(subset(samples, select = c(stove,fuel, id)), stove, fuel),
                              stove != "Background" & fuel != "Background")

    # remove deplicate combinations
      stove_fuel_combos <- distinct(select(stove_fuel_list, -id))
      stove_fuel_combos$id <- NA
        for(i in 1:nrow(stove_fuel_combos)){
          reps <- subset(stove_fuel_list, 
                    fuel == stove_fuel_combos$fuel[i] &
                    stove  == stove_fuel_combos$stove[i],
                    select = id)
          stove_fuel_combos$id[i] <- paste0(reps$id,collapse="", sep = ".")
        }
    # type
      stove_fuel_combos$id <- as.factor(stove_fuel_combos$id)
      
    # relevel NAs
      levels(stove_fuel_combos$id)[levels(stove_fuel_combos$id)=="NA."] <- NA
      
    # plot
      p <- ggplot(stove_fuel_combos, aes(y=stove, x=fuel)) + 
            geom_tile(colour="white",width=.9, height=.9,aes(fill = id)) +
            scale_fill_discrete(na.value = 'grey95') +
            theme_minimal() +
            theme(legend.position="none") +
            theme(axis.text.x = element_text(angle = 35, hjust = 0.95)) +
            theme(axis.text = element_text(size = 18)) +
            geom_text(aes(label = stove_fuel_combos$id)) +
            xlab("") + ylab("")
      p
``` 

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  head(stove_fuel_combos,2)
```

## Batch sample weights

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', fig.width=20, fig.height=10, cache=FALSE}
  df <- batch_wgts
  df <- filter(df, !is.na(value))
  df <- dplyr::group_by(df, var) %>% 
        dplyr::mutate(outlier = ifelse(is_outlier(value), as.character(id), NA))
  
  p <-  ggplot(df, aes(x = var, y = value)) +
        geom_boxplot() +
        theme_bw() +
        theme(axis.text = element_text(size = 14)) +
        geom_text(aes(label = outlier), na.rm = TRUE, nudge_y = 0, nudge_x = 0.3, size = 4) +
        xlab("") + ylab("")
  
  p
```

## Batch fuel schedule

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', fig.width=20, fig.height=10, cache=FALSE}
  df <- subset(batch_times, grepl("[a-z]$", batch_times$var))
  df$value <- df$value/60
  ggplot(df, aes(x=id, y=value, fill=var)) +
          geom_bar(stat = "identity") +
          xlab("") + ylab("duration (minutes") +
          theme(axis.text = element_text(size = 14))

```

## Batch pot schedule

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', fig.width=20, fig.height=10, cache=FALSE}
  df <- subset(batch_times, grepl("[0-9]$", batch_times$var))
  df$value <- df$value/60
  ggplot(df, aes(x=id, y=value, fill=var)) +
          geom_bar(stat = "identity") +
          xlab("") + ylab("duration (minutes") +
          theme(axis.text = element_text(size = 14))
```

## Batch lab conditions

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', fig.width=20, fig.height=10, cache=FALSE}
  p <- ggplot(batch_lab, aes(x=date, y=value)) + 
        geom_line() +
        geom_point() +
        facet_wrap(~var, ncol = 1, scales = "free")
  p
```
