---
title: "Data file summary"
author: "Nicholas Good"
date: "11/10/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
  library(tidyverse)
  library(GGally)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
  source("../r_scripts/R_functions.R")
  source("../r_scripts/R_plots.R")
  source("../r_scripts/R_tidy.R")
```  
  
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=TRUE}
 # time-resolved data data
  load("../r_files/pax.Rda")         # pax data
  load("../r_files/co2_lab.Rda")     # lab co2 levels
  load("../r_files/co2_sample.Rda")  # sample co2 levels
  load("../r_files/scale.Rda")       # scale data
  load("../r_files/temp.Rda")        # temperature data
  load("../r_files/smps.Rda")        # smps data

 # load time-integrated data
  load("../r_files/ecoc.Rda")        # ecoc filter analysis
  load("../r_files/ions.Rda")        # ion filter data
  load("../r_files/trans.Rda")       # transmissometer data
  load("../r_files/voc.Rda")         # voc canister data
  load("../r_files/samples.Rda")     # sample info
  load("../r_files/batch.Rda")       # batch sampling log
  load("../r_files/wood.Rda")        # wood sampling log
  load("../r_files/cal_1.Rda")       # meta data
  load("../r_files/data_1.Rda")      # meta data
  load("../r_files/cal_2.Rda")       # meta data
  load("../r_files/data_2.Rda")      # meta data
```

## Clean and save metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=TRUE}
  batch_wgts <- dplyr::filter(tidy_id_date(batch, "^wgt_", "wgt_"), !is.na(value))

  save(batch_wgts, file="../r_files/batch_wgts.Rda")

  batch_times <- dplyr::filter(tidy_id_date(batch, "^time_", "time_"), !is.na(value))
  
  save(batch_times, file="../r_files/batch_times.Rda")

  batch_lab <- dplyr::filter(tidy_id_date(batch, "^lab_", "lab_"), !is.na(value))

  save(batch_lab, file="../r_files/batch_lab.Rda")
    
  carb_flows <- dplyr::filter(tidy_id_date(data_2, "_carb_.*[^avg]$", "_carb_"), !is.na(value))
  
  carb_flows <- split_flows(carb_flows)
  
  save(carb_flows, file="../r_files/carb_flows.Rda")
  
  carb_times <- dplyr::filter(tidy_id_date(data_2, "^time_.*carb$", "^time_"), !is.na(value))
  
  carb_times <- split_times(carb_times)
  
  save(carb_times, file="../r_files/carb_times.Rda")
 
  iso_flows <- dplyr::filter(tidy_id_date(data_2, "_iso_.*[^avg]$", "_iso_"), !is.na(value))
  
  iso_flows <- split_flows(iso_flows)
  
  save(iso_flows, file="../r_files/iso_flows.Rda")
    
  wood_wgts <- dplyr::filter(tidy_id_date(wood, "^wgt_", "wgt_"), !is.na(value))
  
  save(wood_wgts, file="../r_files/wood_wgts.Rda")
  
  wood_times <- dplyr::filter(tidy_id_date(wood, "^time_", "time_"), !is.na(value))
  
  save(wood_times, file="../r_files/wood_times.Rda")

  wood_lab <- dplyr::filter(tidy_id_date(wood, "^lab_", "lab_"), !is.na(value))
  
  save(wood_lab, file="../r_files/wood_lab.Rda")
    
  fivegas_cal_conc <- dplyr::filter(tidy_date(cal_1, "^conc_", "conc_"), !is.na(value))
  
  save(fivegas_cal_conc, file="../r_files/fivegas_cal_conc.Rda")
    
  fivegas_cal_times <- dplyr::filter(tidy_date(cal_1, "^time_", "time_"), !is.na(value))
  
  fivegas_cal_times <- split_fivegas_cal_times(fivegas_cal_times)
  
  save(fivegas_cal_times, file="../r_files/fivegas_cal_times.Rda")
    
  filter_flows <- dplyr::filter(tidy_date(data_1, "^preflow_.*[^avg]$|postflow.*[^avg]$", ""), !is.na(value))
  
  filter_flows <- split_filter_flows(filter_flows)

  save(filter_flows, file="../r_files/filter_flows.Rda")
    
  filter_times <- dplyr::filter(tidy_date(data_1, "^time_.*cart", "time_"), !is.na(value))
  
  filter_times <- split_filter_times(filter_times)
  
  save(filter_times, file="../r_files/filter_times.Rda")
    
  co2_lab_cal <- dplyr::filter(tidy_date(cal_2, "^sensor_1", "sensor_1_"), !is.na(value))

  co2_lab_cal <- split_co2_cal(co2_lab_cal)

  save(co2_lab_cal, file="../r_files/co2_lab_cal.Rda")
    
  co2_sample_cal <- dplyr::filter(tidy_date(cal_2, "^sensor_2", "sensor_2_"), !is.na(value))
  
  co2_sample_cal <- split_co2_cal(co2_sample_cal)
  
  save(co2_sample_cal, file="../r_files/co2_sample_cal.Rda")
    
  pax_flows <- dplyr::filter(tidy_date(cal_2, "^preflow_pax|^postflow_pax", ""), !is.na(value))
  
  pax_flows <- split_pax_flows(pax_flows)
  
  save(pax_flows, file="../r_files/pax_flows.Rda")

  smps_flows <- dplyr::filter(tidy_date(cal_2, "^preflow_smps|^postflow_smps", ""), !is.na(value))
  
  smps_flows <- split_flows(smps_flows)
  
  save(smps_flows, file="../r_files/smps_flows.Rda")
```

## Stove-fuel combinations

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  count_samples <- dplyr::tally(group_by(samples, type))

  count_stoves <- dplyr::tally(group_by(samples, stove)) 
  
  count_fuels <- dplyr::tally(group_by(samples, fuel))
```

A total of `r (as.numeric(count_samples[1,2]) + as.numeric(count_samples[2,2]))` tests were performed between `r min(samples$date)` and `r max(samples$date)`. A total of `r nrow(count_stoves)` stoves types were tested using `r nrow(count_fuels)` fuels. The graphic below, shows the stove-fuel combinations tested. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=TRUE, fig.width=20, fig.height=14}
 # update to dplyr
  stove_fuel_list <- subset(complete(subset(samples, select = c(stove,fuel, id)), stove, fuel),
                              stove != "Background" & fuel != "Background")

  stove_fuel_combos <- dplyr::distinct(select(stove_fuel_list, -id)) %>%
                       dplyr::mutate(id = NA)

  for(i in 1:nrow(stove_fuel_combos)){
    reps <- dplyr::filter(stove_fuel_list, fuel == fuel[i] & stove == stove[i]) %>%
            dplyr::select(id)

    stove_fuel_combos$id[i] <- paste0(reps$id,collapse="", sep = ".")
  }

  stove_fuel_combos <- dplyr::mutate(stove_fuel_combos, id = as.factor(id))

  levels(stove_fuel_combos$id)[levels(stove_fuel_combos$id)=="NA."] <- NA
```

### Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=TRUE, fig.width=20, fig.height=14}      
  p <- ggplot(stove_fuel_combos, aes(y=stove, x=fuel)) + 
              geom_tile(colour="white",width=.9, height=.9,aes(fill = id)) +
              scale_fill_discrete(na.value = 'grey95') +
              theme_minimal() +
              theme(legend.position="none") +
              theme(axis.text.x = element_text(angle = 35, hjust = 0.95)) +
              theme(axis.text = element_text(size = 18)) +
              geom_text(aes(label = stove_fuel_combos$id)) +
              xlab("") + ylab("")
  p
``` 

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=TRUE}
  head(stove_fuel_combos,2)
```

## Batch sample weights

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  df <- batch_wgts
  df <- filter(df, !is.na(value))
  df <- dplyr::group_by(df, var) %>% 
        dplyr::mutate(outlier = ifelse(is_outlier(value), as.character(id), NA))
```

### Plot
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', fig.width=20, fig.height=10, cache=FALSE}
  p <-  ggplot(df, aes(x = var, y = value)) +
               geom_boxplot() +
               theme_bw() +
               theme(axis.text = element_text(size = 14)) +
               geom_text(aes(label = outlier), na.rm = TRUE, nudge_y = 0, nudge_x = 0.3, size = 4) +
               xlab("") + ylab("")
  p
```

## Batch fuel schedule

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  df <- dplyr::filter(batch_times, grepl("[a-z]$", var) == TRUE) %>%
        dplyr::mutate(value = value/60)
```

### Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', fig.width=20, fig.height=10, cache=FALSE}
  ggplot(df, aes(x=id, y=value, fill=var)) +
         geom_bar(stat = "identity") +
         xlab("") + ylab("duration (minutes") +
         theme(axis.text = element_text(size = 14))
```

## Batch pot schedule

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  df <- dplyr::filter(batch_times, grepl("[0-9]$", var) == TRUE) %>%
        dplyr::mutate(value = value/60)
```

### Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', fig.width=20, fig.height=10, cache=FALSE}
  ggplot(df, aes(x=id, y=value, fill=var)) +
         geom_bar(stat = "identity") +
         xlab("") + ylab("duration (minutes") +
         theme(axis.text = element_text(size = 14))
```

## Batch lab conditions

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', fig.width=20, fig.height=10, cache=FALSE}
  p <- ggplot(batch_lab, aes(x=date, y=value)) + 
              geom_line() +
              geom_point() +
              facet_wrap(~var, ncol = 1, scales = "free")
  p
```
