---
title: "scale"
author: "Nicholas Good"
date: "11/23/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

## Load scale data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/scale.Rda")    # scale dataset
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(scale)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")    # sample info
```

## Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=40}
  ggplot(scale, aes(datetime, wgt_stove)) +
  geom_line() +
  geom_point() +
  facet_wrap(~id, ncol = 3, scales = "free") +
  theme_minimal()
```

## QC

Load notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")    # notes
  notes <- dplyr::filter(notes, grepl("^scale$|^all$", notes$inst) == TRUE)
```

Flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc <- "bad"
  notes$qc[1:9] <- "maybe"
  notes$qc[15:16] <- "maybe"
  notes$qc[17:22] <- "ok"
  notes$qc[23:27] <- "maybe"
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  scale_merged <- dplyr::left_join(scale, select(notes, id, qc), by = "id")
  scale_merged$id <- as.factor(scale_merged$id)
  scale_merged$qc <- ifelse(is.na(scale_merged$qc), "ok", as.character(scale_merged$qc))
  scale_merged$qc <- as.factor(scale_merged$qc)
```

Additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # scale_merged$qc[scale_merged$id == ""] <- "bad"
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(scale_merged, file = "../r_files/scale_merged.Rda")
```

## Summary

$Scale$ data was collected during `r length(unique(scale_merged$id))` experiments between `r min(scale_merged$date, na.rm = TRUE)` and `r max(scale_merged$date, na.rm = TRUE)`. There is no $Scale$ data for tests: `r setdiff(levels(samples$id), levels(scale_merged$id))`.