---
title: "scale"
author: "Nicholas Good"
date: "11/23/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

## Load scale data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/scale.Rda")    # scale dataset
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(scale)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")    # sample info
```

## Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=40}
  ggplot(scale, aes(datetime, wgt_stove)) +
         geom_line() +
         geom_point() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         theme_minimal()
```

## QC

Load notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")

  notes <- dplyr::filter(notes, grepl("scale|all", notes$inst) == TRUE)
```

Set one flag per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  scale_merged <- dplyr::left_join(scale, flags, by = "id") %>%
                  dplyr::mutate(id = as.factor(id)) %>%
                  dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

Additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # scale_merged$qc[scale_merged$id == ""] <- "bad"
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(scale_merged, file = "../r_files/scale_merged.Rda")
```

## Summary

$Scale$ data was collected during `r length(unique(scale_merged$id))` experiments between `r min(scale_merged$date, na.rm = TRUE)` and `r max(scale_merged$date, na.rm = TRUE)`. There is no $Scale$ data for tests: `r setdiff(as.character(samples$id), as.character(scale_merged$id))`.

$scale$ data is expected to be missing for:

All "G" background tests: no scale needed as no stove used.

Non-batchfed stoves (e.g., wood stoves): tests 1-15, 24, and 25 (A, B, and C reps)

21A, 22A: these are batch-fed stoves. However, data is known missing. Possible that we did not use the real-time scale on these tests due to conflict with other teams- as long as weight data is available at critical time points as recorded on the transcribed testing logs, the "scale" data is OK to be missing. 
