---
title: "smps"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(gridExtra)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
  source("../r_scripts/R_functions.R")
```

# Load data

* smps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/smps.Rda")
```

* metadata

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
  load("../r_files/cal_2.Rda")
  load("../r_files/notes.Rda")
  load("../r_files/test_times.Rda")
  load("../r_files/smps_pax_bg_times.Rda")
```

# Organize

* extract measured flows.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_flows <- tidy_date(cal_2, "^preflow_smps|^postflow_smps", "")

  smps_flows <- split_flows(smps_flows)
```

* average flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_flows <- dplyr::arrange(group_by(smps_flows, date, type), date) %>%
                dplyr::mutate(mean = mean(value, na.rm = TRUE),
                              sd = sd(value, na.rm = TRUE)) %>%
                dplyr::filter(rep == 1) %>%
                dplyr::select(-rep, -value) %>%
                tidyr::gather(temp, val, mean, sd) %>%
                tidyr::unite(temp1, type, temp, sep = "_") %>%
                tidyr::spread(temp1, val)
```

* merge

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged <- dplyr::left_join(smps, smps_flows, "date") %>%
                 dplyr::mutate(sample = as.factor(sample))
```

# QC

* notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("smps|all", inst))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  knitr::kable(dplyr::filter(notes, grepl(".*smps.*", inst)) %>%
               dplyr::select(-inst, -date),
               "markdown", digits = 2)
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

* merge

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged <- dplyr::left_join(smps_merged, flags, by = "id") %>%
                 dplyr::mutate(id = as.factor(id)) %>%
                 dplyr::mutate(qc = as.factor(ifelse(is.na(qc),
                                                     "ok",
                                                     as.character(qc)))) %>%
                 dplyr::mutate(type = ifelse(grepl("^[0-9]", id), "test", NA),
                               type = ifelse(grepl("^G", id), "bg", type))
```

* error from instrument

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged <- dplyr::mutate(smps_merged,
                               qc = ifelse(grepl("^Normal Scan$", instrument_errors),
                                           as.character(qc),
                                           "bad"))
```

* additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # smps_merged$qc[smps_merged$id == ""] <- "bad"
```

# Ultrafine number

## per scan 

* select sub-100nm data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_ultrafine <- dplyr::select(smps_merged, id, value, size, time, sample, qc) %>%
                    dplyr::filter(size <= 100)
```

* get timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times <- dplyr::filter(test_times, var == "start_1" | var == "shutdown") %>%
           tidyr::spread(var, value) %>%
           dplyr::select(-date) %>%
           dplyr::rename(start = start_1, end = shutdown)
```

* filter by test time window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_ultrafine <- filter_times(times, smps_ultrafine)
```

* calculate number concentration over each scan

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_ultrafine <- dplyr::group_by(smps_ultrafine, id, sample) %>%
                    dplyr::summarise(number_conc = sum(value) * 0.015625355,
                                     n_bad = sum(qc == "bad"),
                                     qc = first(qc))
```

* plot timeseries for each experiment

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80} 
  ggplot(smps_ultrafine, aes(sample, number_conc, colour = qc)) +
         geom_point() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         scale_y_log10() +
         theme_minimal() + 
         ylab("number per cc") +
         theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
         theme(legend.position = "top") +
         scale_colour_manual(values = c("ok" = "mediumaquamarine",
                                        "maybe" = "mediumorchid",
                                        "bad" = "mistyrose4"))
         
```

## test average

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_ultrafine_avg <- dplyr::filter(smps_ultrafine) %>%
                    dplyr::group_by(id) %>%
                    dplyr::summarise(mean_number_conc = mean(number_conc),
                                     qc = first(qc))
```

* plot test average

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8} 
  ggplot(smps_ultrafine_avg, aes(id, mean_number_conc, colour = qc)) +
         geom_point() +
         scale_y_log10() +
         theme_minimal() + 
         ylab("sub 100 nm - mean number per cc") +
         theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
         theme(legend.position = "top") +
         scale_colour_manual(values = c("ok" = "mediumaquamarine",
                                        "maybe" = "mediumorchid",
                                        "bad" = "mistyrose4"))
```

# Background (G tests)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_ultrafine_bg <-  dplyr::filter(smps_merged, size <= 100, type == "bg") %>%
                        dplyr::select(id, value, size, time, sample, qc, type)
```

* reformat time data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_bg <- dplyr::filter(smps_pax_bg_times, grepl("^G", id), type == "sample") %>%
              tidyr::spread(time_point, value) %>%
              dplyr::select(-date)
```

* filter by test time window

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_ultrafine_bg <- filter_times(times_bg, smps_ultrafine_bg)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_ultrafine_bg <- dplyr::group_by(smps_ultrafine_bg, id, sample) %>%
                       dplyr::summarise(number_conc = sum(value) * 0.015625355,
                                        n_bad = sum(qc == "bad"),
                                        qc = first(qc)) %>%
                       dplyr::ungroup()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=10}
  p_id <- ggplot(smps_ultrafine_bg, aes(sample, number_conc, group_by(id), colour = id)) +
            geom_point() +
            theme_minimal() +
            scale_y_log10() +
            theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
            theme(legend.position = "top")

  p_qc <- ggplot(smps_ultrafine_bg, aes(sample, number_conc, group_by(id), colour = qc)) +
            geom_point() +
            theme_minimal() +
            scale_y_log10() +
            theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
            theme(legend.position = "top") +
            scale_colour_manual(values = c("ok" = "mediumaquamarine",
                                           "maybe" = "mediumorchid",
                                           "bad" = "mistyrose4"))

  grid.arrange(p_id, p_qc, ncol = 1)
```

# Save files

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(smps_merged, file = "../r_files/smps_merged.Rda")

  save(smps_ultrafine, file = "../r_files/smps_ultrafine.Rda")

  save(smps_ultrafine_bg, file = "../r_files/smps_ultrafine_bg.Rda")
```

# Summary

The SMPS measured during `r length(unique(smps_merged$id))` experiments between `r min(smps_merged$date, na.rm = TRUE)` and `r max(smps_merged$date, na.rm = TRUE)`. There is no $SMPS$ data for tests: `r setdiff(as.character(samples$id), as.character(smps_merged$id))`.

SMPS data is expected to be missing for: 
13A, 17A, 18A, 19A, 20A, G3, G4. "no data" was noted on QC log for these tests. 

G1, 23A: test date 1/7/2016: data lost from this day. 
29B: test date 6/29/2016; for some reason there are 2 files, part 1 and 2, where file name is 20160622_29B(1)_SMPS and 20160622_29B(2)_SMPS. Part 1 only contains 24 scans, start times from 13:14 through 14:06. Part 2 starts at 14:22, only 8 scans through 14:45 start. Test start, according to log sheet, was 13:33 - electrical problem with stove at 14:14:00 forced test shutdown early. Need to combine files, or it's possible that this whole test should be scrapped? 

28A: test date 6/23/2016; file is there with correct naming convention…only 9 scans from 9:46 through 10:13; log sheet says test time was from 10:02 through 11:28…suspect an issue occurred with SMPS?

11A: test date 7/19/2016; file is there with correct naming convention…42 scans from 9:53 to 12:09 - but test time in log listed as 12:12 through 13:43…suspect an issue occurred with SMPS?

# Plots

## qc (optional)

* all scans

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80, eval=FALSE}
  ggplot(smps_merged, aes(size, value, group = sample, colour = qc)) +
         geom_line() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         theme_minimal() +
         scale_x_log10() + scale_y_log10() +
         ylab("dN/dlogDp") +
         xlab("nm") +
         theme(legend.position = "none")
```
