---
title: "smps"
author: "Nicholas Good"
date: "11/21/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
  source("../r_scripts/R_functions.R")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")    # sample info
```

## Load smps data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/smps.Rda")    # smps dataset
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(smps)
```

## Load flow data and test times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/cal_2.Rda")         # measured flow
```

Extract measured flows.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_flows <- tidy.date(cal_2, "^preflow_smps|^postflow_smps", "")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(smps_flows, 2)
```

Separate measurement location and replicate

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_flows <- split.flows(smps_flows)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(smps_flows, 2)
```

## Average measured flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # group by date and type, and then order
    smps_flows <- dplyr::arrange(group_by(smps_flows, date, type), date)
  
  # calculate average and sdev of measured flows
    smps_flows <- dplyr::mutate(smps_flows, mean = mean(value, na.rm = TRUE),
                                sd = sd(value, na.rm = TRUE))
    
  # drop duplicates
    smps_flows <- dplyr::select(filter(smps_flows, rep == 1), -rep, -value)
    
  # spread
    smps_flows <- gather(smps_flows, temp, val, mean, sd) %>% 
              unite(temp1, type, temp, sep = "_") %>% 
              spread(temp1, val)
```

## Append measured flows to smps data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged <- dplyr::left_join(smps, smps_flows, "date")
```

## Classes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged$sample <- as.factor(smps_merged$sample)
```

## QC

Load notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")    # notes
  notes <- dplyr::filter(notes, grepl("smps|all", notes$inst) == TRUE)
```

Set one flag per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
            group_by(id) %>%
            arrange(qc) %>%
            summarise(qc = first(qc))
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged <- dplyr::left_join(smps_merged, flags, by = "id")
  smps_merged$id <- as.factor(smps_merged$id)
  smps_merged$qc <- ifelse(is.na(smps_merged$qc), "ok", as.character(smps_merged$qc))
  smps_merged$qc <- as.factor(smps_merged$qc)
```

Additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # smps_merged$qc[smps_merged$id == ""] <- "bad"
```

## Ultrafine number

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
# filter smps data
  smps_ultrafine <- dplyr::select(smps_merged, id, value, size, time, sample)
  smps_ultrafine <- dplyr::filter(smps_merged, size <= 100)

# load time windows
  load("../r_files/times_integrated.Rda")
  
# filter by test time window
  smps_ultrafine <- filter_times(times, smps_ultrafine)
  
  smps_ultrafine <- dplyr::group_by(smps_ultrafine, id, sample) %>%
                      dplyr::summarise(number_conc = sum(value)*0.015625355)
```

Plot timeseries for each experiment

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80} 
  ggplot(smps_ultrafine, aes(sample, number_conc)) +
    geom_point() +
    facet_wrap(~id, ncol = 3, scales = "free") +
    scale_y_log10() +
    theme_minimal() + 
    ylab("number per cc")
```

Average number

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_ultrafine <- dplyr::group_by(smps_ultrafine, id) %>%
                      dplyr::summarise(mean_number_conc = mean(number_conc))
```

Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8} 
  ggplot(smps_ultrafine, aes(id, mean_number_conc)) +
    geom_point() +
    scale_y_log10() +
    theme_minimal() + 
    ylab("sub 100 nm - mean number per cc")
```

## Save files

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(smps_merged, file = "../r_files/smps_merged.Rda")
  save(smps_ultrafine, file = "../r_files/smps_ultrafine.Rda")
```

## Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80, eval=TRUE}
  ggplot(smps_merged, aes(size, value, group = sample)) +
    geom_line(aes(colour = sample)) +
    facet_wrap(~id, ncol = 3, scales = "free") +
    theme_minimal() +
    scale_x_log10() + scale_y_log10() +
    ylab("dN/dlogDp") +
    xlab("nm") +
    theme(legend.position="none")
```

## Plot QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80, eval=TRUE}
  ggplot(smps_merged, aes(size, value, group = sample, colour = qc)) +
    geom_line() +
    facet_wrap(~id, ncol = 3, scales = "free") +
    theme_minimal() +
    scale_x_log10() + scale_y_log10() +
    ylab("dN/dlogDp") +
    xlab("nm") +
    theme(legend.position="none")
```

## Data summary

The $SMPS$ measured during `r length(unique(smps_merged$id))` experiments between `r min(smps_merged$date, na.rm = TRUE)` and `r max(smps_merged$date, na.rm = TRUE)`. There is no $SMPS$ data for tests: `r setdiff(as.character(samples$id), as.character(smps_merged$id))`.
