---
title: "smps"
author: "Nicholas Good"
date: "11/21/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(dplyr)
  library(tidyr)
  library(ggplot2)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("r_scripts/R_tidy.R")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("r_files/samples.Rda")    # sample info
```

## Load smps data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("r_files/smps.Rda")    # smps dataset
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(smps)
```

## Load flow data and test times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("r_files/cal_2.Rda")         # measured flow
```

Extract measured flows.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_flows <- tidy.date(cal_2, "^preflow_smps|^postflow_smps", "")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(smps_flows, 2)
```

Separate measurement location and replicate

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_flows <- split.flows(smps_flows)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(smps_flows, 2)
```

## Average measured flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # group by date and type, and then order
    smps_flows <- dplyr::arrange(group_by(smps_flows, date, type), date)
  
  # calculate average and sdev of measured flows
    smps_flows <- dplyr::mutate(smps_flows, mean = mean(value, na.rm = TRUE),
                                sd = sd(value, na.rm = TRUE))
    
  # drop duplicates
    smps_flows <- dplyr::select(filter(smps_flows, rep == 1), -rep, -value)
    
  # spread
    smps_flows <- gather(smps_flows, temp, val, mean, sd) %>% 
              unite(temp1, type, temp, sep = "_") %>% 
              spread(temp1, val)
```

## Append measured flows to smps data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged <- dplyr::left_join(smps, smps_flows, "date")
```

## Classes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged$sample <- as.factor(smps_merged$sample)
```

## QC

Load notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("r_files/notes.Rda")    # notes
  notes <- dplyr::filter(notes, grepl("^smps$|^all$", notes$inst) == TRUE)
```

Flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc <- "bad"
  notes$qc[1:9] <- "maybe"
  notes$qc[15:16] <- "maybe"
  notes$qc[17:22] <- "ok"
  notes$qc[23:27] <- "maybe"
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged <- dplyr::left_join(smps_merged, select(notes, id, qc), by = "id")
  smps_merged$id <- as.factor(smps_merged$id)
  smps_merged$qc <- ifelse(is.na(smps_merged$qc), "ok", as.character(smps_merged$qc))
  smps_merged$qc <- as.factor(smps_merged$qc)
```

Addtional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  smps_merged$qc[smps_merged$id == ""] <- "bad"
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(smps_merged, file="r_files/smps_merged.Rda")
```

## Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80}
  ggplot(smps_merged, aes(size, value, group = sample)) +
    geom_line(aes(colour = sample)) +
    facet_wrap(~id, ncol = 3, scales = "free") +
    theme_minimal() +
    scale_x_log10() + scale_y_log10() +
    ylab("dN/dlogDp") +
    xlab("nm") +
    theme(legend.position="none")
```

## Plot QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80}
  ggplot(smps_merged, aes(size, value, group = sample, colour = qc)) +
    geom_line() +
    facet_wrap(~id, ncol = 3, scales = "free") +
    theme_minimal() +
    scale_x_log10() + scale_y_log10() +
    ylab("dN/dlogDp") +
    xlab("nm") +
    theme(legend.position="none")
```

## Data summary

The $SMPS$ measured during `r length(unique(smps_merged$id))` experiments between `r min(smps_merged$date, na.rm = TRUE)` and `r max(smps_merged$date, na.rm = TRUE)`. There is no $SMPS$ data for tests: `r setdiff(levels(samples$id), levels(smps_merged$id))`.
