---
title: "temp"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

# Load data

* pot and hood temperature

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp<- readRDS("../r_files/temp.RDS")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  #load(file = "../r_files/samples.Rda")
  #load("../r_files/notes.Rda")
```

# Tidy

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_merged <- tidyr::gather(temp, loc, t_oc,
                               t_1, t_2, t_3, t_4) %>%
                 dplyr::arrange(id)
```

# Switch probes

* standard set-up is `t_1 = t_pot` and `t_4 = flue`
* test `10x, 11x and 12x` have two pots and prob location locies
* some tests have inconsistant prob locations

* filter out `NAs`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, eval=TRUE}
  temp_merged <- dplyr::filter(temp_merged, !is.na(t_oc))
```

* one pot stoves

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, eval=TRUE}
  flipped_25a <- dplyr::filter(temp_merged,
                           id =="25A") %>%
             dplyr::mutate(loc = ifelse(loc == "t_1", "flue", loc),
                           loc = ifelse(loc == "t_4", "front", loc))

  flipped_25bc <- dplyr::filter(temp_merged,
                           id =="25B" |id =="25C") %>%
             dplyr::mutate(loc = ifelse(loc == "t_1", "flue", loc),
                           loc = ifelse(loc == "t_2", "front", loc))

  regular <- dplyr::filter(temp_merged,
                          id !="25A" & id !="25B" & id !="25C" &
                          grepl("^10[A-Z]|^11[A-Z]|^12[A-Z]", id) == FALSE) %>%
             dplyr::mutate(loc = ifelse(loc == "t_1", "front", loc),
                           loc = ifelse(loc == "t_4", "flue", loc))

```

* two pot stoves

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, eval=TRUE}
  temp_10a <- dplyr::filter(temp_merged, id == "10A") %>%
              dplyr::mutate(loc = ifelse(loc == "t_1", "front", loc),
                            loc = ifelse(loc == "t_2", "back", loc),
                            loc = ifelse(loc == "t_4", "flue", loc))

  temp_10b <- dplyr::filter(temp_merged, id == "10B") %>%
              dplyr::mutate(loc = ifelse(loc == "t_1", "flue", loc),
                            loc = ifelse(loc == "t_3", "front", loc),
                            loc = ifelse(loc == "t_4", "back", loc))

  temp_10c <- dplyr::filter(temp_merged, id == "10C") %>%
              dplyr::mutate(loc = ifelse(loc == "t_1", "front", loc),
                            loc = ifelse(loc == "t_2", "back", loc),
                            loc = ifelse(loc == "t_4", "flue", loc))
  
  temp_11a <- dplyr::filter(temp_merged, id == "11A") %>%
              dplyr::mutate(loc = ifelse(loc == "t_1", "front", loc),
                            loc = ifelse(loc == "t_2", "back", loc),
                            loc = ifelse(loc == "t_4", "flue", loc))

  temp_11b <- dplyr::filter(temp_merged, id == "11B") %>%
              dplyr::mutate(loc = ifelse(loc == "t_1", "front", loc),
                            loc = ifelse(loc == "t_2", "back", loc),
                            loc = ifelse(loc == "t_4", "flue", loc))

  temp_11c <- dplyr::filter(temp_merged, id == "11C") %>%
              dplyr::mutate(loc = ifelse(loc == "t_1", "front", loc),
                            loc = ifelse(loc == "t_2", "back", loc),
                            loc = ifelse(loc == "t_4", "flue", loc))

  temp_12a <- dplyr::filter(temp_merged, id == "12A") %>%
              dplyr::mutate(loc = ifelse(loc == "t_1", "front", loc),
                            loc = ifelse(loc == "t_2", "back", loc),
                            loc = ifelse(loc == "t_4", "flue", loc))

  temp_12b <- dplyr::filter(temp_merged, id == "12B") %>%
              dplyr::mutate(loc = ifelse(loc == "t_1", "flue", loc),
                            loc = ifelse(loc == "t_3", "back", loc),
                            loc = ifelse(loc == "t_4", "front", loc))

  temp_12c <- dplyr::filter(temp_merged, id == "12C") %>%
              dplyr::mutate(loc = ifelse(loc == "t_1", "front", loc),
                            loc = ifelse(loc == "t_2", "back", loc),
                            loc = ifelse(loc == "t_4", "flue", loc))
```

* combine

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, eval=TRUE}
  temp_merged <- dplyr::bind_rows(regular,
                                  flipped_25a,
                                  flipped_25bc,
                                  temp_10a, temp_10b, temp_10c,
                                  temp_11a, temp_11b, temp_11c,
                                  temp_12a, temp_12b, temp_12c)

```

# QC

## flags from notes

* extract notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("temp|all", notes$inst) == TRUE)
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

* merge flags with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_merged <- dplyr::left_join(temp_merged, flags, by = "id") %>%
                 dplyr::mutate(id = as.factor(id)) %>%
                 dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

## additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # temp_merged$qc[temp_merged$id == ""] <- "bad"
```

# Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(temp_merged, file = "../r_files/temp_merged.Rda")
```

# Summary

Temperature was measured for `r length(unique(temp_merged$id))` experiments between `r min(temp_merged$date, na.rm = TRUE)` and `r max(temp_merged$date, na.rm = TRUE)`. There is no temperature data for tests: `r setdiff(as.character(samples$id), as.character(temp_merged$id))`.

Temperature data is expected to be missing for: 
All "G" tests (1-18). We did not collect stove/pot temperature for background tests as no stove was operated.

13A, 2A, and 6A (the three tests conducted on 1/5/2016), 18B, 21A (conducted on 2/10/2016), 22A (conducted on 1/13/2016), and 3A, and 7A (conducted on 2/26/2016): test data was lost. It is possible that there were some thermocouple malfunctions resulting in this data loss (check "notes" on transcribed testing sheets).

# Plots

## all

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=50}
  ggplot(temp_merged, aes(datetime, t_oc, color = loc)) +
         geom_line() +
         geom_point() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         theme_minimal() +
         theme(legend.position = "top") +
         ylab("degress celsius")
```

## pot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=15}
  p_data <- dplyr::filter(temp_merged, (loc == "front" | loc == "back")
                                       & (grepl("^10[A-Z]|^11[A-Z]|^12[A-Z]", id)))

  ggplot(p_data, aes(datetime, t_oc, color = loc)) +
         geom_line() +
         geom_point() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         theme_minimal() +
         theme(legend.position = "top") +
         ylab("degress celsius")
```