---
title: "temp"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
  source("../r_scripts/R_functions.R")
```

# Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp <- readRDS("../r_files/temp.RDS")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  test_info <- readRDS("../r_files/test_info.RDS")
  times_fuel <- readRDS("../r_files/times_fuel.RDS")
  times_bg <- readRDS("../r_files/times_bg.RDS")
  times_test <- readRDS("../r_files/times_test.RDS")
  flow <- readRDS("../r_files/flows.RDS")
  notes <- readRDS("../r_files/notes.RDS")
```

# Organize

## sample times

* full test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_full <- filter_times_startup(times_test, temp)
```

* background levels

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_bg <- filter_times_background(times_bg, temp)
```

* levels by replicate

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_rep <- filter_times_rep(times_fuel, temp)
```

## Info

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  info <- dplyr::select(test_info, id_test, id, fuel_type, fuel_quant)

  temp <- dplyr::left_join(temp, info, "id")
```

# QC

## flags from notes

* extract notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("temp|all", notes$inst) == TRUE)
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

* merge flags with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_merged <- dplyr::left_join(temp_merged, flags, by = "id") %>%
                 dplyr::mutate(id = as.factor(id)) %>%
                 dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

## additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # temp_merged$qc[temp_merged$id == ""] <- "bad"
```

# Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(temp_merged, file = "../r_files/temp_merged.Rda")
```

# Summary

Temperature was measured for `r length(unique(temp_merged$id))` experiments between `r min(temp_merged$date, na.rm = TRUE)` and `r max(temp_merged$date, na.rm = TRUE)`. There is no temperature data for tests: `r setdiff(as.character(samples$id), as.character(temp_merged$id))`.

Temperature data is expected to be missing for: 
All "G" tests (1-18). We did not collect stove/pot temperature for background tests as no stove was operated.

13A, 2A, and 6A (the three tests conducted on 1/5/2016), 18B, 21A (conducted on 2/10/2016), 22A (conducted on 1/13/2016), and 3A, and 7A (conducted on 2/26/2016): test data was lost. It is possible that there were some thermocouple malfunctions resulting in this data loss (check "notes" on transcribed testing sheets).

# Plots

## all

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=50}
  ggplot(temp_merged, aes(datetime, t_oc, color = loc)) +
         geom_line() +
         geom_point() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         theme_minimal() +
         theme(legend.position = "top") +
         ylab("degress celsius")
```

## pot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=15}
  p_data <- dplyr::filter(temp_merged, (loc == "front" | loc == "back")
                                       & (grepl("^10[A-Z]|^11[A-Z]|^12[A-Z]", id)))

  ggplot(p_data, aes(datetime, t_oc, color = loc)) +
         geom_line() +
         geom_point() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         theme_minimal() +
         theme(legend.position = "top") +
         ylab("degress celsius")
```