---
title: "temp"
author: "Nicholas Good"
date: "11/23/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

## Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/temp.Rda")           # temperature dataset
  load(file = "../r_files/samples.Rda") # load sample
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(temp)
```

## Tidy

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
temp_merged <- dplyr::select(temp, -ch2, -ch3) %>%
          tidyr::gather(var, val, tamb, thood) %>%
          dplyr::arrange(id)
```

## QC

Load notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")    # notes
  notes <- dplyr::filter(notes, grepl("temp|all", notes$inst) == TRUE)
```

Set one flag per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
            group_by(id) %>%
            arrange(qc) %>%
            summarise(qc = first(qc))
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_merged <- dplyr::left_join(temp_merged, flags, by = "id")
  temp_merged$id <- as.factor(temp_merged$id)
  temp_merged$qc <- ifelse(is.na(temp_merged$qc), "ok", as.character(temp_merged$qc))
  temp_merged$qc <- as.factor(temp_merged$qc)
```

Additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # temp_merged$qc[temp_merged$id == ""] <- "bad"
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(temp_merged, file = "../r_files/temp_merged.Rda")
```

## Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=40}
  ggplot(temp_merged, aes(datetime, val, color = var)) +
  geom_line() +
  geom_point() +
  facet_wrap(~id, ncol = 3, scales = "free") +
  theme_minimal()
```

## Summary

$Temperature$ was measured for `r length(unique(temp_merged$id))` experiments between `r min(temp_merged$date, na.rm = TRUE)` and `r max(temp_merged$date, na.rm = TRUE)`. There is no $temperature$ data for tests: `r setdiff(as.character(samples$id), as.character(temp_merged$id))`.
