---
title: "temp"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

# Load data

* pot and hood temperature

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp <- readRDS("../r_files/temp.RDS")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  samples <- readRDS(file = "../r_files/samples.RDS")
  notes <- readRDS("../r_files/notes.RDS")
  test_times <- readRDS("../r_files/test_times_all.RDS")
```

# Tidy

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_merged <- tidyr::gather(temp, loc, t_oc,
                               t_1, t_2, t_3, t_4) %>%
                 dplyr::arrange(id)
```

# Switch probes

* filter out `NAs`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, eval=TRUE}
  temp_merged <- dplyr::filter(temp_merged, !is.na(t_oc))
```

* one pot stoves

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, eval=TRUE}
  flipped_tests <- dplyr::filter(temp_merged,
                           id =="FL2" |id =="FM2" |id == "LL2" |id == "LL4" |id == "LM4" |id == "FH4") %>%
             dplyr::mutate(loc = ifelse(loc == "t_1", "stove", loc),
                           loc = ifelse(loc == "t_2", "pot", loc))

  regular <- dplyr::filter(temp_merged, id !="FL2" & id !="FM2" & id !="LL2" & id !="LL4" & id !="LM4" & id !="FH4") %>%
             dplyr::mutate(loc = ifelse(loc == "t_1", "pot", loc),
                           loc = ifelse(loc == "t_2", "stove", loc))

```

* combine

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, eval=TRUE}
  temp_merged <- dplyr::bind_rows(regular,
                                  flipped_tests)

```

# QC

## flags from notes

* extract notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("temp|all", notes$inst) == TRUE)
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

* merge flags with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_merged <- dplyr::left_join(temp_merged, flags, by = "id") %>%
                 dplyr::mutate(id = as.factor(id)) %>%
                 dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

## additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # temp_merged$qc[temp_merged$id == ""] <- "bad"
```

# Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  saveRDS(temp_merged, file = "../r_files/temp_merged.RDS")
```

# Summary

Temperature was measured for `r length(unique(temp_merged$id))` experiments between `r min(temp_merged$date, na.rm = TRUE)` and `r max(temp_merged$date, na.rm = TRUE)`. There is no temperature data for tests: `r setdiff(as.character(samples$id), as.character(temp_merged$id))`.

Temperature data is expected to be missing for: 
All "G" and "BG" tests. We did not collect stove/pot temperature for background tests as no stove was operated.


# Plots

## all

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=50}
  ggplot(temp_merged, aes(datetime, t_oc, color = loc)) +
         geom_line() +
         geom_point() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         theme_minimal() +
         theme(legend.position = "top") +
         ylab("degrees celsius")
```

```{r}
times <- rename(test_times, time = value) %>%
         select(id, var, time)
#temp_merged1 <-

temps_times <- filter(temp_merged, loc == "stove") %>%
               select(time, id, t_oc) %>%
               left_join(times, by = c("id", "time"))

temps_times <- filter_times(times, temp_merged)
```

