---
title: "temp"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

# Load data

* pot and hood temperature

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/temp.Rda")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load(file = "../r_files/samples.Rda")
  load("../r_files/notes.Rda")
```

# Tidy

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_merged <- dplyr::select(temp, -ch2, -ch3) %>%
                 tidyr::gather(var, val, t_pot, t_hood) %>%
                 dplyr::arrange(id)
```

# Switch probes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, eval=TRUE}
  id_flipped <- c("10B", "12B", "25A", "25B", "25C")

  flipped <- dplyr::filter(temp_merged,
                           id == "10B"| id =="12B"|
                           id =="25A" | id =="25B" |
                           id =="25C") %>%
             dplyr::mutate(var = ifelse(var == "t_pot", "t_hood", "t_pot"))

  not_flipped <- dplyr::filter(temp_merged,
                           id != "10B" & id !="12B" &
                           id !="25A" & id !="25B" &
                           id !="25C")
  
  temp_merged <- dplyr::bind_rows(flipped, not_flipped) %>%
                 dplyr::arrange(id) %>%
                 dplyr::mutate(var = as.factor(var))
```

# QC

## flags from notes

* extract notes

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("temp|all", notes$inst) == TRUE)
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

* merge flags with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  temp_merged <- dplyr::left_join(temp_merged, flags, by = "id") %>%
                 dplyr::mutate(id = as.factor(id)) %>%
                 dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

## additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # temp_merged$qc[temp_merged$id == ""] <- "bad"
```

# Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(temp_merged, file = "../r_files/temp_merged.Rda")
```

# Summary

Temperature was measured for `r length(unique(temp_merged$id))` experiments between `r min(temp_merged$date, na.rm = TRUE)` and `r max(temp_merged$date, na.rm = TRUE)`. There is no temperature data for tests: `r setdiff(as.character(samples$id), as.character(temp_merged$id))`.

Temperature data is expected to be missing for: 
All "G" tests (1-18). We did not collect stove/pot temperature for background tests as no stove was operated.

13A, 2A, and 6A (the three tests conducted on 1/5/2016), 18B, 21A (conducted on 2/10/2016), 22A (conducted on 1/13/2016), and 3A, and 7A (conducted on 2/26/2016): test data was lost. It is possible that there were some thermocouple malfunctions resulting in this data loss (check "notes" on transcribed testing sheets).

# Plots

## timeseries

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=40}
  ggplot(temp_merged, aes(datetime, val, color = var)) +
         geom_line() +
         geom_point() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         theme_minimal() +
         theme(legend.position = "top") +
         ylab("degress celsius")
```
