---
title: "test times"
author: "Nicholas Good"
date: "11/21/2016"
output: html_document
---

## Libraries

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

## Source files

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../R_scripts/R_load_data.R")
  source("../R_scripts/R_load_metadata.R")
  source("../R_scripts/R_tidy.R")
```

## Load data

Load sample, wood and batch test meta data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
  load("../r_files/batch_times.Rda")
  load("../r_files/wood_times.Rda")
```

Load pre and post experiment background times for online instruments

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/data_1.Rda")
  load("../r_files/data_2.Rda")
```

# Combine time data

* wood and batch times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  wood_batch_times <- dplyr::bind_rows(dplyr::select(wood_times, -date),
                                       dplyr::select(batch_times, -date)) %>%
                      dplyr::inner_join(dplyr::select(samples, id, date),
                                                      test_times, by = "id") %>%
                      dplyr::filter(is.na(value)==FALSE) %>%
                      dplyr::mutate(id = as.factor(id))
```

* fivegas background times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_bg_fivegas <- dplyr::select(data_1, id, 
                                    time_start_fivegas_prebg,
                                    time_end_fivegas_prebg,
                                    time_start_fivegas_post_bg,
                                    time_end_fivegas_post_bg) %>%
                      dplyr::rename(bg_pre_start_fivegas = time_start_fivegas_prebg,
                                    bg_pre_end_fivegas = time_end_fivegas_prebg,
                                    bg_post_start_fivegas = time_start_fivegas_post_bg,
                                    bg_post_end_fivegas = time_end_fivegas_post_bg) %>%
                      tidyr::gather("var", "value", 2:5) %>%
                      dplyr::inner_join(dplyr::select(samples, id, date),
                                        by = "id") %>%
                      dplyr::filter(is.na(value)==FALSE) %>%
                      dplyr::mutate(id = as.factor(id))
```

* pax/smps background times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times_bg_pax_smps <- dplyr::select(data_2, id, 
                                     time_start_smps_pax_bg_pre,
                                     time_end_smps_pax_bg_pre,
                                     time_start_smps_pax_bg_post,
                                     time_end_smps_pax_bg_post) %>%
                       dplyr::rename(bg_pre_start_pax_smps = time_start_smps_pax_bg_pre,
                                     bg_pre_end_pax_smps = time_end_smps_pax_bg_pre,
                                     bg_post_start_pax_smps = time_start_smps_pax_bg_post,
                                     bg_post_end_pax_smps = time_end_smps_pax_bg_post) %>%
                      tidyr::gather("var", "value", 2:5) %>%
                      dplyr::inner_join(dplyr::select(samples, id, date),
                                        by = "id") %>%
                      dplyr::filter(is.na(value)==FALSE) %>%
                      dplyr::mutate(id = as.factor(id))
```

* combine times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  test_times <- dplyr::bind_rows(wood_batch_times,
                                 times_bg_fivegas,
                                 times_bg_pax_smps) %>%
                dplyr::mutate(id = as.factor(id)) %>%
                dplyr::mutate(var = as.factor(var)) %>%
                tibble::as_data_frame()
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(test_times)
```

## Save data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(test_times, file = "../r_files/test_times.Rda")
```

## Summary

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  samples <- dplyr::select(samples, id, stove, fuel)
```

Test times were measured for `r length(unique(test_times$id))` experiments between `r min(test_times$date, na.rm = TRUE)` and `r max(test_times$date, na.rm = TRUE)`. There are times for tests: `r setdiff(levels(samples$id), levels(test_times$id))`.
