---
title: "VOCs"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(forcats)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
  source("../r_scripts/R_functions.R")
```

# Load data

 * voc (ppbv)

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- readRDS("../r_files/voc.RDS")
```

* metadata

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  samples <- readRDS("../r_files/samples.RDS")     # load sample meta data
  fuel_burnt <- readRDS("../r_files/fuel_burnt.RDS")  # fuel use meta data
  emissions <- readRDS("../r_files/emissions_long.RDS")   # load pollutant emissions data
  data_1 <- readRDS("../r_files/data_tester_one.RDS")
  notes <- readRDS("../r_files/notes.RDS")
``` 

* The columns `stove` and `fuel` contain the exact stove and fuel type (e.g., "Ceramic Charcoal (Artisan), Coconut Charcoal") that corresponds with the number of the `id` (3 replicates per id number, specified by A, B, and C (sometimes additional replicates, [i.e. D or E], is also included). The columns `stovecat` and `fuelcat` contain broader categories of stoves and fuels (e.g., "improved, charcoal").

# Organize

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # voc_merged <- dplyr::filter(voc, grepl("^[0-9]|^G[0-9]", id) == TRUE)
```

## longer format

* make longer and remove `NA`.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc_merged <- tidyr::gather(voc, "pol", "ppb", - c(id, date, start_time, stop_time,
                                                     start_reading, stop_reading,
                                                     flow_rate_cc_per_min)) %>%
                dplyr::filter(!is.na(id) & id != "NA")
```

## merge with meta data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc_merged <- dplyr::left_join(voc_merged,
                                 dplyr::select(data_1,
                                               matches("_voc|voc_|^id$|^date$")))
```

# LOD

* Add LOD flag
* Replace pollutant values below LOD with `NA`.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc_merged <- dplyr::mutate(voc_merged,
                              lod = as.factor(ifelse(ppb == -8888,
                                                     "below",
                                                     "above"))) %>%
                dplyr::mutate(ppb = ifelse(ppb == -8888,
                                           NA,
                                           ppb))
```

# QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes <- dplyr::filter(notes, grepl("voc|all", inst))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  knitr::kable(dplyr::filter(notes, grepl(".*voc.*", inst)) %>%
               dplyr::select(-inst, -date),
               "markdown", digits = 2)
```

* flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # notes$qc[33:36] <- "maybe"
```

* apply flags: `bad` preceeds `maybe` preceeds `good`

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

* merge

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc_merged <- dplyr::left_join(voc_merged, flags, by = "id") %>%
                dplyr::mutate(id = as.factor(id)) %>%
                dplyr::mutate(qc = as.factor(ifelse(is.na(qc),
                                                    "ok",
                                                    as.character(qc))))
```

* additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # voc_merged$qc[grav_merged$id == ""] <- "bad"
```

# Background

* extract good background data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg_voc <- dplyr::filter(voc_merged, grepl("^BG", id)) %>%
        dplyr::mutate(ppb = ifelse(is.na(ppb), 0, ppb)) %>%
        na.omit()
        #dplyr::select(-id_can, -id_voc)

```

* plot vocs with a background test higher than 10 ppb (note box plots contain all data points)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  p_data <-  dplyr::group_by(bg_voc, pol) %>%
             dplyr::mutate(max = max(ppb)) %>%
             dplyr::ungroup() %>%
             dplyr::filter(max > 10) %>%
             dplyr::group_by(pol) %>%
             dplyr::mutate(outlier = ifelse(is_outlier(ppb), as.character(id), NA),
                           mean = mean(ppb, na.rm = TRUE)) %>%
             dplyr::ungroup() %>%
             dplyr::mutate(pct = round(percent_rank(mean), 1))

  ggplot(p_data, aes(pol, ppb)) +
    geom_boxplot(outlier.size = 0) +
    geom_point() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ylab("ppb") + xlab("pollutant") +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 4)

```

## average background

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg_voc <- dplyr::group_by(bg_voc, pol) %>%
        dplyr::summarise(mean = mean(ppb),
                         sd = sd(ppb),
                         min = min(ppb),
                         max = max(ppb),
                         n = n())
```

* background values (ppb)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  knitr::kable(bg_voc, "markdown", digits = 2)
```

## merge background

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  bg_voc <- dplyr::select(bg_voc, pol, mean) %>%
        dplyr::rename(ppb_bg = mean)

  voc_merged <- dplyr::left_join(voc_merged, bg_voc, by = "pol") %>%
                dplyr::mutate(ppb_bg = ifelse(is.na(ppb_bg), 0, ppb_bg))
```

## rename variables

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc_merged <- dplyr::rename(voc_merged,
                                         start = time_start_voc,
                                         end = time_end_voc)
```

# Save

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  saveRDS(voc_merged, file="../r_files/voc_merged.RDS")
```

# Summary

VOCs were measured for `r length(unique(voc_merged$id))` experiments between `r min(voc_merged$date, na.rm = TRUE)` and `r max(voc_merged$date, na.rm = TRUE)`. There is no VOC data for tests: `r setdiff(as.character(samples$id), as.character(voc_merged$id))`.

VOCs data is expected to be missing for: G18 only. We did not collect a canister for this test due to supply shortage.

# Plots

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  samples <- dplyr::select(samples, id, stove, fuel)
```

## plot by fuel type

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  p_data <- dplyr::left_join(voc_merged, samples, by = "id") %>%
            dplyr::mutate(pol = sub("^voc_", "", pol)) %>%
            dplyr::mutate(id = as.factor(id)) %>%
            dplyr::filter(ppb > 0 &
                          grepl("^P", id) == FALSE &
                          grepl("^G", id)==FALSE)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=6, fig.height=8}
  ggplot(p_data, aes(id, ppb, colour = fuel)) + 
         geom_point() +
         facet_wrap(~pol, ncol = 1, scales = "free") +
         theme_minimal() +
         theme(legend.position = "top") +
         ylab("ug/m3") +
         xlab("") +
         theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
