---
title: "$VOCs$"
author: "Nicholas Good"
date: "11/23/2016"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_tidy.R")
```

## Load $VOC$ data

The $VOC$ data is contain in one file. Units are $ppbv$. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/voc.Rda")    # voc dataset
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  colnames(voc)
```

## Remove non-study data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- dplyr::filter(voc, grepl("^[0-9]|^G[0-9]", id) == TRUE)
```

## Convert to longer format

Make longer and remove n/a.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc <- tidyr::gather(voc, "pol", "val", starts_with("voc_")) %>%
         dplyr::filter(!is.na(id) & id != "NA")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  head(voc, 2)
```

## Merge with meta data

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load(file="../r_files/data_1.Rda")
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # match voc metadata
  meta <- dplyr::select(data_1, matches("_voc|voc_|^id$|^date$"))
 # merge
  voc_merged <- dplyr::left_join(voc, meta)
```

## QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/notes.Rda")

  notes <- dplyr::filter(notes, grepl("voc|all", notes$inst) == TRUE)
```

Flag bad data from notes and plots

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  notes$qc[33:36] <- "maybe"
```

Set one flag per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  flags <- dplyr::select(notes, id, qc) %>%
           dplyr::group_by(id) %>%
           dplyr::arrange(qc) %>%
           dplyr::summarise(qc = first(qc))
```

Merge with data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  voc_merged <- dplyr::left_join(voc_merged, flags, by = "id") %>%
                dplyr::mutate(id = as.factor(id)) %>%
                dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

Additional bad tests

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  # voc_merged$qc[grav_merged$id == ""] <- "bad"
```

## Save merged file

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  save(voc_merged, file="../r_files/voc_merged.Rda")
```

## Plot (coloured by fuel type)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load(file="../r_files/samples.Rda")

  samples <- dplyr::select(samples, id, stove, fuel)
  
  voc_p <- dplyr::left_join(voc_merged, samples, by = "id") %>%
           dplyr::mutate(pol = sub("^voc_", "", pol)) %>%
           dplyr::mutate(id = as.factor(id))
  
 # filter out background data and non-detects 
  voc_p <- dplyr::filter(voc_p, val > 0 & grepl("^P", id)==FALSE & grepl("^G", id)==FALSE)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=20, fig.height=250}
  ggplot(voc_p, aes(id, val, colour = fuel)) + 
         geom_point() +
         facet_wrap(~pol, ncol = 1, scales = "free") +
         theme_minimal() +
         theme(legend.position = "top") +
         ylab("ug/m3") + xlab("")
```

## Plot QC

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=20, fig.height=250}
  ggplot(voc_p, aes(id, val, colour = qc)) + 
         geom_point() +
         facet_wrap(~pol, ncol = 1, scales = "free") +
         theme_minimal()
```

$VOCs$ were measured for `r length(unique(voc_merged$id))` experiments between `r min(voc_merged$date, na.rm = TRUE)` and `r max(voc_merged$date, na.rm = TRUE)`. There is no $VOC$ data for tests: `r setdiff(as.character(samples$id), as.character(voc_merged$id))`.

$VOCs$ data is expected to be missing for: G18 only. We did not collect a canister for this test due to supply shortage.
