---
title: "Energy applied"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  library(tidyverse)
  library(knitr)
  library(lubridate)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  source("../r_scripts/R_functions.R")
```

# Load

* water (pot) temperature

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/temp_merged.Rda")
```

* constants

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/calc_constants.Rda")
```

* metadata

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  load("../r_files/samples.Rda")
  load("../r_files/test_times.Rda")
  load("../r_files/batch_wgts.Rda")
  load("../r_files/batch_pot.Rda")
  load("../r_files/wood_wgts.Rda")
  load("../r_files/wood_pot.Rda")
```

# Daylight saving

* Correct for summertime

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  summer_time <- as.POSIXct("2016/03/13 02:00:00", tz = "MST") 
  temp_merged <- dplyr::mutate(temp_merged,
                               datetime = if_else(datetime > summer_time, (datetime + 60 * 60), datetime),
                               datetime = as.POSIXct(datetime, origin = "1970-01-01"),
                               time = if_else(datetime > summer_time, (time + 60 * 60), time))
```

# Test timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  times <- dplyr::filter(test_times, var == "start_1" | var == "shutdown") %>%
           tidyr::spread(var, value) %>%
           dplyr::select(-date) %>%
           dplyr::rename(start = start_1, end = shutdown)
```

# Energy

## joules applied

* timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_times <- dplyr::filter(test_times, grepl("^start_[0-9]|^end_[0-9]", var)) %>%
             dplyr::select(-date) %>%
             tidyr::separate(var, c("var", "rep")) %>%
             tidyr::spread(var, value)

 # remove reps added for 2 pot stoves
  e_times <- group_by(e_times, id, start) %>% filter(row_number(rep) == 1) %>%
  	         ungroup()
```

* which pot

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_pot <- dplyr::select(wood_pot, -date) %>%
           dplyr::bind_rows(dplyr::select(batch_pot, -date))
```

* pot mass

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_pot_mass <- dplyr::filter(wood_wgts,
                              grepl("^pot",
                              var)) %>%
                tidyr::separate(var, c("var", "pot")) %>%
                dplyr::select(-var, -date) %>%
                tidyr::spread(pot, value)

  e_pot_mass <- dplyr::bind_rows(e_pot_mass,
                                 dplyr::bind_rows(e_pot_mass,
                                 dplyr::filter(batch_wgts,
                                               grepl("^pot",
                                               var)) %>%
                                 tidyr::separate(var, c("var", "pot")) %>%
                                 dplyr::select(-var, -date) %>%
                                 tidyr::spread(pot, value)))
```

* pot + water mass

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_pot_water_mass <- dplyr::filter(wood_wgts,
                                    grepl("^on_[0-9]|^off_[0-9]", var)) %>%
                      tidyr::separate(var, c("var", "rep")) %>%
                      tidyr::spread(var, value) %>%
                      dplyr::select(-date) %>%
                      dplyr::rename(mass_on = on, mass_off = off)

  e_pot_water_mass <- dplyr::bind_rows(e_pot_water_mass,
                                       dplyr::filter(batch_wgts,
                                       grepl("^on_[0-9]|^off_[0-9]", var)) %>%
                                       tidyr::separate(var, c("var", "rep")) %>%
                                       tidyr::spread(var, value) %>%
                                       dplyr::select(-date) %>%
                                       dplyr::rename(mass_on = on,
                                                     mass_off = off))
```

* Tweak
  tweak times
  mind min
  remove leading tail

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
 # time tweak
  time_tweak <- data_frame(id = unique(temp_merged$id), dt = 0)
  time_tweak$dt[which(time_tweak$id == "13C")] <- 70
  time_tweak$dt[which(time_tweak$id == "13B")] <- 15
  time_tweak$dt[which(time_tweak$id == "16A")] <- 100
  time_tweak$dt[which(time_tweak$id == "25A")] <- -90
  time_tweak$dt[which(time_tweak$id == "25B")] <- -90
  time_tweak$dt[which(time_tweak$id == "20C")] <- -30
  time_tweak$dt[which(time_tweak$id == "21B")] <- 30
  time_tweak$dt[which(time_tweak$id == "25A")] <- -90
  time_tweak$dt[which(time_tweak$id == "25B")] <- -90
  time_tweak$dt[which(time_tweak$id == "26B")] <- -90
  time_tweak$dt[which(time_tweak$id == "2E")] <- -90
  time_tweak$dt[which(time_tweak$id == "6E")] <- -90
  time_tweak$dt[which(time_tweak$id == "8D")] <- -90
  time_tweak$dt[which(time_tweak$id == "8C")] <- -50
  time_tweak$dt[which(time_tweak$id == "9B")] <- -50
  time_tweak$dt[which(time_tweak$id == "5A")] <- 60
  time_tweak$dt[which(time_tweak$id == "3D")] <- -60
  time_tweak$dt[which(time_tweak$id == "29A")] <- -60
  temp_merged <- dplyr::left_join(temp_merged, time_tweak, by = "id") %>%
                 dplyr::mutate(time = time - 130,
                                datetime = datetime - 130,
                                time = time + dt,
                                datetime = datetime + dt)
```

```{r, eval=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  tmp <- dplyr::filter(temp_merged, id == "9B")
  tmp <- dplyr::filter(tmp, loc != "flue")
  tmp <- filter_temp(e_times, tmp)

 ggplot(tmp, aes(datetime, t_oc, group = loc, colour = rep)) +
    geom_point() +
    theme_minimal() +
    theme(legend.position = "top") +
    ylab("temperature oC")
```

* extract temperature for periods of interest

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_temp <- dplyr::filter(temp_merged, loc != "flue")

  e_temp <- filter_temp(e_times, e_temp) # adds rep
```

* filter out unrealistic values

Data should be betwen 10 and 100 degrees C.

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_temp <- dplyr::filter(e_temp, t_oc < 100, t_oc > 10)
```


* temperature change

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  t_range <- dplyr::group_by(e_temp, id, rep, loc) %>%
             dplyr::summarise(t_min = quantile(t_oc, 0.025, na.rm = TRUE) - 2,
                              t_max = quantile(t_oc, 0.975, na.rm = TRUE) + 2)
```

* tweak temp values

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # missing data from start of 2nd rep
  t_range$t_min[which(t_range$id == "1B" & t_range$rep == 2)] <- 20.18
```

* merge data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
 # issue with left join on e_pot_mass creating duplicate rows...
 # inconsistant class(id) perhaps an issue?
  e_data <- dplyr::left_join(e_times, e_pot_water_mass, by = c("id", "rep")) %>%
            dplyr::left_join(t_range, by = c("id", "rep")) %>%
            dplyr::left_join(e_pot, by = c("id", "rep")) %>%
            dplyr::left_join(e_pot_mass, by = "id") %>%
            dplyr::mutate(mass_pot = ifelse(pot == "a", a, NA),
                          mass_pot = ifelse(pot == "b", b, mass_pot),
                          id = as.factor(id)) %>%
            dplyr::select(-a, -b) %>%
            dplyr::distinct()
```

* energy applied

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  e_data <- dplyr::mutate(e_data,
                          mass_evap = mass_on - mass_off,
                          mass_h2o = mass_on - mass_pot,
                          dt = t_max - t_min,
                          j_heat = mass_h2o * c_water * dt, # g * J/(g.oC) * oC
                          j_vap = mass_evap * hfg_water,
                          joules = j_heat + j_vap)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  kable(dplyr::slice(e_data, 1:5),
        align = 'c', caption = "Energy Numbers")
```

* energy applied per test

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  energy <- dplyr::select(e_data, id, joules) %>%
            dplyr::group_by(id) %>%
            dplyr::summarise(joules = sum(joules, na.rm = TRUE))
```

# QC

* add flag: default is "ok"

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  energy <- 
    energy %>%
    dplyr::mutate(qc = "ok")
```

* add bad flags

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  energy <- 
    energy %>%
    dplyr::mutate(qc = ifelse(joules <= 0, "bad", qc))
```

# Plots

## temperature (id & loc)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=50}
  ggplot(dplyr::filter(e_temp, loc != "flue"),
         aes(datetime, t_oc, color = loc)) +
         geom_line() +
         geom_point() +
         facet_wrap(~id, ncol = 3, scales = "free") +
         theme_minimal() +
         theme(legend.position = "top") +
         ylab("degress celsius") +
         theme(legend.position = "top")
```

## temperature (id & rep)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=80}
  p_data <- dplyr::left_join(e_temp, t_range, by = c("id", "rep", "loc"))

  ggplot(p_data, aes(datetime, t_oc, group = loc, colour = rep)) +
    geom_point() +
    geom_line(aes(datetime, t_min, colour = rep)) +
    geom_line(aes(datetime, t_max, colour = rep)) +
    facet_wrap(~id, scales = "free", ncol = 3) +
    theme_minimal() +
    theme(legend.position = "top") +
    ylab("temperature oC")
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
 # list of experiment ids
  ids <- unique(e_temp$id)
 # loop ids
  for(i in 1:length(ids)){
   # extract data for id
    tmp <- dplyr::filter(e_temp, id == ids[i])
    reps <- unique(tmp$rep)
     for(j in 1:length(reps)){
      #print(reps[j])
     }
  }
```

## energy delivered

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
  p_data <- dplyr::left_join(dplyr::filter(samples, type == "SF"),
                             energy,
                             by = "id") %>%
	         dplyr::filter(joules > 0)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE, fig.width=12, fig.height=8}
  ggplot(p_data, aes(id, joules, colour = qc)) +
  geom_point() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = "top") +
  ylab("energy delivered to pot (J)") +
  theme(legend.position = "top")
```

# Save (not implemented)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  saveRDS(energy, "../r_files/energy.RDS")
```
