---
title: "Process metadata"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE, warning = FALSE, results='hide'}
  packages <- c("tidyverse", "GGally", "gridExtra")

  new <- packages[!(packages %in% installed.packages()[,"Package"])]
  if(length(new)) install.packages(new)
  lapply(packages, library, character.only = TRUE)
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/', warning = FALSE, message = FALSE, cache = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
  source("../r_scripts/R_functions.R")
  source("../r_scripts/R_plots.R")
  source("../r_scripts/R_tidy.R")
```  
 
# Load data

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  pax <- readRDS("../r_files/pax.RDS")         # pax data
  co2_lab <- readRDS("../r_files/co2_lab.RDS")     # lab co2 levels
  co2_sample <- readRDS("../r_files/co2_sample.RDS")  # diluted sample line co2 levels
  scale <- readRDS("../r_files/scale.RDS")       # scale data
  temp <- readRDS("../r_files/temp.RDS")        # temperature data
  smps <- readRDS("../r_files/smps.RDS")        # smps data

  ecoc <- readRDS("../r_files/ecoc.RDS")        # ecoc analysis
  ions <- readRDS("../r_files/ions.RDS")        # ion and carbonyls
  trans <- readRDS("../r_files/trans.RDS")       # transmissometer data
  voc <- readRDS("../r_files/voc.RDS")         # voc canister data

  samples <- readRDS("../r_files/samples.RDS")     # sample info
  batch <- readRDS("../r_files/batch.RDS")       # batch sampling log
  wood <- readRDS("../r_files/wood.RDS")        # wood sampling log
  cal_1 <- readRDS("../r_files/cal_1.RDS")       # meta data
  data_1 <- readRDS("../r_files/data_1.RDS")      # meta data
  cal_2 <- readRDS("../r_files/cal_2.RDS")       # meta data
  data_2 <- readRDS("../r_files/data_2.RDS")      # meta data
```

# Stove-fuel combinations

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  test_list <- dplyr::filter(samples, type == "SF") %>%
               dplyr::select(id, stove, fuel, type) %>%
               dplyr::group_by(stove, fuel) %>%
               dplyr::summarise(id = paste(id, collapse = ","))

  ggplot(test_list, aes(y=stove, x=fuel)) + 
    geom_tile(colour = "white", width= 0.9, height = 0.9, aes(fill = id)) +
    scale_fill_discrete(na.value = 'grey95') +
    theme_minimal() +
    theme(legend.position="none") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95)) +
    theme(axis.text = element_text(size = 14)) +
    geom_text(aes(label = id, size = 8)) +
    xlab("") + ylab("")
```

# Clean and save metadata

## batch stove weights

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  batch_wgts <- dplyr::filter(tidy_id_date(batch, "^wgt_", "wgt_"), !is.na(value))

  saveRDS(batch_wgts, file="../r_files/batch_wgts.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(batch_wgts, var) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  ggplot(p_data, aes(x = factor(var), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 4) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

## batch stove timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  batch_times <- dplyr::filter(tidy_id_date(batch, "^time_", "time_"), !is.na(value))

  saveRDS(batch_times, file="../r_files/batch_times.RDS")
```
  
 * unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(batch_times, value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

Outliers shown on graph (18A, 16B, 17D, 23D, 17A, 25B, 25C, 16C) were all  checked by KF in February 2017; no transcription errors exist nor unusual (out of expected range) values.

## batch lab conditions

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  batch_lab <- dplyr::filter(tidy_id_date(batch, "^lab_", "lab_"), !is.na(value))

  saveRDS(batch_lab, file="../r_files/batch_lab.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=12, fig.height=6}
  p_data <- dplyr::group_by(batch_lab, var) %>%
            dplyr::mutate(value_norm =
                         (value - mean(value, na.rm = TRUE)) / sd(value, na.rm = TRUE),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  p_box <- ggplot(p_data, aes(x = factor(var), y = value_norm)) +
           geom_boxplot() +
           geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
           theme_minimal() +
           ylab("z score normalized value") +
           xlab("") +
           theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
           theme(axis.text.y = element_text(size = 30),
           axis.title=element_text(size=40))

  p_hist <- ggplot(p_data, aes(x = value)) +
            geom_histogram(binwidth = 15) +
            theme_minimal() +
            facet_wrap(~var, scales = "free") +
            xlab("")

  grid.arrange(p_hist, p_box, ncol = 2)
```

Outliers shown on graph (22A, 16C, 23A [p] / 29A, 30A [rh] / 25A, 25B [temperature]) were all  checked by KF in February 2017; no transcription errors exist nor unusual (out of expected range) values.

## batch pot number

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  batch_pot <- dplyr::filter(tidy_id_date(batch, "^pot_", "pot_"),
                             !is.na(as.numeric(value))) %>%
               dplyr::mutate(value = as.factor(value)) %>%
               dplyr::rename(pot = value, rep = var)

  batch_pot$pot <- forcats::fct_recode(batch_pot$pot,
                                   a = "1", b = "2", c = "3", d = "4")

  saveRDS(batch_pot, file="../r_files/batch_pot.RDS")
```

## carbonyl flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  carb_flows <- dplyr::filter(tidy_id_date(data_2, "_carb_.*[^avg]$", "_carb_"), !is.na(value))
  
  carb_flows <- split_flows(carb_flows)
  
  saveRDS(carb_flows, file="../r_files/carb_flows.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(carb_flows, type) %>%
            dplyr::mutate(value_norm =
                           (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value),
                                    as.character(id), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

Outlier 16B checked by KF in February 2017. There is no transcription error, but this measure is QC-flagged as bad due to an issue with the carbonyl line being clogged/having known sampling issues that impacted flow rates. 

## carbonyl timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  carb_times <- dplyr::filter(tidy_id_date(data_2, "^time_.*carb$", "^time_"), !is.na(value))
  
  carb_times <- split_times(carb_times)
  
  saveRDS(carb_times, file="../r_files/carb_times.RDS")
```

* unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(batch_times, value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

## isokinetic sampler flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  iso_flows <- dplyr::filter(tidy_id_date(data_2, "_iso_.*[^avg]$", "_iso_"), !is.na(value))
  
  iso_flows <- split_flows(iso_flows)
  
  saveRDS(iso_flows, file="../r_files/iso_flows.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(iso_flows, type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

Outlier 1C checked by KF in February 2017. There is no transcription error and the value is within an expected normal range. 

## wood stove weights

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  wood_wgts <- dplyr::filter(tidy_id_date(wood, "^wgt_", "wgt_"), !is.na(value))
  
  saveRDS(wood_wgts, file="../r_files/wood_wgts.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(wood_wgts, var) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  ggplot(p_data, aes(x = factor(var), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 4) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

Outliers 10A, 6D, 5A, 2A, 1A checked by KF in February 2017. There are no transcription errors. 

Outliers 6A and 13A were transcription errors that were fixed in wood sampling log on 2/18/17 so should not appear in subsequent re-run versions of this file.

## wood stove times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  wood_times <- dplyr::filter(tidy_id_date(wood, "^time_", "time_"), !is.na(value))

  saveRDS(wood_times, file="../r_files/wood_times.RDS")
```

 * unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(wood_times, value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```


## wood lab conditions

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  wood_lab <- dplyr::filter(tidy_id_date(wood, "^lab_", "lab_"), !is.na(value))
  
  saveRDS(wood_lab, file="../r_files/wood_lab.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=12, fig.height=6}
  p_data <- dplyr::group_by(wood_lab, var) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(id), NA))

  p_box <- ggplot(p_data, aes(x = factor(var), y = value_norm)) +
           geom_boxplot() +
           geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
           theme_minimal() +
           ylab("z score normalized value") +
           xlab("") +
           theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
           theme(axis.text.y = element_text(size = 30),
           axis.title=element_text(size=40))

  p_hist <- ggplot(p_data, aes(x = value)) +
            geom_histogram(binwidth = 15) +
            theme_minimal() +
            facet_wrap(~var, scales = "free") +
            xlab("")

  grid.arrange(p_hist, p_box, ncol = 2)
```

All outliers checked by KF in February 2017. There are no transcription errors. 

## wood pot number

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  wood_pot <- dplyr::filter(tidy_id_date(wood, "^pot_", "pot_"),
                             !is.na(as.numeric(value))) %>%
               dplyr::mutate(value = as.factor(value)) %>%
               dplyr::rename(pot = value, rep = var)

  wood_pot$pot <- forcats::fct_recode(wood_pot$pot,
                                   a = "1", b = "2", c = "3", d = "4")
  
  saveRDS(wood_pot, file="../r_files/wood_pot.RDS")
```


## fivegas calibration standards

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}    
  fivegas_cal_conc <- dplyr::filter(tidy_date(cal_1, "^conc_", "conc_"), !is.na(value))
  
  saveRDS(fivegas_cal_conc, file="../r_files/fivegas_cal_conc.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(fivegas_cal_conc, var) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(var), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

Outlier date of 3-11-16: No transcription error, but KF thinks there was a mistake in the original data recording. It should have been written as 49.9 but instead was written as 49.5.

## fivegas calibration times

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}    
  fivegas_cal_times <- dplyr::filter(tidy_date(cal_1, "^time_", "time_"),
                                     !is.na(value))

  fivegas_cal_times <- split_fivegas_cal_times(fivegas_cal_times)

  saveRDS(fivegas_cal_times, file="../r_files/fivegas_cal_times.RDS")
```

 * unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(fivegas_cal_times,
                            value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

## filter cartridge flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  filter_flows <- dplyr::filter(tidy_id_date(data_1,
                                  "^preflow_.*[^avg]$|postflow.*[^avg]$", ""),
                                  !is.na(value))

  filter_flows <- split_filter_flows(filter_flows)

  saveRDS(filter_flows, file="../r_files/filter_flows.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=16}
  p_data <- dplyr::group_by(filter_flows, type, colour) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40)) +
    facet_wrap(~colour, scales = "free")
```

## filter cartridge timestamps

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  filter_times <- dplyr::filter(tidy_id_date(data_1, "^time_.*cart", "time_"), !is.na(value))

  filter_times <- split_filter_times(filter_times)

  saveRDS(filter_times, file="../r_files/filter_times.RDS")
```

 * unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(filter_times,
                            value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

Outlier 23A was a transcription error (AM vs PM) that was fixed on 2/18/17. G3 is correct (~5-7pm test).

## co2 (dilution) calibration

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  co2_lab_cal <- dplyr::filter(tidy_date(cal_2, "^sensor_1", "sensor_1_"), !is.na(value))

  co2_lab_cal <- split_co2_cal(co2_lab_cal)

  saveRDS(co2_lab_cal, file="../r_files/co2_lab_cal.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(co2_lab_cal, type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40)) +
    ggtitle("lab")
```

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  co2_sample_cal <- dplyr::filter(tidy_date(cal_2, "^sensor_2", "sensor_2_"), !is.na(value))
  
  co2_sample_cal <- split_co2_cal(co2_sample_cal)
  
  saveRDS(co2_sample_cal, file = "../r_files/co2_sample_cal.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(co2_sample_cal, type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40)) +
    ggtitle("sample line")
```

## pax flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  pax_flows <- dplyr::filter(tidy_date(cal_2, "^preflow_pax|^postflow_pax", ""),
                             !is.na(value))
  
  pax_flows <- split_pax_flows(pax_flows)
  
  saveRDS(pax_flows, file = "../r_files/pax_flows.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::filter(pax_flows, loc == "inlet") %>%
            dplyr::group_by(type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

## smps flows

```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  smps_flows <- dplyr::filter(tidy_date(cal_2, "^preflow_smps|^postflow_smps", ""),
                              !is.na(value))

  smps_flows <- split_flows(smps_flows)
  
  saveRDS(smps_flows, file="../r_files/smps_flows.RDS")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=FALSE, fig.width=20, fig.height=12}
  p_data <- dplyr::group_by(smps_flows, type) %>%
            dplyr::mutate(value_norm = (value - mean(value)) / sd(value),
                          outlier = ifelse(is_outlier(value), as.character(date), NA))

  ggplot(p_data, aes(x = factor(type), y = value_norm)) +
    geom_boxplot() +
    geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 8) +
    theme_minimal() +
    ylab("z score normalized value") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 35, hjust = 0.95, size = 30)) +
    theme(axis.text.y = element_text(size = 30),
          axis.title=element_text(size=40))
```

## smps / pax times


```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide', cache=FALSE}
  smps_pax_bg_times <- dplyr::filter(tidy_id_date(data_2,
                                                  "^time_.*smps.*", "^time_"),
                                                  !is.na(value))

  smps_pax_bg_times <- split_times_smps_pax(smps_pax_bg_times)

  saveRDS(smps_pax_bg_times, file="../r_files/smps_pax_bg_times.RDS")
```

* unexpected values:

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  outliers <- dplyr::filter(smps_pax_bg_times,
                            value < 60 * 60 * 4 | value > 60 * 60 * 19)

  knitr::kable(outliers, "markdown", digits = 2)
```

# Summary

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='markup', cache=FALSE}
  count_samples <- dplyr::tally(group_by(samples, type))

  count_stoves <- dplyr::tally(group_by(samples, stove)) 
  
  count_fuels <- dplyr::tally(group_by(samples, fuel))
```

A total of `r (as.numeric(count_samples[1,2]) + as.numeric(count_samples[2,2]))` tests were performed between `r min(samples$date)` and `r max(samples$date)`. A total of `r nrow(count_stoves)` stoves types were tested. A total of `r nrow(count_fuels)` fuels were tested (on the appropriate stove).
