---
title: "Process carbonyl data"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  if (!require("pacman")) install.packages("pacman")
  pacman::p_load(tidyverse, kableExtra)
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/',
                        fig.width = 8, fig.height = 8,
                        warning = FALSE, message = FALSE, cache = FALSE)
```

# tidy data

## load files

load scripts

```{r}
  source("../../scripts/functions.R")
```

load data

```{r}
  carbonyls <- readRDS("../../r_data/ions.RDS")
  carb_flows <- readRDS("../../r_data/carb_flows.RDS")
  carb_times <- readRDS("../../r_data/carb_times.RDS")
  notes <- readRDS("../../r_data/notes.RDS")
  samples <- readRDS("../../r_data/samples.RDS")
```

load schemes

```{r}
  qc_colors <- readRDS("../../r_data/qc_colors.RDS")
  lod_shapes <- readRDS("../../r_data/lod_shapes.RDS")
```

## extract carbonyls 

## remove empty measurement

```{r}
  carbonyls <- carbonyls[-46,]
```

```{r}
  carbs <- 
    carbonyls %>%
    dplyr::select(27:42, type, id) %>%
    dplyr::filter(!grepl("^27|^28|^29|^30", id)) %>%
    dplyr::filter(type == "test" | type == "bg")
```

## reformat

convert to long format

```{r}
  carbs <- 
    carbs %>%
    tidyr::gather("pol", "val",
                  -id, -type) %>%
    dplyr::filter(!is.na(pol)) %>%
    dplyr::mutate(pol = as.factor(pol)) %>%
    dplyr::select(id, pol, val, type)
        
```

## flags

extract flags

```{r}
  notes <- 
    notes %>%
    dplyr::filter(grepl("carb|carbonyl|carbonyls|all", inst) == TRUE)
```

bad preceeds maybe preceeds good

```{r}
  flags <-
    notes %>%
    dplyr::select(id, qc) %>%
    dplyr::group_by(id) %>%
    dplyr::arrange(qc) %>%
    dplyr::summarise(qc = first(qc))
```

add flag by id

```{r}
  carbs <-
    carbs %>%
    dplyr::left_join(flags, by = "id") %>%
    dplyr::mutate(id = as.factor(id)) %>%
    dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

## test type 

define test type variable

```{r}
  carbs <- 
    carbs %>%
    dplyr::rename("test_type" = "type")
```

# missing and bad data

## table: missing data

```{r, echo = FALSE}
  dplyr::select(samples, id, stove, fuel) %>%
    dplyr::anti_join(carbs %>%
                     dplyr::select(id) %>%
                     dplyr::distinct(),
                   by = "id") %>%
    dplyr::filter(!grepl("^27|^28|^29|^30", id)) %>%
    knitr::kable("markdown", align = 'c')
```

## table: bad data

```{r, echo=FALSE}
  notes %>%
    dplyr::filter(grepl("carb|carbonyl|carbonyls|all", inst) == TRUE,
                  qc != "ok") %>%
    dplyr::select(-inst) %>%
    dplyr::filter(!grepl("^27|^28|^29|^30", id)) %>%
    knitr::kable("markdown", align = 'c')
```

notes: 

* 1C: dilution line and carbonyl clogged - bad
* 2B: ozone scrubber behind carbonyl - bad
* 8C: clog in sample line - bad
* 16B: flow issues - bad

* G2: unclear notes - check if outlier
* 2A: fire went out during test - check if outlier
* 2C: start up period twice - check if outlier
* 6B: possible clog in sample line - check if outlier
* 6C: pumps may have turned off during test - check if outlier
* 8A: possible flow issue - check if outlier
* 23A: fire relit with kerosene and match during refuel - check if outlier

## remove bad tests

```{r}
  carbs_merged <-
    carbs %>% 
    dplyr::filter(qc != "bad")
```

# concentration calculations

## extract flow data and average

```{r}
  flows <- 
    carb_flows %>%
    dplyr::select(-date) %>%
    dplyr::group_by(id, type) %>%
    dplyr::summarise(flow = mean(value, na.rm = TRUE)) %>%
    dplyr::group_by(id) %>%
    dplyr::summarise(flow = mean(flow, na.rm = TRUE))
```

notes: no flow data appear to be missing

## extract time data

```{r}
  times <- 
    carb_times %>%
    dplyr::select(-date) %>%
    tidyr::spread(type, value)
```

notes: no time data appear to be missing

## merge and calculate

merge data with flows and times

```{r}
  carbs_merged <-
    carbs_merged %>%
    dplyr::left_join(flows, by = "id") %>%
    dplyr::left_join(times, by = "id")
```

## flag filters below lod

define lod

```{r}
  lod <- 0.1 # micrograms per cubic meter
```

```{r}
  carbs_merged <- 
    carbs_merged %>%
    dplyr::mutate(lod = ifelse(is.na(val), "below", "above"))
```

## calculate concentrations

```{r}
  carbs_merged <-
    carbs_merged %>%
    dplyr::mutate(dur = (end - start) / 60,
                  val = val * 1000 / (flow * dur))
```

## replace NA values with lod / sqrt(2)

```{r}
  carbs_merged <- 
    carbs_merged %>%
    dplyr::mutate(val = ifelse(grepl("below", lod), 0.1 / sqrt(2), val))
```

## save lod

```{r}
  carbs_lod <- 
    carbs_merged %>%
    dplyr::select(pol) %>%
    dplyr::distinct() %>%
    dplyr::mutate(lod = lod)

  saveRDS(carbs_lod, file = "../../r_data/carbonyls_lod.RDS")
```

# backgrounds

```{r}
  bg <-
    carbs_merged %>%
    dplyr::filter(test_type == "bg")
```

## plot: lab background outliers

outliers (z > 3) when backgrounds are grouped by pollutant

```{r, echo=FALSE}
  bg %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(pol) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, pol, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    dplyr::rename("carbonyl type" = "pol") %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

notes: none are outliers of interest - proceed

## plot: time series of lab backgrounds

```{r carbonyls_bg_timeseries, echo=FALSE}
  bg %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::left_join(dplyr::select(samples, id, date), by = 'id') %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(color = qc, shape = lod), size = 2) +
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      theme_bw() +
      theme(text = element_text(size = 12),
            legend.position = "top") +
      facet_wrap(~pol, ncol = 4, scales = "free") +
      ylab("concentration (micrograms per meters cubed)") +
      xlab("")
```

conclusions: 

* acetone has the highest background concentration by far
* backgrounds look good! 
* no obvious trends in time

## table: percent lab backgrounds with NA values

```{r, echo=FALSE}
  bg %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

## table: background summary

summarize backgrounds

```{r}
  bg <-
    bg %>%
    dplyr::group_by(pol) %>%
    dplyr::summarise(mean = mean(val, na.rm = TRUE),
                     median = median(val, na.rm = TRUE),
                     stdev = sd(val, na.rm = TRUE),
                     min = min(val, na.rm = TRUE),
                     max = max(val, na.rm = TRUE),
                     count = n()) %>%
    dplyr::mutate(mean = ifelse(is.nan(mean), NA, mean),
                  median = ifelse(is.nan(median), NA, median),
                  stdev = ifelse(is.nan(stdev), NA, stdev),
                  min = ifelse(is.infinite(min), NA, min),
                  max = ifelse(is.infinite(max), NA, max),
                  count = ifelse(is.nan(count), NA, count))
```

## table: summary of backgrounds (ug per meter cubed)

```{r, echo=FALSE}
  bg %>%
    knitr::kable(digits = 2, align = 'c')
```

## save backgrounds

```{r}
  saveRDS(bg, file = "../../r_data/carbonyls_bg.RDS")
```

## plot: background and burn concentrations

```{r carbonyls_test_bg_conc, echo=FALSE}
  carbs_merged %>%
    dplyr::group_by(id, pol, test_type) %>%
    ggplot(aes_string(y = "val", x = "test_type")) +
      geom_jitter(aes(color = qc, shape = lod), size = 2) + 
      geom_boxplot(outlier.colour = NA, alpha = 0.0) +
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      scale_y_log10() +
      theme_bw() +
      facet_wrap(~pol, ncol = 4, scales = "free") +
      theme(text = element_text(size = 12),
            legend.position = "top") +
      ylab("concentration (micrograms per cubic meter)") +
      xlab("")
```

conclusions: 

* some overlap between burn and background concentrations
* need to investigate ways to account for background measurements (for now, subtract the median)

## subract background

```{r}
  carbs_merged <-
    carbs_merged %>%
    dplyr::filter(test_type == "test")
```

```{r}
  carbs_merged <-
    carbs_merged %>%
    dplyr::left_join(bg, by = c("pol"))
```

```{r}
  carbs_merged <-
    carbs_merged %>%
    dplyr::mutate(val = val - median) %>%
    dplyr::select(-min, -max, -count, -mean, -median, -stdev)
```

## table: percentage of measurements below the background median

```{r, echo=FALSE}
  carbs_merged %>%
    dplyr::left_join(dplyr::select(samples, id, stove, fuel), by = "id") %>%
    #dplyr::filter(stove == "lpg") %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., val <= 0))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```
## create a new below background variable

```{r}
  carbs_merged <-
    carbs_merged %>%
    dplyr::mutate(bg = ifelse(val <= 0, "below", "above"),
                  val = ifelse(bg == "below", NA, val))
```

# burn analysis

## plot: time series of burns

```{r carbonyls_samples_timeseries, echo=FALSE}
 carbs_merged %>%
    dplyr::filter(test_type == "test") %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::left_join(dplyr::select(samples, id, date), by = 'id') %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(color = qc, shape = lod), size = 2) +
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      theme_bw() +
      theme(text = element_text(size = 12),
            legend.position = "top") +
      facet_wrap(~pol, ncol = 4, scales = "free") +
      ylab("concentration (micrograms per cubic meter)") +
      xlab("")
```

notes:

## plot: concentrations by stove type

```{r carbonyls_stovetype, echo=FALSE, fig.height=25}
  carbs_merged_p <-
    carbs_merged %>%
    dplyr::left_join(dplyr::select(samples, id, stove, fuel, stovecat, fuelcat), by = "id")

  carbs_merged_p %>%
    ggplot(aes_string(x = "stove", y = "val")) +
      geom_jitter(aes(color = qc, shape = lod), size = 2) + 
      geom_boxplot(outlier.colour = NA, alpha = 0.0) +
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      theme_bw() +
      #scale_y_log10() +
      theme(text = element_text(size = 12),
            axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            legend.position = "top") +
      facet_grid(pol~fuelcat, scales = "free", space = "free_x") +
      ylab("concentration (micrograms per cubic meter)") +
      xlab("")
```

notes:

## table: outliers by stove type

outliers (z > 3) when samples are grouped by pollutant and stove type

```{r, echo=FALSE}
  carbs_merged_p %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(pol, stove) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE))/ sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, pol, stove, fuel, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

notes: not outlier of interest - proceed

## plot: concentrations by fuel type

```{r carbonyls_fueltype, echo=FALSE, fig.height=25}
  carbs_merged_p %>%
    ggplot(aes_string(x = "fuel", y = "val")) +
      geom_jitter(aes(color = qc, shape = lod), size = 2) + 
      geom_boxplot(outlier.colour = NA, alpha = 0.0) +
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      theme_bw() +
      #scale_y_log10() +
      theme(text = element_text(size = 12),
            axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            legend.position = "top") +
      facet_grid(pol~fuelcat, scales = "free", space = "free_x") +
      ylab("concentration (micrograms per cubic meter)") +
      xlab("")
```

conclusions: 

## table: outliers by fuel type

outliers (z > 3) when samples are grouped by pollutant and fuel type

```{r, echo=FALSE}
  carbs_merged_p %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(pol, fuel) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE))/ sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, pol, stove, fuel, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

notes: 

* 1A seems to be a consistent outlier across pollutants - proceed for now
* 8A appears to be an outlier as well - proceed for now

## table: percent of burns with NA values

```{r, echo=FALSE}
  carbs_merged %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## save data

```{r}
  carbs_merged <-
    carbs_merged %>%
    dplyr::select(id, pol, val, dur, lod, qc, bg) %>%
    dplyr::mutate(pol_category = "carbonyls")

  saveRDS(carbs_merged, file = "../../r_data/carbonyls_merged.RDS")
```

## table: missing data at end

```{r, echo = FALSE}
  dplyr::select(samples, id, stove, fuel) %>%
  dplyr::anti_join(carbs_merged %>%
                     dplyr::distinct(),
                   by = "id") %>%
  dplyr::filter(!grepl("^G", id)) %>%
  dplyr::filter(!grepl("^27|^28|^29|^30", id)) %>%
  knitr::kable("markdown", align = 'c')
```

## table: percent of data below background

```{r, echo=FALSE}
  carbs_merged %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., bg == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```