---
title: "Process voc data"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  if (!require("pacman")) install.packages("pacman")
  pacman::p_load(tidyverse, kableExtra)
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/', 
                        fig.width = 12, fig.height = 5,
                        warning = FALSE, message = FALSE, cache = FALSE)
```

# tidy data

## load files

load scripts

```{r}
  source("../../scripts/functions.R")
```

load data

```{r}
  fivegas_conc <- readRDS("../../r_data/fivegas_conc.RDS") 
  fivegas_volts <- readRDS("../../r_data/fivegas_volts.RDS")
  fivegas_calibration <- readRDS("../../r_data/fivegas_calibration.RDS")
  cal_1 <- readRDS("../../r_data/cal_1.RDS")
  notes <- readRDS("../../r_data/notes.RDS")
  samples <- readRDS("../../r_data/samples.RDS")
  pol_properties <- readRDS("../../r_data/pol_properties.RDS")
```

## concentration data

convert concentration data to longer format

```{r}
  fivegas_conc <- 
    fivegas_conc %>%
    dplyr::select(-datetime_secs, -time_str) %>%
    tidyr::gather("pol", "val", ch4:co)
```

## extract vocs 

```{r}
  vocs <- 
    vocs %>%
    dplyr::select(id, type, 5:59) %>%
    dplyr::filter(type == "test" | type == "bg", id != "P8") %>%
    dplyr::rename("voc_n_nonane" = "voc_n_no_8888ne")
```

## reformat

* convert to long format

```{r}
  vocs <- 
    vocs %>%
    tidyr::gather("pol", "val",
                  -id, -type) %>%
    dplyr::filter(!is.na(pol)) %>%
    dplyr::mutate(pol = as.factor(pol)) %>%
    dplyr::select(id, pol, val, type) %>%
    dplyr::mutate(pol = gsub("voc_", "", pol))
        
```

## lod

```{r}
  vocs <-
    vocs %>%
    dplyr::left_join(voc_lod, by = c("pol" = "voc")) %>%
    dplyr::rename("voc_lod" = "lod") %>%
    dplyr::mutate(lod = as.factor(ifelse(val == -8888, "below", "above"))) %>%
    dplyr::mutate(val = ifelse(val == -8888, voc_lod / sqrt(2), val)) %>%
    dplyr::select(-voc_lod)
```

## flags

* extract flags

```{r}
  notes <- 
    notes %>%
    dplyr::filter(grepl("vocs|all", inst) == TRUE)
```

* bad preceeds maybe preceeds good

```{r}
  flags <-
    notes %>%
    dplyr::select(id, qc) %>%
    dplyr::group_by(id) %>%
    dplyr::arrange(qc) %>%
    dplyr::summarise(qc = first(qc))
```

* add flag by id

```{r}
  vocs <-
    vocs %>%
    dplyr::left_join(flags, by = "id") %>%
    dplyr::mutate(id = as.factor(id)) %>%
    dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

## test type 

define test type variable

```{r}
  vocs <- 
    vocs %>%
    dplyr::rename("test_type" = "type")
```

# missing and bad data

## table: missing data

```{r, echo = FALSE}
  dplyr::select(samples, id, stove, fuel) %>%
    dplyr::anti_join(vocs %>%
                     dplyr::select(id) %>%
                     dplyr::distinct(),
                   by = "id") %>%
    knitr::kable("markdown", align = 'c')
```

## table: tests with bad measurements

```{r, echo=FALSE}
  notes %>%
    dplyr::filter(grepl("voc|all", inst) == TRUE,
                  qc != "ok") %>%
    dplyr::select(-inst) %>%
    knitr::kable("markdown", align = 'c')
```

## remove bad tests

```{r}
  vocs_merged <-
    vocs %>% 
    dplyr::filter(qc != "bad")
```

# concentration calculations

## extract data

extract time data

```{r}
  times <- 
    times %>%
    dplyr::mutate(dur = difftime(datetime_end, datetime_start)) %>%
    dplyr::select(id, dur)
```

## merge and calculate

merge data with flows and times

```{r}
  vocs_merged <-
    vocs_merged %>%
    dplyr::mutate(flow = flow) %>%
    dplyr::left_join(times, by = "id")
```

calculate concentrations

```{r}
  vocs_merged <-
    vocs_merged %>%
    dplyr::left_join(dplyr::select(pol_properties, pol, mw), by = "pol") %>%
    dplyr::mutate(val = convert_ppmv_ugpmc(val / 1000, mw, 25, 85),
                  units = "ug per meter cubed") %>%
    dplyr::select(-mw)
```

# backgrounds

extract background

```{r}
  bg <-
    vocs_merged %>%
    dplyr::filter(test_type == "bg")
```

## plot: lab background outliers

outliers (z > 3) when backgrounds are grouped by pollutant and filter type

```{r, echo=FALSE, fig.height = 12}
  bg %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(pol) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, pol, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    dplyr::rename("voc type" = "pol") %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

conclusions: 

* 

## plot: time series of lab backgrounds

background concentrations (ug per meter cubed)

* shape indicates if sample was above or below instrument lod

```{r vocs_bg_timeseries, echo=FALSE, fig.height = 30, fig.width =16}
 bg %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::left_join(dplyr::select(samples, id, date), by = 'id') %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(color = qc), size = 4) +
      scale_shape(solid = FALSE) +
      guides(fill = FALSE) +
      theme_bw() +
      theme(text = element_text(size = 20),
            legend.position = "top") +
      facet_wrap(~pol, ncol = 4, scales = "free") +
      ylab("") +
      xlab("")
```

conclusions: 

* 

## table: lab blanks below lod for chemical analysis

percent of lab blanks below the lod for chemical analysis

```{r, echo=FALSE}
  bg %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    dplyr::rename("voc type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

## table: lab backgrounds with NA values

percent of lab backgrounds with NA values

```{r, echo=FALSE}
  bg %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    dplyr::rename("voc type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

## table: background summary

summarize backgrounds

```{r}
  bg <-
    bg %>%
    dplyr::group_by(pol) %>%
    dplyr::summarise(mean = mean(val, na.rm = TRUE),
                     median = median(val, na.rm = TRUE),
                     stdev = sd(val, na.rm = TRUE),
                     min = min(val, na.rm = TRUE),
                     max = max(val, na.rm = TRUE),
                     count = n()) %>%
    dplyr::mutate(mean = ifelse(is.nan(mean), NA, mean),
                  median = ifelse(is.nan(median), NA, median),
                  stdev = ifelse(is.nan(stdev), NA, stdev),
                  min = ifelse(is.infinite(min), NA, min),
                  max = ifelse(is.infinite(max), NA, max),
                  count = ifelse(is.nan(count), NA, count))
```

summary of backgrounds (ug per meter cubed)

```{r, echo=FALSE}
  bg %>%
    knitr::kable(digits = 2, align = 'c')
```

## save backgrounds

```{r}
  saveRDS(bg, file = "../../r_data/vocs_bg.RDS")
```


# sample analysis

extract burn data

```{r}
  vocs_samples <-
    vocs_merged %>%
    dplyr::filter(test_type == "test")
```

```{r}
  vocs_samples <-
    vocs_samples %>%
    dplyr::left_join(bg, by = c("pol"))
```

## plot: time series of burns

burn concentrations (ug per meter cubed)

* samples are not background corrected

```{r vocs_samples_timeseries, echo=FALSE, fig.height = 30, fig.width = 16}
 vocs_samples %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::left_join(dplyr::select(samples, id, date), by = 'id') %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(color = qc), size = 4) +
      scale_shape(solid = FALSE) +
      guides(fill = FALSE) +
      theme_bw() +
      theme(text = element_text(size = 20),
            legend.position = "top") +
      facet_wrap(~pol, ncol = 4, scales = "free") +
      ylab("") +
      xlab("")
```

conclusions:

*

## plot: comparing background and sample concentrations

test and background concentrations (ug per meter cubed)

```{r vocs_test_bg_val, echo=FALSE, fig.height = 50, fig.width = 16}
  vocs_merged %>%
  dplyr::filter(pol != "n_nonane") %>%
    dplyr::group_by(id, pol, test_type) %>%
    ggplot(aes_string(y = "val", x = "test_type")) +
      geom_boxplot(outlier.colour = NA) +
      geom_jitter(aes(color = qc), size = 3) +
      theme_bw() +
      scale_y_log10() +
      theme(text = element_text(size = 20),
            legend.position = "top") +
      ylab("concentration") +
      xlab("") + 
      facet_wrap(~pol, ncol = 4, scales = "free")
```

conclusions: 

* need to investigate ways to account for background measurements (for now, subtract the median)

```{r}
  vocs_samples <-
    vocs_samples %>%
    dplyr::mutate(val = val - median)
```

## table: measurements below the background median

percentage of test concentrations below background concentration

```{r, echo=FALSE}
  vocs_samples %>%
    dplyr::left_join(dplyr::select(samples, id, stove, fuel), by = "id") %>%
    #dplyr::filter(stove == "lpg") %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., val < 0))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    dplyr::rename("voc type" = "pol") %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

## plot: concentrations by stove type

concentrations (ug per meter cubed)

* samples are background corrected

```{r vocs_stovetype, echo=FALSE, fig.height = 70, fig.width = 16}
  vocs_samples_p <-
    vocs_samples %>%
    dplyr::left_join(dplyr::select(samples, id, stove, fuel, stovecat, fuelcat), by = "id") %>%
    dplyr::mutate(stove = 
                    stove %>% 
                    tolower %>%
                    gsub(" [:(:].*[:):]", "", .) %>%
                    forcats::fct_relevel("three stone fire",
                                         "clay ring",
                                         "ceramic charcoal",
                                         "rocket elbow",
                                         "built-in justa",
                                         "metal charcoal",
                                         "fan rocket elbow",
                                         "gasifier",
                                         "wick kerosene",
                                         "pressure kerosene",
                                         "lpg"))

  vocs_samples_p %>%
    ggplot(aes_string(x = "stove", y = "val")) +
      geom_boxplot(outlier.colour = NA) +
      geom_jitter(aes(color = qc), size = 3,  width = 0.2) + 
      scale_shape(solid = FALSE) +
      guides(fill = FALSE) +
      theme_bw() +
      theme(text = element_text(size = 20),
            axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            legend.position = "top") +
      facet_wrap(~pol, ncol = 2, scales = "free") +
      ylab("") +
      xlab("")
```

conclusions:

* 

## table: outliers by stove type

outliers (z > 3) when samples are grouped by pollutant and stove type

```{r, echo=FALSE}
  vocs_samples_p %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(pol, stove) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE))/ sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, pol, stove, fuel, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    dplyr::rename("voc type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

conclusions:

* no repeat tests recommended

## plot: concentrations by fuel type

concentrations (ug per meter cubed):

* samples are background corrected

```{r vocs_fueltype, echo=FALSE, fig.height = 70, fig.width = 16}
  vocs_samples_p %>%
    ggplot(aes_string(x = "fuel", y = "val")) +
      geom_boxplot(outlier.colour = NA) +
      geom_jitter(aes(color = qc), size = 3,  width = 0.2) + 
      scale_shape(solid = FALSE) +
      guides(fill = FALSE) +
      theme_bw() +
      theme(text = element_text(size = 20),
            axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            legend.position = "top") +
      facet_wrap(~pol, ncol = 2, scales = "free") +
      ylab("") +
      xlab("")
```

conclusions: 

* 

## table: outliers by fuel type

outliers (z > 3) when samples are grouped by pollutant and fuel type

```{r, echo=FALSE}
  vocs_samples_p %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(pol, fuel) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE))/ sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, pol, stove, fuel, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    dplyr::rename("voc type" = "pol") %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

conclusions: 

* 

## table: burns with NA values

percent of burn samples with NA values

```{r, echo=FALSE}
  vocs_samples %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    dplyr::rename("voc type" = "pol") %>%
    knitr::kable("markdown", align = 'c')
```

## save data

```{r}
  saveRDS(vocs_merged, file = "../../r_data/vocs_merged.RDS")
```

# conclusions
