---
title: "Process gravimetric data"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  if (!require("pacman")) install.packages("pacman")
  pacman::p_load(tidyverse, kableExtra)
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/', 
                        warning = FALSE, message = FALSE, cache = FALSE)
```

# tidy data

## load files

load scripts

```{r}
  source("../../scripts/functions.R")
```

load data

```{r}
  grav <- readRDS("../../r_data/grav.RDS")
  times <- readRDS("../../r_data/filter_times.RDS")
  flows <- readRDS("../../r_data/filter_flows.RDS")
  notes <- readRDS("../../r_data/notes.RDS")
  samples <- readRDS("../../r_data/samples.RDS")
```

load schemes

```{r}
  qc_colors <- readRDS("../../r_data/qc_colors.RDS")
  lod_shapes <- readRDS("../../r_data/lod_shapes.RDS")
```

## flags

extract flags

```{r}
  notes <- 
    notes %>%
    dplyr::filter(grepl("grav|all", inst) == TRUE)
```

bad preceeds maybe preceeds good

```{r}
  flags <-
    notes %>%
    dplyr::select(id, qc) %>%
    dplyr::group_by(id) %>%
    dplyr::arrange(qc) %>%
    dplyr::summarise(qc = first(qc))
```

add flag by id

```{r}
  grav <-
    grav %>%
    dplyr::left_join(flags, by = "id") %>%
    dplyr::mutate(id = as.factor(id)) %>%
    dplyr::mutate(qc = as.factor(ifelse(is.na(qc), "ok", as.character(qc))))
```

## test type 

define test type variable

```{r}
  grav <- 
    grav %>%
    dplyr::filter(!grepl("^27|^28|^29|^30", id)) %>%
    dplyr::mutate(test_type = ifelse(!grepl("^B|^G", id), "test", "bg"),
                  test_type = ifelse(grepl("^B", id), "blank", test_type)) %>%
    dplyr::filter(!grepl("MC", id))
```

## calculate mass

pre-weights

```{r}
  grav_pre <- 
    grav %>%
    dplyr::select(wgt_pre_1, wgt_pre_2, wgt_pre_3, id) %>%
    tidyr::gather("var", "val", wgt_pre_1, wgt_pre_2, wgt_pre_3) %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(id) %>%
    dplyr::summarise(wgt_pre = mean(val))
```

post-weights

```{r}
  grav_post <- 
    grav %>%
    dplyr::select(wgt_post_1, wgt_post_2, wgt_post_3, id) %>%
    tidyr::gather("var", "val", wgt_post_1, wgt_post_2, wgt_post_3) %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(id) %>%
    dplyr::summarise(wgt_post = mean(val))
```

* merge weights and calculate change in mass
* correct for weigh room blank filter weight

```{r}
  grav <- 
    grav %>%
    dplyr::left_join(grav_pre, by = "id") %>%
    dplyr::left_join(grav_post, by = "id") %>%
    dplyr::mutate(wgt_delta = wgt_post - wgt_pre,
                  wgt_cal_delta = wgt_cal_post - wgt_cal_pre,
                  wgt_blank_delta = wgt_blank_avg_post - wgt_blank_avg_pre,
                  val = wgt_delta - wgt_blank_delta) %>%
    dplyr::filter(!is.na(id))
```

## plot: calibration weight by date

```{r pm_calibration_wgt_timeseries}
  grav %>%
    ggplot(aes(x = date_pre, y = wgt_cal_pre)) +
      geom_point(shape = 1) +
      geom_point(aes(x = date_post, y = wgt_cal_post), shape = 1, size = 2) +
      theme_bw() +
      theme(text = element_text(size = 12),
            legend.position = "top") +
    ylab("mass (micrograms)") +
    xlab("")
```

## plot: scale room blank by date

```{r pm_wgtoom_filter_timeseries, echo=FALSE}
  grav %>%
    ggplot(aes(x = date_pre, y = wgt_blank_avg_pre)) +
      geom_point(shape = 1) +
      geom_point(aes(x = date_post, y = wgt_blank_avg_post), shape = 1, size = 2) +
      theme_bw() +
      theme(text = element_text(size = 12),
            legend.position = "top") +
    ylab("mass (micrograms)") +
    xlab("")
```

add pollutant type

```{r}
  grav <-
    grav %>%
    dplyr::mutate(pol = "pm_mass") %>%
    dplyr::select(id, val, qc, test_type, pol)
```

# missing and bad data

## table: missing data

```{r, echo = FALSE}
  dplyr::select(samples, id, stove, fuel) %>%
    dplyr::anti_join(grav %>%
                     dplyr::select(id) %>%
                     dplyr::distinct(),
                   by = "id") %>%
    dplyr::filter(!grepl("^27|^28|^29|^30", id)) %>%
    knitr::kable("markdown", align = 'c')
```

notes: never post weighed, because filter ruptured

## table: bad data

```{r, echo=FALSE}
  notes %>%
    dplyr::filter(grepl("grav|all", inst) == TRUE,
                  qc != "ok") %>%
    dplyr::select(-inst) %>%
    dplyr::filter(!grepl("^27|^28|^29|^30", id)) %>%
    knitr::kable("markdown", align = 'c')
```

notes: 

* 2A: fire went out - check if outlier
* 2C: filter switched but should be ok now, also had flow issues, also start up period twice - check if outlier
* 2E: possible filter clog - check if outlier
* 3A: white rotometer malfunctioning - check if outlier
* 4C: flow problem at end of test - check if outlier
* 7A: possible contamination - check if outlier
* 8C: lots of test problems - check if outlier
* 9A: possible transcription error - check if outlier
* 14C: filter switched but should be ok now - check if outlier
* 23A: fire went out - check if outlier
* G2: filter ruptured, not sure which one - check if outlier - ok
* G8: possibly switched - check if outlier - ok

* 15A: contaminated - mark as bad
* 2D: filter contamination - mark as bad
* 3A: flow unknown for most of test - mark as bad
* 17A: filter ruptured - mark as bad
* G5: filter contamination or switched - mark as bad
* G7: unresolved transcription error - mark as bad
* G14: filter contamination - mark as bad

* 1A: filter ruptured, not sure which one - ecoc artifact ruptured - ignore

## mark additional tests as bad

```{r}
  grav <- 
    grav %>%
    dplyr::mutate(qc = as.character(qc),
                  qc = ifelse(grepl("^2D|^3A|^17A|G5|G14|G7|^15A", id), "bad", qc))
```

## remove bad tests

```{r}
  grav_merged <-
    grav %>% 
    dplyr::filter(qc != "bad")
```

# blank analysis 

extract blanks

```{r}
  blanks <-
    grav_merged %>%
    dplyr::filter(test_type == "blank")
```

## table: lab blank outliers

outliers (z > 3) when blanks are grouped by pollutant and filter type

```{r, echo=FALSE}
  blanks %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::select(id, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    knitr::kable("markdown", align = 'c')
```

notes: not that much of an outlier - proceed

## plot: time series of lab blanks

```{r pm_blank_timeseries, echo=FALSE}
 blanks %>%
    dplyr::left_join(dplyr::select(samples, blank, date), by = c('id' = 'blank')) %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(color = qc), size = 2) +
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      theme_bw() +
      theme(text = element_text(size = 12),
            legend.position = "top") +
      ylab("mass (micrograms)")
```

```{r, echo=FALSE}
  acf(blanks$val)
```

notes:

## table: lab blanks with NA values

percent of lab blanks with NA values

```{r, echo=FALSE}
  blanks %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## summarize blanks

```{r}
  blanks <-
    blanks %>%
    dplyr::summarise(mean = mean(val, na.rm = TRUE),
                     median = median(val, na.rm = TRUE),
                     stdev = sd(val, na.rm = TRUE),
                     lod = stdev * 3,
                     loq = stdev * 10)
```

## table: blank summary

```{r, echo=FALSE}
  blanks %>%
    knitr::kable(digits = 2, align = 'c')
```

## save blanks

```{r}
  saveRDS(blanks, file = "../../r_data/pm_lod.RDS")
```

remove lab blanks from merged file

```{r}
  grav_merged <-
    grav_merged %>%
    dplyr::filter(test_type != "blank") # remove blanks
```

## subtract blank mean

```{r}
  grav_merged <-
    grav_merged %>%
    dplyr::mutate(val = val - blanks$mean)
```

## lod

add lod variable and replace values below lod with lod/sqrt(2)

```{r}
  grav_merged <-
    grav_merged %>%
    dplyr::mutate(lod = ifelse(val <= blanks$lod, "below", "above"),
                  val = ifelse(val <= blanks$lod, blanks$lod / sqrt(2), val))
```

# calculate concentrations

## extract flow data

```{r}
  flows <- 
    flows %>%
    dplyr::select(-date) %>%
    dplyr::filter(colour == "white") %>%
    dplyr::group_by(id, type, colour) %>%
    dplyr::summarise(flow = mean(value, na.rm = TRUE)) %>%
    dplyr::group_by(id, colour) %>%
    dplyr::summarise(flow = mean(flow, na.rm = TRUE)) %>%
    dplyr::select(-colour)
```

notes: no flow data appear to be missing

## extract time data

```{r}
  times <-
    times %>%
    dplyr::select(-date) %>%
    dplyr::filter(color == "white") %>%
    tidyr::spread(type, value) %>%
    dplyr::select(-color)
```

notes: no time data appear to be missing

## merge data with flows and times

```{r}
  grav_merged <-
    grav_merged %>%
    dplyr::left_join(flows, by = "id") %>%
    dplyr::left_join(times, by = "id")
```

notes: all weights matched up with times and flows - no data missing

## calculate concentrations

```{r}
  grav_merged <-
    grav_merged %>%
    dplyr::mutate(dur = (end - start) / 60,
                  val = val * 1000 / (flow * dur))
```

# backgrounds

```{r}
  bg <-
    grav_merged %>%
    dplyr::filter(test_type == "bg")
```

## plot: lab background outliers

outliers (z > 3) when backgrounds are grouped by pollutant

```{r echo=FALSE}
  bg %>%
    dplyr::group_by(pol) %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::mutate(z_score = (val- mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    knitr::kable("markdown", align = 'c')
```

notes: no outliers

## plot: time series of lab backgrounds

```{r pm_bg_timeseries, echo=FALSE}
 bg %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::left_join(dplyr::select(samples, id, date), by = 'id') %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(shape = lod, color = qc), size = 2) +
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      theme_bw() +
      theme(text = element_text(size = 12),
            legend.position = "top") +
      ylab("concentrations (micrograms per meter cubed)") +
      xlab("")
```

notes: G10 is high, but no notes regarding measurement - remove

```{r}
  bg <-
    bg %>%
    dplyr::filter(id != "G10")
```

```{r, echo=FALSE}
  acf(bg$val)
```

notes: no trend in time

## table: percent of lab backgrounds below study lod

```{r, echo=FALSE}
  bg %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## table: percent of lab backgrounds with NA values

```{r, echo=FALSE}
  bg %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## table: background summary

summarize backgrounds

```{r}
  bg <-
    bg %>%
    dplyr::group_by(pol) %>%
    dplyr::summarise(mean = mean(val, na.rm = TRUE),
                     median = median(val, na.rm = TRUE),
                     stdev = sd(val, na.rm = TRUE),
                     min = min(val, na.rm = TRUE),
                     max = max(val, na.rm = TRUE),
                     count = n()) %>%
    dplyr::mutate(mean = ifelse(is.nan(mean), NA, mean),
                  median = ifelse(is.nan(median), NA, median),
                  stdev = ifelse(is.nan(stdev), NA, stdev),
                  min = ifelse(is.infinite(min), NA, min),
                  max = ifelse(is.infinite(max), NA, max),
                  count = ifelse(is.nan(count), NA, count))
```

## table: background summary

```{r, echo=FALSE}
  bg %>%
    knitr::kable(digits = 2, align = 'c')
```

notes: reasonable levels for indoor air in a combustion lab

## save backgrounds

```{r}
  saveRDS(bg, file = "../../r_data/pm_bg.RDS")
```

## plot: background and burn concentrations

```{r ecoc_burn_bg_conc, echo=FALSE, fig.height=4, fig.width=4}
  grav_merged %>%
    dplyr::group_by(id, test_type) %>%
    dplyr::filter(!is.na(val)) %>%
    ggplot(aes_string(y = "val", x = "test_type")) +
      geom_boxplot(outlier.colour = NA) +
      geom_jitter(aes(color = qc, shape = lod), size = 2) +
      theme_bw() +
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      #scale_y_log10() +
      theme(text = element_text(size = 12),
            legend.position = "top") +
      ylab("concentration (microgram per meter cubed)") +
      xlab("")
```

notes: some burns have low concentrations - opt to subtract median background from all burn data

## subtract background

```{r}
  grav_merged <-
    grav_merged %>%
    dplyr::filter(test_type == "test")
```

```{r}
  grav_merged <-
    grav_merged %>%
    dplyr::left_join(bg, by = c("pol"))
```

```{r}
  grav_merged <-
   grav_merged %>%
    dplyr::mutate(val = val - median) %>%
    dplyr::select(-min, -max, -count, -mean, -median, -stdev)
```

## table: percentage of measurements below the background median

```{r, echo=FALSE}
  grav_merged %>%
    dplyr::left_join(dplyr::select(samples, id, stove, fuel), by = "id") %>%
    #dplyr::filter(stove == "lpg") %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., val <= 0))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c', digits = 2)
```

## create a new below background variable

```{r}
  grav_merged <-
   grav_merged %>%
    dplyr::mutate(bg = ifelse(val <= 0, "below", "above"),
                  val = ifelse(bg == "below", NA, val))
```

# burn analysis

## plot: time series of burns

```{r ecoc_burn_timeseries, echo=FALSE}
  grav_merged %>%
    dplyr::filter(test_type == "test") %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::left_join(dplyr::select(samples, id, date), by = 'id') %>%
    ggplot(aes_string(x = "date", y = "val")) +
      geom_point(aes(shape = lod, color = qc), size = 2) +
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      theme_bw() +
      theme(text = element_text(size = 16),
            legend.position = "top") +
      labs(shape = "lod") +
      ylab("concentration (microgram per meter cubed)") +
      xlab("")
```

## plot: concentrations by stove type

```{r pm_stovetype, echo=FALSE, fig.width=8, fig.height=4}
  grav_merged_p <-
    grav_merged %>%
    dplyr::left_join(dplyr::select(samples, id, stove, fuel, stovecat, fuelcat), by = "id")

  grav_merged_p %>%
    ggplot(aes_string(x = "stove", y = "val")) +
      geom_boxplot(outlier.colour = NA) +
      geom_jitter(aes(color = qc, shape = lod), size = 2) + 
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      theme_bw() +
      theme(text = element_text(size = 12),
            axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            legend.position = "top") +
      facet_grid(~fuelcat, scales = "free", space = "free_x") +
      ylab("concentration (micrograms per cubic meter)") +
      xlab("")
```

notes: 

## table: outliers by stove type

outliers (z > 3) when samples are grouped by pollutant and stove type

```{r, echo=FALSE}
  grav_merged_p %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(stove, pol) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, pol, stove, fuel, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    knitr::kable("markdown", align = 'c')
```

notes:

## plot: concentrations by fuel type

```{r pm_fueltype, echo=FALSE, fig.width=8, fig.height=4}
  grav_merged_p %>%
    ggplot(aes_string(x = "fuel", y = "val")) +
      geom_boxplot(outlier.colour = NA) +
      geom_jitter(aes(color = qc, shape = lod), size = 2) + 
      scale_shape_manual(values = lod_shapes) +
      scale_color_manual(values = qc_colors) +
      theme_bw() +
      theme(text = element_text(size = 12),
            axis.text.x = element_text(angle = 15, hjust = 1, vjust = 1),
            legend.position = "top") +
      facet_grid(~fuelcat, scales = "free", space = "free_x") +
      ylab("concentration (micrograms per cubic meter)") +
      xlab("")
```

notes:

## table: outliers by fuel type

outliers (z > 3) when samples are grouped by pollutant and fuel type

```{r, echo=FALSE}
  grav_merged_p %>%
    dplyr::filter(!is.na(val)) %>%
    dplyr::group_by(fuel, pol) %>%
    dplyr::mutate(z_score = (val - mean(val, na.rm = TRUE)) / sd(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::select(id, stove, fuel, z_score) %>%
    dplyr::filter(z_score > 3) %>%
    knitr::kable("markdown", align = 'c')
```

notes:

## table: percent of burns below study lod

```{r, echo=FALSE}
  grav_merged %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., lod == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## table: percent of burns with NA values

```{r, echo=FALSE}
  grav_merged %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., is.na(val) == TRUE))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```

## save data<

```{r}
  grav_merged <-
    grav_merged %>%
    dplyr::select(id, pol, val, dur, lod, bg, qc) %>%
    dplyr::mutate(pol_category = "pm_mass")

  saveRDS(grav_merged, file = "../../r_data/pm_merged.RDS")
```

## table: missing data at end

```{r, echo = FALSE}
  dplyr::select(samples, id, stove, fuel) %>%
  dplyr::anti_join(grav_merged %>%
                     dplyr::distinct(),
                   by = "id") %>%
  dplyr::filter(!grepl("^G", id)) %>%
  dplyr::filter(!grepl("^27|^28|^29|^30", id)) %>%
  knitr::kable("markdown", align = 'c')
```

## table: percent of data below background

```{r, echo=FALSE}
  grav_merged %>%
    dplyr::group_by(pol) %>%
    dplyr::do(val = round(nrow(filter(., bg == "below"))/nrow(.), 2) * 100) %>%
    dplyr::mutate(val = cell_spec(val, "html", color = ifelse(val > 10, "red", "blue"))) %>%
    knitr::kable("markdown", align = 'c')
```
