---
title: "figure 1: particle emissions"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_libraries, echo=FALSE, message=FALSE}
  if (!require("pacman")) install.packages("pacman")
  pacman::p_load(tidyverse, pals, devtools, patchwork)
  #devtools::install_github("thomasp85/patchwork")
```

```{r global_options, echo=FALSE}
  knitr::opts_chunk$set(fig.path ='figures/',
                        warning = FALSE, message = FALSE, cache = FALSE)
```

# tidy data

load data

```{r}
  emissions <- readRDS("../../../r_data/emissions_factors.RDS")
  samples <- readRDS("../../../r_data/samples.RDS")
```

## format stove names

```{r}
  samples <-
    samples %>%
    dplyr::mutate(stove = 
                    stove %>% 
                    factor %>% 
                    gsub(" ", "\n",.) %>% 
                    forcats::fct_relevel("three\nstone\nfire",
                                         "clay\nring",
                                         "ceramic\ncharcoal",
                                         "rocket\nelbow",
                                         "built-in\nplancha",
                                         "metal\ncharcoal",
                                         "fan\nrocket\nelbow",
                                         "gasifier",
                                         "wick\nkerosene",
                                         "pressure\nkerosene",
                                         "lpg"))
```

## add stove fuel info

```{r}
  emissions <-
    emissions %>%
    dplyr::left_join(samples %>%
                       dplyr::select(id, stove, fuel, stovecat, fuelcat), by = "id")
```

## rename pollutants

```{r}
  emissions <-
    emissions %>%
    dplyr::mutate(pol_category = ifelse(grepl("^oc$", pol_category), "organic carbon", pol_category),
                  pol_category = ifelse(grepl("^ec$", pol_category), "elemental carbon", pol_category),
                  pol_category = ifelse(grepl("ions", pol_category), "inorganic ions", pol_category),
                  pol = ifelse(grepl("^co$", pol), "carbon monoxide", pol))
```

## select pollutants

```{r}
  ultrafines <-
    emissions %>%
    dplyr::filter(grepl("ultrafines", pol)) %>%
    dplyr::filter(metric == "energy_delivered_ef") %>%
    dplyr::mutate(pol_category = "ultrafine particles",
                  pol = "ultrafine particles")
```

```{r}
  ec_oc_ions <-
    emissions %>%
    dplyr::filter(grepl("inorganic ions|elemental carbon|organic carbon", pol_category)) %>%
    dplyr::filter(metric == "energy_delivered_ef")
```

```{r}
  pm_mass <-
    emissions %>%
    dplyr::filter(grepl("pm_mass", pol_category)) %>%
    dplyr::filter(metric == "energy_delivered_ef") %>%
    dplyr::mutate(pol_category = ifelse(grepl("pm_mass", pol_category), "PM2.5 mass", pol_category))
```

```{r}
  co <-
    emissions %>%
    dplyr::filter(grepl("monoxide", pol)) %>%
    dplyr::filter(metric == "energy_delivered_ef")
```

# check for outliers

## elemental carbon

## carbon monoxide

```{r, echo=FALSE, fig.width=10, fig.height=8}
  co %>%
    ggplot(aes(x = id, y = val, color = fuel)) +
    geom_point() +
    theme_bw() +
    theme(text = element_text(size = 12),
          legend.position = "top") +
    facet_wrap(~stove, scales = "free")
```

```{r, echo=FALSE, fig.width=10, fig.height=8, eval=FALSE}
  ultrafines %>%
    ggplot(aes(x = id, y = val, color = fuel)) +
    geom_point() +
    theme_bw() +
    theme(text = element_text(size = 12),
          legend.position = "top") +
    facet_wrap(~stove, scales = "free")
```

## elemental carbon

```{r, echo=FALSE, fig.width=10, fig.height=8, eval=FALSE}
  ec_oc_ions %>%
    dplyr::filter(pol_category == "elemental carbon") %>%
    ggplot(aes(x = id, y = val, color = fuel)) +
    geom_point() +
    theme_bw() +
    theme(text = element_text(size = 12),
          legend.position = "top") +
    facet_wrap(~stove, scales = "free")
```

## organic carbon

```{r, echo=FALSE, fig.width=10, fig.height=8, eval=FALSE}
  ec_oc_ions %>%
    dplyr::filter(pol_category == "organic carbon") %>%
    ggplot(aes(x = id, y = val, color = fuel)) +
    geom_point() +
    theme_bw() +
    theme(text = element_text(size = 12),
          legend.position = "top") +
    facet_wrap(~stove, scales = "free_x")
```

## pm mass

```{r, echo=FALSE, fig.width=10, fig.height=8, eval=FALSE}
  pm_mass %>%
    ggplot(aes(x = id, y = val, color = fuel)) +
    geom_point() +
    theme_bw() +
    theme(text = element_text(size = 12),
          legend.position = "top") +
    facet_wrap(~stove, scales = "free_x")
```

## ions

```{r, echo=FALSE, fig.width=10, fig.height=8, eval=FALSE}
  ec_oc_ions %>%
    dplyr::filter(pol_category == "inorganic ions") %>%
    dplyr::group_by(id) %>%
    dplyr::mutate(val = sum(val, na.rm = TRUE)) %>%
      ggplot(aes(x = id, y = val, color = fuel)) +
      geom_point() +
      theme_bw() +
      theme(text = element_text(size = 12),
            legend.position = "top") +
      facet_wrap(~stove, scales = "free")
```

## remove tests

```{r}
  ec_oc_ions <-
    ec_oc_ions %>%
    dplyr::filter(id != "6E")
```

## average pm, ec, oc, and ions by stove type

## calculate range of pm mass

```{r}
  range_pm <-
    pm_mass %>%
    na.omit %>%
    dplyr::group_by(stove, stovecat) %>%
    dplyr::summarise(min = min(val, na.rm = TRUE),
                     max = max(val, na.rm = TRUE)) %>%
    tidyr::gather("var", "val", min, max) %>%
    dplyr::mutate(sym = ifelse(var == "min", "[", "]"))
```

```{r}
  pm_mass <-
    pm_mass %>%
    dplyr::group_by(pol_category, stove, stovecat) %>%
    dplyr::summarise(val = mean(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(pol = "PM2.5 mass",
                  pol_category = "PM2.5 mass")
```

multiply by 1.8 to convert organic carbon to organic matter

```{r}
  ec_oc_ions <-
    ec_oc_ions %>%
    dplyr::mutate(val = ifelse(pol_category == "organic carbon", val * 1.8, val)) %>%
    dplyr::group_by(id, pol_category, stove, fuel, stovecat, fuelcat) %>%
    dplyr::summarise(val = sum(val, na.rm = TRUE)) %>%
    tidyr::spread(pol_category, val) %>%
    na.omit %>%
    tidyr::gather(pol_category, val, 6:8) %>%
    dplyr::group_by(pol_category, stove, stovecat, fuelcat) %>%
    dplyr::summarise(val = mean(val, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(pol = pol_category,
                  pol_category = "PM2.5 mass")
```

```{r}
  ultrafines <-
    ultrafines %>%
    dplyr::group_by(stove, stovecat) %>%
    dplyr::summarise(mean = mean(val, na.rm = TRUE),
                     min = min(val, na.rm = TRUE),
                     max = max(val, na.rm = TRUE)) %>%
    dplyr::mutate(pol_category = "ultrafine particles")
```

```{r}
  co <-
    co %>%
    dplyr::group_by(pol, stove, stovecat) %>%
    dplyr::mutate(val = val / 1000) %>%
    dplyr::summarise(mean = mean(val, na.rm = TRUE),
                     min = min(val, na.rm = TRUE),
                     max = max(val, na.rm = TRUE))
```

# figure 1a: particle mass balance

```{r, echo=FALSE}
  p1 <-
    pm_mass  %>%
    dplyr::filter(grepl("traditional", stovecat)) %>%
    ggplot(aes(x = stove, y = val)) +
    geom_histogram(aes(fill = pol), stat = "identity") +
    geom_histogram(data = ec_oc_ions %>%
                   dplyr::filter(grepl("traditional", stovecat)),
                   aes(x = stove, y = val, fill = pol), stat = "identity", width = 0.75) +
    #scale_fill_discrete(breaks = c("organic carbon", "elemental carbon", "inorganic ions", "total PM mass"))
    scale_fill_manual(values = c("gray30", "#E69F00", "snow", "#999999")) +
    geom_text(data = range_pm %>%
              dplyr::filter(grepl("traditional", stovecat)),
              aes(x = stove, y = val, label = sym), size = 10, angle = 90, nudge_x = -0.02) +
    theme_bw() +
    theme(text = element_text(size = 20),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 18),
          axis.text.x = element_blank(),
          legend.position = "none") +
    guides(fill = guide_legend(title = NULL)) +
    facet_wrap(~stovecat, scales = "free") +
    ylab("mg emitted per MJ delivered") +
    xlab("")

  p2 <-
    pm_mass %>%
    dplyr::filter(grepl("improved", stovecat)) %>%
    ggplot(aes(x = stove, y = val)) +
      geom_histogram(aes(fill = pol), stat = "identity") +
      geom_histogram(data = ec_oc_ions %>%
                     dplyr::filter(grepl("improved", stovecat)),
                     aes(x = stove, y = val, fill = pol), stat = "identity", width = 0.75) +
      scale_fill_manual(values = c("gray30", "#E69F00", "snow", "#999999")) +
      scale_y_continuous(limits = c(0, 1000)) +
      geom_text(data = range_pm %>%
                dplyr::filter(grepl("improved", stovecat)),
                aes(x = stove, y = val, label = sym), size = 10, angle = 90, nudge_x = -0.04) +
      theme_bw() +
      theme(text = element_text(size = 20),
            axis.text.y = element_text(size = 18),
            axis.title.y = element_blank(),
            axis.text.x = element_blank(),
            legend.position = "none") +
      guides(fill = guide_legend(title = NULL)) +
      facet_wrap(~stovecat, scales = "free") +
      xlab("")
  
  p3 <-
    pm_mass %>%
    dplyr::filter(grepl("liquid", stovecat)) %>%
    ggplot(aes(x = stove, y = val)) +
      geom_histogram(aes(fill = pol), stat = "identity") +
      geom_histogram(data = ec_oc_ions %>%
                       dplyr::filter(grepl("liquid", stovecat)),
                     aes(x = stove, y = val, fill = pol),
                     stat = "identity", width = 0.75) +
      scale_fill_manual(values = c("gray30", "#E69F00", "snow", "#999999")) +
      geom_text(data = range_pm %>%
                  dplyr::filter(grepl("liquid", stovecat)),
                aes(x = stove, y = val, label = sym),
                stroke = 2, size = 10, angle = 90, nudge_x = -0.02) +
      theme_bw() +
      scale_y_continuous(limits = c(0, 25)) +
      theme(text = element_text(size = 20),
            axis.text.y = element_text(size = 18),
            axis.title.y = element_blank(),
            axis.text.x = element_blank(),
            legend.position = "right",
            legend.text = element_text(size = 12)) +
      guides(fill = guide_legend(title = NULL)) +
      facet_grid(pol_category~stovecat, scales = "free") +
      xlab("")
```

```{r, echo=FALSE}
  p4 <- 
    ultrafines %>%
    dplyr::filter(grepl("traditional", stovecat)) %>%
      ggplot(aes(x = stove, y = mean)) +
        geom_errorbar(aes(ymin = min, ymax = max), width = 0, size = 1) +
        geom_point(aes(shape = pol_category), size = 5, stroke = 2, color = "#56B4E9") +
        scale_shape_manual(values = c(8)) +
        theme_bw() +
        scale_y_continuous(trans = "log10", limits = c(1e5, 1e8), breaks = c(1e5, 1e6, 1e7, 1e8)) +
        theme(text = element_text(size = 20),
              axis.text.y = element_text(size = 18),
              axis.title.y = element_text(size = 14),
              axis.text.x = element_blank(),
              strip.text.x = element_blank(),
              legend.position = "none") +
        guides(fill = guide_legend(title = NULL)) +
        facet_wrap(~stovecat, scales = "free") +
        xlab("") +
        ylab("num. emitted per MJ delivered")

  p5 <- 
    ultrafines %>%
    dplyr::filter(grepl("improved", stovecat)) %>%
      ggplot(aes(x = stove, y = mean)) +
        geom_errorbar(aes(ymin = min, ymax = max), width = 0, size = 1) +
        geom_point(aes(shape = pol_category), size = 5, stroke = 2, color = "#56B4E9") +
        scale_shape_manual(values = c(8)) +
        theme_bw() +
        scale_y_continuous(trans = "log10", limits = c(1e5, 1e8), breaks = c(1e5, 1e6, 1e7, 1e8)) +
        theme(text = element_text(size = 20),
              axis.text.y = element_text(size = 18),
              axis.text.x = element_blank(),
              strip.text.x = element_blank(),
              axis.title.y = element_blank(),
              legend.position = "none") +
        guides(fill = guide_legend(title = NULL)) +
        facet_wrap(~stovecat, scales = "free") +
        xlab("")
  p6 <- 
    ultrafines %>%
    dplyr::filter(grepl("liquid", stovecat)) %>%
      ggplot(aes(x = stove, y = mean)) +
        geom_errorbar(aes(ymin = min, ymax = max), width = 0, size = 1) +
        geom_point(aes(shape = pol_category), size = 5, stroke = 2, color = "#56B4E9") +
        scale_shape_manual(values = c(8)) +
        theme_bw() +
        scale_y_continuous(trans = "log10", limits = c(1e5, 1e8), breaks = c(1e5, 1e6, 1e7, 1e8)) +
        theme(text = element_text(size = 20),
              axis.text.y = element_text(size = 18),
              strip.text.x = element_blank(),
              axis.text.x = element_blank(),
              legend.title = element_blank(),
              axis.title.y = element_blank(),
              legend.position = "none",
              legend.text = element_text(size = 12)) +
        guides(fill = guide_legend(title = NULL)) +
        facet_grid(pol_category~stovecat, scales = "free") +
        xlab("")
```

```{r, echo=FALSE}
  p7 <- 
    co %>%
    dplyr::filter(grepl("traditional", stovecat)) %>%
      ggplot(aes(x = stove, y = mean)) +
        geom_errorbar(aes(ymin = min, ymax = max), width = 0, size = 1) +
        geom_point(aes(shape = pol), size = 5, color = "#009E73") +
        scale_shape_manual(values = c(19)) +
        scale_y_continuous(limits = c(0, 30)) +
        theme_bw() +
        theme(text = element_text(size = 20),
              axis.text.y = element_text(size = 18),
              axis.title.y = element_text(size = 14),
              strip.text.x = element_blank(),
              legend.position = "none") +
        guides(fill = guide_legend(title = NULL)) +
        facet_wrap(~stovecat, scales = "free") +
        xlab("") +
        ylab("g emitted per MJ delivered")

  p8 <- 
    co %>%
    dplyr::filter(grepl("improved", stovecat)) %>%
      ggplot(aes(x = stove, y = mean)) +
        geom_errorbar(aes(ymin = min, ymax = max), width = 0, size = 1) +
        geom_point(aes(shape = pol), size = 5, color = "#009E73") +
        scale_shape_manual(values = c(19)) +
        theme_bw() +
        scale_y_continuous(limits = c(0, 30)) +
        theme(text = element_text(size = 20),
              axis.text.y = element_text(size = 18),
              strip.text.x = element_blank(),
              axis.title.y = element_blank(),
              legend.position = "none") +
        guides(fill = guide_legend(title = NULL)) +
        facet_wrap(~stovecat, scales = "free") +
        xlab("")
  
  p9 <- 
    co %>%
    dplyr::filter(grepl("liquid", stovecat)) %>%
      ggplot(aes(x = stove, y = mean)) +
        geom_errorbar(aes(ymin = min, ymax = max), width = 0, size = 1) +
        geom_point(aes(shape = pol), size = 5, color = "#009E73") +
        scale_shape_manual(values = c(19)) +
        scale_y_continuous(limits = c(0, 4)) +
        theme_bw() +
        theme(text = element_text(size = 20),
              axis.text.y = element_text(size = 18),
              strip.text.x = element_blank(),
              legend.title = element_blank(),
              axis.title.y = element_blank(),
              legend.position = "none",
              legend.text = element_text(size = 12)) +
        guides(fill = guide_legend(title = NULL)) +
        facet_grid(pol~stovecat, scales = "free") +
        xlab("")
```

```{r figure_1b, echo=FALSE, fig.width=16, fig.height=10}
 (p1 + p2 + p3) + p4 + p5 + p6 + p7 + p8 + p9 + plot_layout(ncol = 3, widths = c(1, 1.5, 1), heights = c(1.25, 1, 1))
```